### 2021-01-26 notes/notes.tex
Take home.

\documentclass{report}
\usepackage{graphicx}
\usepackage{breakcites}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage[margin=1.5cm]{geometry}
\bibliographystyle{apalike}

% Remember to start reftex-mode

\begin{document}

\begin{chapter}{Random ideas}
\section{Straight movements as a sign of learning}
According to Gurevich\_93 (see \cite{Wolpert_Are_1995}) and \cite{Shadmehr_Adaptive_1994}, the more a person practices a movement, the straighter this movement becomes. Could we use the curvature of a movement as a measure of how much they've learned this movement?

This could be useful for the context-dependent learning paradigm. The question to answer is: if a participant learns context A fully (and their movements become more straight), when switching back to A, will the first movement be straight or curved? If it's straight, then we can link this to the idea of context-based learning.

NOTE: Gurevich\_93 is a doctoral thesis that doesn't seem to be available.
\end{chapter}

\begin{chapter}{Paper}
In this chapter, I describe the everything related to the paper that should come out of this project, including modeling and experimental findings, as well as the direction of the paper and things needed to get there.

\section{Project description}
The idea of the project is the following: We will create an agent that learns to control an arm in environment A (say, the real world carrying nothing), and learns how to control the arm in environment B (say, with high winds). The agent then learns to recognize environments A and B and to deploy the right controller for each environment depending on this recognition.

We will argue that this is how humans do it and present evidence (based on literature research) for the predictions this model makes, as well as try to make new predictions.

After this methods paper is done, model validation can be carried out in future studies.

\section{Planning}
Here I enumerate the things I need to learn and do in order to get this project moving and to finish it.

\begin{enumerate}
\item (1.5 weeks) Learn about published models on arm-movement planning, especially those that can be used as generative models.
\item (0.5 weeks) Learn about measures for how practiced a movement is. Reaction times? Speed profiles? Trajectories?
\item (2-3 weeks)Implement the best model/s and do some testing as to its capabilities.
\item (1 weeks) Add the context inference as another level in the hierarchy.
\item (2 weeks) Run some exploratory simulations
\item (1 week + meeting)Talk about possible predictions that can be made with the model for later experimental verification.
\item (Months?) Write the paper
\item Profit!

\end{enumerate}

\end{chapter}

\begin{chapter}{Modeling}
Notes from \cite{Harris_Signaldependent_1998}:
\begin{enumerate}
\item Check the following models: Nelson 83, Todorov and Jordan 02, Hamilton and Wolpert 02, van Beers at al 02, 04, Enderle and Wolfe 87, Happee 92
\end{enumerate}

I should read Knill and Pouget 2004 for Bayesian calculations in neurons.

\end{chapter}

\begin{chapter}{Physiology and neurons}
A few notes from \cite{Wolpert_Are_1995}:
\begin{enumerate}
\item Hand paths become straight with practice (Gurevich 93, Shadmehr and Musa-Ivaldi 94)
\item Deafferented patients move differently (Ghez 90)
\item Bell-shaped speed emerges from dynamics of control, not from planning (Jordan 94)
\item Curvature matches the participant's messed up perception of straight lines (wolpert 94)

\end{enumerate}

Notes from \cite{Harris_Signaldependent_1998}:
\begin{enumerate}
\item Arm stiffness can be adapted (Burdet et al 01.
\end{enumerate}

Notes from \cite{Bays_Computational_2007}:
\begin{enumerate}
\item We have the ability to recall previously learned dynamics even after months! (Bashers-Krug et al 96, Gandolfo et al 96).
\end{enumerate}

Notes from \cite{Shadmehr_Adaptive_1994}:
\begin{enumerate}
\item When planning movement, the target is transformed from a retinocentric vector to a head-centered and finally a shoulder-centered one. (Andersen\_85, Soechting and Flanders\_91).
\item Gordon\_93 says that the target is finally represented as a vector from the current hand positoin (or whatever the effector is) to the goal position.
\item From Lackner\_and\_Dizio\_92, aftereffects exist when introduced Coriolis forces are withdrawn. They also exist when perception is mucked with with prism glasses (Held\_59, 62, 63).
\end{enumerate}

\end{chapter}

\begin{chapter}{Notes on papers}
Conclusions and useful things about different papers I read. This differs from the comments above in that the notes here are holistic.

\section{\cite{Shadmehr_Adaptive_1994}}
In this paper they show a number of things. First, the show that people can adapt the dynamics of their control states to match changing dynamics of the observers. They also show that this adaptation takes the form of a straightening of trajectories, which are very twisted when they are first introduced to new system dynamics (viscosity of the environment); participants learn to do straight lines in the new environment after some trials, and the learing is monotonically increasing. They show a model of the dynamics of the arm, the control states and the adaptation that happens when exposed to the new dynamics. Finally, they show, both experimentally and in their model, the aftereffects of the new dynamics, whereby they display bad behavior when they go back to the natural environment, but eventually adapt to it as well.

Interesting paper that cound be the foundation of our models, assuming that it hasn't been improved upon and maybe entirely replaced by newer stuff.

\end{chapter}

\bibliography{../MyLibrary}

\end{document}

### 2021-02-17 notes/notes.tex
Added description of new model

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{tabularx}
 \usepackage[margin=1.5cm]{geometry}
@@
 \begin{document}
+
+\begin{chapter}{Glossary}
+\begin{enumerate}
+\item Error-clamp: Refers to the phase of a simulation in the models of motor adaptation where the error that drives learning is set to zero. This emulates a no-feedback experimental condition.
+\end{enumerate}
+\end{chapter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 NOTE: Gurevich\_93 is a doctoral thesis that doesn't seem to be available.
+
+\section{Quick adaptation to perturbations in the irrelevant manifold.}
+Models for motor control can be divided into two categories: (1) optimal trajectory and (2) end point.
+Optimal trajectory are those in which the controller tries to stick to a previously-calculated optimal trajectory and any perturbation (internal, external) will be dealt with by trying to go back to the optimal trajectory.
+Endpoint models, on the other hand, do not care that much about the trajectory itself and, with models that follow the ``minimal intervention'' principle, will only correct for perturbations if they mess with the endpoint.
+
+Endpoint models are more recent and arguably better than optimal trajectory ones.
+However, there is one thorn on their side: if during a movement an arm is perturbed, it will immediately try to return to the optimal trajectory \citep{Bizzi_Posture_1984}, even if the perturbation does not detract from the endpoint (this last part is my speculation).
+\cite{Todorov_Optimality_2004} argues that this is due to impedance, as a sort of immediate dumb fighback.
+
+A nice low-hanging fruit project would be to test whether \cite{Todorov_Optimality_2004} is right by having participants perform a reaching task in real life, where a motor creates random perturbations (in the irrelevant manifold, for good measure) and compare their reactions to the same task in virtual reality where the perturbation happens in the VR environment, but not in real time.
+If participants adapt in the same way in both environments, then \cite{Todorov_Optimality_2004} is wrong.
+
 \end{chapter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\begin{enumerate}
-\item (1.5 weeks) Learn about published models on arm-movement planning, especially those that can be used as generative models.
-\item (0.5 weeks) Learn about measures for how practiced a movement is. Reaction times? Speed profiles? Trajectories?
-\item (2-3 weeks)Implement the best model/s and do some testing as to its capabilities.
-\item (1 weeks) Add the context inference as another level in the hierarchy.
-\item (2 weeks) Run some exploratory simulations
-\item (1 week + meeting)Talk about possible predictions that can be made with the model for later experimental verification.
-\item (Months?) Write the paper
-\item Profit!
+\begin{tabularx}{\textwidth}{
+  l|
+  >{\hsize=.4\hsize\linewidth=\hsize}X|
+  >{\hsize=.2\hsize\linewidth=\hsize}X|
+  >{\hsize=.2\hsize\linewidth=\hsize}X}
+% {l|{\hsize=.4\hsize}X|{\hsize=0.3\hsize}X|{\hsize=0.3\hsize}X}
+End Date & Milestone & A priori comments & A posteriori comments \\ \hline \hline
+12.02.2021 & Learn about published models on arm-movement planning, especially those that can be used as generative models & & \\ \hline
+19.02.2021 & Learn about measures for how practiced a movement is. Reaction times? Speed profiles? Trajectories? & & \\\hline
+12.03.2021 & Implement the best model/s and do some testing as to its capabilities. & Without inference &   \\ \hline
+19.03.2021 & Add the sequential goals as another level in the hierarchy. & Still no inference. & \\ \hline
+02.04.2021 & Run some exploratory simulations & How well does the model behave? Are there bad cases? & \\ \hline
+09.04.2021 & Talk a  bout possible predictions that can be made with the model. & This includes a meeting. & \\ \hline
+--- & Run final simulations and write paper. & & \\ \hline
 
-\end{enumerate}
+\end{tabularx}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{chapter}{Modeling}
-Notes from \cite{Harris_Signaldependent_1998}:
+\section{Random thoughts}
+I should read Knill and Pouget 2004 for Bayesian calculations in neurons.
+
+\section{Categorizing models}
+
+\section{Models to implement}
+\subsection{Optimal feedback control from \cite{Todorov_Optimal_2002}}
+\label{subsec:todorov-2002}
+This model, very well specified in the supplementary information of \cite{Todorov_Optimal_2002}, is an optimal control model, where some cost function is established and solved alongside the dynamical system.
+
+The model is based on what they call the ``minimal intervention principle'', where trajectories are only corrected when they interfere with the goals.
+This is fancy talk for the idea of using cost functions that are based entirely on the final goal (and subgoals, if any), as opposed to calculating optimal trajectories and trying to stick to them.
+The cost function also must be written in a way that establishes what is relevant and what isn't.
+For example, they talk about a cost function of the form $x_1 + x_2 - x*$, which does not specify what the end positions of $x_1$ and $x_2$ have to be, but only that their sum must be close to $x*$.
+
+Once the cost function is created, the system of differential equations (see below), cost function and observation feedback functions are all solved by casting the problem in terms of a Kalman filter.
+
+It seems like the model is ready to be applied as-is, with modifications only at the level of the dynamical system and the cost function.
+
+From \cite{Wolpert_Principles_2011}:
 \begin{enumerate}
-\item Check the following models: Nelson 83, Todorov and Jordan 02, Hamilton and Wolpert 02, van Beers at al 02, 04, Enderle and Wolfe 87, Happee 92
+\item ``The minimum intervention principle has now
+been demonstrated in a number of tasks including the seemingly simple task of
+generating a target force with the tip of the index finger[35]''
+\item ``Recently, it has been pro- posed that such impedance control can be
+brought within an optimal control framework by formulating impedance control as
+the optimal response in the face of uncertainty about the dynamics of the body
+and environment[39].''
 \end{enumerate}
 
-I should read Knill and Pouget 2004 for Bayesian calculations in neurons.
+\subsection{Adaptive dynamics from \cite{Shadmehr_Adaptive_1994}}
+\label{subsec:shadmehr-1994}
+The gist of this model is as follows: first, determine an optimal trajectory
+$q^*(t)$, then set up the controller of each joint such that $q^*(t)$ is an
+attractor (for $0 \le t \le T$) of the system and then just let it rip. The
+optimal trajectory in this paper is obtained using the minimum jerk model
+\citep{Hogan_organizing_1984}.
+
+Because of the use of an optimal trajectory that the controlled system is trying
+to stick to, this model differs from that presented by
+\cite{Todorov_Optimal_2002} (see \ref{subsec:todorov-2002}). In addition to this
+essential difference, this model also does not have the possibility of feedback
+signals (visual or proprioceptive) and, due to its complexity, is probably less
+viable as a generative model for inversion, though that might be a problem for
+future studies and not this paper.
+
+\subsection{Object avoidance from \cite{Hamilton_Controlling_2002}}
+In this paper they're touting the awesomeness of the TOPS model
+\citep{Harris_Signaldependent_1998} and applying it to a reaching problem with
+an obstacle in between.
+
+The implementation of this model is way too handcrafted. The gist of it is as
+follows: They put an obstacle between the starting and ending positions and
+assume that the optimal trajectory is a polynomial of fifth order (maybe even
+just $y = Ax^5$, I'm not sure). To optimize its parameters they created a cost
+function with two parts: the first part is the predicted covariance of the end
+point and the second part is a term penalizing ``visiting'' the obstacle. They
+basically discard any trajectory that gets close to the obstacle.
+
+While it works, I find it too ad hoc to be useful.
+
+\subsection{2-timescale motor learning from \cite{Lee_Dual_2009}}
+\label{subsec:lee-dual}
+This is a motor learning model, i.e. a model for how we can learn to perform
+movements in an environment that is different from the default. For example,
+under strong winds. It's an extension of the model presented by
+\cite{Smith_Interacting_2006} to the problem of ``dual adaptation'', i.e. the
+adaptation to different contexts and their force fields, and switching between
+them.
+
+While they present many models for dual adaptation, through model comparison (of
+the qualitative type) they arrive at the conclusion that the only good one is
+the parallel 1-fast N-slow model, which I describe here.
+
+Consider an abstract motor system (for example, an arm) that has learned to
+perform its movements in a baseline environment. Let $y(t) \in \mathbb{R}$ be
+the outcome for some motor signal given at time $t$ by the controller in the
+baseline environment. Now, assume that in a new environment, with some force
+field that is unknown to the controller, the controller issues the command that
+under the baseline environment would lead to $y(t)$, but, due to the force
+field, it leads to the outcome $f(t)$. Now, define a prediction error between
+the expected outcome and the observed outcome as $e(t) = f(t) - y(t)$. The model
+from \cite{Lee_Dual_2009} postulates two adaptation systems, one fast and one
+slow, which learn from this prediction error.
+
+Both slow and fast systems adapt with a similar rule:
+\[
+x_{f/s}(t + 1) = A_{f/s}x_{f/s}(t) + B_{f/s}e(n)c(n)^T
+\]
+where $x$ is a ``learning process'', $f/s$ refers to fast/slow (one equation
+each), and $A$ and $B$ are learning rates. $c(n)$ is the environment variable
+that informs the controller which environment is currently active.
+$x_s,c, B \in \mathbb{R}^M$ to account for multiple environments. In the case
+where the learning of one environment does not interfere with the learning of
+others (as in \cite{Lee_Dual_2009}), $c = (0, ..., 1, 0, ..., 0)$, where the 1
+is in the position of the current environment. Note that the fast process is
+single-dimensional in this model (see \citep{Lee_Dual_2009} for the reasoning).
+
+The controller is then given by:
+\[
+y(t) = x_f(t) + x_s(t)c(t)^T
+\]
+
+\subsection{Proposal: Bayesian motor learner}
+This model is a Bayesian alternative to the N-state learners from
+\cite{Lee_Dual_2009}. In this model, there is no need for explicit different
+timescales and multiple states; instead, the model relies on Bayesian learning
+of the best control, combined with inference over the current context. This
+model will, hopefully, reproduce all the experimental findings from the
+literature on motor learning (e.g. saving, interference).
+
+The model consists of three parts: (1) context inference, (2) decision making,
+(3) updating of the decision-making rules.
+
+Before describing these parts, I present some nomenclature:
+\begin{itemize}
+\item $t$: Time at the moment an observation is made, before inference is done.
+\item $\theta_t$: Observation. Note that this is assumed to be a noisy
+observation of the generalized state (see below) given by $\theta_t = h(z_t)$.
+\item $s_t$: Hidden state, e.g. the position of the hand. It is assumed not to
+depend on context. E.g. the physical position of the hand.
+\item $a_t$: Action taken after observing $\theta_t$ and inferring the context.
+\item $c_t$: Context inferred after observing $\theta_t$.
+\item $\omega_t$: Contextual cue.
+\item $z_t$: Generalized state, $z_t = \{s_t, c_t, \omega_t\}$. Observed noisily
+as $\theta_t$.
+\end{itemize}
+
+Context inference can work through context-specific (possibly unreliable) cues,
+or through the dynamics themselves, using prediction error as the force behind
+inference. Let us assume that both parts are present in the observation
+$\theta_t$. In this case, the inferred generalized state is given by
+$q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t |
+z_t)p(z_t|\theta_{t-1}, a_{t-1})$.
+
+Having inferred the generalized state (including context), a decision is made
+sampling from $p(a_t | z_t, \beta)$, which is a distribution that is assumed to
+be parametrized by $\beta$. These parameters can be updated via
+$q(\beta | z_t, \theta_{t+1}, a_t) \propto p(\theta_{t+1} | \beta, z_t,
+a_t)p(\beta)$
+
+\subsubsection{Expected results}
+I predict that this model will display all the experimental findings that other
+models do, without invoking a possibly-infinite hierarchy of time scales, each
+with their own learning rate. Here I briefly go through the experimental
+findings and why I expect the model to reproduce them.
+
+Savings. This refers to the effect that, while learning a new context, if
+learning is interrupted by going back to a baseline context (e.g. no force
+field) and the new context is later retaken, the new dynamics do not need to be
+learned from scratch. Instead, learning continues where it was left off, with
+the possibility of some performance having been lost (i.e. learning having been
+forgotten). The Bayesian lerner displays this behavior simply by virtue of the
+learning of contex-specific action rules (learned dynamics), which are ``stored
+away'' once the context is no longer the new one (i.e. it has returned to
+baseline).
+
+Retrograde interference. This refers to the learning of context A interfering
+with the learning of context B. This occurs, for example, if context A has a
+force field that pushes right, while context B has a force field that pushes
+left. In this case, learning B is made slower by having learned Abefore,
+compared to learning B without having encountered A. This effect will be
+observed in the behavior of the model if context inference needs to accumulate
+before the agent is sure that a context switch has occured. In the meantime,
+context A's learned dynamics will interfere with context B's because actions are
+taken with a rule that depends on context (via $z_t$). Related to this, the
+model makes three predictions: (1) Contextual cues (that can be trusted) will
+reduce interference. (2) The amount of learning on contexts A and B will affect
+intereference such that highly trained contexts will interfere more
+strongly. (3) The duration of interference will be reduced in the case of
+highly-learned, opposing contexts, as the prediction error of the outcomes of
+actions will quickly drive contexts to switch to the current one.
+
+Rapid de-adaptation. This is the phenomenon in which forgetting an adaptation is
+faster than learning it. This would happen simply through context-based
+learning, where getting out of a context means that adaptations are no longer
+used (and eventually are forgotten). Note: I need to read more on this
+phenomenon, as I might be misinterpreting what it means.
+
+Bicycle effect. We never forget how to ride a bicycle. Existing models would
+need to incorporate an almost-infinite hierarchy of time scales to account for
+this, but my Bayesian learner naturally accounts for it by making use of
+probability distributions: the longer you practice something, the more
+entrenched the adaptation becomes by virtue of the posterior probability
+becoming very peaky.
 
@@
 
-\begin{chapter}{Physiology and neurons}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{chapter}{Behavioral findings, Physiology and neurons}
 A few notes from \cite{Wolpert_Are_1995}:
@@
 
-Notes from \cite{Harris_Signaldependent_1998}:
+\section{Notes from \cite{Harris_Signaldependent_1998}}
 \begin{enumerate}
@@
 
-Notes from \cite{Bays_Computational_2007}:
+\section{Notes from \cite{Bays_Computational_2007}}
 \begin{enumerate}
@@
 
-Notes from \cite{Shadmehr_Adaptive_1994}:
+\section{Notes from \cite{Shadmehr_Adaptive_1994}}
 \begin{enumerate}
@@
 
+\section{Notes from \cite{deC.Hamilton_scaling_2004}}
+\begin{enumerate}
+\item ``Muscle activity in specific tasks shows stereotypy whether subjects generate force with the fingertips (Valero-Cuevas et al.
+1998), the wrist (Hoffman and Strick 1999), the neck (Vasavada et al.
+2002) or the arm (van Zuylen et al.
+1988; Flanders and Soechting 1990; Buchanan et al.
+1993; van Bolhuis and Gielen 1997).''
+\item ``Primates also show repeatable patterns of muscle activation when grasping the same object repeatedly, and different activation patterns for different objects (Brochier et al.
+2001).
+This suggests that stereotypy is a general characteristic of movement and is not unique to humans.''
+\item There is a paragraph that discusses different cost functions.
+For example, one is based on the total torque, another on fatigue and endurance, and other mathematical alchemy.
+Lots of references there too.
+
+\end{enumerate}
+
+\section{Notes from \cite{Kojima_Memory_2004}}
+\begin{enumerate}
+\item ``Previous studies have shown that adaptation is relatively specific to the size and direction of target eccentricity (Deubel et al., 1987; Semmlow et al., 1989; Frens and van Opstal, 1994; Straube et al., 1997; Noto et al., 1999; Watanabe et al., 2000).''
+\item ``It has been demonstrated that adaptation progresses more slowly when the subject is exposed to adapting target steps of a variety of sizes and directions in humans (Miller et al., 1981) and monkeys (Scudder et al., 1998).''
+\end{enumerate}
+
+\section{Notes from \cite{Lee_Dual_2009}}
+``When given contextual cues and sufficient trials, humans can simultaneously adapt to two opposite force fields (Osu et al., 2004; Nozaki et al., 2006; Howard et al., 2008), two saccadic gains (Shelhamer et al., 2005), or several visuomotor rotations (Imamizu et al., 2007; Choi et al., 2008).''
+
+\section{Motor adaptation}
+Motor adaptation to modified environments (e.g. force fields) has been shown to display the following characteristics:
+\begin{enumerate}
+\item Savings: Learning a previously-learned environment (after some extinction phase) is faster than the original learning.
+This means that the previously learned model is saved for a while and its learning can be resumed after extinction.
+However, \cite{Kojima_Memory_2004} showed that savings can be ``washed out'' if baseline trials (no force field) are inserted between extinction and the second learning phase.
+\item Anterograde interference: Learning an adaptation (e.g. force field pointing left) slows down learning an opposing adaptation (e.g. force field pointing right).
+\item Rapid de-adaptation: Unlearning an adaptation (even after fully learning it) is faster than tlearning it \citep{Davidson_Scaling_2004}
+\end{enumerate}
+
 \end{chapter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+A description of their model can be seen in \ref{subsec:shadmehr-1994} above.
+
+\section{\cite{Todorov_Optimal_2002}}
+See \ref{subsec:todorov-2002}
+
+\section{\cite{deC.Hamilton_scaling_2004}}
+The authors make use of the TOPS strategy \citep{Harris_Signaldependent_1998} and an experiment where they measure torque at different joints to show that bigger muscles produce less variability than smaller muscles for the same level of voluntary torque (I wonder what the opposite of voluntary is).
+
+It's an experimental paper with some muscle simulations thrown in for good measure.
+Contains a lot of references that might be useful later on the nature and characteristics of human and primate movements, as well as cost functions used in the past for modeling them.
+
+\section{\cite{Beers_When_2002}}
+This paper speaks of a model of optimal adaptation and weighting of visual and proprioceptive cues. Nothing of interest for us as of now.
+
+\section{\cite{Scheidt_Learning_2001}}
+Humans can adapt their movements to stochasticly varying changes in the dynamics of the environment.
+Moreover, they showed that their adaptation is most complatible with the idea that participants adapt to the mean of the dynamics (a one-directional force of changing magnitude) of the environment instead of doing trial-to-trial adaptations.
+
+These views, however, are challenged by the two-time-scales models of learning, like \cite{Smith_Interacting_2006}.
+
+\section{\cite{DeKleijn_Everyday_2014}}
+They cite \cite{Henry_Increased_1960} saying that participants have longer reaction times when preparing complex movements than when preparing simple ones. May be worth checking out.
+
+\section{\cite{Haar_Motor_2020}}
+This is the only paper I know where they do motor learning in a very natural task: playing billiards. It's great and I wish I could do this research.
+
+They tracked the joints of participants who were doing the same pool shot over and over (300 times).
+I do not think there was any variation in initial positions.
+They track performance and variability in joint movement.
+
+\section{\cite{Wolpert_Principles_2011}}
+Good review on motor learning. many relevant citations here.
+
+\section{\cite{Todorov_Optimality_2004}}
+Review of optimal control. Citation goldmine! Lots of papers on optimal control in biological systems are cited here.
+
+In particular, [43] and [44] might be similar to the idea of inferring goals and predicting future movements.
+
+``However, feedback control[25] explains an important additional observation: the increased duration of more accurate movements is due to a prolonged deceleration phase, making the speed profiles significantly skewed[88].''
+
 \end{chapter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-02-17 notes/notes.tex
Improved the model

@@
 $\theta_t$. In this case, the inferred generalized state is given by
-$q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t |
-z_t)p(z_t|\theta_{t-1}, a_{t-1})$.
+\[
+  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
+\]
+
+Additionally, after having inferred the current state, motor adaptation is given by
+\[
+q(\gamma | z_t, z_{t-1}, a_{t-1}) \propto p(z_t | z_{t-1}, a_{t-1}, \gamma)p(z_{t-1})p(\gamma)
+\]
+
+where $\gamma$ are the parameters of the internal model for the dynamics,
+i.e. the function $f: (z_t, a_t) \rightarrow z_{t+1}$, which determines the beliefs of
+the agent as to how the system evolves after having taken action $a_t$.
 
 Having inferred the generalized state (including context), a decision is made
-sampling from $p(a_t | z_t, \beta)$, which is a distribution that is assumed to
-be parametrized by $\beta$. These parameters can be updated via
-$q(\beta | z_t, \theta_{t+1}, a_t) \propto p(\theta_{t+1} | \beta, z_t,
-a_t)p(\beta)$
+sampling from $p(a_t | z_t, \gamma)$, which is the mapping between state and
+action, given the updated value for $\gamma$. Because the model is meant to be
+used to follow a given trajectory (which was created by another system, not
+modeled here), the desired $z_{t+1}$ is known at time $t$. Therefore, having
+$p(z_{t+1} | z_t, a_t, \gamma)$ is enough to create $p(a_t | z_t, \gamma)$.
 

### 2021-02-18 notes/notes.tex
Redefined error clamp

@@
 \begin{enumerate}
-\item Error-clamp: Refers to the phase of a simulation in the models of motor adaptation where the error that drives learning is set to zero. This emulates a no-feedback experimental condition.
+\item Error-clamp: Refers to the phase of a simulation in the models of motor
+adaptation where the error that drives learning is forced to zero. For example,
+when performing a task in which a robotic attachment creates forces perpendicular
+to the direction of movement, an error-clamp phase is one in which the attachment
+pushes the arm constantly towards the desired trajectory, thus making error zero.
 \end{enumerate}
@@
 \section{Straight movements as a sign of learning}
-According to Gurevich\_93 (see \cite{Wolpert_Are_1995}) and \cite{Shadmehr_Adaptive_1994}, the more a person practices a movement, the straighter this movement becomes.
-Could we use the curvature of a movement as a measure of how much they've learned this movement?
+According to Gurevich\_93 (see \cite{Wolpert_Are_1995}) and
+\cite{Shadmehr_Adaptive_1994}, the more a person practices a movement, the
+straighter this movement becomes. Could we use the curvature of a movement as a
+measure of how much they've learned this movement?
 
-This could be useful for the context-dependent learning paradigm.
-The question to answer is: if a participant learns context A fully (and their movements become more straight), when switching back to A, will the first movement be straight or curved?
-If it's straight, then we can link this to the idea of context-based learning.
+This could be useful for the context-dependent learning paradigm. The question
+to answer is: if a participant learns context A fully (and their movements
+become more straight), when switching back to A, will the first movement be
+straight or curved? If it's straight, then we can link this to the idea of
+context-based learning.
 
@@
 \section{Quick adaptation to perturbations in the irrelevant manifold.}
-Models for motor control can be divided into two categories: (1) optimal trajectory and (2) end point.
-Optimal trajectory are those in which the controller tries to stick to a previously-calculated optimal trajectory and any perturbation (internal, external) will be dealt with by trying to go back to the optimal trajectory.
-Endpoint models, on the other hand, do not care that much about the trajectory itself and, with models that follow the ``minimal intervention'' principle, will only correct for perturbations if they mess with the endpoint.
+Models for motor control can be divided into two categories: (1) optimal
+trajectory and (2) end point. Optimal trajectory are those in which the
+controller tries to stick to a previously-calculated optimal trajectory and any
+perturbation (internal, external) will be dealt with by trying to go back to the
+optimal trajectory. Endpoint models, on the other hand, do not care that much
+about the trajectory itself and, with models that follow the ``minimal
+intervention'' principle, will only correct for perturbations if they mess with
+the endpoint.
 
-Endpoint models are more recent and arguably better than optimal trajectory ones.
-However, there is one thorn on their side: if during a movement an arm is perturbed, it will immediately try to return to the optimal trajectory \citep{Bizzi_Posture_1984}, even if the perturbation does not detract from the endpoint (this last part is my speculation).
-\cite{Todorov_Optimality_2004} argues that this is due to impedance, as a sort of immediate dumb fighback.
+Endpoint models are more recent and arguably better than optimal trajectory
+ones. However, there is one thorn on their side: if during a movement an arm is
+perturbed, it will immediately try to return to the optimal trajectory
+\citep{Bizzi_Posture_1984}, even if the perturbation does not detract from the
+endpoint (this last part is my speculation). \cite{Todorov_Optimality_2004}
+argues that this is due to impedance, as a sort of immediate dumb fighback.
 
-A nice low-hanging fruit project would be to test whether \cite{Todorov_Optimality_2004} is right by having participants perform a reaching task in real life, where a motor creates random perturbations (in the irrelevant manifold, for good measure) and compare their reactions to the same task in virtual reality where the perturbation happens in the VR environment, but not in real time.
-If participants adapt in the same way in both environments, then \cite{Todorov_Optimality_2004} is wrong.
+A nice low-hanging fruit project would be to test whether
+\cite{Todorov_Optimality_2004} is right by having participants perform a
+reaching task in real life, where a motor creates random perturbations (in the
+irrelevant manifold, for good measure) and compare their reactions to the same
+task in virtual reality where the perturbation happens in the VR environment,
+but not in real time. If participants adapt in the same way in both
+environments, then \cite{Todorov_Optimality_2004} is wrong.
 
@@
 \begin{chapter}{Paper}
-In this chapter, I describe the everything related to the paper that should come out of this project, including modeling and experimental findings, as well as the direction of the paper and things needed to get there.
+In this chapter, I describe the everything related to the paper that should come
+out of this project, including modeling and experimental findings, as well as
+the direction of the paper and things needed to get there.
 
 \section{Project description}
-The idea of the project is the following: We will create an agent that learns to control an arm in environment A (say, the real world carrying nothing), and learns how to control the arm in environment B (say, with high winds).
-The agent then learns to recognize environments A and B and to deploy the right controller for each environment depending on this recognition.
+The idea of the project is the following: We will create an agent that learns to
+control an arm in environment A (say, the real world carrying nothing), and
+learns how to control the arm in environment B (say, with high winds). The agent
+then learns to recognize environments A and B and to deploy the right controller
+for each environment depending on this recognition.
 
-We will argue that this is how humans do it and present evidence (based on literature research) for the predictions this model makes, as well as try to make new predictions.
+We will argue that this is how humans do it and present evidence (based on
+literature research) for the predictions this model makes, as well as try to
+make new predictions.
 
-After this methods paper is done, model validation can be carried out in future studies.
+After this methods paper is done, model validation can be carried out in future
+studies.
 
 \section{Planning}
-Here I enumerate the things I need to learn and do in order to get this project moving and to finish it.
+Here I enumerate the things I need to learn and do in order to get this project
+moving and to finish it.
 
@@
 \label{subsec:todorov-2002}
-This model, very well specified in the supplementary information of \cite{Todorov_Optimal_2002}, is an optimal control model, where some cost function is established and solved alongside the dynamical system.
+This model, very well specified in the supplementary information of
+\cite{Todorov_Optimal_2002}, is an optimal control model, where some cost
+function is established and solved alongside the dynamical system.
 
-The model is based on what they call the ``minimal intervention principle'', where trajectories are only corrected when they interfere with the goals.
-This is fancy talk for the idea of using cost functions that are based entirely on the final goal (and subgoals, if any), as opposed to calculating optimal trajectories and trying to stick to them.
-The cost function also must be written in a way that establishes what is relevant and what isn't.
-For example, they talk about a cost function of the form $x_1 + x_2 - x*$, which does not specify what the end positions of $x_1$ and $x_2$ have to be, but only that their sum must be close to $x*$.
+The model is based on what they call the ``minimal intervention principle'',
+where trajectories are only corrected when they interfere with the goals. This
+is fancy talk for the idea of using cost functions that are based entirely on
+the final goal (and subgoals, if any), as opposed to calculating optimal
+trajectories and trying to stick to them. The cost function also must be written
+in a way that establishes what is relevant and what isn't. For example, they
+talk about a cost function of the form $x_1 + x_2 - x*$, which does not specify
+what the end positions of $x_1$ and $x_2$ have to be, but only that their sum
+must be close to $x*$.
 
-Once the cost function is created, the system of differential equations (see below), cost function and observation feedback functions are all solved by casting the problem in terms of a Kalman filter.
+Once the cost function is created, the system of differential equations (see
+below), cost function and observation feedback functions are all solved by
+casting the problem in terms of a Kalman filter.
 
-It seems like the model is ready to be applied as-is, with modifications only at the level of the dynamical system and the cost function.
+It seems like the model is ready to be applied as-is, with modifications only at
+the level of the dynamical system and the cost function.
 
@@
 
+Spontaneous recovery. In an A-B-error_clamp paradigm, in which during B the
+opposite adaptation to A is learned, at the beginning of the error-clamp phase
+(after B) participants seem to quickly gravitate towards A (thinking that A is
+now active?). This would be displayed by the model via the state inference, as
+now the visual cue and the observed dynamics point to different contexts. In
+time, updating $p(context | visual cue)$ will lead to spontaneous recovery.
+
 Bicycle effect. We never forget how to ride a bicycle. Existing models would
@@
 \begin{enumerate}
-\item Hand paths become straight with practice (Gurevich 93, Shadmehr and Musa-Ivaldi 94)
+\item Hand paths become straight with practice (Gurevich 93, Shadmehr and
+Musa-Ivaldi 94)
 \item Deafferented patients move differently (Ghez 90)
-\item Bell-shaped speed emerges from dynamics of control, not from planning (Jordan 94)
-\item Curvature matches the participant's messed up perception of straight lines (wolpert 94)
+\item Bell-shaped speed emerges from dynamics of control, not from planning
+(Jordan 94)
+\item Curvature matches the participant's messed up perception of straight lines
+(wolpert 94)
 
@@
 \begin{enumerate}
-\item We have the ability to recall previously learned dynamics even after months! (Bashers-Krug et al 96, Gandolfo et al 96).
+\item We have the ability to recall previously learned dynamics even after
+months! (Bashers-Krug et al 96, Gandolfo et al 96).
 \end{enumerate}
@@
 \begin{enumerate}
-\item When planning movement, the target is transformed from a retinocentric vector to a head-centered and finally a shoulder-centered one. (Andersen\_85, Soechting and Flanders\_91).
-\item Gordon\_93 says that the target is finally represented as a vector from the current hand positoin (or whatever the effector is) to the goal position.
-\item From Lackner\_and\_Dizio\_92, aftereffects exist when introduced Coriolis forces are withdrawn. They also exist when perception is mucked with with prism glasses (Held\_59, 62, 63).
+\item When planning movement, the target is transformed from a retinocentric
+vector to a head-centered and finally a shoulder-centered one. (Andersen\_85,
+Soechting and Flanders\_91).
+\item Gordon\_93 says that the target is finally represented as a vector from
+the current hand positoin (or whatever the effector is) to the goal position.
+\item From Lackner\_and\_Dizio\_92, aftereffects exist when introduced Coriolis
+forces are withdrawn. They also exist when perception is mucked with with prism
+glasses (Held\_59, 62, 63).
 \end{enumerate}
@@
 \begin{enumerate}
-\item ``Muscle activity in specific tasks shows stereotypy whether subjects generate force with the fingertips (Valero-Cuevas et al.
-1998), the wrist (Hoffman and Strick 1999), the neck (Vasavada et al.
-2002) or the arm (van Zuylen et al.
-1988; Flanders and Soechting 1990; Buchanan et al.
-1993; van Bolhuis and Gielen 1997).''
-\item ``Primates also show repeatable patterns of muscle activation when grasping the same object repeatedly, and different activation patterns for different objects (Brochier et al.
-2001).
-This suggests that stereotypy is a general characteristic of movement and is not unique to humans.''
-\item There is a paragraph that discusses different cost functions.
-For example, one is based on the total torque, another on fatigue and endurance, and other mathematical alchemy.
-Lots of references there too.
+\item ``Muscle activity in specific tasks shows stereotypy whether subjects
+generate force with the fingertips (Valero-Cuevas et al. 1998), the wrist
+(Hoffman and Strick 1999), the neck (Vasavada et al. 2002) or the arm (van
+Zuylen et al. 1988; Flanders and Soechting 1990; Buchanan et al. 1993; van
+Bolhuis and Gielen 1997).''
+\item ``Primates also show repeatable patterns of muscle activation when
+grasping the same object repeatedly, and different activation patterns for
+different objects (Brochier et al. 2001). This suggests that stereotypy is a
+general characteristic of movement and is not unique to humans.''
+\item There is a paragraph that discusses different cost functions. For example,
+one is based on the total torque, another on fatigue and endurance, and other
+mathematical alchemy. Lots of references there too.
 
@@
 \begin{enumerate}
-\item ``Previous studies have shown that adaptation is relatively specific to the size and direction of target eccentricity (Deubel et al., 1987; Semmlow et al., 1989; Frens and van Opstal, 1994; Straube et al., 1997; Noto et al., 1999; Watanabe et al., 2000).''
-\item ``It has been demonstrated that adaptation progresses more slowly when the subject is exposed to adapting target steps of a variety of sizes and directions in humans (Miller et al., 1981) and monkeys (Scudder et al., 1998).''
+\item ``Previous studies have shown that adaptation is relatively specific to
+the size and direction of target eccentricity (Deubel et al., 1987; Semmlow et
+al., 1989; Frens and van Opstal, 1994; Straube et al., 1997; Noto et al., 1999;
+Watanabe et al., 2000).''
+\item ``It has been demonstrated that adaptation progresses more slowly when the
+subject is exposed to adapting target steps of a variety of sizes and directions
+in humans (Miller et al., 1981) and monkeys (Scudder et al., 1998).''
 \end{enumerate}
@@
 \section{Notes from \cite{Lee_Dual_2009}}
-``When given contextual cues and sufficient trials, humans can simultaneously adapt to two opposite force fields (Osu et al., 2004; Nozaki et al., 2006; Howard et al., 2008), two saccadic gains (Shelhamer et al., 2005), or several visuomotor rotations (Imamizu et al., 2007; Choi et al., 2008).''
+``When given contextual cues and sufficient trials, humans can simultaneously
+adapt to two opposite force fields (Osu et al., 2004; Nozaki et al., 2006;
+Howard et al., 2008), two saccadic gains (Shelhamer et al., 2005), or several
+visuomotor rotations (Imamizu et al., 2007; Choi et al., 2008).''
 
 \section{Motor adaptation}
-Motor adaptation to modified environments (e.g. force fields) has been shown to display the following characteristics:
+Motor adaptation to modified environments (e.g. force fields) has been shown to
+display the following characteristics:
 \begin{enumerate}
-\item Savings: Learning a previously-learned environment (after some extinction phase) is faster than the original learning.
-This means that the previously learned model is saved for a while and its learning can be resumed after extinction.
-However, \cite{Kojima_Memory_2004} showed that savings can be ``washed out'' if baseline trials (no force field) are inserted between extinction and the second learning phase.
-\item Anterograde interference: Learning an adaptation (e.g. force field pointing left) slows down learning an opposing adaptation (e.g. force field pointing right).
-\item Rapid de-adaptation: Unlearning an adaptation (even after fully learning it) is faster than tlearning it \citep{Davidson_Scaling_2004}
+\item Savings: Learning a previously-learned environment (after some extinction
+phase) is faster than the original learning. This means that the previously
+learned model is saved for a while and its learning can be resumed after
+extinction. However, \cite{Kojima_Memory_2004} showed that savings can be
+``washed out'' if baseline trials (no force field) are inserted between
+extinction and the second learning phase.
+\item Anterograde interference: Learning an adaptation (e.g. force field
+pointing left) slows down learning an opposing adaptation (e.g. force field
+pointing right).
+\item Rapid de-adaptation: Unlearning an adaptation (even after fully learning
+it) is faster than tlearning it \citep{Davidson_Scaling_2004}
 \end{enumerate}
@@
 \begin{chapter}{Notes on papers}
-Conclusions and useful things about different papers I read. This differs from the comments above in that the notes here are holistic.
+Conclusions and useful things about different papers I read. This differs from
+the comments above in that the notes here are holistic.
 
 \section{\cite{Shadmehr_Adaptive_1994}}
-In this paper they show a number of things.
-First, the show that people can adapt the dynamics of their control states to match changing dynamics of the observers.
-They also show that this adaptation takes the form of a straightening of trajectories, which are very twisted when they are first introduced to new system dynamics (viscosity of the environment); participants learn to do straight lines in the new environment after some trials, and the learing is monotonically increasing.
-They show a model of the dynamics of the arm, the control states and the adaptation that happens when exposed to the new dynamics.
-Finally, they show, both experimentally and in their model, the aftereffects of the new dynamics, whereby they display bad behavior when they go back to the natural environment, but eventually adapt to it as well.
+In this paper they show a number of things. First, the show that people can
+adapt the dynamics of their control states to match changing dynamics of the
+observers. They also show that this adaptation takes the form of a straightening
+of trajectories, which are very twisted when they are first introduced to new
+system dynamics (viscosity of the environment); participants learn to do
+straight lines in the new environment after some trials, and the learing is
+monotonically increasing. They show a model of the dynamics of the arm, the
+control states and the adaptation that happens when exposed to the new
+dynamics. Finally, they show, both experimentally and in their model, the
+aftereffects of the new dynamics, whereby they display bad behavior when they go
+back to the natural environment, but eventually adapt to it as well.
 
-Interesting paper that cound be the foundation of our models, assuming that it hasn't been improved upon and maybe entirely replaced by newer stuff.
+Interesting paper that cound be the foundation of our models, assuming that it
+hasn't been improved upon and maybe entirely replaced by newer stuff.
 
@@
 \section{\cite{deC.Hamilton_scaling_2004}}
-The authors make use of the TOPS strategy \citep{Harris_Signaldependent_1998} and an experiment where they measure torque at different joints to show that bigger muscles produce less variability than smaller muscles for the same level of voluntary torque (I wonder what the opposite of voluntary is).
+The authors make use of the TOPS strategy \citep{Harris_Signaldependent_1998}
+and an experiment where they measure torque at different joints to show that
+bigger muscles produce less variability than smaller muscles for the same level
+of voluntary torque (I wonder what the opposite of voluntary is).
 
-It's an experimental paper with some muscle simulations thrown in for good measure.
-Contains a lot of references that might be useful later on the nature and characteristics of human and primate movements, as well as cost functions used in the past for modeling them.
+It's an experimental paper with some muscle simulations thrown in for good
+measure. Contains a lot of references that might be useful later on the nature
+and characteristics of human and primate movements, as well as cost functions
+used in the past for modeling them.
 
 \section{\cite{Beers_When_2002}}
-This paper speaks of a model of optimal adaptation and weighting of visual and proprioceptive cues. Nothing of interest for us as of now.
+This paper speaks of a model of optimal adaptation and weighting of visual and
+proprioceptive cues. Nothing of interest for us as of now.
 
 \section{\cite{Scheidt_Learning_2001}}
-Humans can adapt their movements to stochasticly varying changes in the dynamics of the environment.
-Moreover, they showed that their adaptation is most complatible with the idea that participants adapt to the mean of the dynamics (a one-directional force of changing magnitude) of the environment instead of doing trial-to-trial adaptations.
+Humans can adapt their movements to stochasticly varying changes in the dynamics
+of the environment. Moreover, they showed that their adaptation is most
+complatible with the idea that participants adapt to the mean of the dynamics (a
+one-directional force of changing magnitude) of the environment instead of doing
+trial-to-trial adaptations.
 
-These views, however, are challenged by the two-time-scales models of learning, like \cite{Smith_Interacting_2006}.
+These views, however, are challenged by the two-time-scales models of learning,
+like \cite{Smith_Interacting_2006}.
 
 \section{\cite{DeKleijn_Everyday_2014}}
-They cite \cite{Henry_Increased_1960} saying that participants have longer reaction times when preparing complex movements than when preparing simple ones. May be worth checking out.
+They cite \cite{Henry_Increased_1960} saying that participants have longer
+reaction times when preparing complex movements than when preparing simple
+ones. May be worth checking out.
 
 \section{\cite{Haar_Motor_2020}}
-This is the only paper I know where they do motor learning in a very natural task: playing billiards. It's great and I wish I could do this research.
+This is the only paper I know where they do motor learning in a very natural
+task: playing billiards. It's great and I wish I could do this research.
 
-They tracked the joints of participants who were doing the same pool shot over and over (300 times).
-I do not think there was any variation in initial positions.
-They track performance and variability in joint movement.
+They tracked the joints of participants who were doing the same pool shot over
+and over (300 times). I do not think there was any variation in initial
+positions. They track performance and variability in joint movement.
 
@@
 \section{\cite{Todorov_Optimality_2004}}
-Review of optimal control. Citation goldmine! Lots of papers on optimal control in biological systems are cited here.
+Review of optimal control. Citation goldmine! Lots of papers on optimal control
+in biological systems are cited here.
 
-In particular, [43] and [44] might be similar to the idea of inferring goals and predicting future movements.
+In particular, [43] and [44] might be similar to the idea of inferring goals and
+predicting future movements.
 
-``However, feedback control[25] explains an important additional observation: the increased duration of more accurate movements is due to a prolonged deceleration phase, making the speed profiles significantly skewed[88].''
+``However, feedback control[25] explains an important additional observation:
+the increased duration of more accurate movements is due to a prolonged
+deceleration phase, making the speed profiles significantly skewed[88].''
 

### 2021-02-24 notes/notes.tex
Description of implementation of adaptation model ready. Code started.

--- 2021-02-17 notes/notes.tex
@@
 \bibliographystyle{apalike}
+
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
+
+% Defining parameters
+\def \parts {60 }                  % Number of participants
+\def \fref #1{Figure \ref{#1}}     % Reference figures
+\def \tref #1{Table \ref{#1}}      % Reference tables
+\def \cents {20 }                  % Cents earned per mini-block
+\def \eref #1{Equation \ref{#1}}   % Reference equations
+\def \sref #1{Section '\nameref{#1}'}    % Reference sections
+\def \supmat {the Supplementary Materials}
+\def \excper {5 }     % Participants excluded because they cannot sort
 
@@
 End Date & Milestone & A priori comments & A posteriori comments \\ \hline \hline
-12.02.2021 & Learn about published models on arm-movement planning, especially those that can be used as generative models & & \\ \hline
+12.02.2021 & Learn about published models on arm-movement planning, especially those that can be used as generative models & & Discovered low-hanging fruit.
+Shifting directions a bit.
+See \ref{subsec:proposal}\\ \hline
 19.02.2021 & Learn about measures for how practiced a movement is. Reaction times? Speed profiles? Trajectories? & & \\\hline
@@
 \[
-x_{f/s}(t + 1) = A_{f/s}x_{f/s}(t) + B_{f/s}e(n)c(n)^T
+x_{f/s}(t + 1) = A_{f/s}x_{f/s}(t) + B_{f/s}e(n)c(n)^T \label{eqn:lee-learn}
 \]
@@
 \subsection{Proposal: Bayesian motor learner}
+\label{subsec:proposal}
 This model is a Bayesian alternative to the N-state learners from
@@
 sampling from $p(a_t | z_t, \beta)$, which is a distribution that is assumed to
-be parametrized by $\beta$. These parameters can be updated via
-$q(\beta | z_t, \theta_{t+1}, a_t) \propto p(\theta_{t+1} | \beta, z_t,
-a_t)p(\beta)$
+be parametrized by $\beta$. These parameters can be updated via:
+\[
+q(\beta | z_t, \theta_{t+1}, a_t) \propto p(\theta_{t+1} | \beta, z_t,
+a_t)p(\beta) \label{eqn:update-parameters}
+\]
 
@@
 
+\begin{chapter}{Testing the model in \ref{subsec:proposal}}
+In this section, I propose a simulated task in which the prowess of the model proposed in \ref{subsec:proposal} is put to the test.
+I additionally present simulation results and some writeup.
+Basically a mini-paper on this model.
+
+\section{The Task}
+The tasks used for models like this one are usually \citep[e.g.][]{Lee_Dual_2009} reaching tasks: a participant sits in front of a table with shoulders immobilized, and has to bring her hand from the starting position A to a designated position B.
+This is done while holding a robotic arm that moves with the hand.
+The robotic arm is used to create force fields in different directions (CW and CCW, typically, but it's basically left and right) for the different contexts.
+The position of the participant's arm is recorded and deviations from the straight line are considered as errors ($e(t)$ in \ref{eqn:lee-learn}).
+
+To simplify the task, I propose the following: participants need to keep their hand still in the starting position for the duration of the trial.
+The robotic arm creates forces that will drive the arm away from its current position by pushing in a constant direction with constant force.
+This experiment has the same elements as the one by \cite{Lee_Dual_2009} but eliminates the necessity of an underlying movement.
+While it works wonders as a thought experiment and helps me see if the proposed model behaves the way I want, it must not be used as a real experiment because participants could simply use impedance for each trial, regardless of context.
+
+In this experiment, every context $C_i$ is characterized by the strength and direction of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ..., N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force measured Cartesianly, and $w_i$ is the intensity of the force.
+
+As in \cite{Lee_Dual_2009}, I will simulate an A-B-clamp paradigm, which starts with a context $C_0 = \{0, 0\}$ in which no force is being applied.
+Then context $C_A$ is shown for a few trials, then context $C_B$, then an error-clamp block in which the hand is forced to stay at the initial position by the robotic arm.
+
+Additionally, a contextual cue $\omega$ is shown to the participant which signals the current context.
+The experiment will be run with different levels of reliability of this cue.
+Different sessions will be run with different cue reliabilities.
+
+An experimental session $\Xi$ is characterized by the reliability of the cue, the contexts avaiable and the order in which they are shown.
+In a given session, each trial begins with the cue and the XXX seconds during which the participant must keep the hand in the starting position.
+During the intertrial time, no forces are applied to the hand, but the hand must remain in the starting position.
+
+\section{The model}
+The model is an implementation of \ref{subsec:proposal}. The generalized state is given by:
+\[
+z_t = \{x, \theta_t, \gamma_t, \omega_t\}
+\]
+where $x$ is the current position of the hand in Cartesian coordinates, where the origin is the starting point.
+We assume that motor commands are issued every $\Delta t$ for simplicity.
+At the beginning of each time interval, the context is inferred combining the cue and the prediction error of the outcome of the previous motor command.
+We assume further than the cue and the force field are uncorrelated, which yields:
+\begin{align}
+  q(C_t) &= q(C_t | \omega_t)q(C_t | s_t, s_{t-1}, a_{t-1}) \\ \label{eqn:estimated-context}
+  q(C_t | \omega_t) &\propto p(\omega_t | C_t)p(C_t) \\
+  q(C_t | x_t, x_{t-1}, a_{t-1}) &\propto p(x_t | C_t, x_{t-1}, a_{t-1})p(x_t)p(x_{t-1})p(C_t)
+\end{align}
+The terms $p(C_t)$ refer to the prior probability of the context at the beginning of the trial: at the first trial, this refers to prior beliefs over which contexts are more common/likely.
+At each subsequent trial, it incorporates the belief that has so far been accumulated, given previous trials.
+
+Once the posterior over contexts has been calculated for the current trial, it is used to select the current context.
+A simple rule is to sample the context directly from the posterior distribution.
+However, a useful extension of this is to use a softmax function to create a distribution:
+
+\begin{equation}
+S(C^{(t)}) = \frac{e^{\kappa q\left(C^{(t)}\right)}}{\displaystyle \sum_{i=1}^{N_c}e^{\kappa q\left(C_i\right)}}
+\end{equation}
+where $q(C_i)$ is given by \eref{eqn:estimated-context}.
+The function $S(\cdot)$ is then used to sample the (inferred) current context.
+This representation has the advantage of adding a parameter with which the sampling behavior can be characterized, ranging from choosing uniformly when $\kappa = 0$ to choosing the context with the highest probability deterministically as $\kappa \rightarrow \infty$.
+
+In the case where the hand position can be determined by the participant with perfect accuracy (ignoring visual and proprioceptive error), we have that $p(x_\tau) = \delta(x_\tau, \hat x)$, where $\hat x$ is the real position of the hand, and $\delta (\cdot, \cdot)$ is the Kronecker delta function.
+We assume this to be the case during this experiment.
+
+For simplicity, we assume that participants know how many distinct contexts will be shown to them, and therefore all the $C_x$ have been created by the agent before the session begins.
+By default, all contexts assume a force with angle and magnitude equal zero, and their true values are learned during the experiment (see below).
+
+For mathematical tractability (i.e.
+to avoid circular distributions), we will limit contexts to one where the force points left (angle $\pi$) and one where the force points right (angle 0), in addition to the baseline context.
+In this scheme, participants need only learn the magnitude of the force for each context.
+We will assume that for each context, the prior distribution over the magnitude of the force is Gaussian:
+\[
+p(\gamma) = N(1, 0.1)
+\]
+We will show that the form of this prior affects learning rate, but not the final learned magnitude.
+
+After the generalized state has been inferred given the information at the beginning of the trial, the next step is to infer the parameters of the context, as in \eref{eqn:update-parameters}:
+\begin{equation}
+q(C_t | x_t, x_{t-1}, a_{t-1}) \propto p(x_t | x_{t-1}, a_{t-1}, C_t)p(C_t) \label{eqn:context-from-x}
+\end{equation}
+
+where $p(C_t)$ is either a prior belief of the likelihood of each context, or the probability inferred up to the previous trial.
+Note that, because we assumed that the position of the hand is directly observable without any noise, there are no terms for the uncertainty of $x_t$ or $x_{t-1}$.
+
+With this, \eref{eqn:context-from-x} becomes:
+\begin{equation}
+q(\gamma | x_t, x_{t-1}, a_{t-1}) \propto N(\hat x_t, \sqrt{\Delta t})p(\gamma) \label{eqn:data-dist}
+\end{equation}
+where $\hat x_t$ is the predicted value of the position $x_t$ given the previous state and action and the force field of the context; see \eref{eqn:dynamics} below.
+Because we chose the priors $p(\gamma)$ to be Gaussian, the posteriors $q(\gamma)$ after every trial will also be a Gaussian distribution with parameters:
+\begin{align}
+  \mu_\gamma^{(t)} &= \frac{\mu_\gamma^{(t-1)}\left(\sigma_\gamma^{(d)}\right)^2 + \mu_\gamma^{(d)}\left(\sigma_\gamma^{(t-1)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2} \\
+  \sigma_\gamma^{(t)} & = \sqrt{\frac{\left(\sigma_\gamma^{(t-1)}\right)^2\left(\sigma_\gamma^{(d)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2}}
+\end{align}
+where the superscripts $(t)$ and $(t-1)$ refer to the inferred values for the current and previous trials, respectively, and $(d)$ to the parameters in \eref{eqn:data-dist}.
+
+The final step is action selection. Action selection is done at the level of desired outcome (as opposed to motor commands) following the equations:
+\begin{align}
+  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta t}\epsilon\\ \label{eqn:dynamics}
+  f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
+  a_t &= \argmax_{a_t}p(x_{t+1} = (0, 0) | x_t, a_t, f_t)
+\end{align}
+where $\gamma_t$ and $\theta_t$ are the participant's estimate of the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian motor error component.
+
+\section{Simulation results}
+
+\section{Conclusions}
+
+\end{chapter}
+
 \bibliography{../MyLibrary}

### 2021-02-24 code/adaptation/model.py

@@ -1,0 +1,77 @@
+# -*- coding: utf-8 -*-
+# ./adaptation/model.py
+
+"""Bayesian motor adaptation model."""
+
+
+class Context(Object):
+    """Stores the force and direction of a context. If called, returns the force in
+    2D Cartesian coordinates.
+
+    """
+    def __init__(self, magnitude=1, angle=0, baseline=False):
+        """Well..."""
+        self.magnitude = magnitude
+        self.angle = angle
+
+    def __call__(self, x=None):
+        """Returns the force."""
+        if baseline:
+            return 0
+        return self.magnitude * np.array([np.cos(self.angle),
+                                          np.sin(self.angle)])
+
+    def update(self, magnitude=None, angle=None):
+        """Replaces the values of the magnitude and the angle, if
+        any provided. If both are None, does nothing but cry and
+        drink itself silly in the dark.
+
+        """
+        if magnitude:
+            self.magnitude = magnitude
+        if angle:
+            self.angle = angle
+
+
+class BayesianAgent(Object):
+    """Base class of the agent that adapts to new motor dynamics."""
+
+    def __init__(self, context_pars=None):
+        self.num_contexts = 3
+        if context_pars is None:
+            context_pars = self._default_pars()
+        contexts = self.init_contexts(context_pars)
+        self.p_contexts = self._prior_context()  # Will be updated
+
+    def _default_pars(self, ):
+        """Default parameters for the contexts, which are assumed to be like the
+        baseline."""
+        return [[0, 0]] * self.num_contexts
+
+    def _prior_context(self, ):
+        """Prior probabilities of observing all contexts. The baseline context
+        is arbitrarily set to be five times as likely as any other.
+
+        """
+        return np.array([5, *np.ones(self.num_contexts - 1)])
+
+    def _prior_force_magnitudes(self, ):
+        """Defines the parameters for the normal distribution over the magnitude
+        parameter of the force for each context. The priors are chosen to be
+        centered around 2 with a standard deviation of 1.
+
+        """
+        return [[0, 0.01]] + [[2, 1]] * (self.num_contexts - 1)
+
+    def init_contexts(self, context_pars=None):
+        "Initializes a context with default values, or those in --context_pars--, if
+        given.
+
+        Returns a list of instances of the class Context, where the
+        zero-th entry is the baseline.
+
+        """
+        contexts = [Context(baseline=True), ]
+        for pars in context_pars:
+            contexts.append(Context(pars))
+        return contexts

### 2021-02-24 notes/notes.tex

@@
 $\theta_t$. In this case, the inferred generalized state is given by
-$q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t |
-z_t)p(z_t|\theta_{t-1}, a_{t-1})$.
+\[
+  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
+\]
+
+Additionally, after having inferred the current state, motor adaptation is given by
+\[
+q(\gamma | z_t, z_{t-1}, a_{t-1}) \propto p(z_t | z_{t-1}, a_{t-1}, \gamma)p(z_{t-1})p(\gamma)
+\]
+
+where $\gamma$ are the parameters of the internal model for the dynamics,
+i.e. the function $f: (z_t, a_t) \rightarrow z_{t+1}$, which determines the beliefs of
+the agent as to how the system evolves after having taken action $a_t$.
 

### 2021-02-24 notes/notes.tex
Added RandomVariable. Not much else...

@@
 
-Spontaneous recovery. In an A-B-error_clamp paradigm, in which during B the
+Spontaneous recovery. In an A-B-error\_clamp paradigm, in which during B the
 opposite adaptation to A is learned, at the beginning of the error-clamp phase
@@
 
-\begin{chapter}{Testing the model in \ref{subsec:proposal}}
+\begin{chapter}{Implementing the adaptation model}
 In this section, I propose a simulated task in which the prowess of the model proposed in \ref{subsec:proposal} is put to the test.

### 2021-02-24 code/adaptation/model.py

@@ -4,11 +4,82 @@
 """Bayesian motor adaptation model."""
 
 
+class RandomVariable(Object):
+    """Defines a random variable, with a pdf and cdf, as well as a __call__ method
+    that samples from the pdf.
+
+    The distribution parameters can be updated via the method update_parameters.
+
+    Can be initiated with its own pdf and cdf. If None provided, it defaults to
+    numpy's implementation of a Gaussian.
+
+    """
+
+    def __init__(self,):
+        """Does nothing!"""
+        pass
+
+    def __call__(self, ):
+        return self.sample()
+
+    def sample(self, ):
+        """Samples a value using the pdf."""
+        rand = np.random.rand()
+        if self.var_type == 'continuous':
+            solution = sp.optimize.minimize(fun=lambda x: self.cdf(x) - rand,
+                                            x0=0)
+            sample = solution.x
+        else:
+            sample = np.argmin(abs(self.cdf - rand))
+        return sample
+
+    def plot_pdf(self, num_points=100, support=None):
+        """Plots a smooth approximation of the pdf, with --num_points-- points and
+        an x-range given by the elements of --support--.
+        """
+        raise NotImplementedError('Not yet!')
+
+
+class GaussGaussInference(RandomVariable):
+    """Performs Bayesian inference using Gaussian priors and likelihood. """
+    def __init__(self, mu=None, sd=None):
+        """Initiates the Gaussian priors with the given parameters or defaults.
+        """
+        if mu is None:
+            mu = 0
+        if sd is None:
+            sd = 1
+
+        self.pars = np.array([mu, sd])
+        self.pars_history = [self.pars]
+
+    def pdf(self, x):
+        """Evaluates the Gaussian pdf on the current parameters."""
+        return sp.stats.norm.pdf(x, loc=self.pars[0], scale=self.pars[1])
+
+    def cdf(self, x):
+        """Evaluates the Gaussian cdf on the current parameters."""
+        return sp.stats.norm.cdf(x, loc=self.pars[0], scale=self.pars[1])
+
+    def inference(self, likelihood_pars):
+        """Updates the parameters of the posterior."""
+        if priors_pars is None:
+            priors_pars = self.pars
+        mu_l, sd_l = likelihood_pars
+        mu_p, sd_p = self.pars
+        mu_post = (mu_l * sd_p ** 2 + mu_p * sd_p ** 2) / \
+            (sd_l ** 2 + sd_p ** 2)
[…]

### 2021-03-12 notes/notes.tex
*** empty log message ***

@@
 \usepackage[margin=1.5cm]{geometry}
+\usepackage{todonotes}
 \bibliographystyle{apalike}
@@
 
-\begin{chapter}{Implementing the adaptation model}
-In this section, I propose a simulated task in which the prowess of the model proposed in \ref{subsec:proposal} is put to the test.
-I additionally present simulation results and some writeup.
-Basically a mini-paper on this model.
+\begin{chapter}{Implementing the adaptation model} In this section, I propose a
+simulated task in which the prowess of the model proposed in
+\ref{subsec:proposal} is put to the test. I additionally present simulation
+results and some writeup. Basically a mini-paper on this model.
 
-\section{The Task}
-The tasks used for models like this one are usually \citep[e.g.][]{Lee_Dual_2009} reaching tasks: a participant sits in front of a table with shoulders immobilized, and has to bring her hand from the starting position A to a designated position B.
-This is done while holding a robotic arm that moves with the hand.
-The robotic arm is used to create force fields in different directions (CW and CCW, typically, but it's basically left and right) for the different contexts.
-The position of the participant's arm is recorded and deviations from the straight line are considered as errors ($e(t)$ in \ref{eqn:lee-learn}).
+\section{The Task} The tasks used for models like this one are usually
+\citep[e.g.][]{Lee_Dual_2009} reaching tasks: a participant sits in front of a
+table with shoulders immobilized, and has to bring her hand from the starting
+position A to a designated position B. This is done while holding a robotic arm
+that moves with the hand. The robotic arm is used to create force fields in
+different directions (CW and CCW, typically, but it's basically left and right)
+for the different contexts. The position of the participant's arm is recorded
+and deviations from the straight line are considered as errors ($e(t)$ in
+\ref{eqn:lee-learn}).
 
-To simplify the task, I propose the following: participants need to keep their hand still in the starting position for the duration of the trial.
-The robotic arm creates forces that will drive the arm away from its current position by pushing in a constant direction with constant force.
-This experiment has the same elements as the one by \cite{Lee_Dual_2009} but eliminates the necessity of an underlying movement.
-While it works wonders as a thought experiment and helps me see if the proposed model behaves the way I want, it must not be used as a real experiment because participants could simply use impedance for each trial, regardless of context.
+To simplify the task, I propose the following: participants need to keep their
+hand still in the starting position for the duration of the trial. The robotic
+arm creates forces that will drive the arm away from its current position by
+pushing in a constant direction with constant force. This experiment has the
+same elements as the one by \cite{Lee_Dual_2009} but eliminates the necessity
+of an underlying movement. While it works wonders as a thought experiment and
+helps me see if the proposed model behaves the way I want, it must not be used
+as a real experiment because participants could simply use impedance for each
+trial, regardless of context.
 
-In this experiment, every context $C_i$ is characterized by the strength and direction of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ..., N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force measured Cartesianly, and $w_i$ is the intensity of the force.
+In this experiment, every context $C_i$ is characterized by the strength and
+direction of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ...,
+N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force
+measured Cartesianly, and $w_i$ is the intensity of the force.
 
-As in \cite{Lee_Dual_2009}, I will simulate an A-B-clamp paradigm, which starts with a context $C_0 = \{0, 0\}$ in which no force is being applied.
-Then context $C_A$ is shown for a few trials, then context $C_B$, then an error-clamp block in which the hand is forced to stay at the initial position by the robotic arm.
+As in \cite{Lee_Dual_2009}, I will simulate an A-B-clamp paradigm, which starts
+with a context $C_0 = \{0, 0\}$ in which no force is being applied. Then
+context $C_A$ is shown for a few trials, then context $C_B$, then an
+error-clamp block in which the hand is forced to stay at the initial position
+by the robotic arm.
 
-Additionally, a contextual cue $\omega$ is shown to the participant which signals the current context.
-The experiment will be run with different levels of reliability of this cue.
-Different sessions will be run with different cue reliabilities.
+Additionally, a contextual cue $\omega$ is shown to the participant which
+signals the current context. The experiment will be run with different levels
+of reliability of this cue. Different sessions will be run with different cue
+reliabilities.
 
-An experimental session $\Xi$ is characterized by the reliability of the cue, the contexts avaiable and the order in which they are shown.
-In a given session, each trial begins with the cue and the XXX seconds during which the participant must keep the hand in the starting position.
-During the intertrial time, no forces are applied to the hand, but the hand must remain in the starting position.
+An experimental session $\Xi$ is characterized by the reliability of the cue,
+the contexts avaiable and the order in which they are shown. In a given
+session, each trial begins with the cue and the XXX seconds during which the
+participant must keep the hand in the starting position. During the intertrial
+time, no forces are applied to the hand, but the hand must remain in the
+starting position.
 
 \section{The model}
-The model is an implementation of \ref{subsec:proposal}. The generalized state is given by:
-\[
+The model is an implementation of \ref{subsec:proposal}. The generalized state
+is given by:
+\begin{equation}
 z_t = \{x, \theta_t, \gamma_t, \omega_t\}
-\]
-where $x$ is the current position of the hand in Cartesian coordinates, where the origin is the starting point.
-We assume that motor commands are issued every $\Delta t$ for simplicity.
-At the beginning of each time interval, the context is inferred combining the cue and the prediction error of the outcome of the previous motor command.
-We assume further than the cue and the force field are uncorrelated, which yields:
+\end{equation}
+where $x$ is the current position of the hand in Cartesian coordinates,
+where the origin is the starting point. We assume that motor commands are
+issued every $\Delta t$ for simplicity. At the beginning of each time interval,
+the context is inferred combining the cue and the prediction error of the
+outcome of the previous motor command. We assume further than the cue and the
+force field are uncorrelated, which yields:
 \begin{align}
@@
 \end{align}
-The terms $p(C_t)$ refer to the prior probability of the context at the beginning of the trial: at the first trial, this refers to prior beliefs over which contexts are more common/likely.
-At each subsequent trial, it incorporates the belief that has so far been accumulated, given previous trials.
+The terms $p(C_t)$ refer to the prior probability of the context at
+the beginning of the trial: at the first trial, this refers to prior beliefs
+over which contexts are more common/likely. At each subsequent trial, it
+incorporates the belief that has so far been accumulated, given previous
+trials.
 
-Once the posterior over contexts has been calculated for the current trial, it is used to select the current context.
-A simple rule is to sample the context directly from the posterior distribution.
-However, a useful extension of this is to use a softmax function to create a distribution:
+Once the posterior over contexts has been calculated for the current trial, it
+is used to select the current context. A simple rule is to sample the context
+directly from the posterior distribution. However, a useful extension of this
+is to use a softmax function to create a distribution:
 
 \begin{equation}
-S(C^{(t)}) = \frac{e^{\kappa q\left(C^{(t)}\right)}}{\displaystyle \sum_{i=1}^{N_c}e^{\kappa q\left(C_i\right)}}
+S(C^{(t)}) = \frac{e^{\kappa
+q\left(C^{(t)}\right)}}{\displaystyle \sum_{i=1}^{N_c}e^{\kappa
+q\left(C_i\right)}}
 \end{equation}
-where $q(C_i)$ is given by \eref{eqn:estimated-context}.
-The function $S(\cdot)$ is then used to sample the (inferred) current context.
-This representation has the advantage of adding a parameter with which the sampling behavior can be characterized, ranging from choosing uniformly when $\kappa = 0$ to choosing the context with the highest probability deterministically as $\kappa \rightarrow \infty$.
+where $q(C_i)$ is given by \eref{eqn:estimated-context}. The
+function $S(\cdot)$ is then used to sample the (inferred) current context. This
+representation has the advantage of adding a parameter with which the sampling
+behavior can be characterized, ranging from choosing uniformly when $\kappa =
+0$ to choosing the context with the highest probability deterministically as
+$\kappa \rightarrow \infty$.
 
-In the case where the hand position can be determined by the participant with perfect accuracy (ignoring visual and proprioceptive error), we have that $p(x_\tau) = \delta(x_\tau, \hat x)$, where $\hat x$ is the real position of the hand, and $\delta (\cdot, \cdot)$ is the Kronecker delta function.
-We assume this to be the case during this experiment.
+In the case where the hand position can be determined by the participant with
+perfect accuracy (ignoring visual and proprioceptive error), we have that
+$p(x_\tau) = \delta(x_\tau, \hat x)$, where $\hat x$ is the real position of
+the hand, and $\delta (\cdot, \cdot)$ is the Kronecker delta function. We
+assume this to be the case during this experiment.
 
-For simplicity, we assume that participants know how many distinct contexts will be shown to them, and therefore all the $C_x$ have been created by the agent before the session begins.
-By default, all contexts assume a force with angle and magnitude equal zero, and their true values are learned during the experiment (see below).
+For simplicity, we assume that participants know how many distinct contexts
+will be shown to them, and therefore all the $C_x$ have been created by the
+agent before the session begins. By default, all contexts assume a force with
+angle and magnitude equal zero, and their true values are learned during the
+experiment (see below).
 
-For mathematical tractability (i.e.
-to avoid circular distributions), we will limit contexts to one where the force points left (angle $\pi$) and one where the force points right (angle 0), in addition to the baseline context.
-In this scheme, participants need only learn the magnitude of the force for each context.
-We will assume that for each context, the prior distribution over the magnitude of the force is Gaussian:
-\[
-p(\gamma) = N(1, 0.1)
-\]
-We will show that the form of this prior affects learning rate, but not the final learned magnitude.
+For mathematical tractability (i.e. to avoid circular distributions), we will
+limit contexts to one where the force points left (angle $\pi$) and one where
+the force points right (angle 0), in addition to the baseline context. In this
+scheme, participants need only learn the magnitude of the force for each
+context. We will assume that for each context, the agent's belief over the
+magnitude of the force is given by a normal distribution:
+\begin{equation}
+p(\gamma) = N(\mu_f, \sigma_f) \label{eqn:data-dist}
+\end{equation}
+where $\mu_f$ and $\sigma_f$ are parameters to be estimated at each
+trial. Before the experiment begins, the agent will have a prior distribution
+over these parameters. A standard Bayesian approach is to choose NormalGamma
+priors:
+\begin{equation}
+p(\mu_f, \sigma_f) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
+\end{equation}
+where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
+model. Because participants are told that there are three contexts (baseline,
+left and right), we assume the following hyperparameter values:
+\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
+  \nu_f^{(0)} &= 0.1 \\
+  \alpha_f^{(0)} &= 1 \\
+  \beta_f^{(0)} &= 0.5
+\end{align}
+Note that $\mu_f^{(0)}$ is different for each context and is thus
+provided as a vector $(0, -1, 1)$ representing the baseline, left and right
+contexts, respectively. For all other hyperparameters, all contexts have the
+same value. These values ensure that the prior over the hyperparameters
+$(\mu_f, \sigma_f)$ has the mode at $(\mu_f^{(0)}, 1)$. We will show that the
+choice of these hyperparameters affects learning rate, but not the final
+learned magnitude \todo{Check this}.
 
-After the generalized state has been inferred given the information at the beginning of the trial, the next step is to infer the parameters of the context, as in \eref{eqn:update-parameters}:
+The update equations for the magnitud parameters are given by:
 \begin{equation}
-q(C_t | x_t, x_{t-1}, a_{t-1}) \propto p(x_t | x_{t-1}, a_{t-1}, C_t)p(C_t) \label{eqn:context-from-x}
+q(\mu_f, \sigma_f | x_t, x_{t-1}, a_{t-1}) \propto p(x_t |
+x_{t-1}, a_{t-1}, C_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
 \end{equation}
+where $p(x_t | x_{t-1}, a_{t-1}, C_t)$ is a Gaussian distribution centered
+around $\mu_a$ with standard deviation $\sigma_a$, which is a free parameter of
+the model. It is assumed that the standard deviation over outcomes ($\sigma_a$)
+is related to that of the parameter $\mu_f$ as follows:
+\begin{equation}
+\sigma_f = \sigma_a / \nu_f
+\end{equation}
+which makes $\nu_f$ a parameter to estimate. Because we chose the priors to be
+NormalGamma and the likelihood to be Gaussian, the posteriors are also
+NormalGamma, and the parameters are updated after one time step via:
+\begin{align}
+  \mu_f^{(t)} &= \frac{\nu_f^{(t-1)} \mu_f^{(t-1)} + x_t}{\nu_f + 1} \\
+  \nu_f^{(t)} &= \nu_f^{(t-1)} + 1 \\
+  \alpha_f^{(t)} &= \alpha_f^{(t-1)} + 0.5 \\
+  \beta_f^{(t)} &= \beta_f^{(t-1)} + \frac{\nu_f^{(t-1)}}{\nu_f^{(t-1)} +
+                  1}\frac{\left(x - \mu_f^{(t-1)}\right)^2}{2}
+\end{align}
 
-where $p(C_t)$ is either a prior belief of the likelihood of each context, or the probability inferred up to the previous trial.
-Note that, because we assumed that the position of the hand is directly observable without any noise, there are no terms for the uncertainty of $x_t$ or $x_{t-1}$.
-
-With this, \eref{eqn:context-from-x} becomes:
-\begin{equation}
-q(\gamma | x_t, x_{t-1}, a_{t-1}) \propto N(\hat x_t, \sqrt{\Delta t})p(\gamma) \label{eqn:data-dist}
-\end{equation}
-where $\hat x_t$ is the predicted value of the position $x_t$ given the previous state and action and the force field of the context; see \eref{eqn:dynamics} below.
-Because we chose the priors $p(\gamma)$ to be Gaussian, the posteriors $q(\gamma)$ after every trial will also be a Gaussian distribution with parameters:
+The final step is action selection. Action selection is done at the level of
+desired outcome (as opposed to motor commands) following the equations:
 \begin{align}
-  \mu_\gamma^{(t)} &= \frac{\mu_\gamma^{(t-1)}\left(\sigma_\gamma^{(d)}\right)^2 + \mu_\gamma^{(d)}\left(\sigma_\gamma^{(t-1)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2} \\
-  \sigma_\gamma^{(t)} & = \sqrt{\frac{\left(\sigma_\gamma^{(t-1)}\right)^2\left(\sigma_\gamma^{(d)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2}}
-\end{align}
-where the superscripts $(t)$ and $(t-1)$ refer to the inferred values for the current and previous trials, respectively, and $(d)$ to the parameters in \eref{eqn:data-dist}.
-
-The final step is action selection. Action selection is done at the level of desired outcome (as opposed to motor commands) following the equations:
-\begin{align}
-  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta t}\epsilon\\ \label{eqn:dynamics}
+  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta
+            t}\epsilon\\ \label{eqn:dynamics}
   f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
@@
 \end{align}
-where $\gamma_t$ and $\theta_t$ are the participant's estimate of the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian motor error component.
+where $\gamma_t$ and $\theta_t$ are the participant's estimate of
+the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian
+motor error component.
 

### 2021-03-15 notes/notes.tex
Added Cue and Clamp to both task and models.

@@
 In this experiment, every context $C_i$ is characterized by the strength and
-direction of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ...,
+8'rection of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ...,
 N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force
@@
 is to use a softmax function to create a distribution:
-
+ !''¿
 \begin{equation}
@@
 experiment (see below).
-
+1234568¿|
 For mathematical tractability (i.e. to avoid circular distributions), we will
@@
 scheme, participants need only learn the magnitude of the force for each
-context. We will assume that for each context, the agent's belief over the
-magnitude of the force is given by a normal distribution:
+context.
+
+The generative model for the force can take on different shapes for the prior
+and posterior distributions. In this work, we explore three different forms of
+this generative model which differ on the assumptions the agent makes regarding
+how the force fields are being generated by the environment. Each version of
+the model is discussed separately below and their pros and cons discussed;
+additionally, simulations are presented in the following sections.
+
+To infer the parameters of the posterior distributions, we make use of Bayes'
+theorem. In the following equations, we drop the dependence on the context and
+observations for notational simplicity:
+\begin{equation}
+q(\theta | x_{1:t}) \propto p(x_{1:t} | \theta)p(\beta)
+\end{equation}
+where $\theta$ are the parameters of the force field to be estimated and
+$x_{1:t}$ are the data observed until time $t$. Across all variants of the
+model presented below, the likelihood will be a Gaussian:
+\begin{align}
+  p(x_t | \theta) &= N(\mu_x, \sigma_x)  \\
+  \mu_x &= \theta_{t-1} + x \\
+  \sigma_x &= \xi_0 1
+\end{align}
+This likelihood is the probability of observing $x$ given the previous estimate
+of the parameters of the force field $\theta_{t-1}$, and given the dynamics of
+the system (see below).
+
+The final part of the model to establish is action selection. Action selection
+is done at the level of desired outcome (as opposed to motor commands)
+following the equations:
+\begin{align}
+  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta
+            t}\epsilon\\ \label{eqn:dynamics}
+  f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
+  a_t &= \argmax_{a_t}p(x_{t+1} = (0, 0) | x_t, a_t, f_t)
+\end{align}
+where $\gamma_t$ and $\theta_t$ are the participant's estimate of
+the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian
+motor error component.
+
+\subsection{Fixed force value}
+In this version, the force field is assumed to have a fixed value which the
+agent tries to infer. The agent's uncertainty on this inference is reflected on
+the parameters of the posterior over the force directly. For simplicity, we
+assume that the priors (and posteriors) are normal with mean and standard
+deviation to be updated at every time step. Therefore, the posterior over the
+magnitude of the force for each context is given by:
+\begin{equation}
+q_t(\gamma) = N(\mu_{\gamma}, \sigma_\gamma)
+\end{equation}
+where $q_t(\gamma)$ is the posterior of the magnitude $\gamma$ of the force
+field distribution after the observations at time $t$, and $\mu_\gamma$ and
+$\sigma_\gamma$ are the inferred parameters of the distribution.
+
+Because the likelihood is a Gaussian with support $\mu_\gamma$, the update
+equations are given by:
+\begin{align}
+  \mu_\gamma^{(t)} &= \frac{\mu_\gamma^{(t-1)}\left(\sigma_\gamma^{(d)}\right)^2 + \mu_\gamma^{(d)}\left(\sigma_\gamma^{(t-1)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2} \\
+  \sigma_\gamma^{(t)} & = \sqrt{\frac{\left(\sigma_\gamma^{(t-1)}\right)^2\left(\sigma_\gamma^{(d)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2}}
+\end{align}
+where the superscripts $(t)$ and $(t-1)$ refer to the inferred values for the current and previous trials, respectively, and $(d)$ to the parameters in \eref{eqn:data-dist}.
+
+\subsection{Fixed-$\sigma$ model}
+In this model, the force at each trial is assumed to be sampled from a normal
+distribution and the mean of this distribution is inferred from the
+observations, while the standard deviation is assumed to be known. The mean of
+the force is assumed to also be normal with unknown mean and standard
+deviation:
+\begin{align}
+  p(\gamma) &= N(\mu_t, \sigma) \\
+  p(\mu_t) &= N(\mu_mu, \sigma_\mu)
+\end{align}
+In this manner, the uncertainty for this agent is at the level of the estimate
+of the parameters of the force field, as opposed to being directly on the
+force. Because the priors for $\mu_\mu$ are normal and the standard deviation
+$\sigma$ (not $\sigma_mu$) is known, the update equations are given by:
+\begin{align}
+  \mu_\mu &= \frac{\sigma^2}{\sigma^2 + \sigma_\mu^2}\mu_\mu + \frac{\sigma_\mu^2}{\sigma^2 + \sigma_\mu^2} x \\
+  \sigma_\mu &= \frac{\sigma^2\sigma_\mu^2}{\sigma^2 + \sigma_\mu^2}
+\end{align}
+where $x$ is the observation.
+
+\subsection{Full uncertainty}
+In this version, we will assume that for each context, the agent's belief over
+the magnitude of the force is given by a normal distribution:
 \begin{equation}
@@
 
-The final step is action selection. Action selection is done at the level of
-desired outcome (as opposed to motor commands) following the equations:
-\begin{align}
-  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta
-            t}\epsilon\\ \label{eqn:dynamics}
-  f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
-  a_t &= \argmax_{a_t}p(x_{t+1} = (0, 0) | x_t, a_t, f_t)
-\end{align}
-where $\gamma_t$ and $\theta_t$ are the participant's estimate of
-the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian
-motor error component.
-
 \section{Simulation results}

### 2021-03-15 code/adaptation/model.py

@@ -1,10 +1,17 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/model.py
 
+import ipdb
+
+import numpy as np
+import scipy as sp
+from scipy import stats
+import pandas as pd
+
 """Bayesian motor adaptation model."""
 
 
-class RandomVariable(Object):
+class RandomVariable(object):
     """Defines a random variable, with a pdf and cdf, as well as a __call__ method
     that samples from the pdf.
 
@@ -55,16 +62,14 @@ class GaussGaussInference(RandomVariable):
 
     def pdf(self, x):
         """Evaluates the Gaussian pdf on the current parameters."""
-        return sp.stats.norm.pdf(x, loc=self.pars[0], scale=self.pars[1])
+        return stats.norm.pdf(x, loc=self.pars[0], scale=self.pars[1])
 
     def cdf(self, x):
         """Evaluates the Gaussian cdf on the current parameters."""
-        return sp.stats.norm.cdf(x, loc=self.pars[0], scale=self.pars[1])
+        return stats.norm.cdf(x, loc=self.pars[0], scale=self.pars[1])
 
     def inference(self, likelihood_pars):
         """Updates the parameters of the posterior."""
-        if priors_pars is None:
-            priors_pars = self.pars
         mu_l, sd_l = likelihood_pars
         mu_p, sd_p = self.pars
         mu_post = (mu_l * sd_p ** 2 + mu_p * sd_p ** 2) / \
@@ -74,7 +79,7 @@ class GaussGaussInference(RandomVariable):
         self.pars_history.append(self.pars)
 
 
-class Context(Object):
+class Context(object):
     """Stores the force and direction of a context. If called, returns the force in
     2D Cartesian coordinates.
 
@@ -84,10 +89,11 @@ class Context(Object):
         """Well..."""
         self.magnitude = magnitude
         self.angle = angle
+        self.baseline = baseline
 
     def __call__(self, x=None):
         """Returns the force."""
-        if baseline:
+        if self.baseline:
             return 0
         return self.magnitude * np.array([np.cos(self.angle),
                                           np.sin(self.angle)])
@@ -104,14 +110,13 @@ class Context(Object):
             self.angle = angle
 
 
-class BayesianAgent(Object):
+class BayesianAgent(object):
     """Base class of the agent that adapts to new motor dynamics."""
 
     def __init__(self, context_pars=None):
         self.num_contexts = 3
         if context_pars is None:
[…]

### 2021-03-15 code/adaptation/model.py
Minor correction on likelihood with max force

@@ -399,21 +399,26 @@ class LeftRightAgent(object):
     # Fixed parameters
     angles = np.array([0, 1, -1])
     force_sds = np.array([0.01, 0.2, 0.2])
-    
+
     # Free parameters:
     action_sd = 0.01  # SD of a Gaussian for action uncertainty
-    cue_noise = 0.1  # If ix_cue is observed, the posterior over the corresponding
-                     # context is 1 - 2 * cue_noise.
+    cue_noise = 0.1  # If ix_cue is observed, the posterior over the
+                     # corresponding context is 1 - 2 * cue_noise.
 
     sample_context_mode = 'mode'
     sample_force_mode = 'mode'
 
+    max_force = 3.5  # Maximum force (abs value) the agent can exert.
+
     context_noise = 0.0001
     context_sd_constant = 0.1
     context_sd_exponent = 3
-    
+
     def __init__(self, obs_sd=None):
-        """Initializes the known left and right contexts, as well as baseline."""
+        """Initializes the known left and right contexts, as well as
+        baseline.
+
+        """
         self.num_contexts = 3
         _, self.magnitudes = self.__init_contexts()
         self.magnitude_history = [self.magnitudes]
@@ -456,6 +461,8 @@ class LeftRightAgent(object):
         c_context = self.sample_context('mode')
         c_force = self.magnitudes[c_context][0]
         action = -c_pos - c_force
+        if abs(action) > self.max_force:
+            action = action / abs(action) * self.max_force
         self.action = action
         self.action_history.append(action)
         return action
@@ -476,12 +483,9 @@ class LeftRightAgent(object):
         """
         ix_context = self.sample_context(t=-2)
         mu_p, sd_p = self.magnitudes[ix_context]
-        p_context = np.exp(self.log_context[ix_context])
         mu_l, sd_l = self.predict_outcome()[1][ix_context]
-        mu_l = mu_p + self.hand_position
-        sd_con = self._sd_context(p_context)
-        # sd_p = np.sqrt(sd_p ** 2 * sd_con ** 2 / (sd_p ** 2 + sd_con ** 2))
-        mu_post = (mu_l * sd_p ** 2 + mu_p * sd_l ** 2 ) / \
+        mu_l = mu_p + self.hand_position - mu_l
+        mu_post = (mu_l * sd_p ** 2 + mu_p * sd_l ** 2) / \
[…]

### 2021-03-17 code/adaptation/model.py
Added MuSD model

@@ -154,242 +154,6 @@ class BayesianAgent(object):
         return contexts
 
 
-class LeftRightGenAgent(object):
-    """Subclass that knows two contexts(plus baseline): one pointing left and one
-    pointing right. The magnitude of these forces are inferred during the
-    experiment.
-
-    This model assumes that the environment generates the force field 
-    stochastically, sampling from a normal distribution with unknown parameters
-    that are estimated after every observation. It uses NormalGamma priors for
-    these parameters and drinks cold coffee. Psycho.
-
-    """
-    def __init__(self,):
-        """Initializes the known left and right contexts, as well as baseline."""
-        self.num_contexts = 3
-        self.angles, self.mags_hyperpars = self.__init_contexts()
-        self.magnitude_history = [self.mags_hyperpars]
-        self.decision_history = []
-        self.hand_position_history = [0]
-        self.hand_position = 0
-        self.action_history = [0]
-        self.log_context = np.log([0.8, 0.1, 0.1])
-        self.log_context_history = [self.log_context]
-        self.action_sd = 0.01  # SD of a Gaussian for action uncertainty
-        self.obs_sd = 1  # fake observation noise. Good for inference.
-        
-        self.sample_context_mode = 'mode'
-        self.sample_force_mode = 'mode'
-
-    def __init_contexts(self, ):
-        r"""Creates baseline, left and right contexts, with a known angle (no force,
-        right, left), as well as the hyperparameters for the magnitudes,
-        assumed to be NormalGamma with mu, nu, alpha and beta parameters.
-
-        """
-        angles = np.array([0, 1, -1])
-        magnitudes = np.array([[angles[0], 1, 1, 1], [angles[1], 1, 1, 1],
-                               [angles[2], 1, 1, 1]])  # mu, nu, alpha, beta
-        return angles, magnitudes
-
-    def make_decision(self, ):
-        """Makes a decision after all the inferences for the trial have been
-        carried out.
-
-        The logic is to choose an action such that
-        p(pos) - p(force) + p(s_(t+1)| action) = 0
-
-        A simple proxy is to use the mode of each one of this, which is what
-        this function does for now.
-
-        The decision is returned and also saved into self.decision_history.
-
-        """
-        c_pos = self.hand_position
-        c_context = self.sample_context('mode')
-        c_force = self.mags_hyperpars[c_context][0]
-        action = -c_pos - c_force
-        self.action = action
-        self.action_history.append(action)
[…]

### 2021-03-17 code/adaptation/pars.py

@@ -1,0 +1,6 @@
+pars = {'obs_noise': [0, 0.01],  # Observation noise N(mu, sd)
+        'force_noise': [[0, 0.0001], [0, 0.2], [0, 0.2]],  # Noise of the force process
+        'forces': [[0, 0], [1, 3], [-1, 2]],  # [angle, magnitude] for each force
+        'num_trials': 10,  # Trials per miniblock
+        'context_seq': [0, 1, 2, 'clamp', 0]  # Sequence of context type to use.
+        }

### 2021-03-19 notes/notes.tex
Added simulations, including some parallel stuff to run.
Also, embarassingly, I had forgotten to add the task file to the
git, even though it's been around for a while.

@@
 \section{Simulation results}
+\subsection{Effects of observation noise on inference}
+Observation noise refers to the noise on the position of the hand.
+The generating process is a Gaussian with $\sigma = 0.1$, which is determined in ./code/adaptation/pars.py.
+However, the agent has a belief on this value, which is fixed for the agent.
+This value (obs_sd, which can be passed on instanciation) can be used in lieu of an explicit $\Delta t$ parameter which controls the size of the time step
 

### 2021-03-19 code/adaptation/model.py

@@ -7,6 +7,7 @@ import numpy as np
 import scipy as sp
 from scipy import stats
 import pandas as pd
+from matplotlib import pyplot as plt
 
 """Bayesian motor adaptation model."""
 
@@ -110,64 +111,20 @@ class Context(object):
             self.angle = angle
 
 
-class BayesianAgent(object):
-    """Base class of the agent that adapts to new motor dynamics."""
-
-    def __init__(self, context_pars=None):
-        self.num_contexts = 3
-        if context_pars is None:
-            context_pars = self._default_pars()
-        self.p_contexts = self._prior_context()  # Will be updated
-        self.gamma = GaussGaussInference()
-
-    def _default_pars(self, ):
-        """Default parameters for the contexts, which are assumed to be like the
-        baseline."""
-        return [[0, 0]] * self.num_contexts
-
-    def _prior_context(self, ):
-        """Prior probabilities of observing all contexts. The baseline context
-        is arbitrarily set to be five times as likely as any other.
-
-        """
-        return np.array([5, *np.ones(self.num_contexts - 1)])
-
-    def _prior_force_magnitudes(self, ):
-        """Defines the parameters for the normal distribution over the magnitude
-        parameter of the force for each context. The priors are chosen to be
-        centered around 2 with a standard deviation of 1.
-
-        """
-        return [[0, 0.01]] + [[2, 1]] * (self.num_contexts - 1)
-
-    def init_contexts(self, context_pars=None):
-        """Initializes a context with default values, or those in --context_pars--, if
-        given.
-
-        Returns a list of instances of the class Context, where the
-        zero-th entry is the baseline.
-
-        """
-        contexts = [Context(baseline=True), ]
-        for pars in context_pars:
-            contexts.append(Context(pars))
-        return contexts
-
-
 class LeftRightAgent(object):
     """Subclass that knows two contexts(plus baseline): one pointing left and one
     pointing right. The magnitude of these forces are inferred during the
     experiment.
 
     """
-    # Fixed parameters
-    angles = np.array([0, 1, -1])
-    force_sds = np.array([0.01, 0.2, 0.2])
-
-    # Free parameters:
+    name = 'LR'
+    # Free parameters with default values. Can be overwritten by __init__
[…]

### 2021-03-19 code/adaptation/pars.py

@@ -1,6 +1,6 @@
-pars = {'obs_noise': [0, 0.01],  # Observation noise N(mu, sd)
+pars = {'obs_noise': [0, 0.1],  # Observation noise N(mu, sd)
         'force_noise': [[0, 0.0001], [0, 0.2], [0, 0.2]],  # Noise of the force process
-        'forces': [[0, 0], [1, 3], [-1, 2]],  # [angle, magnitude] for each force
-        'num_trials': 10,  # Trials per miniblock
+        'forces': [[0, 0], [1, 2], [-1, 2]],  # [angle, magnitude] for each force
+        'num_trials': 300,  # Trials per miniblock
         'context_seq': [0, 1, 2, 'clamp', 0]  # Sequence of context type to use.
         }

### 2021-03-19 code/adaptation/simulations.py

@@ -1,0 +1,95 @@
+# -*- coding: utf-8 -*-
+# ./adaptation/simulations.py
+
+"""Some simulations for the agent and the task."""
+
+import multiprocessing as mp
+from itertools import product
+
+import numpy as np
+from matplotlib import pyplot as plt
+
+import model
+import task_hold_hand as thh
+
+
+def test_obsnoise(fignum=1):
+    """Runs the game with increasingly high observation noise, for all three
+    agents. Plots the results in a big platter for consumption.
+
+    """
+    cue_noise = 0.001
+    obs_noises = np.arange(0, 0.3, 0.1)
+    fig, axes = plt.subplots(3, len(obs_noises), num=fignum, clear=True,
+                             sharex=True, sharey=True)
+    for ix_noise, obs_noise in enumerate(obs_noises):
+        kwargs = {'angles': [0, 0, 0], 'obs_sd': obs_noise,
+                  'cue_noise': cue_noise}
+        agents = [model.LeftRightAgent(**kwargs),
+                  model.LRMean(**kwargs),
+                  model.LRMeanSD(**kwargs)]
+        for ix_agent, agent in enumerate(agents):
+            thh.run(agent)
+            thh.run(agent, cue_key=[0, 2, 1])
+            agent.plot_mu(axis=axes[ix_agent, ix_noise])
+            agent.plot_position(axis=axes[ix_agent, ix_noise])
+            if ix_agent == 0:
+                axes[ix_agent, ix_noise].set_title('Obs Noise: {}'.format(obs_noise))
+    plt.ylabel('Adaptation')
+    plt.xlabel('Trial')
+    plt.draw()
+    plt.show(block=False)
+
+
+def test_deadaptation(fignum=2):
+    """Runs the game once with some cues and once again with cues that have
+    the opposite meaning.
+
+    """
+    kwargs = {'angles': [0, 0, 0], 'obs_sd': 1,
+              'cue_noise': 0.001}
+    fig, axes = plt.subplots(3, 1, num=fignum, clear=True,
+                             sharex=True, sharey=True, squeeze=True)
+    agents = [model.LeftRightAgent(**kwargs),
+              model.LRMean(**kwargs),
+              model.LRMeanSD(**kwargs)]
+    for ix_agent, agent in enumerate(agents):
+        thh.run(agent)
+        thh.run(agent, cue_key=[0, 2, 1])
[…]

### 2021-03-19 code/adaptation/task_hold_hand.py

@@ -1,0 +1,162 @@
+# -*- coding: utf-8 -*-
+# ./adaptation/task_hold_hand.py
+import ipdb
+from datetime import datetime
+import argparse
+from importlib import import_module
+
+import numpy as np
+from scipy import stats
+import pandas as pd
+
+import model
+
+"""Task described in section 7.1 of notes.pdf in which the participant
+must hold her hand in the starting position in different force fields.
+The game is meant to be played by model.LeftRightAgent or its children.
+
+"""
+
+# Import task parameters from a file, if provided:
+PARSER = argparse.ArgumentParser(description='Run experiment')
+PARSER.add_argument('-p', '--pars', dest='pars', default=None,
+                    help='Filename for the parameters')
+ARGS = PARSER.parse_args()
+parsfile = ARGS.pars
+if parsfile is None:
+    parsfile = 'pars'
+if parsfile.endswith('.py'):
+    parsfile = parsfile[:-3]
+pars = import_module(parsfile).pars
+
+
+def run(agent=None, save=False, filename=None, **trial_kwargs):
+    """Runs an entire game of the holding-hand task."""
+    if agent is None:
+        agent = model.LeftRightAgent(obs_sd=pars['obs_noise'][1])
+    outs = []
+    for ix_miniblock in pars['context_seq']:
+        out = miniblock(ix_miniblock, 0, agent, **trial_kwargs)
+        outs.append(out)
+    data = np.stack(outs, axis=0)
+    pandata = pd.DataFrame(data.reshape((-1, 5)),
+                           columns=('action', 'force', 'pos(t)', 'pos(t+1)',
+                                    'ix_context'))
+    pandata.rename_axis('trial', inplace=True)
+    pandagent = agent.pandify_data()
+    if save:
+        save_pandas(pandata, pandagent, filename)
+    return pandata, pandagent, agent
+
+
+def miniblock(ix_context, hand_position, agent, **trial_kwargs):
+    """Runs a miniblock of the task."""
+    outs = []
+    for ix_trial in range(pars['num_trials']):
+        out = trial(ix_context, hand_position, agent, **trial_kwargs)
+        outs.append(out)
+        hand_position = out[3]
+    outs = np.stack(outs, axis=0)
+    return outs
+
+
+def trial(ix_context, hand_position, agent, cue_key=None):
+    """Runs a single trial of the task.
+
+    Parameters
+    ----------
+    ix_context : int or string
+    Index that indicates the current context. Should index everything
[…]

### 2021-03-19 code/adaptation/simulations.py
Fixed minor bug with simulations. Should run now

@@ -82,14 +82,14 @@ def grid_sims():
     seeds = np.unique(seeds)
 
     for ix_pars, pars in enumerate(all_pars):
-        my_robots.apply_async(_one_sim, pars)
+        my_robots.apply_async(_one_sim, list(pars))
 
     my_robots.close()
     my_robots.join()
 
 
-def _one_sim(obs_noise, cue_noise, agent):
+def _one_sim(obs_noise, cue_noise, agent_fun):
     """Runs one iteration for grid_sims(). """
+    agent = agent_fun(obs_sd=obs_noise, cue_noise=cue_noise)
     filename = 'grid_sims_{}_{}_{}.pi'.format(obs_noise, cue_noise, agent.name)
-    thh.run(agent=agent, obs_noise=obs_noise, cue_noise=cue_noise,
-            save=True, filename=filename)
+    thh.run(agent=agent, save=True, filename=filename)

### 2021-03-19 code/adaptation/model.py
Minor bug fix with pandification of lrms agent

@@ -579,5 +579,6 @@ class LRMeanSD(LeftRightAgent):
                                         'mag_alpha_0', 'mag_alpha_1', 'mag_alpha_2',
                                         'mag_beta_0', 'mag_beta_1', 'mag_beta_2',
                                         'trial'])
+        pandata.set_index('trial', inplace=True)
         return pandata
         

### 2021-03-19 code/adaptation/simulations.py

@@ -5,9 +5,11 @@
 
 import multiprocessing as mp
 from itertools import product
+import glob
 
 import numpy as np
 from matplotlib import pyplot as plt
+import pandas as pd
 
 import model
 import task_hold_hand as thh
@@ -93,3 +95,27 @@ def _one_sim(obs_noise, cue_noise, agent_fun):
     agent = agent_fun(obs_sd=obs_noise, cue_noise=cue_noise)
     filename = 'grid_sims_{}_{}_{}.pi'.format(obs_noise, cue_noise, agent.name)
     thh.run(agent=agent, save=True, filename=filename)
+
+
+def read_data():
+    """Reads the data files with the format in _one_sim() and creates
+    a panda with it, including the parameter values as columns.
+
+    """
+    data_folder = './sim_data/'
+    files = glob.iglob(data_folder + 'grid_sims_*.pi')
+    pandas = []
+    for file in files:
+        split = file.split('_')
+        panda = pd.read_pickle(file)
+        len_panda = len(panda)
+        obs_noise = float(split[3]) * np.ones(len_panda)
+        cue_noise = float(split[4]) * np.ones(len_panda)
+        agent = [split[5][:-3]] * len_panda
+        panda['obs_noise'] = obs_noise
+        panda['cue_noise'] = cue_noise
+        panda['agent'] = agent
+        pandas.append(panda)
+    # pandatron = pd.concat(pandas, axis=0)
+    # return pandatron
+    return pandas

### 2021-03-22 code/adaptation/simulations.py
Fixed bug with pandification of LRMD agent

@@ -72,6 +72,7 @@ def grid_sims():
     """
     obs_noises = np.arange(0, 2, 0.1)
     cue_noises = np.arange(0, 0.05, 0.005)
+
     agents = [model.LeftRightAgent,
               model.LRMean,
               model.LRMeanSD]
@@ -116,6 +117,5 @@ def read_data():
         panda['cue_noise'] = cue_noise
         panda['agent'] = agent
         pandas.append(panda)
-    # pandatron = pd.concat(pandas, axis=0)
-    # return pandatron
-    return pandas
+    pandatron = pd.concat(pandas, axis=0)
+    return pandatron

### 2021-03-22 code/adaptation/model.py
Another minor bug with pandification fixed

@@ -492,6 +492,7 @@ class LRMeanSD(LeftRightAgent):
 
     """
     name = 'LRMS'
+
     def __init__(self, *args, **kwargs):
         """Initializes hyperparameters and calls LeftRightAgent.__init__ """
         super().__init__(*args, **kwargs)
@@ -566,19 +567,24 @@ class LRMeanSD(LeftRightAgent):
         mag_mu_sd = magnitudes[..., 1]
         mag_sd_mu = magnitudes[..., 2]
         mag_sd_sd = magnitudes[..., 3]
+        hand = self.hand_position_history
+        actions = self.aciton_history
         trial_number = np.arange(context.shape[0]) - 1
         aggregate = np.stack([*context.T, max_context,
                               *mag_mu_mu.T, *mag_mu_sd.T,
-                              *mag_sd_mu.T, *mag_sd_sd.T, trial_number],
+                              *mag_sd_mu.T, *mag_sd_sd.T, hand, actions,
+                              trial_number],
                              axis=1)
         pandata = pd.DataFrame(aggregate,
                                columns=['con0', 'con1', 'con2',
                                         'con_t',
                                         'mag_mu_0', 'mag_mu_1', 'mag_mu_2',
                                         'mag_nu_0', 'mag_nu_1', 'mag_nu_2',
-                                        'mag_alpha_0', 'mag_alpha_1', 'mag_alpha_2',
-                                        'mag_beta_0', 'mag_beta_1', 'mag_beta_2',
-                                        'trial'])
+                                        'mag_alpha_0', 'mag_alpha_1',
+                                        'mag_alpha_2',
+                                        'mag_beta_0', 'mag_beta_1',
+                                        'mag_beta_2',
+                                        'hand', 'action', 'trial'])
+        pandata.reset_index(drop=True, inplace=True)
         pandata.set_index('trial', inplace=True)
         return pandata
-        

### 2021-03-23 code/adaptation/model.py
Changed the way cues are generated for more flexibility

@@ -568,7 +568,7 @@ class LRMeanSD(LeftRightAgent):
         mag_sd_mu = magnitudes[..., 2]
         mag_sd_sd = magnitudes[..., 3]
         hand = self.hand_position_history
-        actions = self.aciton_history
+        actions = self.action_history
         trial_number = np.arange(context.shape[0]) - 1
         aggregate = np.stack([*context.T, max_context,
                               *mag_mu_mu.T, *mag_mu_sd.T,

### 2021-03-23 code/adaptation/pars.py

@@ -1,6 +1,9 @@
 pars = {'obs_noise': [0, 0.1],  # Observation noise N(mu, sd)
         'force_noise': [[0, 0.0001], [0, 0.2], [0, 0.2]],  # Noise of the force process
         'forces': [[0, 0], [1, 2], [-1, 2]],  # [angle, magnitude] for each force
-        'num_trials': 300,  # Trials per miniblock
-        'context_seq': [0, 1, 2, 'clamp', 0]  # Sequence of context type to use.
+        'num_trials': 50,  # Trials per miniblock
+        'context_seq': [0, 1, 2, 'clamp', 0, 1, 2, 0],  # Sequence of context type to use.
+        'cues': [0, 1, 2, 0, 0, 2, 1, 0]
         }
+if len(pars['context_seq']) != len(pars['cues']):
+    raise ValueError('Number of cues and number of miniblocks do not match.')

### 2021-03-23 code/adaptation/simulations.py

@@ -118,4 +118,80 @@ def read_data():
         panda['agent'] = agent
         pandas.append(panda)
     pandatron = pd.concat(pandas, axis=0)
+    pandatron.reset_index(inplace=True)
+    pandatron['index'] = np.arange(len(pandatron))
+    pandatron.set_index('index', inplace=True)
     return pandatron
+
+
+def performance(pandata):
+    """Calculates the performance of each agent and each parameter combination in
+    --pandata--. Performance is calculated as the sum of the absolute value of hand
+    position.
+
+    Returns
+    performance : pd.DataFrame
+
+    """
+    pandata = pandata.copy()
+    pandata['abs_dev'] = pandata['hand'].apply(abs)
+    return pandata.groupby(['agent', 'obs_noise', 'cue_noise']).sum()['abs_dev']
+
+
+def context_inference(pandata):
+    """Calculates how well each agent in --pandata-- inferred the context. For every
+    trial that the agent mis-identified the context, a 1 is added to the score. Lower
+    scores mean better inference.
+
+    """
+    pandata = pandata.copy()
+    ix_delete = np.array(pandata['ix_context'] == 'clamp')
+    pandata.drop(np.nonzero(ix_delete)[0], inplace=True)
+    pandata.dropna(axis=0, how='any', inplace=True)
+    pandata['context_error'] = pandata['ix_context'].astype(float) != pandata['con_t']
+    return pandata.groupby(['agent', 'obs_noise', 'cue_noise']).sum()['context_error']
+
+
+def interactive_plot(pandata, axis=None, fignum=3):
+    """Generator to be able to navigate through the different agents in --pandata--.
+
+    Send an signed int N to go back or forth N agents. Send a list with parameter
+    values [obs_noise, cue_noise, agent] to go to that agent.
+
+    """
+    if axis is None:
+        fig, axis = plt.subplots(1, 1, clear=True, num=fignum)
+    plt.show(block=False)
+    pandata = pandata.set_index(['obs_noise', 'cue_noise', 'agent'])
+    indices = list(pandata.index.unique())
+    num_indices = len(indices)
+    ix_ix = 0
+    while 1:
+        pandatum = pandata.loc[indices[ix_ix]]
+        num_trials = len(pandatum)
+        real_con = np.zeros((num_trials, 4))
[…]

### 2021-03-23 code/adaptation/task_hold_hand.py

@@ -1,9 +1,7 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/task_hold_hand.py
-import ipdb
+# import ipdb
 from datetime import datetime
-import argparse
-from importlib import import_module
 
 import numpy as np
 from scipy import stats
@@ -17,26 +15,16 @@ The game is meant to be played by model.LeftRightAgent or its children.
 
 """
 
-# Import task parameters from a file, if provided:
-PARSER = argparse.ArgumentParser(description='Run experiment')
-PARSER.add_argument('-p', '--pars', dest='pars', default=None,
-                    help='Filename for the parameters')
-ARGS = PARSER.parse_args()
-parsfile = ARGS.pars
-if parsfile is None:
-    parsfile = 'pars'
-if parsfile.endswith('.py'):
-    parsfile = parsfile[:-3]
-pars = import_module(parsfile).pars
 
-
-def run(agent=None, save=False, filename=None, **trial_kwargs):
+def run(agent=None, save=False, filename=None, pars=None):
     """Runs an entire game of the holding-hand task."""
+    if pars is None:
+        from pars import pars
     if agent is None:
         agent = model.LeftRightAgent(obs_sd=pars['obs_noise'][1])
     outs = []
-    for ix_miniblock in pars['context_seq']:
-        out = miniblock(ix_miniblock, 0, agent, **trial_kwargs)
+    for ix_miniblock, _ in enumerate(pars['context_seq']):
+        out = miniblock(ix_miniblock, 0, agent, pars)
         outs.append(out)
     data = np.stack(outs, axis=0)
     pandata = pd.DataFrame(data.reshape((-1, 5)),
@@ -49,18 +37,18 @@ def run(agent=None, save=False, filename=None, **trial_kwargs):
     return pandata, pandagent, agent
 
 
-def miniblock(ix_context, hand_position, agent, **trial_kwargs):
+def miniblock(ix_context, hand_position, agent, pars):
     """Runs a miniblock of the task."""
     outs = []
     for ix_trial in range(pars['num_trials']):
-        out = trial(ix_context, hand_position, agent, **trial_kwargs)
+        out = trial(ix_context, hand_position, agent, pars)
         outs.append(out)
         hand_position = out[3]
     outs = np.stack(outs, axis=0)
     return outs
 
 
-def trial(ix_context, hand_position, agent, cue_key=None):
+def trial(ix_miniblock, hand_position, agent, pars):
     """Runs a single trial of the task.
 
     Parameters
[…]

### 2021-03-24 notes/notes.tex
Changed pars a bit for big sim in office

@@
 However, the agent has a belief on this value, which is fixed for the agent.
-This value (obs_sd, which can be passed on instanciation) can be used in lieu of an explicit $\Delta t$ parameter which controls the size of the time step
+This value (obs\_sd, which can be passed on instanciation) can be used in lieu of an explicit $\Delta t$ parameter which controls the size of the time step
+
+\subsection{Parameter values to look into}
+By visual inspection of the results of simulations, these are the parameter values that seem promising:
+
+\begin{tabularx}{\textwidth}{
+  >{\hsize=.2\hsize\linewidth=\hsize}X|
+  >{\hsize=.1\hsize\linewidth=\hsize}X|
+  >{\hsize=.1\hsize\linewidth=\hsize}X|
+  l}
+% {l|{\hsize=.4\hsize}X|{\hsize=0.3\hsize}X|{\hsize=0.3\hsize}X}
+Agent & Obs noise & Cue noise & Notes \\ \hline \hline
+LR    & 1.2       & 0.04      & Slow fake-context inference. Maybe the priors coming from the previous stage are messing things up. \\ \hline
+1 & 1 & 1 & 1
+\end{tabularx}
 

### 2021-03-24 code/adaptation/model.py

@@ -136,7 +136,7 @@ class LeftRightAgent(object):
     context_sd_exponent = 3
 
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
-                 angles=None):
+                 angles=None, context_noise=None):
         """Initializes the known left and right contexts, as well as
         baseline.
 
@@ -149,6 +149,8 @@ class LeftRightAgent(object):
             self.cue_noise = cue_noise
         if angles:
             self.angles = np.array(angles)
+        if context_noise:
+            self.context_noise = context_noise
 
         self.num_contexts = 3
         _, self.magnitudes = self.__init_contexts()
@@ -501,21 +503,22 @@ class LRMeanSD(LeftRightAgent):
         self.mag_hypers_history = [self.mag_hypers]
 
     def update_magnitudes(self):
-        """Updates all four hyperparameters of the force magnitudes, using
+        r"""Updates all four hyperparameters of the force magnitudes, using
         NormalGamma priors and posteriors.
 
-        Parameters
-        ----------
-        likelihood_pars : iterable size=(2,)
-        Mean and standard deviation of the current observation, used as the
-        likelihood to update the posteriors over magnitudes.
+        Note: for simplicity, a single value for the mean and the standard
+        deviation is saved to self.magnitudes (and the rest of the code will
+        treat these values as the mean and standard deviation of a
+        Gaussian). The mean is simply the first hyperparameter and the standard
+        deviation is $\beta / (\nu \alpha)$, where the hyperparameters are
+        $\mu, \nu, \alpha, \beta$.
 
         """
         ix_context = self.sample_context()
         mu_l, sd_l = self.predict_outcome()[1][ix_context]
         mag_pars = self.mag_hypers[ix_context]
         mu_l = mag_pars[0] + self.hand_position - mu_l
-        post_mag = [(mag_pars[0] + mag_pars[1] * mu_l) / (mag_pars[1] + 1),
+        post_mag = [(mag_pars[0] * mag_pars[1] + mu_l) / (mag_pars[1] + 1),
                     mag_pars[1] + 1,
                     mag_pars[2] + 0.5,
                     mag_pars[3] + mag_pars[1] *
@@ -526,7 +529,7 @@ class LRMeanSD(LeftRightAgent):
         self.mag_hypers_history.append(mag_hypers)
         magnitudes = np.array(self.magnitudes)
[…]

### 2021-03-24 code/adaptation/pars.py

@@ -1,6 +1,6 @@
 pars = {'obs_noise': [0, 0.1],  # Observation noise N(mu, sd)
         'force_noise': [[0, 0.0001], [0, 0.2], [0, 0.2]],  # Noise of the force process
-        'forces': [[0, 0], [1, 2], [-1, 2]],  # [angle, magnitude] for each force
+        'forces': [[0, 0], [1, 20], [-1, 20]],  # [angle, magnitude] for each force
         'num_trials': 50,  # Trials per miniblock
         'context_seq': [0, 1, 2, 'clamp', 0, 1, 2, 0],  # Sequence of context type to use.
         'cues': [0, 1, 2, 0, 0, 2, 1, 0]

### 2021-03-24 code/adaptation/simulations.py

@@ -2,7 +2,6 @@
 # ./adaptation/simulations.py
 
 """Some simulations for the agent and the task."""
-
 import multiprocessing as mp
 from itertools import product
 import glob
@@ -13,7 +12,6 @@ import pandas as pd
 
 import model
 import task_hold_hand as thh
-
 
 def test_obsnoise(fignum=1):
     """Runs the game with increasingly high observation noise, for all three
@@ -57,7 +55,6 @@ def test_deadaptation(fignum=2):
               model.LRMeanSD(**kwargs)]
     for ix_agent, agent in enumerate(agents):
         thh.run(agent)
-        thh.run(agent, cue_key=[0, 2, 1])
         agent.plot_mu(axis=axes[ix_agent])
         agent.plot_contexts(axis=axes[ix_agent])
         agent.plot_position(axis=axes[ix_agent])
@@ -70,19 +67,16 @@ def test_deadaptation(fignum=2):
 def grid_sims():
     """Makes the agent(s) perform the task for all the parameter values.
     """
+
     obs_noises = np.arange(0, 2, 0.1)
     cue_noises = np.arange(0, 0.05, 0.005)
+    con_noises = np.arange(0, 0.3, 0.1)
 
     agents = [model.LeftRightAgent,
               model.LRMean,
               model.LRMeanSD]
-    all_pars = product(obs_noises, cue_noises, agents)
-    num_pars = len(obs_noises) * len(cue_noises) * len(agents)
+    all_pars = product(obs_noises, cue_noises, con_noises, agents)
     my_robots = mp.Pool()
-    num_seeds = num_pars
-
-    seeds = np.random.randint(low=0, high=10000, size=num_seeds * 10)
-    seeds = np.unique(seeds)
 
     for ix_pars, pars in enumerate(all_pars):
         my_robots.apply_async(_one_sim, list(pars))
@@ -91,11 +85,15 @@ def grid_sims():
     my_robots.join()
 
 
-def _one_sim(obs_noise, cue_noise, agent_fun):
+def _one_sim(obs_noise, cue_noise, con_noise, agent_fun):
     """Runs one iteration for grid_sims(). """
-    agent = agent_fun(obs_sd=obs_noise, cue_noise=cue_noise)
-    filename = 'grid_sims_{}_{}_{}.pi'.format(obs_noise, cue_noise, agent.name)
-    thh.run(agent=agent, save=True, filename=filename)
+    from pars import pars as task_pars
+    task_pars['obs_noise'][1] = obs_noise
[…]

### 2021-03-24 code/adaptation/task_hold_hand.py

@@ -25,9 +25,9 @@ def run(agent=None, save=False, filename=None, pars=None):
     outs = []
     for ix_miniblock, _ in enumerate(pars['context_seq']):
         out = miniblock(ix_miniblock, 0, agent, pars)
-        outs.append(out)
-    data = np.stack(outs, axis=0)
-    pandata = pd.DataFrame(data.reshape((-1, 5)),
+        for one_out in out:
+            outs.append(one_out)
+    pandata = pd.DataFrame(outs,
                            columns=('action', 'force', 'pos(t)', 'pos(t+1)',
                                     'ix_context'))
     pandata.rename_axis('trial', inplace=True)
@@ -44,7 +44,7 @@ def miniblock(ix_context, hand_position, agent, pars):
         out = trial(ix_context, hand_position, agent, pars)
         outs.append(out)
         hand_position = out[3]
-    outs = np.stack(outs, axis=0)
+    # outs = np.stack(outs, axis=0)
     return outs
 
 
@@ -126,6 +126,8 @@ def join_pandas(pandata, pandagent):
     into one big panda, aligned by trial number. Returns the panda.
 
     """
+    pandagent = pandagent.drop(['action', 'hand'], axis=1)
+    pandagent = pandagent.iloc[pandagent.index != -1]
     return pd.concat([pandata, pandagent], axis=1)
 
 

### 2021-03-24 code/adaptation/model.py
Increased max force for big sim in office

@@ -129,7 +129,7 @@ class LeftRightAgent(object):
     sample_context_mode = 'mode'
     sample_force_mode = 'mode'
 
-    max_force = 3.5  # Maximum force (abs value) the agent can exert.
+    max_force = 30  # Maximum force (abs value) the agent can exert.
 
     context_noise = 0.0001
     context_sd_constant = 0.1

### 2021-03-24 code/adaptation/simulations.py

@@ -205,4 +205,4 @@ def interactive_plot(pandata, axis=None, fignum=3):
 
 
 if __name__ == '__main__':
-    grid_sims()
+    print('yo')

### 2021-03-29 notes/notes.tex
Take home

@@
 \DeclareMathOperator*{\argmax}{argmax}
+
+% Model and task parameter macros
+\input{parameters}
 
@@
 
+\section{Parametrization of the task and the model}
+Some of the parameters of the model can be fixed by the physics and scales of
+the simulated experiment.
+
+Henceforth, distances are measured in meters and time in seconds. As a
+consequence, forces should be measured in Newtons, but bare in mind that what
+the model calls ``force'' is no such thing. Instead, ``forces'' in the model
+are the distance (in meters) that the real force (in Newtons) would make the
+arm move in some $\Delta t$, assuming a mass of 1Kg and zero initial
+velocity. This simplification of the terms might come back to bite us in the
+ass and might be replaced later. But first, let us speak of real forces.
+
+The force exerted by the mechanical arm depends on the context. In the baseline
+context, the force is 0N. Taking inspiration from \todo{Citation needed}, we
+take the forces of the other two contexts to be \force and -\force,
+respectively. We will assume that $\Delta t = \dt$ \todo{Need to find a good
+  way of fixing this}, which represents reaction, processing and decision
+times.
+
+Back to the lazy encoding of forces in the model, one can calculate the effects
+of these forces on the arm. With a force of 10N and a $\Delta t$ of 50ms, the
+total distance the arm will be moved by the force (again, assuming mass of 1Kg)
+is given by:
+\begin{align}
+  d_\plusminus &= \plusminus \frac{1}{2}(\force / 1Kg)((\dt)^2)\\
+               &= \plusminus 0.0125m
+\end{align}
+The agent will try to estimate $d_\plusminus$ during the experiment.
+
+We will assume \todo{This is arbitrary. Might want to justify it later.} that
+the participant observes the position of her own hand (visual + proprioceptive)
+with Gaussian observation noise with $\mu = 0m$ and $\sigma = 0.005m$.
+
 \section{Simulation results}
 \subsection{Effects of observation noise on inference}
-Observation noise refers to the noise on the position of the hand.
-The generating process is a Gaussian with $\sigma = 0.1$, which is determined in ./code/adaptation/pars.py.
-However, the agent has a belief on this value, which is fixed for the agent.
-This value (obs\_sd, which can be passed on instanciation) can be used in lieu of an explicit $\Delta t$ parameter which controls the size of the time step
+Observation noise refers to the noise on the position of the hand. The
+generating process is a Gaussian with $\sigma = 0.1$, which is determined in
+./code/adaptation/pars.py. However, the agent has a belief on this value, which
+is fixed for the agent. This value (obs\_sd, which can be passed on
+instanciation) can be used in lieu of an explicit $\Delta t$ parameter which
+controls the size of the time step
 

### 2021-03-29 code/adaptation/model.py

@@ -9,6 +9,8 @@ from scipy import stats
 import pandas as pd
 from matplotlib import pyplot as plt
 
+from pars import agent as pars
+
 """Bayesian motor adaptation model."""
 
 
@@ -119,21 +121,19 @@ class LeftRightAgent(object):
     """
     name = 'LR'
     # Free parameters with default values. Can be overwritten by __init__
-    action_sd = 0.01  # SD of a Gaussian for action uncertainty
+    action_sd = pars['action_sd']  # SD of a Gaussian for action uncertainty
     cue_noise = 0.1  # If ix_cue is observed, the posterior over the
                      # corresponding context is 1 - 2 * cue_noise.
-    obs_sd = 1
+    obs_sd = pars['obs_noise']
     angles = np.array([0, 1, -1])
-    force_sds = np.array([0.01, 0.2, 0.2])
+    force_sds = pars['force_sd']  # np.array([0.01, 0.2, 0.2])
 
     sample_context_mode = 'mode'
     sample_force_mode = 'mode'
 
-    max_force = 30  # Maximum force (abs value) the agent can exert.
+    max_force = pars['max_force']  # Maximum force the agent can exert
 
     context_noise = 0.0001
-    context_sd_constant = 0.1
-    context_sd_exponent = 3
 
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None):
@@ -266,15 +266,6 @@ class LeftRightAgent(object):
         def predicted_hand(pos, context):
             return funs[context](pos)
         return predicted_hand, posterior_pars
-
-    def _sd_context(self, prob):
-        """Function that determines the effect of the probability of the context
-        on the standard deviation of the outcome prediction.
-
-        """
-        if prob == 0:
-            return np.inf
-        return self.context_sd_constant / prob ** self.context_sd_exponent
 
     def infer_context(self, hand_position, cue=None):
         """Infers the current context given the current observation, as well
@@ -445,6 +436,7 @@ class LRMean(LeftRightAgent):
 
     """
     name = 'LRM'
+
     def __init__(self, *args, **kwargs):
         super().__init__(*args, **kwargs)
         self.mag_hypers = self.magnitudes

### 2021-03-29 code/adaptation/pars.py

@@ -1,9 +1,32 @@
-pars = {'obs_noise': [0, 0.1],  # Observation noise N(mu, sd)
-        'force_noise': [[0, 0.0001], [0, 0.2], [0, 0.2]],  # Noise of the force process
-        'forces': [[0, 0], [1, 20], [-1, 20]],  # [angle, magnitude] for each force
+# -*- coding: utf-8 -*-
+
+import numpy as np
+
+"""Parameters for the model and the task. See /notes/notes.pdf for an
+explanation of each one of them and how these values were found.
+
+All values are in the international system of units (m, g, s, N, ...).
+"""
+# Parameters common to agent and task:
+delta_t = 0.05
+obs_noise = 0.0001
+forces = np.array([0, 10, 10])
+fake_mags = 0.5 * forces * delta_t ** 2  # Really, see the notes.pdf
+force_sd = 0.0001
+
+# Values for the parameters of the task in task_hold_hand.py
+task = {'obs_noise': obs_noise,  # Observation noise N(mu, sd)
+        'force_noise': force_sd * np.ones(3),  # Noise of the force process
+        'forces': [[0, 0], [1, fake_mags[1]], [-1, fake_mags[2]]],  # [angle, magnitude]
         'num_trials': 50,  # Trials per miniblock
         'context_seq': [0, 1, 2, 'clamp', 0, 1, 2, 0],  # Sequence of context type to use.
         'cues': [0, 1, 2, 0, 0, 2, 1, 0]
         }
-if len(pars['context_seq']) != len(pars['cues']):
-    raise ValueError('Number of cues and number of miniblocks do not match.')
+
+# Default values for the parameters of the agents in models.py
+agent = {'obs_noise': obs_noise,
+         'delta_t': delta_t,
+         'max_force': 2 * max(fake_mags),
+         'action_sd': max(fake_mags) / 5,
+         'force_sd': force_sd * np.ones(3)
+         }

### 2021-03-29 code/adaptation/simulations.py

@@ -12,6 +12,8 @@ import pandas as pd
 
 import model
 import task_hold_hand as thh
+from pars import task as task_pars
+
 
 def test_obsnoise(fignum=1):
     """Runs the game with increasingly high observation noise, for all three
@@ -46,8 +48,9 @@ def test_deadaptation(fignum=2):
     the opposite meaning.
 
     """
-    kwargs = {'angles': [0, 0, 0], 'obs_sd': 1,
-              'cue_noise': 0.001}
+    kwargs = {'angles': [0, 0, 0], 'obs_sd': task_pars['obs_noise'],
+              'context_noise': 0.1,
+              'cue_noise': 0.01}
     fig, axes = plt.subplots(3, 1, num=fignum, clear=True,
                              sharex=True, sharey=True, squeeze=True)
     agents = [model.LeftRightAgent(**kwargs),
@@ -66,9 +69,13 @@ def test_deadaptation(fignum=2):
 
 def grid_sims():
     """Makes the agent(s) perform the task for all the parameter values.
+
+    All your processors are belong to us.
+
+    Outputs are saved (callback to _one_sim()) to pickled files
     """
 
-    obs_noises = np.arange(0, 2, 0.1)
+    obs_noises = task_pars['obs_noise']
     cue_noises = np.arange(0, 0.05, 0.005)
     con_noises = np.arange(0, 0.3, 0.1)
 
@@ -87,8 +94,8 @@ def grid_sims():
 
 def _one_sim(obs_noise, cue_noise, con_noise, agent_fun):
     """Runs one iteration for grid_sims(). """
-    from pars import pars as task_pars
-    task_pars['obs_noise'][1] = obs_noise
+    from pars import task as task_pars
+    task_pars['obs_noise'] = obs_noise
     agent = agent_fun(obs_sd=obs_noise, cue_noise=cue_noise,
                       context_noise=con_noise, angles=[0, 0, 0])
     filename = 'grid_sims_{}_{}_{}_{}.pi'.format(obs_noise, cue_noise,

### 2021-03-29 code/adaptation/task_hold_hand.py

@@ -19,9 +19,9 @@ The game is meant to be played by model.LeftRightAgent or its children.
 def run(agent=None, save=False, filename=None, pars=None):
     """Runs an entire game of the holding-hand task."""
     if pars is None:
-        from pars import pars
+        from pars import task as pars
     if agent is None:
-        agent = model.LeftRightAgent(obs_sd=pars['obs_noise'][1])
+        agent = model.LeftRightAgent(obs_sd=pars['obs_noise'])
     outs = []
     for ix_miniblock, _ in enumerate(pars['context_seq']):
         out = miniblock(ix_miniblock, 0, agent, pars)
@@ -104,8 +104,8 @@ def sample_observation(hand_position, pars):
                                       'n-dimensional spaces.')
     except TypeError:
         pass
-    loc, scale = pars['obs_noise']
-    loc += hand_position
+    scale = pars['obs_noise']
+    loc = hand_position
     distri = stats.norm(loc=loc, scale=scale)
     return distri.rvs()
 
@@ -115,8 +115,8 @@ def sample_force(context, pars):
     given the --hand_position-- and the context.
 
     """
-    loc, scale = pars['force_noise'][context]
-    distri = stats.norm(loc=loc, scale=scale)
+    scale = pars['force_noise'][context]
+    distri = stats.norm(loc=0, scale=scale)
     magnitude = np.prod(pars['forces'][context])
     return magnitude + distri.rvs()
 

### 2021-03-29 code/adaptation/pars.py
MP doesn't work from console

@@ -28,5 +28,5 @@ agent = {'obs_noise': obs_noise,
          'delta_t': delta_t,
          'max_force': 2 * max(fake_mags),
          'action_sd': max(fake_mags) / 5,
-         'force_sd': force_sd * np.ones(3)
+         'force_sd': force_sd * np.ones(3) * 3
          }

### 2021-03-29 code/adaptation/simulations.py

@@ -74,7 +74,6 @@ def grid_sims():
 
     Outputs are saved (callback to _one_sim()) to pickled files
     """
-
     obs_noises = task_pars['obs_noise']
     cue_noises = np.arange(0, 0.05, 0.005)
     con_noises = np.arange(0, 0.3, 0.1)
@@ -82,12 +81,11 @@ def grid_sims():
     agents = [model.LeftRightAgent,
               model.LRMean,
               model.LRMeanSD]
-    all_pars = product(obs_noises, cue_noises, con_noises, agents)
+    all_pars = product([obs_noises], cue_noises, con_noises, agents)
     my_robots = mp.Pool()
 
     for ix_pars, pars in enumerate(all_pars):
         my_robots.apply_async(_one_sim, list(pars))
-
     my_robots.close()
     my_robots.join()
 
@@ -181,24 +179,25 @@ def interactive_plot(pandata, axis=None, fignum=3):
         pandatum = pandata.loc[indices[ix_ix]]
         num_trials = len(pandatum)
         real_con = np.zeros((num_trials, 4))
-        real_con[pandatum['ix_context'] == 0, 0] = 1
-        real_con[pandatum['ix_context'] == 1, 1] = 1
-        real_con[pandatum['ix_context'] == 2, 2] = 1
-        real_con[pandatum['ix_context'] == 'clamp', 3] = 1
         axis.clear()
         axis.plot(np.array(pandatum['pos(t)']), color='black', alpha=0.4)
         axis.plot(np.array(pandatum['mag_mu_0']), color='black')
         axis.plot(np.array(pandatum['mag_mu_1']), color='red')
         axis.plot(np.array(pandatum['mag_mu_2']), color='blue')
         ylim = axis.get_ylim()
-        conx = np.array(pandatum.loc[:, ['con0', 'con1', 'con2']])
+        height = np.abs(np.max(ylim) - np.min(ylim)) * 0.2
+        real_con[pandatum['ix_context'] == 0, 0] = height
+        real_con[pandatum['ix_context'] == 1, 1] = height
+        real_con[pandatum['ix_context'] == 2, 2] = height
+        real_con[pandatum['ix_context'] == 'clamp', 3] = height
+        conx = np.array(pandatum.loc[:, ['con0', 'con1', 'con2']]) * height
         axis.plot(ylim[0] + conx[:, 0], color='black')
[…]

### 2021-03-29 code/adaptation/task_hold_hand.py

@@ -11,7 +11,7 @@ import model
 
 """Task described in section 7.1 of notes.pdf in which the participant
 must hold her hand in the starting position in different force fields.
-The game is meant to be played by model.LeftRightAgent or its children.
+The game is meant to be played by model.LeftRightAgent or its childrenlambda.
 
 """
 

### 2021-03-31 notes/notes.tex
Added new likelihood description to notes

@@
 
+\subsection{Context inference requires a new likelihood}
+\label{subsec:new-likelihood}
+It became clear by running preliminary simulations that, regardless of the
+agent, context inference happened too quickly and too ``reliably''. The context
+was identified during the first trial at first, and during deadaptation, when
+the cue and the forces do not match, it took at most a couple of trials to
+recognize the real context.
+
+Since the model proposes that deadaptation is nothing but a consequence of slow
+recognition of the context, lightning-fast context recognition will just not do
+it.
+
+The problem stems from the fact that the estimate on the magnitudes becomes so
+good (e.g. low standard deviation on the estimate) that the likelihood
+$p(x | C_t)$ becomes a delta function. Because of this, the cue itself is being
+ignored and the context is immediately identified.
+
+The best fix to this problem is to modify the way the likelihood works. As it
+stands, the likelihood is the probability of observing the current hand position
+given the current context and the previous hand position. The more ``separated''
+the different estimates become from each other (the means are too far apart
+and/or the precisions are too high), the more this is going to be a delta
+function.
+
+The best fix for this is to modify the likelihood function and introduce in it a
+measurement of the reliability of each of the modalities of perception (force
+feedback and visual cue). Such moderators are common in the literature
+\citep[e.g.][]{Lee_Neural_2014}.
+
+The simplest implementation I can think of is
+one in which additive noise is added to the posterior estimate for each
+modality:
+\begin{equation}
+p(C | \omega_t) \leftarrow (p(C | \omega_t) + \chi_{\omega})
+p(C | s_{t-1}, s_t, a_{t-1}) \leftarrow (p(C |  s_{t-1}, s_t, a_{t-1}) + \chi_{\text{pred}})
+\end{equation}
+where $\chi_{\omega}$ and $\chi_{\text{pred}}$ are constants which encode the
+reliability of each modality.
+
+A more convoluted approach would be to bring all modalities to the same
+level. In the current implementation, the likelihood uses the standard deviation
+of the inference over the magnitudes to calculate $p(C | s, a)$. Instead, a
+standard deviation could be used that is proportional to the mean, which would
+put the cue and predictive modalities on the same playing ground, depending on
+the proportionality constant chosen.
+
 \subsection{Parameter values to look into}

### 2021-03-31 code/adaptation/model.py

@@ -1,7 +1,7 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/model.py
 
-import ipdb
+# import ipdb
 
 import numpy as np
 import scipy as sp

### 2021-03-31 code/adaptation/simulations.py

@@ -83,9 +83,9 @@ def grid_sims():
               model.LRMeanSD]
     all_pars = product([obs_noises], cue_noises, con_noises, agents)
     my_robots = mp.Pool()
-
     for ix_pars, pars in enumerate(all_pars):
         my_robots.apply_async(_one_sim, list(pars))
+
     my_robots.close()
     my_robots.join()
 
@@ -210,27 +210,5 @@ def interactive_plot(pandata, axis=None, fignum=3):
         ix_ix = ix_ix % num_indices
 
 
-def dummy(x):
-    print('yo')
-
-
-def _error(x):
-    print(x)
-
-
 if __name__ == '__main__':
-    obs_noises = task_pars['obs_noise']
-    cue_noises = np.arange(0, 0.05, 0.005)
-    con_noises = np.arange(0, 0.3, 0.1)
-
-    agents = [model.LeftRightAgent,
-              model.LRMean,
-              model.LRMeanSD]
-    all_pars = product([obs_noises], cue_noises, con_noises, agents)
-    my_robots = mp.Pool()
-
-    # for ix_pars, pars in enumerate(all_pars):
-    #     my_robots.apply_async(_one_sim, list(pars), callback=dummy)
-    # my_robots.close()
-    # my_robots.join()
-    my_robots.map_async(dummy, list(all_pars)).get()
+    grid_sims()

### 2021-04-01 code/adaptation/model.py
Added reset at new mblock

@@ -1,8 +1,6 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/model.py
 
-import ipdb
-
 import numpy as np
 import scipy as sp
 from scipy import stats
@@ -127,6 +125,8 @@ class LeftRightAgent(object):
     obs_sd = pars['obs_noise']
     angles = np.array([0, 1, -1])
     force_sds = pars['force_sd']  # np.array([0.01, 0.2, 0.2])
+    prediction_noise = pars['prediction_noise']
+    reset_after_change = pars['reset_after_change']
 
     sample_context_mode = 'mode'
     sample_force_mode = 'mode'
@@ -136,7 +136,8 @@ class LeftRightAgent(object):
     context_noise = 0.0001
 
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
-                 angles=None, context_noise=None):
+                 angles=None, context_noise=None, prediction_noise=None,
+                 reset_after_change=None):
         """Initializes the known left and right contexts, as well as
         baseline.
 
@@ -151,6 +152,10 @@ class LeftRightAgent(object):
             self.angles = np.array(angles)
         if context_noise:
             self.context_noise = context_noise
+        if prediction_noise:
+            self.prediction_noise = prediction_noise
+        if reset_after_change:
+            self.reset_after_change = reset_after_change
 
         self.num_contexts = 3
         _, self.magnitudes = self.__init_contexts()
@@ -174,6 +179,11 @@ class LeftRightAgent(object):
                                self.force_sds[:, None]])
         return self.angles, magnitudes
 
+    def reset(self, ):
+        if self.reset_after_change:
+            self.log_context = np.log(np.ones(self.num_contexts) /
+                                      self.num_contexts)
+
     def make_decision(self, ):
         """Makes a decision after all the inferences for the trial have been
         carried out.
@@ -187,7 +197,6 @@ class LeftRightAgent(object):
         The decision is returned and also saved into self.decision_history.
 
         """
-        # raise NotImplementedError('obs noise!')
         c_pos = self.hand_position
         c_context = self.sample_context('mode')
         c_force = self.magnitudes[c_context][0]
@@ -276,7 +285,12 @@ class LeftRightAgent(object):
         # Hand position likelihood:
[…]

### 2021-04-01 code/adaptation/task_hold_hand.py

@@ -40,6 +40,7 @@ def run(agent=None, save=False, filename=None, pars=None):
 def miniblock(ix_context, hand_position, agent, pars):
     """Runs a miniblock of the task."""
     outs = []
+    agent.reset()
     for ix_trial in range(pars['num_trials']):
         out = trial(ix_context, hand_position, agent, pars)
         outs.append(out)

### 2021-04-06 code/adaptation/model.py
Before overhaul of task structure

@@ -128,6 +128,10 @@ class LeftRightAgent(object):
     prediction_noise = pars['prediction_noise']
     reset_after_change = pars['reset_after_change']
 
+    is_reset = True  # Whether a new miniblock started
+
+    preferred = 1
+
     sample_context_mode = 'mode'
     sample_force_mode = 'mode'
 
@@ -179,10 +183,18 @@ class LeftRightAgent(object):
                                self.force_sds[:, None]])
         return self.angles, magnitudes
 
-    def reset(self, ):
-        if self.reset_after_change:
-            self.log_context = np.log(np.ones(self.num_contexts) /
-                                      self.num_contexts)
+    def reset(self, priors=None):
+        """Resets the priors over context to a uniform distribution.
+        If --priors-- are provided, they are used instead.
+
+        """
+        self.is_reset = True
+        if priors is None:
+            self.log_context = - np.log(self.num_contexts) * np.ones(self.num_contexts)
+        else:
+            if not np.isclose(np.sum(priors), 1):
+                raise ValueError('Provided priors do not add to 1')
+            self.log_context = np.log(priors)
 
     def make_decision(self, ):
         """Makes a decision after all the inferences for the trial have been
@@ -281,15 +293,18 @@ class LeftRightAgent(object):
         as the previous action and outcome.
 
         """
-        lh_fun, lh_pars = self.predict_outcome()
-        # Hand position likelihood:
-        log_li_hand = np.array([lh_fun(hand_position, ix_context)
-                                for ix_context in range(self.num_contexts)])
-        li_hand = np.exp(log_li_hand - log_li_hand.max())
-        p_hand = li_hand / li_hand.sum()
-        p_hand += self.prediction_noise
-        p_hand /= p_hand.sum()
-        log_li_hand = np.log(p_hand)
+        if not self.is_reset:
+            lh_fun, lh_pars = self.predict_outcome()
+            # Hand position likelihood:
+            log_li_hand = np.array([lh_fun(hand_position, ix_context)
+                                    for ix_context in range(self.num_contexts)])
+            li_hand = np.exp(log_li_hand - log_li_hand.max())
+            p_hand = li_hand / li_hand.sum()
+            p_hand += self.prediction_noise
+            p_hand /= p_hand.sum()
+            log_li_hand = np.log(p_hand)
+        else:
+            log_li_hand = -np.log(self.num_contexts) * np.ones(self.num_contexts)
 
         # Cue likelihood:
[…]

### 2021-04-06 code/adaptation/pars.py

@@ -26,7 +26,16 @@ task = {'obs_noise': obs_noise,  # Observation noise N(mu, sd)
 # Default values for the parameters of the agents in models.py
 agent = {'obs_noise': obs_noise,
          'delta_t': delta_t,
-         'max_force': 2 * max(fake_mags),
-         'action_sd': max(fake_mags) / 5,
-         'force_sd': force_sd * np.ones(3) * 3
+         'max_force': 1.2 * max(fake_mags),
+         'action_sd': max(fake_mags) / 10,
+         'force_sd': 5 * force_sd * np.ones(3),
+         'prediction_noise': 0.01,
+         'reset_after_change': False,  # Whether to reset priors on new miniblock
          }
+
+# The following parameters follow the experiment in Smith_2006
+task_smith = task.copy()
+task_smith['num_trials'] = 10
+task_smith['context_seq'] = [0] * 12 + [1] * 40 + [2] * 2 + ['clamp'] * 18
+task_smith['cues'] = [0] * 12 + [1] * 40 + [1] * 2 + [1] * 18
+task_smith['reset_after_change'] = True

### 2021-04-06 code/adaptation/simulations.py

@@ -36,7 +36,8 @@ def test_obsnoise(fignum=1):
             agent.plot_mu(axis=axes[ix_agent, ix_noise])
             agent.plot_position(axis=axes[ix_agent, ix_noise])
             if ix_agent == 0:
-                axes[ix_agent, ix_noise].set_title('Obs Noise: {}'.format(obs_noise))
+                axes[ix_agent, ix_noise].set_title(
+                    'Obs Noise: {}'.format(obs_noise))
     plt.ylabel('Adaptation')
     plt.xlabel('Trial')
     plt.draw()
@@ -72,32 +73,57 @@ def grid_sims():
 
     All your processors are belong to us.
 
-    Outputs are saved (callback to _one_sim()) to pickled files
+    Outputs are saved to pickled files
     """
-    obs_noises = task_pars['obs_noise']
+    labels, grid_elements = _define_grid()
+    labels = labels + ['agent']
+    all_pars = product(*grid_elements)
+    my_robots = mp.Pool()
+
+    for pars in all_pars:
+        inputs = {label: value
+                  for label, value in zip(labels, pars)}
+        # _one_sim(inputs)
+        my_robots.apply_async(_one_sim, args=[inputs])
+    my_robots.close()
+    my_robots.join()
+
+
+def _define_grid():
+    """Defines the grid for grid_sims and read_data.  This function is meant to be
+    edited manually to do different simulations. Keep in mind that the labels
+    must match those expected by the agents' instantiation functions.
+
+    """
+    from sys import version_info
+    if version_info[0] < 3 or version_info[1] < 7:
+        raise NotImplementedError('The functions that use this function '
+                                  'rely on dictionaries being ordered and '
+                                  'I was too lazy to implement it, so '
+                                  'python 3.7+ is required.')
+    pred_noises = np.arange(0, 1, 0.1)
     cue_noises = np.arange(0, 0.05, 0.005)
     con_noises = np.arange(0, 0.3, 0.1)
 
     agents = [model.LeftRightAgent,
               model.LRMean,
               model.LRMeanSD]
-    all_pars = product([obs_noises], cue_noises, con_noises, agents)
-    my_robots = mp.Pool()
+    labels = ['prediction_noise', 'cue_noise', 'context_noise']
+    try:  # Test labels on agent
+        agents[0](**{label: 0.1 for label in labels})
[…]

### 2021-04-06 code/adaptation/task_hold_hand.py

@@ -45,7 +45,6 @@ def miniblock(ix_context, hand_position, agent, pars):
         out = trial(ix_context, hand_position, agent, pars)
         outs.append(out)
         hand_position = out[3]
-    # outs = np.stack(outs, axis=0)
     return outs
 
 
@@ -68,11 +67,6 @@ def trial(ix_miniblock, hand_position, agent, pars):
     agent : model.LeftRightAgent instance
     Agent that makes decisions. Needs one_trial(hand_position, cue) method as
     well as log_context attribute.
-
-    cue_key : iterable
-    Iterable to fake cues. If in a trial the context is 2, cue_key[2] will
-    be used. If cue_key=[0, 1, 2] (the default), this has no effect. If
-    something else, the agent will be lied to. LIES!
 
     """
     context = pars['context_seq'][ix_miniblock]

### 2021-04-06 code/adaptation/model.py
Finished removing miniblocks from task

@@ -116,6 +116,31 @@ class LeftRightAgent(object):
     pointing right. The magnitude of these forces are inferred during the
     experiment.
 
+    The class takes many values from the pars.py module, which must be
+    importable and must contain a dictionary called --agent--. For the default
+    values, see pars.py.
+
+    The necessary entries of the dictionary are the following:
+    'action_sd' : float
+    Std for the Gaussian from which actions should be sampled.
+
+    'obs_noise' : float
+    Assumed Std for the noise in the observations.
+
+    'force_sd' : float
+    Assumed Std for the force applied by the environment
+
+    'prediction_noise' : float
+    Inverse precision given to the predictions of the current state given the
+    last state and the last action. See self.update_magnitudes for more.
+
+    'reset_after_change' : bool
+    Whether to reset the priors over context when self.reset() is called by
+    external actors (i.e. the task).
+
+    'max_force': float
+    Maximum force (in units of distance (Newtons times delta t)). See
+    the model implementation in the notes.pdf file for an explanation of this.
     """
     name = 'LR'
     # Free parameters with default values. Can be overwritten by __init__

### 2021-04-06 code/adaptation/pars.py

@@ -7,6 +7,61 @@ explanation of each one of them and how these values were found.
 
 All values are in the international system of units (m, g, s, N, ...).
 """
+CLAMP_INDEX=150
+
+def _populate(nonzeros, vector=None, size=None, element=1):
+    """Creates a vector of zeros of size --size-- and puts an --element-- in each
+    of the indices in --nonzeros--.
+
+    If --vector-- is provided, it is used instead of creating zeros.
+
+    Beware of automatic casting!
+
+    """
+    if vector is None:
+        vector = np.zeros(size, dtype=int)
+    for index in nonzeros:
+        vector[int(index)] = element
+    return vector
+
+
+def _define_contexts_base():
+    """Defines the context of each trial. This is meant to be edited by
+    hand. """
+    # baseline = np.arange(50)
+    first_adaptation = np.arange(50, 200)
+    second_adaptation = np.arange(200, 350)
+    deadaptation = np.arange(350, 400)
+    clamp = np.arange(400, 500)
+    vector = _populate(size=500, nonzeros=first_adaptation, element=1)
+    _populate(vector=vector, nonzeros=second_adaptation, element=2)
+    _populate(vector=vector, nonzeros=deadaptation, element=2)
+    _populate(vector=vector, nonzeros=clamp, element=CLAMP_INDEX)
+    return vector
+
+
+def _define_cues_base():
+    """Defines the cues of each trial. This is meant to be edited by hand. """
+    # baseline = np.arange(50)
+    first_adaptation = np.arange(50, 200)
+    second_adaptation = np.arange(200, 350)
+    deadaptation = np.arange(350, 400)
+    clamp = np.arange(400, 500)
+    vector = _populate(size=500, nonzeros=first_adaptation, element=1)
+    _populate(vector=vector, nonzeros=second_adaptation, element=2)
+    _populate(vector=vector, nonzeros=deadaptation, element=1)
+    _populate(vector=vector, nonzeros=clamp, element=1)
+    return vector
+
+
+def _define_breaks():
+    """Defines the agent breaks for the entire session. This is meant to be edited
+    by hand."""
+    breaks = [50, 200, 350, 400]
+    vector = _populate(size=500, nonzeros=breaks, element=1)
+    return vector
+
+
 # Parameters common to agent and task:
 delta_t = 0.05
 obs_noise = 0.0001
[…]

### 2021-04-06 code/adaptation/task_hold_hand.py

@@ -11,7 +11,7 @@ import model
 
 """Task described in section 7.1 of notes.pdf in which the participant
 must hold her hand in the starting position in different force fields.
-The game is meant to be played by model.LeftRightAgent or its childrenlambda.
+The game is meant to be played by model.LeftRightAgent or its children.
 
 """
 
@@ -23,10 +23,14 @@ def run(agent=None, save=False, filename=None, pars=None):
     if agent is None:
         agent = model.LeftRightAgent(obs_sd=pars['obs_noise'])
     outs = []
-    for ix_miniblock, _ in enumerate(pars['context_seq']):
-        out = miniblock(ix_miniblock, 0, agent, pars)
-        for one_out in out:
-            outs.append(one_out)
+    hand_position = 0
+    for c_trial, do_break in enumerate(pars['breaks']):
+        if do_break:
+            agent.reset()
+            hand_position = 0
+        out = trial(c_trial, hand_position, agent, pars)
+        outs.append(out)
+        hand_position = out[3]
     pandata = pd.DataFrame(outs,
                            columns=('action', 'force', 'pos(t)', 'pos(t+1)',
                                     'ix_context'))
@@ -48,7 +52,7 @@ def miniblock(ix_context, hand_position, agent, pars):
     return outs
 
 
-def trial(ix_miniblock, hand_position, agent, pars):
+def trial(c_trial, hand_position, agent, pars):
     """Runs a single trial of the task.
 
     Parameters
@@ -69,12 +73,12 @@ def trial(ix_miniblock, hand_position, agent, pars):
     well as log_context attribute.
 
     """
-    context = pars['context_seq'][ix_miniblock]
-    cue = pars['cues'][ix_miniblock]
+    context = pars['context_seq'][c_trial]
+    cue = pars['cues'][c_trial]
     c_hand_position = hand_position
     c_obs = sample_observation(hand_position, pars)
     action = agent.one_trial(c_obs, cue=cue)
-    if context == 'clamp':
+    if context == pars['clamp_index']:
         n_hand_position = 0
         force = -action
     else:

### 2021-04-07 notes/notes.tex
Minor changes to pars. Added some notes on simulations

--- 2021-03-29 notes/notes.tex
@@
 
-\subsection{Fixed force value}
+\subsection{Left-right with fixed force value (LR)}
 In this version, the force field is assumed to have a fixed value which the
@@
 
-\subsection{Fixed-$\sigma$ model}
+\subsection{Left-right with unknown standard deviation (LRS)}
 In this model, the force at each trial is assumed to be sampled from a normal
@@
 
-\subsection{Full uncertainty}
+\subsection{Left-right with unknown mean and standard deviation (LRMS)}
 In this version, we will assume that for each context, the agent's belief over
@@
 \section{Simulation results}
-\subsection{Effects of observation noise on inference}
-Observation noise refers to the noise on the position of the hand. The
-generating process is a Gaussian with $\sigma = 0.1$, which is determined in
-./code/adaptation/pars.py. However, the agent has a belief on this value, which
-is fixed for the agent. This value (obs\_sd, which can be passed on
-instanciation) can be used in lieu of an explicit $\Delta t$ parameter which
-controls the size of the time step
+\subsection{Deadaptation and context inference}
+During the deadaptation phase, not all parametrized agents detect the correct context.
+For each agent type (LR, LRS, LRMS), different parameter values allow the information from the predicted hand positions to overcome the information from the cue, which enables the correct identification of the current context.
 
-\subsection{Parameter values to look into}
-By visual inspection of the results of simulations, these are the parameter values that seem promising:
+It is unclear which parameter values lead to this, but, as expected, prediction noise (see XXX) and context noise (see XXX) play an important role.
+In particular, keeping the prediction noise below 0.2 seems to be necessary for the correct identification of the context, for all the used values of context noise.
+It could be good to check this.
 
-\begin{tabularx}{\textwidth}{
-  >{\hsize=.2\hsize\linewidth=\hsize}X|
-  >{\hsize=.1\hsize\linewidth=\hsize}X|
-  >{\hsize=.1\hsize\linewidth=\hsize}X|
-  l}
-% {l|{\hsize=.4\hsize}X|{\hsize=0.3\hsize}X|{\hsize=0.3\hsize}X}
-Agent & Obs noise & Cue noise & Notes \\ \hline \hline
-LR    & 1.2       & 0.04      & Slow fake-context inference. Maybe the priors coming from the previous stage are messing things up. \\ \hline
-1 & 1 & 1 & 1
-\end{tabularx}
-
+\subsection{Adaptation}
+Adaptation happens mostly very quickly for all agents.
+For the current $\Delta t = \dt$, adaptation happens within 20 to 30 time steps, although many times (depending on agent and parametrization) it happens much more quickly.
+For example, for the LR agent with prediction noise of 0.15, cue noise and context noise of 0, and assumed force variability of 0.25, within 15-20 time steps the adaptation is complete.
 \section{Conclusions}

### 2021-04-07 code/adaptation/model.py

@@ -2,7 +2,6 @@
 # ./adaptation/model.py
 
 import numpy as np
-import scipy as sp
 from scipy import stats
 import pandas as pd
 from matplotlib import pyplot as plt
@@ -10,105 +9,6 @@ from matplotlib import pyplot as plt
 from pars import agent as pars
 
 """Bayesian motor adaptation model."""
-
-
-class RandomVariable(object):
-    """Defines a random variable, with a pdf and cdf, as well as a __call__ method
-    that samples from the pdf.
-
-    The distribution parameters can be updated via the method update_parameters.
-
-    Can be initiated with its own pdf and cdf. If None provided, it defaults to
-    numpy's implementation of a Gaussian.
-
-    """
-
-    def __init__(self,):
-        """Does nothing!"""
-        pass
-
-    def __call__(self, ):
-        return self.sample()
-
-    def sample(self, ):
-        """Samples a value using the pdf."""
-        rand = np.random.rand()
-        if self.var_type == 'continuous':
-            solution = sp.optimize.minimize(fun=lambda x: self.cdf(x) - rand,
-                                            x0=0)
-            sample = solution.x
-        else:
-            sample = np.argmin(abs(self.cdf - rand))
-        return sample
-
-    def plot_pdf(self, num_points=100, support=None):
-        """Plots a smooth approximation of the pdf, with --num_points-- points and
-        an x-range given by the elements of --support--.
-        """
-        raise NotImplementedError('Not yet!')
-
-
-class GaussGaussInference(RandomVariable):
-    """Performs Bayesian inference using Gaussian priors and likelihood. """
-    def __init__(self, mu=None, sd=None):
-        """Initiates the Gaussian priors with the given parameters or defaults.
-        """
-        if mu is None:
-            mu = 0
-        if sd is None:
-            sd = 1
-
-        self.pars = np.array([mu, sd])
-        self.pars_history = [self.pars]
-
-    def pdf(self, x):
-        """Evaluates the Gaussian pdf on the current parameters."""
-        return stats.norm.pdf(x, loc=self.pars[0], scale=self.pars[1])
-
-    def cdf(self, x):
-        """Evaluates the Gaussian cdf on the current parameters."""
-        return stats.norm.cdf(x, loc=self.pars[0], scale=self.pars[1])
-
-    def inference(self, likelihood_pars):
[…]

### 2021-04-07 code/adaptation/pars.py

@@ -7,7 +7,9 @@ explanation of each one of them and how these values were found.
 
 All values are in the international system of units (m, g, s, N, ...).
 """
-CLAMP_INDEX=150
+
+CLAMP_INDEX = 150
+
 
 def _populate(nonzeros, vector=None, size=None, element=1):
     """Creates a vector of zeros of size --size-- and puts an --element-- in each
@@ -85,7 +87,7 @@ agent = {'obs_noise': obs_noise,
          'delta_t': delta_t,
          'max_force': 1.2 * max(fake_mags),
          'action_sd': max(fake_mags) / 10,
-         'force_sd': 5 * force_sd * np.ones(3),
+         'force_sd': 2 * force_sd * np.ones(3),
          'prediction_noise': 0.01,
          'reset_after_change': True,  # Whether to reset priors on new miniblock
          }

### 2021-04-07 code/adaptation/simulations.py

@@ -101,29 +101,40 @@ def _define_grid():
                                   'rely on dictionaries being ordered and '
                                   'I was too lazy to implement it, so '
                                   'python 3.7+ is required.')
-    pred_noises = np.arange(0, 1, 0.1)
+    pred_noises = np.arange(0, 0.2, 0.03)
     cue_noises = np.arange(0, 0.05, 0.005)
     con_noises = np.arange(0, 0.3, 0.1)
+    force_sds = np.tile(np.arange(0.0001, 0.002, 0.0002), (3, 1)).T
 
     agents = [model.LeftRightAgent,
               model.LRMean,
               model.LRMeanSD]
-    labels = ['prediction_noise', 'cue_noise', 'context_noise']
+    values = [pred_noises, cue_noises, con_noises, force_sds]
+    labels = ['prediction_noise', 'cue_noise', 'context_noise', 'force_sds']
     try:  # Test labels on agent
-        agents[0](**{label: 0.1 for label in labels})
+        test_values = [value[0] for value in values]
+        agents[0](**{label: value for label, value in zip(labels, test_values)})
     except TypeError:
         raise ValueError('Testing the labels on the agent failed. '
                          'Maybe the labels are misspelled?')
-    return labels, [pred_noises, cue_noises, con_noises, agents]
+    values.append(agents)
+    return labels, values
 
 
 def _one_sim(kwargs):
     """Runs one iteration for grid_sims(). """
-    from pars import task_smith as task_pars
+    from pars import task as task_pars
     agent = kwargs.pop('agent')
     agent = agent(**kwargs, angles=[0, 0, 0])
     file_stem = 'grid_sims' + '_{}' * (len(kwargs) + 1) + '.pi'
-    filename = file_stem.format(*list(kwargs.values()), agent.name)
+    values = []
+    for value in kwargs.values():
+        if np.size(value) > 1:
+            c_value = value[0]
+        else:
+            c_value = value
+        values.append(c_value)
+    filename = file_stem.format(*values, agent.name)
     thh.run(agent=agent, save=True, filename=filename, pars=task_pars)
 
 
@@ -141,7 +152,12 @@ def read_data():
         panda = pd.read_pickle(file)
         len_panda = len(panda)
         for ix_part, part in enumerate(split[:-1]):
-            panda[labels[ix_part]] = float(part) * np.ones(len_panda)
+            split_split = part.split(' ')
[…]

### 2021-04-07 notes/notes.tex

--- 2021-03-31 notes/notes.tex
@@
 \section{Simulation results}
-\subsection{Effects of observation noise on inference}
-Observation noise refers to the noise on the position of the hand. The
-generating process is a Gaussian with $\sigma = 0.1$, which is determined in
-./code/adaptation/pars.py. However, the agent has a belief on this value, which
-is fixed for the agent. This value (obs\_sd, which can be passed on
-instanciation) can be used in lieu of an explicit $\Delta t$ parameter which
-controls the size of the time step
+\subsection{Deadaptation and context inference}
+During the deadaptation phase, not all parametrized agents detect the correct
+context. For each agent type (LR, LRS, LRMS), different parameter values allow
+the information from the predicted hand positions to overcome the information
+from the cue, which enables the correct identification of the current context.
+
+It is unclear which parameter values lead to this, but, as expected, prediction
+noise (see XXX) and context noise (see XXX) play an important role. In
+particular, keeping the prediction noise below 0.2 seems to be necessary for
+the correct identification of the context, for all the used values of context
+noise. It could be good to check this.
 
@@
 \end{tabularx}
+>>>>>>> adc91ff996e45d6eaf8974cd33ea69b5d380ed07
 
+\subsection{Adaptation}
+Adaptation happens mostly very quickly for all agents.
+For the current $\Delta t = \dt$, adaptation happens within 20 to 30 time steps, although many times (depending on agent and parametrization) it happens much more quickly.
+For example, for the LR agent with prediction noise of 0.15, cue noise and context noise of 0, and assumed force variability of 0.25, within 15-20 time steps the adaptation is complete.
 \section{Conclusions}

### 2021-04-08 notes/notes.tex
End of day. Adaptation visualization complete.

@@
 
-\subsection{Left-right with unknown standard deviation (LRS)}
+\subsection{Left-right with unknown mean (LRS)}
 In this model, the force at each trial is assumed to be sampled from a normal
@@
 \begin{align}
-  d_\plusminus &= \plusminus \frac{1}{2}(\force / 1Kg)((\dt)^2)\\
-               &= \plusminus 0.0125m
+  d_\pm &= \pm \frac{1}{2}(\force / 1Kg)((\dt)^2)\\
+               &= \pm 0.0125m
 \end{align}
-The agent will try to estimate $d_\plusminus$ during the experiment.
+The agent will try to estimate $d_\pm$ during the experiment.
 

### 2021-04-08 code/adaptation/model.py

@@ -66,6 +66,8 @@ class LeftRightAgent(object):
 
     context_noise = 0.0001
 
+    sample_norm = stats.norm().rvs
+
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None, prediction_noise=None,
                  reset_after_change=None, force_sds=None):
@@ -99,8 +101,7 @@ class LeftRightAgent(object):
         self.action_history = [0]
         self.log_context = np.log([0.8, 0.1, 0.1])
         self.log_context_history = [self.log_context]
-
-        self.sample_norm = stats.norm().rvs
+        self.cue_history = [-1]
 
     def __init_contexts(self, ):
         r"""Creates baseline, left and right contexts, with a known angle (no force,
@@ -142,12 +143,49 @@ class LeftRightAgent(object):
         c_context = self.sample_context('mode')
         c_force = self.magnitudes[c_context][0]
         action_mu = -c_pos - c_force
-        action = self.sample_norm() * self.action_sd + action_mu
+        # action = self.sample_norm() * self.action_sd + action_mu
+        action = self._make_decision_mixed()
         if abs(action) > self.max_force:
             action = action / abs(action) * self.max_force
         self.action = action
         self.action_history.append(action)
         return action
+
+    def _make_decision_mixed(self, sampled=True):
+        r"""Creates a distribution which is the weighted sum of a Gaussian distribution
+        for each context:
+
+        \begin{equation}
+        p(a_t) = \sum_{i=1}^{N_c}p(c_t=c_i) N(\hat{a}_{c_i}, \sigma)
+        \end{equation}
+
+        where $\hat{a}_{c_i}$ is the best action given the context $c_i$, and
+        $\sigma$ is self.action_sd.
+
+        If --sampled-- is True, a decision is sampled from this
+        distribution. Otherwise, the decision is the mean of the distribution.
+
+        """
+        bests = np.array([self._best(ix_context)
+                          for ix_context in range(self.num_contexts)])
+        probs = np.exp(self.log_context - self.log_context.max())
+        probs /= probs.sum()
+        if not sampled:
+            return (bests * probs).sum() / probs.sum()
+        sampled_context = np.random.choice(np.arange(len(probs)), p=probs)
+        sampled_decision = bests[sampled_context] + \
+            self.action_sd * self.sample_norm()
+        return sampled_decision
+
+    def _best(self, context):
+        """Returns the best possible decision given the current context in
+        --context--. The best decision is that that counteracts the estimated
[…]

### 2021-04-08 code/adaptation/simulations.py

@@ -202,7 +202,7 @@ def context_inference(pandata):
     return pandata.groupby(labels).sum()['context_error']
 
 
-def interactive_plot(pandata, axis=None, fignum=3):
+def interactive_plot(pandata, axes=None, fignum=3):
     """Generator to be able to navigate through the different agents in
     --pandata--.
 
@@ -211,40 +211,35 @@ def interactive_plot(pandata, axis=None, fignum=3):
 
     """
     labels, _ = _define_grid()
-    if axis is None:
-        fig, axis = plt.subplots(1, 1, clear=True, num=fignum)
-    plt.show(block=False)
-    pandata = pandata.set_index(labels + ['agent'])
+    if axes is None:
+        fig, axes = plt.subplots(2, 1, clear=True, num=fignum)
+    else:
+        if len(axes) != 2:
+            error_text = 'Number of axes provided is {}, should be {}.'
+            raise ValueError(error_text.format(len(axes), 2))
+        plt.show(block=False)
+    flag_noindex = False
+    try:
+        pandata = pandata.set_index(labels + ['agent'])
+    except KeyError:
+        print('Warning: the given panda does not have at least some '
+              'of the columns in _define_grid(); using the indices '
+              'present in the panda instead.')
+        flag_noindex = True
     indices = list(pandata.index.unique())
     num_indices = len(indices)
     ix_ix = 0
     while 1:
         pandatum = pandata.loc[indices[ix_ix]]
-        num_trials = len(pandatum)
-        real_con = np.zeros((num_trials, 4))
-        axis.clear()
-        axis.plot(np.array(pandatum['pos(t)']), color='black', alpha=0.4)
-        axis.plot(np.array(pandatum['mag_mu_0']), color='black')
-        axis.plot(np.array(pandatum['mag_mu_1']), color='red')
-        axis.plot(np.array(pandatum['mag_mu_2']), color='blue')
-        ylim = axis.get_ylim()
-        height = np.abs(np.max(ylim) - np.min(ylim)) * 0.2
-        real_con[pandatum['ix_context'] == 0, 0] = height
-        real_con[pandatum['ix_context'] == 1, 1] = height
-        real_con[pandatum['ix_context'] == 2, 2] = height
-        real_con[pandatum['ix_context'] == 'clamp', 3] = height
[…]

### 2021-04-21 notes/notes.tex
Added new time table

@@
 
+The previous time table was interrupted by the possibility of the new project. Now that this new project has taken shape, this new time table applies:
+
+\begin{tabularx}{\textwidth}{
+  l|
+  >{\hsize=0.4\hsize\linewidth=\hsize}X|
+  >{\hsize=0.2\hsize\linewidth=\hsize}X|
+  >{\hsize=0.2\hsize\linewidth=\hsize}X}
+End Date & Result & A priori comments & A posteriori comments \\ \hline \hline
+30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & \\ \hline
+07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
+14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
+--- & Start writing the paper. & & \\ \hline
+
+\end{tabularx}
 \end{chapter}
@@
 
+\section{Notes from }
+There is extensive evidence that fast trial-by-trial error-based learning relies on the cerebellum.
+Patients with cerebellar lesions show substantial impairment in fast adaptation across many task domains51–55
+
 \section{Motor adaptation}
-Motor adaptation to modified environments (e.g. force fields) has been shown to
-display the following characteristics:
+In typical motor adaptation experiments \citep[e.g.][]{Medina_Mechanism_2001,
+  Davidson_Scaling_2004,Kojima_Memory_2004}, the goal of the participant is to
+move some squishy appendage (e.g. eyes, hand) from point A to point B,
+repeatedly, under different circumstances. These experiments consist of a
+number of phases in different order (and some of them not present in some
+experiments). These phases are the following:
+\begin{enumerate}
+\item Baseline (O): Participants move their appendage from A to B without being
+hindered, as they normally would in their life.
+\item Adaptation (A, B, ...): In this phase, the circumstances in which the movement
+happens (still from A to B) are changed. For example, in human experiments with
+arm movement, the arm will be subjected to forces by an external actor (e.g. a
+mechanical arm attached to the participant's hand). The participant's goal is
+to learn to counteract these forces and perform the movement from A to B as
+well as in Baseline.
+\item De-adaptation: In this phase, the circumstances are changed, either
+reverted to Baseline or to some other circumstance. For example, the forces
+applied by the mechanical arm are removed, or their direction is changed to the
+opposite.
+\item Error-clamp (E): In this phase, the error in the movement is forced to
+zero. For example, the mechanical arm applies a force that forces the
+participant's hand to move in a straight line from A to B, even if the
+participant applies perpendicular forces.
+\end{enumerate}
+
+Using these phases, several phenomena has been shown:
 \begin{enumerate}
@@
 ``washed out'' if baseline trials (no force field) are inserted between
-extinction and the second learning phase.
+extinction and the second learning phase; their findings are perhaps confounded
+by the passage of time, as this inserted baseline makes a lot of time pass.
 \item Anterograde interference: Learning an adaptation (e.g. force field
@@
 \item Rapid de-adaptation: Unlearning an adaptation (even after fully learning
-it) is faster than tlearning it \citep{Davidson_Scaling_2004}
+it) is faster than learning it \citep{Davidson_Scaling_2004}
+\item Spontaneous recovery: After a de-adaptation phase, an error-clamp phase
+is introduced and participants show a rapid adaptation in the direction of A,
+which slowly goes back to zero across hundreds (?) of trials.
 \end{enumerate}
 
+The experiment carried out with monkeys by \cite{Kojima_Memory_2004} used
+different structures (see their figure 1C). For example, to show savings they
+used O-A-B-A, where in A and B the saccade target moved during the saccade to
+make the monkey think that their saccades were off. During A and B, opposite
+directions were used (e.g. in A the monkeys had to learn to increase their
+gain, and in B learn to decrease it). They found savings in the form of a quick
+learning period during the second A block that took the monkey close to the
+adaptation at the end of the first A block (see their figure 3A); after that,
+learning proceeded as before.
+
+Using an O-A-B-O-A paradigm, they showed that savings were no longer
+present. On the other hand, using an O-A-B-E-A paradigm (where the error-clamp
+phase consisted on putting the monkey in the dark) they found savings (their
+figures 5A and 7A).
+
+\cite{Smith_Interacting_2006} showed spontaneous recovery in a O-A-B-E
+paradigm. From their data it is unclear what exactly happens during B, but
+during E it seems like participants adapted a bit in the direction of B and
+then quickly jumped to the direction of A, from which they slowly went to zero.
+
+Similarly, \cite{Ethier_Spontaneous_2008} who spontaneous recovery with
+saccades in an E-A-B-E paradigm. From their results (e.g. their figure 2) it
+seems possible that some participants go back to A's adaptation and some others
+to O's. Additionally, they introduced 30s breaks during adaptation and showed
+that these breaks had a ``forgetting'' effect, i.e. adaptation was set back by
+a small amount after each break. The further posit that introducing such a
+break before the second E phase would eliminate spontaneous recovery entirely.
+
+\cite{Huang_Persistence_2009} showed that if participants are placed in an
+adaptation environment where the force gradually ramps up (across many trials),
+participants show slower de-adaptation during an error-clamp phase, compared to
+participants in which the force was suddenly increased during adaptation. They
+also showed that participants alter their learning rates to better fit the
+volatility of the environment, regardless of the direction of adaptation.
 \end{chapter}

### 2021-04-21 timetable.tex
Update on Overleaf.

--- 2021-04-08 notes/notes.tex
@@
 \usepackage{amsmath}
-\usepackage{amsfonts}
+\usepackage[margin=1.5cm]{geometry}
 \usepackage{tabularx}
-\usepackage[margin=1.5cm]{geometry}
-\usepackage{todonotes}
 \bibliographystyle{apalike}
-
-\DeclareMathOperator*{\argmin}{argmin}
-\DeclareMathOperator*{\argmax}{argmax}
-
-% Model and task parameter macros
-\input{parameters}
-
-% Defining parameters
-\def \parts {60 }                  % Number of participants
-\def \fref #1{Figure \ref{#1}}     % Reference figures
-\def \tref #1{Table \ref{#1}}      % Reference tables
-\def \cents {20 }                  % Cents earned per mini-block
-\def \eref #1{Equation \ref{#1}}   % Reference equations
-\def \sref #1{Section '\nameref{#1}'}    % Reference sections
-\def \supmat {the Supplementary Materials}
-\def \excper {5 }     % Participants excluded because they cannot sort
 
@@
 \begin{document}
+\section*{Brief project description}
+The idea of the project is the following: We will create an agent that learns to control an arm in environment A (say, the real world carrying nothing), and learns how to control the arm in environment B (say, with high winds).
+The agent then learns to recognize environments A and B and to deploy the right controller for each environment depending on this recognition.
 
-\begin{chapter}{Glossary}
-\begin{enumerate}
-\item Error-clamp: Refers to the phase of a simulation in the models of motor
-adaptation where the error that drives learning is forced to zero. For example,
-when performing a task in which a robotic attachment creates forces perpendicular
-to the direction of movement, an error-clamp phase is one in which the attachment
-pushes the arm constantly towards the desired trajectory, thus making error zero.
-\end{enumerate}
-\end{chapter}
+We will argue that this is how humans do it and present evidence (based on literature research) for the predictions this model makes, as well as try to make new predictions.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+After this methods paper is done, model validation can be carried out in future studies.
 
-\begin{chapter}{Random ideas}
-\section{Straight movements as a sign of learning}
-According to Gurevich\_93 (see \cite{Wolpert_Are_1995}) and
-\cite{Shadmehr_Adaptive_1994}, the more a person practices a movement, the
-straighter this movement becomes. Could we use the curvature of a movement as a
-measure of how much they've learned this movement?
-
-This could be useful for the context-dependent learning paradigm. The question
-to answer is: if a participant learns context A fully (and their movements
-become more straight), when switching back to A, will the first movement be
-straight or curved? If it's straight, then we can link this to the idea of
-context-based learning.
-
-NOTE: Gurevich\_93 is a doctoral thesis that doesn't seem to be available.
-
-\section{Quick adaptation to perturbations in the irrelevant manifold.}
-Models for motor control can be divided into two categories: (1) optimal
-trajectory and (2) end point. Optimal trajectory are those in which the
-controller tries to stick to a previously-calculated optimal trajectory and any
-perturbation (internal, external) will be dealt with by trying to go back to the
-optimal trajectory. Endpoint models, on the other hand, do not care that much
-about the trajectory itself and, with models that follow the ``minimal
-intervention'' principle, will only correct for perturbations if they mess with
-the endpoint.
-
-Endpoint models are more recent and arguably better than optimal trajectory
-ones. However, there is one thorn on their side: if during a movement an arm is
-perturbed, it will immediately try to return to the optimal trajectory
-\citep{Bizzi_Posture_1984}, even if the perturbation does not detract from the
-endpoint (this last part is my speculation). \cite{Todorov_Optimality_2004}
-argues that this is due to impedance, as a sort of immediate dumb fighback.
-
-A nice low-hanging fruit project would be to test whether
-\cite{Todorov_Optimality_2004} is right by having participants perform a
-reaching task in real life, where a motor creates random perturbations (in the
-irrelevant manifold, for good measure) and compare their reactions to the same
-task in virtual reality where the perturbation happens in the VR environment,
-but not in real time. If participants adapt in the same way in both
-environments, then \cite{Todorov_Optimality_2004} is wrong.
-
-\end{chapter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{chapter}{Paper}
-In this chapter, I describe the everything related to the paper that should come
-out of this project, including modeling and experimental findings, as well as
-the direction of the paper and things needed to get there.
-
-\section{Project description}
-The idea of the project is the following: We will create an agent that learns to
-control an arm in environment A (say, the real world carrying nothing), and
-learns how to control the arm in environment B (say, with high winds). The agent
-then learns to recognize environments A and B and to deploy the right controller
-for each environment depending on this recognition.
-
-We will argue that this is how humans do it and present evidence (based on
-literature research) for the predictions this model makes, as well as try to
-make new predictions.
-
-After this methods paper is done, model validation can be carried out in future
-studies.
-
-\section{Planning}
-Here I enumerate the things I need to learn and do in order to get this project
-moving and to finish it.
+\section*{Time plan for the arm movement project.}
 
@@
 End Date & Milestone & A priori comments & A posteriori comments \\ \hline \hline
-12.02.2021 & Learn about published models on arm-movement planning, especially those that can be used as generative models & & Discovered low-hanging fruit.
-Shifting directions a bit.
-See \ref{subsec:proposal}\\ \hline
+12.02.2021 & Learn about published models on arm-movement planning, especially those that can be used as generative models & & \\ \hline
 19.02.2021 & Learn about measures for how practiced a movement is. Reaction times? Speed profiles? Trajectories? & & \\\hline
@@
 
-\end{chapter}
+\section*{Time plan for the motor adaptation project.}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The previous time table was interrupted by the possibility of the new project. Now that this new project has taken shape, this new time table applies:
 
-\begin{chapter}{Modeling}
-\section{Random thoughts}
-I should read Knill and Pouget 2004 for Bayesian calculations in neurons.
-
-\section{Categorizing models}
-
-\section{Models to implement}
-\subsection{Optimal feedback control from \cite{Todorov_Optimal_2002}}
-\label{subsec:todorov-2002}
-This model, very well specified in the supplementary information of
-\cite{Todorov_Optimal_2002}, is an optimal control model, where some cost
-function is established and solved alongside the dynamical system.
-
-The model is based on what they call the ``minimal intervention principle'',
-where trajectories are only corrected when they interfere with the goals. This
-is fancy talk for the idea of using cost functions that are based entirely on
-the final goal (and subgoals, if any), as opposed to calculating optimal
-trajectories and trying to stick to them. The cost function also must be written
-in a way that establishes what is relevant and what isn't. For example, they
-talk about a cost function of the form $x_1 + x_2 - x*$, which does not specify
-what the end positions of $x_1$ and $x_2$ have to be, but only that their sum
-must be close to $x*$.
-
-Once the cost function is created, the system of differential equations (see
-below), cost function and observation feedback functions are all solved by
-casting the problem in terms of a Kalman filter.
-
-It seems like the model is ready to be applied as-is, with modifications only at
-the level of the dynamical system and the cost function.
-
-From \cite{Wolpert_Principles_2011}:
-\begin{enumerate}
-\item ``The minimum intervention principle has now
-been demonstrated in a number of tasks including the seemingly simple task of
-generating a target force with the tip of the index finger[35]''
-\item ``Recently, it has been pro- posed that such impedance control can be
-brought within an optimal control framework by formulating impedance control as
-the optimal response in the face of uncertainty about the dynamics of the body
-and environment[39].''
-\end{enumerate}
-
-\subsection{Adaptive dynamics from \cite{Shadmehr_Adaptive_1994}}
-\label{subsec:shadmehr-1994}
-The gist of this model is as follows: first, determine an optimal trajectory
-$q^*(t)$, then set up the controller of each joint such that $q^*(t)$ is an
-attractor (for $0 \le t \le T$) of the system and then just let it rip. The
-optimal trajectory in this paper is obtained using the minimum jerk model
-\citep{Hogan_organizing_1984}.
-
-Because of the use of an optimal trajectory that the controlled system is trying
-to stick to, this model differs from that presented by
-\cite{Todorov_Optimal_2002} (see \ref{subsec:todorov-2002}). In addition to this
-essential difference, this model also does not have the possibility of feedback
-signals (visual or proprioceptive) and, due to its complexity, is probably less
-viable as a generative model for inversion, though that might be a problem for
-future studies and not this paper.
-
-\subsection{Object avoidance from \cite{Hamilton_Controlling_2002}}
-In this paper they're touting the awesomeness of the TOPS model
-\citep{Harris_Signaldependent_1998} and applying it to a reaching problem with
-an obstacle in between.
-
-The implementation of this model is way too handcrafted. The gist of it is as
-follows: They put an obstacle between the starting and ending positions and
-assume that the optimal trajectory is a polynomial of fifth order (maybe even
-just $y = Ax^5$, I'm not sure). To optimize its parameters they created a cost
-function with two parts: the first part is the predicted covariance of the end
-point and the second part is a term penalizing ``visiting'' the obstacle. They
-basically discard any trajectory that gets close to the obstacle.
-
-While it works, I find it too ad hoc to be useful.
-
-\subsection{2-timescale motor learning from \cite{Lee_Dual_2009}}
-\label{subsec:lee-dual}
-This is a motor learning model, i.e. a model for how we can learn to perform
-movements in an environment that is different from the default. For example,
-under strong winds. It's an extension of the model presented by
-\cite{Smith_Interacting_2006} to the problem of ``dual adaptation'', i.e. the
-adaptation to different contexts and their force fields, and switching between
-them.
-
-While they present many models for dual adaptation, through model comparison (of
-the qualitative type) they arrive at the conclusion that the only good one is
-the parallel 1-fast N-slow model, which I describe here.
-
-Consider an abstract motor system (for example, an arm) that has learned to
-perform its movements in a baseline environment. Let $y(t) \in \mathbb{R}$ be
-the outcome for some motor signal given at time $t$ by the controller in the
-baseline environment. Now, assume that in a new environment, with some force
-field that is unknown to the controller, the controller issues the command that
-under the baseline environment would lead to $y(t)$, but, due to the force
-field, it leads to the outcome $f(t)$. Now, define a prediction error between
-the expected outcome and the observed outcome as $e(t) = f(t) - y(t)$. The model
-from \cite{Lee_Dual_2009} postulates two adaptation systems, one fast and one
-slow, which learn from this prediction error.
-
-Both slow and fast systems adapt with a similar rule:
-\[
-x_{f/s}(t + 1) = A_{f/s}x_{f/s}(t) + B_{f/s}e(n)c(n)^T \label{eqn:lee-learn}
-\]
-where $x$ is a ``learning process'', $f/s$ refers to fast/slow (one equation
-each), and $A$ and $B$ are learning rates. $c(n)$ is the environment variable
-that informs the controller which environment is currently active.
-$x_s,c, B \in \mathbb{R}^M$ to account for multiple environments. In the case
-where the learning of one environment does not interfere with the learning of
-others (as in \cite{Lee_Dual_2009}), $c = (0, ..., 1, 0, ..., 0)$, where the 1
-is in the position of the current environment. Note that the fast process is
-single-dimensional in this model (see \citep{Lee_Dual_2009} for the reasoning).
-
-The controller is then given by:
-\[
-y(t) = x_f(t) + x_s(t)c(t)^T
-\]
-
-\subsection{Proposal: Bayesian motor learner}
-\label{subsec:proposal}
-This model is a Bayesian alternative to the N-state learners from
-\cite{Lee_Dual_2009}. In this model, there is no need for explicit different
-timescales and multiple states; instead, the model relies on Bayesian learning
-of the best control, combined with inference over the current context. This
-model will, hopefully, reproduce all the experimental findings from the
-literature on motor learning (e.g. saving, interference).
-
-The model consists of three parts: (1) context inference, (2) decision making,
-(3) updating of the decision-making rules.
-
-Before describing these parts, I present some nomenclature:
-\begin{itemize}
-\item $t$: Time at the moment an observation is made, before inference is done.
-\item $\theta_t$: Observation. Note that this is assumed to be a noisy
-observation of the generalized state (see below) given by $\theta_t = h(z_t)$.
-\item $s_t$: Hidden state, e.g. the position of the hand. It is assumed not to
-depend on context. E.g. the physical position of the hand.
-\item $a_t$: Action taken after observing $\theta_t$ and inferring the context.
-\item $c_t$: Context inferred after observing $\theta_t$.
-\item $\omega_t$: Contextual cue.
-\item $z_t$: Generalized state, $z_t = \{s_t, c_t, \omega_t\}$. Observed noisily
-as $\theta_t$.
-\end{itemize}
-
-Context inference can work through context-specific (possibly unreliable) cues,
-or through the dynamics themselves, using prediction error as the force behind
-inference. Let us assume that both parts are present in the observation
-$\theta_t$. In this case, the inferred generalized state is given by
-\[
-  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
-\]
-
-Additionally, after having inferred the current state, motor adaptation is given by
-\[
-q(\gamma | z_t, z_{t-1}, a_{t-1}) \propto p(z_t | z_{t-1}, a_{t-1}, \gamma)p(z_{t-1})p(\gamma)
-\]
-
-where $\gamma$ are the parameters of the internal model for the dynamics,
-i.e. the function $f: (z_t, a_t) \rightarrow z_{t+1}$, which determines the beliefs of
-the agent as to how the system evolves after having taken action $a_t$.
-
-Having inferred the generalized state (including context), a decision is made
-sampling from $p(a_t | z_t, \beta)$, which is a distribution that is assumed to
-be parametrized by $\beta$. These parameters can be updated via:
-\[
-q(\beta | z_t, \theta_{t+1}, a_t) \propto p(\theta_{t+1} | \beta, z_t,
-a_t)p(\beta) \label{eqn:update-parameters}
-\]
-
-\subsubsection{Expected results}
-I predict that this model will display all the experimental findings that other
-models do, without invoking a possibly-infinite hierarchy of time scales, each
-with their own learning rate. Here I briefly go through the experimental
-findings and why I expect the model to reproduce them.
-
-Savings. This refers to the effect that, while learning a new context, if
-learning is interrupted by going back to a baseline context (e.g. no force
-field) and the new context is later retaken, the new dynamics do not need to be
-learned from scratch. Instead, learning continues where it was left off, with
-the possibility of some performance having been lost (i.e. learning having been
-forgotten). The Bayesian lerner displays this behavior simply by virtue of the
-learning of contex-specific action rules (learned dynamics), which are ``stored
-away'' once the context is no longer the new one (i.e. it has returned to
-baseline).
-
-Retrograde interference. This refers to the learning of context A interfering
-with the learning of context B. This occurs, for example, if context A has a
-force field that pushes right, while context B has a force field that pushes
-left. In this case, learning B is made slower by having learned Abefore,
-compared to learning B without having encountered A. This effect will be
-observed in the behavior of the model if context inference needs to accumulate
-before the agent is sure that a context switch has occured. In the meantime,
-context A's learned dynamics will interfere with context B's because actions are
-taken with a rule that depends on context (via $z_t$). Related to this, the
-model makes three predictions: (1) Contextual cues (that can be trusted) will
-reduce interference. (2) The amount of learning on contexts A and B will affect
-intereference such that highly trained contexts will interfere more
-strongly. (3) The duration of interference will be reduced in the case of
-highly-learned, opposing contexts, as the prediction error of the outcomes of
-actions will quickly drive contexts to switch to the current one.
-
-Rapid de-adaptation. This is the phenomenon in which forgetting an adaptation is
-faster than learning it. This would happen simply through context-based
-learning, where getting out of a context means that adaptations are no longer
-used (and eventually are forgotten). Note: I need to read more on this
-phenomenon, as I might be misinterpreting what it means.
-
-Spontaneous recovery. In an A-B-error\_clamp paradigm, in which during B the
-opposite adaptation to A is learned, at the beginning of the error-clamp phase
-(after B) participants seem to quickly gravitate towards A (thinking that A is
-now active?). This would be displayed by the model via the state inference, as
-now the visual cue and the observed dynamics point to different contexts. In
-time, updating $p(context | visual cue)$ will lead to spontaneous recovery.
-
-Bicycle effect. We never forget how to ride a bicycle. Existing models would
-need to incorporate an almost-infinite hierarchy of time scales to account for
-this, but my Bayesian learner naturally accounts for it by making use of
-probability distributions: the longer you practice something, the more
-entrenched the adaptation becomes by virtue of the posterior probability
-becoming very peaky.
-
-\end{chapter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{chapter}{Behavioral findings, Physiology and neurons}
-A few notes from \cite{Wolpert_Are_1995}:
-\begin{enumerate}
-\item Hand paths become straight with practice (Gurevich 93, Shadmehr and
-Musa-Ivaldi 94)
-\item Deafferented patients move differently (Ghez 90)
-\item Bell-shaped speed emerges from dynamics of control, not from planning
-(Jordan 94)
-\item Curvature matches the participant's messed up perception of straight lines
-(wolpert 94)
-
-\end{enumerate}
-
-\section{Notes from \cite{Harris_Signaldependent_1998}}
-\begin{enumerate}
-\item Arm stiffness can be adapted (Burdet et al 01.
-\end{enumerate}
-
-\section{Notes from \cite{Bays_Computational_2007}}
-\begin{enumerate}
-\item We have the ability to recall previously learned dynamics even after
-months! (Bashers-Krug et al 96, Gandolfo et al 96).
-\end{enumerate}
-
-\section{Notes from \cite{Shadmehr_Adaptive_1994}}
-\begin{enumerate}
-\item When planning movement, the target is transformed from a retinocentric
-vector to a head-centered and finally a shoulder-centered one. (Andersen\_85,
-Soechting and Flanders\_91).
-\item Gordon\_93 says that the target is finally represented as a vector from
-the current hand positoin (or whatever the effector is) to the goal position.
-\item From Lackner\_and\_Dizio\_92, aftereffects exist when introduced Coriolis
-forces are withdrawn. They also exist when perception is mucked with with prism
-glasses (Held\_59, 62, 63).
-\end{enumerate}
-
-\section{Notes from \cite{deC.Hamilton_scaling_2004}}
-\begin{enumerate}
-\item ``Muscle activity in specific tasks shows stereotypy whether subjects
-generate force with the fingertips (Valero-Cuevas et al. 1998), the wrist
-(Hoffman and Strick 1999), the neck (Vasavada et al. 2002) or the arm (van
-Zuylen et al. 1988; Flanders and Soechting 1990; Buchanan et al. 1993; van
-Bolhuis and Gielen 1997).''
-\item ``Primates also show repeatable patterns of muscle activation when
-grasping the same object repeatedly, and different activation patterns for
-different objects (Brochier et al. 2001). This suggests that stereotypy is a
-general characteristic of movement and is not unique to humans.''
-\item There is a paragraph that discusses different cost functions. For example,
-one is based on the total torque, another on fatigue and endurance, and other
-mathematical alchemy. Lots of references there too.
-
-\end{enumerate}
-
-\section{Notes from \cite{Kojima_Memory_2004}}
-\begin{enumerate}
-\item ``Previous studies have shown that adaptation is relatively specific to
-the size and direction of target eccentricity (Deubel et al., 1987; Semmlow et
-al., 1989; Frens and van Opstal, 1994; Straube et al., 1997; Noto et al., 1999;
-Watanabe et al., 2000).''
-\item ``It has been demonstrated that adaptation progresses more slowly when the
-subject is exposed to adapting target steps of a variety of sizes and directions
-in humans (Miller et al., 1981) and monkeys (Scudder et al., 1998).''
-\end{enumerate}
-
-\section{Notes from \cite{Lee_Dual_2009}}
-``When given contextual cues and sufficient trials, humans can simultaneously
-adapt to two opposite force fields (Osu et al., 2004; Nozaki et al., 2006;
-Howard et al., 2008), two saccadic gains (Shelhamer et al., 2005), or several
-visuomotor rotations (Imamizu et al., 2007; Choi et al., 2008).''
-
-\section{Motor adaptation}
-Motor adaptation to modified environments (e.g. force fields) has been shown to
-display the following characteristics:
-\begin{enumerate}
-\item Savings: Learning a previously-learned environment (after some extinction
-phase) is faster than the original learning. This means that the previously
-learned model is saved for a while and its learning can be resumed after
-extinction. However, \cite{Kojima_Memory_2004} showed that savings can be
-``washed out'' if baseline trials (no force field) are inserted between
-extinction and the second learning phase.
-\item Anterograde interference: Learning an adaptation (e.g. force field
-pointing left) slows down learning an opposing adaptation (e.g. force field
-pointing right).
-\item Rapid de-adaptation: Unlearning an adaptation (even after fully learning
-it) is faster than tlearning it \citep{Davidson_Scaling_2004}
-\end{enumerate}
-
-\end{chapter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{chapter}{Notes on papers}
-Conclusions and useful things about different papers I read. This differs from
-the comments above in that the notes here are holistic.
-
-\section{\cite{Shadmehr_Adaptive_1994}}
-In this paper they show a number of things. First, the show that people can
-adapt the dynamics of their control states to match changing dynamics of the
-observers. They also show that this adaptation takes the form of a straightening
-of trajectories, which are very twisted when they are first introduced to new
-system dynamics (viscosity of the environment); participants learn to do
-straight lines in the new environment after some trials, and the learing is
-monotonically increasing. They show a model of the dynamics of the arm, the
-control states and the adaptation that happens when exposed to the new
-dynamics. Finally, they show, both experimentally and in their model, the
-aftereffects of the new dynamics, whereby they display bad behavior when they go
-back to the natural environment, but eventually adapt to it as well.
-
-Interesting paper that cound be the foundation of our models, assuming that it
-hasn't been improved upon and maybe entirely replaced by newer stuff.
-
-A description of their model can be seen in \ref{subsec:shadmehr-1994} above.
-
-\section{\cite{Todorov_Optimal_2002}}
-See \ref{subsec:todorov-2002}
-
-\section{\cite{deC.Hamilton_scaling_2004}}
-The authors make use of the TOPS strategy \citep{Harris_Signaldependent_1998}
-and an experiment where they measure torque at different joints to show that
-bigger muscles produce less variability than smaller muscles for the same level
-of voluntary torque (I wonder what the opposite of voluntary is).
-
-It's an experimental paper with some muscle simulations thrown in for good
-measure. Contains a lot of references that might be useful later on the nature
-and characteristics of human and primate movements, as well as cost functions
-used in the past for modeling them.
-
-\section{\cite{Beers_When_2002}}
-This paper speaks of a model of optimal adaptation and weighting of visual and
-proprioceptive cues. Nothing of interest for us as of now.
-
-\section{\cite{Scheidt_Learning_2001}}
-Humans can adapt their movements to stochasticly varying changes in the dynamics
-of the environment. Moreover, they showed that their adaptation is most
-complatible with the idea that participants adapt to the mean of the dynamics (a
-one-directional force of changing magnitude) of the environment instead of doing
-trial-to-trial adaptations.
-
-These views, however, are challenged by the two-time-scales models of learning,
-like \cite{Smith_Interacting_2006}.
-
-\section{\cite{DeKleijn_Everyday_2014}}
-They cite \cite{Henry_Increased_1960} saying that participants have longer
-reaction times when preparing complex movements than when preparing simple
-ones. May be worth checking out.
-
-\section{\cite{Haar_Motor_2020}}
-This is the only paper I know where they do motor learning in a very natural
-task: playing billiards. It's great and I wish I could do this research.
-
-They tracked the joints of participants who were doing the same pool shot over
-and over (300 times). I do not think there was any variation in initial
-positions. They track performance and variability in joint movement.
-
-\section{\cite{Wolpert_Principles_2011}}
-Good review on motor learning. many relevant citations here.
-
-\section{\cite{Todorov_Optimality_2004}}
-Review of optimal control. Citation goldmine! Lots of papers on optimal control
-in biological systems are cited here.
-
-In particular, [43] and [44] might be similar to the idea of inferring goals and
-predicting future movements.
-
-``However, feedback control[25] explains an important additional observation:
-the increased duration of more accurate movements is due to a prolonged
-deceleration phase, making the speed profiles significantly skewed[88].''
-
-\end{chapter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{chapter}{Implementing the adaptation model} In this section, I propose a
-simulated task in which the prowess of the model proposed in
-\ref{subsec:proposal} is put to the test. I additionally present simulation
-results and some writeup. Basically a mini-paper on this model.
-
-\section{The Task} The tasks used for models like this one are usually
-\citep[e.g.][]{Lee_Dual_2009} reaching tasks: a participant sits in front of a
-table with shoulders immobilized, and has to bring her hand from the starting
-position A to a designated position B. This is done while holding a robotic arm
-that moves with the hand. The robotic arm is used to create force fields in
-different directions (CW and CCW, typically, but it's basically left and right)
-for the different contexts. The position of the participant's arm is recorded
-and deviations from the straight line are considered as errors ($e(t)$ in
-\ref{eqn:lee-learn}).
-
-To simplify the task, I propose the following: participants need to keep their
-hand still in the starting position for the duration of the trial. The robotic
-arm creates forces that will drive the arm away from its current position by
-pushing in a constant direction with constant force. This experiment has the
-same elements as the one by \cite{Lee_Dual_2009} but eliminates the necessity
-of an underlying movement. While it works wonders as a thought experiment and
-helps me see if the proposed model behaves the way I want, it must not be used
-as a real experiment because participants could simply use impedance for each
-trial, regardless of context.
-
-In this experiment, every context $C_i$ is characterized by the strength and
-8'rection of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ...,
-N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force
-measured Cartesianly, and $w_i$ is the intensity of the force.
-
-As in \cite{Lee_Dual_2009}, I will simulate an A-B-clamp paradigm, which starts
-with a context $C_0 = \{0, 0\}$ in which no force is being applied. Then
-context $C_A$ is shown for a few trials, then context $C_B$, then an
-error-clamp block in which the hand is forced to stay at the initial position
-by the robotic arm.
-
-Additionally, a contextual cue $\omega$ is shown to the participant which
-signals the current context. The experiment will be run with different levels
-of reliability of this cue. Different sessions will be run with different cue
-reliabilities.
-
-An experimental session $\Xi$ is characterized by the reliability of the cue,
-the contexts avaiable and the order in which they are shown. In a given
-session, each trial begins with the cue and the XXX seconds during which the
-participant must keep the hand in the starting position. During the intertrial
-time, no forces are applied to the hand, but the hand must remain in the
-starting position.
-
-\section{The model}
-The model is an implementation of \ref{subsec:proposal}. The generalized state
-is given by:
-\begin{equation}
-z_t = \{x, \theta_t, \gamma_t, \omega_t\}
-\end{equation}
-where $x$ is the current position of the hand in Cartesian coordinates,
-where the origin is the starting point. We assume that motor commands are
-issued every $\Delta t$ for simplicity. At the beginning of each time interval,
-the context is inferred combining the cue and the prediction error of the
-outcome of the previous motor command. We assume further than the cue and the
-force field are uncorrelated, which yields:
-\begin{align}
-  q(C_t) &= q(C_t | \omega_t)q(C_t | s_t, s_{t-1}, a_{t-1}) \\ \label{eqn:estimated-context}
-  q(C_t | \omega_t) &\propto p(\omega_t | C_t)p(C_t) \\
-  q(C_t | x_t, x_{t-1}, a_{t-1}) &\propto p(x_t | C_t, x_{t-1}, a_{t-1})p(x_t)p(x_{t-1})p(C_t)
-\end{align}
-The terms $p(C_t)$ refer to the prior probability of the context at
-the beginning of the trial: at the first trial, this refers to prior beliefs
-over which contexts are more common/likely. At each subsequent trial, it
-incorporates the belief that has so far been accumulated, given previous
-trials.
-
-Once the posterior over contexts has been calculated for the current trial, it
-is used to select the current context. A simple rule is to sample the context
-directly from the posterior distribution. However, a useful extension of this
-is to use a softmax function to create a distribution:
- !''¿
-\begin{equation}
-S(C^{(t)}) = \frac{e^{\kappa
-q\left(C^{(t)}\right)}}{\displaystyle \sum_{i=1}^{N_c}e^{\kappa
-q\left(C_i\right)}}
-\end{equation}
-where $q(C_i)$ is given by \eref{eqn:estimated-context}. The
-function $S(\cdot)$ is then used to sample the (inferred) current context. This
-representation has the advantage of adding a parameter with which the sampling
-behavior can be characterized, ranging from choosing uniformly when $\kappa =
-0$ to choosing the context with the highest probability deterministically as
-$\kappa \rightarrow \infty$.
-
-In the case where the hand position can be determined by the participant with
-perfect accuracy (ignoring visual and proprioceptive error), we have that
-$p(x_\tau) = \delta(x_\tau, \hat x)$, where $\hat x$ is the real position of
-the hand, and $\delta (\cdot, \cdot)$ is the Kronecker delta function. We
-assume this to be the case during this experiment.
-
-  For simplicity, we assume that participants know how many distinct contexts
-will be shown to them, and therefore all the $C_x$ have been created by the
-agent before the session begins. By default, all contexts assume a force with
-angle and magnitude equal zero, and their true values are learned during the
-experiment (see below).
-1234568¿|
-For mathematical tractability (i.e. to avoid circular distributions), we will
-limit contexts to one where the force points left (angle $\pi$) and one where
-the force points right (angle 0), in addition to the baseline context. In this
-scheme, participants need only learn the magnitude of the force for each
-context.
-
-The generative model for the force can take on different shapes for the prior
-and posterior distributions. In this work, we explore three different forms of
-this generative model which differ on the assumptions the agent makes regarding
-how the force fields are being generated by the environment. Each version of
-the model is discussed separately below and their pros and cons discussed;
-additionally, simulations are presented in the following sections.
-
-To infer the parameters of the posterior distributions, we make use of Bayes'
-theorem. In the following equations, we drop the dependence on the context and
-observations for notational simplicity:
-\begin{equation}
-q(\theta | x_{1:t}) \propto p(x_{1:t} | \theta)p(\beta)
-\end{equation}
-where $\theta$ are the parameters of the force field to be estimated and
-$x_{1:t}$ are the data observed until time $t$. Across all variants of the
-model presented below, the likelihood will be a Gaussian:
-\begin{align}
-  p(x_t | \theta) &= N(\mu_x, \sigma_x)  \\
-  \mu_x &= \theta_{t-1} + x \\
-  \sigma_x &= \xi_0 1
-\end{align}
-This likelihood is the probability of observing $x$ given the previous estimate
-of the parameters of the force field $\theta_{t-1}$, and given the dynamics of
-the system (see below).
-
-The final part of the model to establish is action selection. Action selection
-is done at the level of desired outcome (as opposed to motor commands)
-following the equations:
-\begin{align}
-  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta
-            t}\epsilon\\ \label{eqn:dynamics}
-  f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
-  a_t &= \argmax_{a_t}p(x_{t+1} = (0, 0) | x_t, a_t, f_t)
-\end{align}
-where $\gamma_t$ and $\theta_t$ are the participant's estimate of
-the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian
-motor error component.
-
-\subsection{Left-right with fixed force value (LR)}
-In this version, the force field is assumed to have a fixed value which the
-agent tries to infer. The agent's uncertainty on this inference is reflected on
-the parameters of the posterior over the force directly. For simplicity, we
-assume that the priors (and posteriors) are normal with mean and standard
-deviation to be updated at every time step. Therefore, the posterior over the
-magnitude of the force for each context is given by:
-\begin{equation}
-q_t(\gamma) = N(\mu_{\gamma}, \sigma_\gamma)
-\end{equation}
-where $q_t(\gamma)$ is the posterior of the magnitude $\gamma$ of the force
-field distribution after the observations at time $t$, and $\mu_\gamma$ and
-$\sigma_\gamma$ are the inferred parameters of the distribution.
-
-Because the likelihood is a Gaussian with support $\mu_\gamma$, the update
-equations are given by:
-\begin{align}
-  \mu_\gamma^{(t)} &= \frac{\mu_\gamma^{(t-1)}\left(\sigma_\gamma^{(d)}\right)^2 + \mu_\gamma^{(d)}\left(\sigma_\gamma^{(t-1)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2} \\
-  \sigma_\gamma^{(t)} & = \sqrt{\frac{\left(\sigma_\gamma^{(t-1)}\right)^2\left(\sigma_\gamma^{(d)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2}}
-\end{align}
-where the superscripts $(t)$ and $(t-1)$ refer to the inferred values for the current and previous trials, respectively, and $(d)$ to the parameters in \eref{eqn:data-dist}.
-
-\subsection{Left-right with unknown mean (LRS)}
-In this model, the force at each trial is assumed to be sampled from a normal
-distribution and the mean of this distribution is inferred from the
-observations, while the standard deviation is assumed to be known. The mean of
-the force is assumed to also be normal with unknown mean and standard
-deviation:
-\begin{align}
-  p(\gamma) &= N(\mu_t, \sigma) \\
-  p(\mu_t) &= N(\mu_mu, \sigma_\mu)
-\end{align}
-In this manner, the uncertainty for this agent is at the level of the estimate
-of the parameters of the force field, as opposed to being directly on the
-force. Because the priors for $\mu_\mu$ are normal and the standard deviation
-$\sigma$ (not $\sigma_mu$) is known, the update equations are given by:
-\begin{align}
-  \mu_\mu &= \frac{\sigma^2}{\sigma^2 + \sigma_\mu^2}\mu_\mu + \frac{\sigma_\mu^2}{\sigma^2 + \sigma_\mu^2} x \\
-  \sigma_\mu &= \frac{\sigma^2\sigma_\mu^2}{\sigma^2 + \sigma_\mu^2}
-\end{align}
-where $x$ is the observation.
-
-\subsection{Left-right with unknown mean and standard deviation (LRMS)}
-In this version, we will assume that for each context, the agent's belief over
-the magnitude of the force is given by a normal distribution:
-\begin{equation}
-p(\gamma) = N(\mu_f, \sigma_f) \label{eqn:data-dist}
-\end{equation}
-where $\mu_f$ and $\sigma_f$ are parameters to be estimated at each
-trial. Before the experiment begins, the agent will have a prior distribution
-over these parameters. A standard Bayesian approach is to choose NormalGamma
-priors:
-\begin{equation}
-p(\mu_f, \sigma_f) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
-\end{equation}
-where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
-model. Because participants are told that there are three contexts (baseline,
-left and right), we assume the following hyperparameter values:
-\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
-  \nu_f^{(0)} &= 0.1 \\
-  \alpha_f^{(0)} &= 1 \\
-  \beta_f^{(0)} &= 0.5
-\end{align}
-Note that $\mu_f^{(0)}$ is different for each context and is thus
-provided as a vector $(0, -1, 1)$ representing the baseline, left and right
-contexts, respectively. For all other hyperparameters, all contexts have the
-same value. These values ensure that the prior over the hyperparameters
-$(\mu_f, \sigma_f)$ has the mode at $(\mu_f^{(0)}, 1)$. We will show that the
-choice of these hyperparameters affects learning rate, but not the final
-learned magnitude \todo{Check this}.
-
-The update equations for the magnitud parameters are given by:
-\begin{equation}
-q(\mu_f, \sigma_f | x_t, x_{t-1}, a_{t-1}) \propto p(x_t |
-x_{t-1}, a_{t-1}, C_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
-\end{equation}
-where $p(x_t | x_{t-1}, a_{t-1}, C_t)$ is a Gaussian distribution centered
-around $\mu_a$ with standard deviation $\sigma_a$, which is a free parameter of
-the model. It is assumed that the standard deviation over outcomes ($\sigma_a$)
-is related to that of the parameter $\mu_f$ as follows:
-\begin{equation}
-\sigma_f = \sigma_a / \nu_f
-\end{equation}
-which makes $\nu_f$ a parameter to estimate. Because we chose the priors to be
-NormalGamma and the likelihood to be Gaussian, the posteriors are also
-NormalGamma, and the parameters are updated after one time step via:
-\begin{align}
-  \mu_f^{(t)} &= \frac{\nu_f^{(t-1)} \mu_f^{(t-1)} + x_t}{\nu_f + 1} \\
-  \nu_f^{(t)} &= \nu_f^{(t-1)} + 1 \\
-  \alpha_f^{(t)} &= \alpha_f^{(t-1)} + 0.5 \\
-  \beta_f^{(t)} &= \beta_f^{(t-1)} + \frac{\nu_f^{(t-1)}}{\nu_f^{(t-1)} +
-                  1}\frac{\left(x - \mu_f^{(t-1)}\right)^2}{2}
-\end{align}
-
-\section{Parametrization of the task and the model}
-Some of the parameters of the model can be fixed by the physics and scales of
-the simulated experiment.
-
-Henceforth, distances are measured in meters and time in seconds. As a
-consequence, forces should be measured in Newtons, but bare in mind that what
-the model calls ``force'' is no such thing. Instead, ``forces'' in the model
-are the distance (in meters) that the real force (in Newtons) would make the
-arm move in some $\Delta t$, assuming a mass of 1Kg and zero initial
-velocity. This simplification of the terms might come back to bite us in the
-ass and might be replaced later. But first, let us speak of real forces.
-
-The force exerted by the mechanical arm depends on the context. In the baseline
-context, the force is 0N. Taking inspiration from \todo{Citation needed}, we
-take the forces of the other two contexts to be \force and -\force,
-respectively. We will assume that $\Delta t = \dt$ \todo{Need to find a good
-  way of fixing this}, which represents reaction, processing and decision
-times.
-
-Back to the lazy encoding of forces in the model, one can calculate the effects
-of these forces on the arm. With a force of 10N and a $\Delta t$ of 50ms, the
-total distance the arm will be moved by the force (again, assuming mass of 1Kg)
-is given by:
-\begin{align}
-  d_\pm &= \pm \frac{1}{2}(\force / 1Kg)((\dt)^2)\\
-               &= \pm 0.0125m
-\end{align}
-The agent will try to estimate $d_\pm$ during the experiment.
-
-We will assume \todo{This is arbitrary. Might want to justify it later.} that
-the participant observes the position of her own hand (visual + proprioceptive)
-with Gaussian observation noise with $\mu = 0m$ and $\sigma = 0.005m$.
-
-\section{Simulation results}
-\subsection{Deadaptation and context inference}
-During the deadaptation phase, not all parametrized agents detect the correct
-context. For each agent type (LR, LRS, LRMS), different parameter values allow
-the information from the predicted hand positions to overcome the information
-from the cue, which enables the correct identification of the current context.
-
-It is unclear which parameter values lead to this, but, as expected, prediction
-noise (see XXX) and context noise (see XXX) play an important role. In
-particular, keeping the prediction noise below 0.2 seems to be necessary for
-the correct identification of the context, for all the used values of context
-noise. It could be good to check this.
-
-\subsection{Context inference requires a new likelihood}
-\label{subsec:new-likelihood}
-It became clear by running preliminary simulations that, regardless of the
-agent, context inference happened too quickly and too ``reliably''. The context
-was identified during the first trial at first, and during deadaptation, when
-the cue and the forces do not match, it took at most a couple of trials to
-recognize the real context.
-
-Since the model proposes that deadaptation is nothing but a consequence of slow
-recognition of the context, lightning-fast context recognition will just not do
-it.
-
-The problem stems from the fact that the estimate on the magnitudes becomes so
-good (e.g. low standard deviation on the estimate) that the likelihood
-$p(x | C_t)$ becomes a delta function. Because of this, the cue itself is being
-ignored and the context is immediately identified.
-
-The best fix to this problem is to modify the way the likelihood works. As it
-stands, the likelihood is the probability of observing the current hand position
-given the current context and the previous hand position. The more ``separated''
-the different estimates become from each other (the means are too far apart
-and/or the precisions are too high), the more this is going to be a delta
-function.
-
-The best fix for this is to modify the likelihood function and introduce in it a
-measurement of the reliability of each of the modalities of perception (force
-feedback and visual cue). Such moderators are common in the literature
-\citep[e.g.][]{Lee_Neural_2014}.
-
-The simplest implementation I can think of is
-one in which additive noise is added to the posterior estimate for each
-modality:
-\begin{equation}
-p(C | \omega_t) \leftarrow (p(C | \omega_t) + \chi_{\omega})
-p(C | s_{t-1}, s_t, a_{t-1}) \leftarrow (p(C |  s_{t-1}, s_t, a_{t-1}) + \chi_{\text{pred}})
-\end{equation}
-where $\chi_{\omega}$ and $\chi_{\text{pred}}$ are constants which encode the
-reliability of each modality.
-
-A more convoluted approach would be to bring all modalities to the same
-level. In the current implementation, the likelihood uses the standard deviation
-of the inference over the magnitudes to calculate $p(C | s, a)$. Instead, a
-standard deviation could be used that is proportional to the mean, which would
-put the cue and predictive modalities on the same playing ground, depending on
-the proportionality constant chosen.
-
-\subsection{Parameter values to look into}
-By visual inspection of the results of simulations, these are the parameter values that seem promising:
-
+\noindent
 \begin{tabularx}{\textwidth}{
+  l|
+  >{\hsize=.4\hsize\linewidth=\hsize}X|
   >{\hsize=.2\hsize\linewidth=\hsize}X|
-  >{\hsize=.1\hsize\linewidth=\hsize}X|
-  >{\hsize=.1\hsize\linewidth=\hsize}X|
-  l}
-% {l|{\hsize=.4\hsize}X|{\hsize=0.3\hsize}X|{\hsize=0.3\hsize}X}
-Agent & Obs noise & Cue noise & Notes \\ \hline \hline
-LR    & 1.2       & 0.04      & Slow fake-context inference. Maybe the priors coming from the previous stage are messing things up. \\ \hline
-1 & 1 & 1 & 1
+  >{\hsize=.2\hsize\linewidth=\hsize}X}
+End Date & Result & A priori comments & A posteriori comments \\ \hline \hline
+30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & \\ \hline
+07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
+14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
+--- & Start writing the paper. & & \\ \hline
 \end{tabularx}
->>>>>>> adc91ff996e45d6eaf8974cd33ea69b5d380ed07
-
-\subsection{Adaptation}
-Adaptation happens mostly very quickly for all agents.
-For the current $\Delta t = \dt$, adaptation happens within 20 to 30 time steps, although many times (depending on agent and parametrization) it happens much more quickly.
-For example, for the LR agent with prediction noise of 0.15, cue noise and context noise of 0, and assumed force variability of 0.25, within 15-20 time steps the adaptation is complete.
-\section{Conclusions}
-
-\end{chapter}
-
-\bibliography{../MyLibrary}
 

### 2021-05-11 notes/notes.tex
Models back to normality. Before context-weighted adaptation

@@
 End Date & Result & A priori comments & A posteriori comments \\ \hline \hline
-30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & \\ \hline
+30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & I found a lot of evidence. Moving on.\\ \hline
 07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
@@
 volatility of the environment, regardless of the direction of adaptation.
+
+\section{Notes on savings, quick deadaptation and contextual cues}
+In \citep{Kojima_Memory_2004}, Figure 1a-b, we can see that de-adaptation was a
+bit faster than adaptation for the gain-increase experiment, and much faster
+for the gain-decrease experiment. It seems like there is no contextual cue
+available to the monkey, as all trials (over 1k) were presented one after the
+other with no breaks. Poor monkey.
+
+\citep{Kojima_Memory_2004} additionally show savings in the form of
+re-adaptation that, for a few hundred trials, is much faster than the original
+adaptation (control adaptation). This can be better seen in Figure3a-b. They
+discuss that there is even a point of inflection in test-adaptation that
+clearly marks the end of savings and resuming of adaptation. The duration of
+the savings phase changed from experiment to experiment, from monkey to
+monkey. The authors do not check whether the past-facilitation rate of learning
+is similar to that of the end of the control adaptation phase, which would
+indicate that the monkey quickly went back to where it had left off.
+
+Finally, they show that savings seem to disappear after a period of
+error-clamp-like trials, in which the ISS is set to zero, thus allowing the
+monkey hundreds of trials in which his adapted saccades (gain != 1) yielded
+correct results. In contrast, savings do not disappear after the monkey head
+was covered with a black cloth, placing him in total darkness (no feedback for
+his saccades).
+
+Given the lack of contextual cues, context identification could be very slow in
+monkeys, which would explain the facilitation period. During normometric
+saccades after adaptation (error-clamp-like trials), context inference could
+have been cimented to the point that it became difficult for the monkey to
+shift away from it once test adaptation began, thus removing savings from the
+equation. During darkness trials, in contrast, there is no contextual evidence
+in any direction, and thus the precision over context does not change. In this
+case, savings are back on the table.
+
+\cite{Ethier_Spontaneous_2008} present a simple experiments with saccade
+adaptation in humans. In their paradigm, humans go through a O-A-B-E paradigm,
+in which the adaptation is forced by making the target move closer or further
+away than it was before, making participants believe that they had
+over/under-shot the target. While this study does not quite talk about savings,
+there is evidence of context-based savings in their paradigm, as adaptation
+trials were done in blocks of 60 trials, with 30s pauses between blocks, in
+which participants closed their eyes. In their figure 2, they show that each
+pause caused a dip in the observed adaptation. While this pause does not
+provide any clues as to the identity of the following block, it does indicate
+that a new block will start, which could make participants revert to using
+their priors over contexts to make the first decision in the following block
+and consequently using the feedback to perform context inference, explaining
+the quick dip after each pause.
+
+Their study does show quick de-adaptation during a second block of trials in
+which the desired gain is opposite of the adaptation trials. However, it is
+difficult to interpret these savings in terms of the context-based learner, as
+learning in the original adaptation is also quite fast, and the B context had
+not been introduced in advance. However, the fact that group 2 (which has
+longer pauses) shows faster de-adaptation supports the idea of context-based
+learning, as the longer pause before de-adaptation causes context inference to
+rely more on priors and thus make the jump from model A to model O almost
+immediately, and then summoning model B and learning at the same speed as model
+A was learned.
+
+Importantly, the effect of these pauses speak for continuity-based priors,
+which are based on previous trials and not on prior beliefs regarding which
+contexts would, in general, be encountered more often in the world.
+
+\cite{Huang_Persistence_2009} presents interesting data, but I can't interpret
+it for now.
+
+\cite{Lee_Dual_2009} present a study that is very useful to us. In it,
+participants have to use a joystick to move a cursor from its initial position
+to a target, under different visuomotor rotations (tasks A, B, C and D). Of
+special significance is the fact that participants were provided with very
+clear contextual cues, as the target area was in a position that uniquely
+identified the rotation. The effects of these cues can be seen in their figure
+6, in which they show how participants very quickly learn the adaptations
+during an A-B-A phase, with very quick jump from A to B and from B to A. Then,
+participants enter a phase in which tasks C and D are randomly chosen every
+trial. Because of the unambiguity of the contextual cues, even during learning
+participants display very clear switching of internal models. This is so far
+the most clear example of contextual cues causing immediate ``deadaptation''
+(which we know is really switching, but don't tell anybody).
+
+\cite{Forano_Timescales_2020} present a study with humans and a mechanical arm,
+in which participants have to perform forward reaching movements under
+different forces. During their experimental sessions, their complications for
+which are endless and unnecessary, there is a baseline phase, in which
+participants perform no-force-field movements in two different workspaces
+(strong contextual cues). Then an adaptation phase starts, in which
+participants do eight trials in context A, eight in B (with random Es thrown in
+for good measure). Awful things happen later on, of which we do not
+speak. During de-adaptation, they follow the same 8-8 structure, but the cues
+are flipped; this phase ends when the adaptation is taken back to zero. Finally
+the error-clamp trials start, with a contextual cue that is the same as the
+first one they got during the adaptation phase. They do not speak of savings,
+and savings cannot be seen in their plot, but quick de-adaptation is clearly
+seen in their figures 6 and 7. However, it is not as quick as it could be, and
+I believe this to be due to the strength of the cues, which cause a lot of
+conflict in the participant as the force-feedback and contextual cues are so at
+odds with each other. I would argue that the slow decay (still fast compared to
+adaptation) is evidence of yet another phenomenon: the devaluation (or
+re-evaluation) of contextual cues, in which the precision with which they are
+treated decays over time as a consequence of the apparently false information
+they provide. I expect this to interact with the following error-clamp trials,
+in which \cite{Forano_Timescales_2020} show small-but-significant spontaneous
+recovery. I believe that if those de-adaptation trials were removed, the
+increased strength of the contextual cues would make the spontaneous recovery
+larger.
+
+\cite{Davidson_Scaling_2004} tested quick de-adaptation more thoroughly than is
+usual, with the explicit goal of checking whether there was something special
+about A-O de-adaptation that made it quick, or if A-B de-adaptation would also
+be quick. Additionally, they explored the question of which one is faster. They
+used the typical curling-force mechanical arm for their experiment with human
+participants. Their results can be groupped into two: (1) De-adaptation from A
+to O is slower than de-adaptation from A to 1/3A. (2) De-adaptation from A to
+3A is perhaps a bit faster than from A to -A, but, more importantly,
+re-adaptation to A is faster coming from 3A than from -A. The first result can
+be explained simply: de-adapting from A to A/3 implies less ``travel''
+(i.e. adaptation) than from A to 0, and thus motor error takes longer to go
+down when de-adapting to O. This is precisely why \cite{Davidson_Scaling_2004}
+performed their second experiment from which (2) was obtained. Their second
+result can be explained by switching internal models, if action selection is
+done via a weighted sum of the actions generated by each possible model (where
+the weights are the posterior probabilities of the models). In effect, the
+assymetry is caused by the existance of O. In this account, going from 3A to A
+more quickly reduces motor error because the contributions of 3A and O to motor
+output ``cancel out'', as they point in opposite directions relative to A,
+leaving only the contributions of A. When coming from -A, the contributions of
+-A and O point in the same (incorrect) direction (relative to the real
+generating process A), thus making motor error be bigger.
+
+Re-adaptation in \cite{Davidson_Scaling_2004} differs greatly between -A and
+2A. While the explanation above could be enough, I believe that an extra factor
+would be the fact that while adapting to forces in different directions,
+different muscles must be used. The transfer from using a model that uses a set
+of muscles to another that uses another set of muscles might come with its own
+perils. More research into this could be useful.
+
+Interestingly, \cite{Davidson_Scaling_2004} calculated that re-adaptation from
+3A to A was not significantly different in experiments 1 and 2, even though
+participants in experiment 2 had already learned the A environment, while those
+in experiment 1 had not. This can perhaps be explained as a fluke in the
+switching mechanism, where the weight of the A model in experiment 2 made its
+contribution to action lie square in the middle of those from 3A and O. This
+might need further consideration.
+
+\cite{Zarahn_Explaining_2008} performed a typical experiment with visuomotor
+modifications (screen, hand position, axis rotation) to show that in their
+O-A-W-A trial, where W are washout trials, savings still existed, even after
+the fast-slow learner predicted that savings should have disappeared. They
+showed the same with O-A-(-A)-A and who knows what else. Their figure 3 is a
+nice poster child. Note that while they did not have any contextual cues, their
+transformations are HUGE. In their O-A-W-A experiment, A = 45°. A = 30° in their
+other experiment. This in stark contrast with many other studies that use
+10°-20° ranges.
+
+\cite{Kim_Neural_2015} present savings with very informative and clear
+cues. Their savings are beautifully portraited in their figure
+2A. Additionally, they present imaging results based on regressors obtained
+from the fast-slow linear models.
+
+\section{Notes on the effects of context inference on learning}
+
+\cite{Herzfeld_memory_2014} present what is possibly (so far) the most
+compelling evidence for the effects of contex-inference on learning itself,
+uncoupled from motor responses. The authors make the point (with a model and
+experiment) that the brain adjusts its sensitivity to errors based on the
+volatility of the environment. Moreover, this sensitivity depends on the size
+of the error, such that a participants' learning might be faster when facing
+small errors than with large ones. In their first experiment, the authors
+divided participants into three groups that differed only on the volatility of
+the environment, group 1 facing the most stable environment, where the
+probability that the perturbation changed from 13 to -13 (N s/m) was 0.1; group
+2 had a probability of 0.5 and group3 of 0.9. \cite{Herzfeld_memory_2014}
+labeled these groups as 0.9, 0.5 and 0.1, respectively. Here I keep with these
+labels.
+
+\cite{Herzfeld_memory_2014} showed in their figure 1D that the stable
+environment of group 0.9 made participants learn more quickly from errors, with
+the effects being lower for groups 0.5 and 0.1, in that order. In my model, the
+interpretation of these results is fairly straightforward: more volatile
+environments lead to less precise (and more often incorrect) context inference,
+which would slow down learning. In particular, in the mathematical
+implementations of the model so far (see \ref{sec:model-instance}), the
+precision with which observations are taken in depends on the precision of the
+context inference, which should create behavior like that observed by
+\cite{Herzfeld_memory_2014}. Note that the authors' results in experiment 2 are
+explained by my model in very much the same vein.
+
+One puzzling result from \citep{Herzfeld_memory_2014} is that sensitivity to
+error depends on the size of the error. With some effort (invested in
+understanding their paper...), this can be very nicely explained by the
+context-inference model. First, to put some of their numbers in perspective,
+their second experiment implies errors roughly in the range of -2 to 2 (which
+matches the x-axis in Figure 2). An error the size of -2 or 2 very clearly
+signals the change from one context to the other (0.9 gain to 1.1, or vice
+versa; it is unclear in the paper). Because the direction of the change is
+clear evidence from the previous and current contexts, context inference right
+after such a switch becomes very clear (the maximum of the posterior $>>$ the
+minimum of the posterior), which makes learning somewhat fast and does not
+depend on the volatility of the environment, at least in the first trial after
+a switch; during the next trials, the uncertainty increases again due to
+volatility. The smaller the error is (in absolute value), the less clearly it
+can be attributed to a switch in the environment, to the extreme where changes
+are minuscule and thus must be attributed to the motor adaptation's error; it
+is because of this that small errors have a higher effect on learning. Now, for
+smaller errors, there is a clear effect of volatility on learning. I believe
+this can be explained by the context-learner: Because of the overall slower
+learning in volatile environments, the internal models are, for any given
+trial, not as well learned for the high-volatility groups, which leads to many
+errors being misinterpreted as context switches (for which no learning might
+occur), making the average look worse than it is (except, as explained above,
+when the error is close to 2 or -2, which makes switches very evident).
+
+In fact, this introduces important confounds: errors of exactly 2 and -2 can
+only ocurr when the participant has learned both internal models perfectly and
+simply predicted the wrong one, thus adapting in the incorrect direction. In
+this case, learning should be exactly 0. However, errors very close to, but not
+exactly, 2 and -2 can be attributed to bad context inference plus the need to
+adapt the ``other'' model, i.e. the model that was not predicted. In this case,
+depending on how smart the participant is, the other model might be updated
+with an error sensitivity as high as that of the near zero error. On the other
+hand, participants observing an error close to zero might interpret this as a
+signal to update an internal model (in which case the sensitivity should be
+high), but they might also falsely (or even correctly) accuse the context of
+having changed, but with low precision, making learning of the ``other'' model
+very low (due to low precision on context), or even zero, if the participant is
+so unsure that she does not want to learn from this. Because of this, a more
+correct version of Figure 2E would take into account how well the internal
+models have been learned at each trial, how many trials were spent in
+perfect-model land, how many times the participant took a near-2 error as a
+switch (correctly and incorrectly), how many times near-2 errors occurred when
+the models been already fully learned, etc.
+
+The third experiment by \cite{Herzfeld_memory_2014} has too many
+confounds. They conclude that they can independently manipulate sensitivity to
+big and small errors, but I believe their experimental design has unwanted
+consequences because 8N and -8N were shown a lot during the fast-switch part,
+but 4N and -4N were not. It's such a mess that I don't want to touch it.
+
+\section{Notes on error-clamp weirdness}
+\cite{Kojima_Memory_2004} have error-clamp-like trials, but not at the end. No
+weirdness here. Also no cues.
+
+\cite{Smith_Interacting_2006} speak of error-clamp weirdness and presents a
+model to explain it. Their data is limited, however, and might not be specially
+useful for me.
+
+\cite{Ethier_Spontaneous_2008} have error-clamp weirdness and their
+within-block pauses might be useful for my interpretations.
 \end{chapter}
+
+\cite{Forano_Timescales_2020} has spontaneous recovery and slow tapering off
+during error-clamp trials, but it's quite small. Maybe because of the large
+de-adaptation that devalues contextual cues?
+
+\cite{Vaswani_Decay_2013} explores in detail the motor decay observed during EC
+trials after adaptation trials. They conclude, among other things, that the
+motor adaptation decays across as many as hundreds of trials, but never goes
+back to zero. Additionally, there is a lag before decay starts (all their
+experiments end in A-E, so no baseline before), and it depends on the
+participant, as well as how noticeable the transition to error-clamp trials
+is. They have some clever experiments.
+
+\section{Notes on implicit vs explicit, and strategies.}
+The idea behind strategies in motor adaptation is simple: participants can be
+told how to perform in a novel environment. Then, when the novel environment
+kicks in, there is no period of adaptation before performance is near-perfect.
+
+For example, \cite{Taylor_Flexible_2011,Mazzoni_Implicit_2006} showed that in a
+visuomotor rotation task, participants can be given the hint to ``do as if you
+were aiming for this other target'', which allows them to complensate almost
+perfectly from trial 1. However, the authors also found interesting effects:
+while at first participants have zero error, they slowly drift away from this,
+overcompensating for the visuomotor rotation. The authors showed that this
+overcompensation goes away, given sufficient trials.
+
+\cite{Taylor_Flexible_2011} present a computational model based on the linear,
+prediction-error based models \citep[e.g.][]{Smith_Interacting_2006} by making
+prediction error incorporate some element of the strategy. This model is fine
+and explains the observed phenomena, but I believe that they overlooked the
+mismatch between propriosceptive and visual feedback. This mismatch, which is
+independent of reward, would explain their results (at least qualitatively). I
+would speculate that the effect of the strategy given to the participants by
+\cite{Taylor_Flexible_2011} makes them not adapt their motor rotations, but
+rather shift the goal of the trial. Then, as trials go by, the propiosceptive
+feedback does not match what is observed on the screen, and that causes motor
+adaptation. Perhaps more recent studies have tested this.
 
@@
 \section{The model}
+\label{sec:model-instance}
 The model is an implementation of \ref{subsec:proposal}. The generalized state
@@
 
+The effect of the hyperparameters of the priors are worth a note or two, as they are complex.
+
+Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
+units as the necessary adaptation. $\nu$ encodes how stable this hyperprior is:
+higher values (e.g. 10000) all but guarantee that the hypermean will not move
+in the face of evidence; In principle, enough evidence should still move it,
+but that won't happen during an experiment. Smaller values (e.g. 1 / force\_sd,
+as is the default on the code) make the hypermean (and thus the mean) follow
+evidence more freely. Note that as more observations are accumulated, $\nu$
+becames bigger and bigger, solidifying the value of the hypermean.
+
+The hyperparameters $\alpha$ and $\beta$ are a bit more complex. Note that the
+mean of a Gamma distribution is $\beta / (\alpha \nu)$; this mean is being used
+as the standard deviation of a Gaussian by the rest of the agent (cheating),
+which makes it an important measure of uncertainty. While setting the default
+hyperparameters, the values used are $\alpha = 0.5 / \sigma_0$ and
+$\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the
+standard deviation of the force exerted by the environment (force\_sds in the
+code), which controls the initial learning rate. This makes the initial
+standard deviation equal $\sigma_0$, which makes it consistent with the
+fixed-force model in its interface. The 0.5 values ensure that uncertainty is
+large at the beginning and during the experiment is greatly reduced, but never
+to a point where it's ``visually'' too small. Changing this 0.5 would make the
+standard deviation change more quickly, making the model more or less precise
+in its predictions, independently of the volatility of the mean of the
+adaptation (via the hypermean).
+
+The baseline model defaults to different values that make it a lot more
+stable. The hyperstd of the mean is set to 10,000, which makes the mean
+entirely stable during the duration of the experiment. The values of $\alpha$
+and $\beta$ are fixed regardless of $\sigma_0$ such that the standard deviation
+is 0.001 (compare that to the size of the adaptations, around 0.0125), and the
+hyperparameters of the standar deviation are stable during the experiment.
+
 \section{Parametrization of the task and the model}
@@
 
+\subsubsection{Example: LR model with bad cues}
+Using the LR models with only two contexts (baseline, red) creates slow
+adaptation (across 400 trials or so), with fast-but-not-immediate switch back
+to baseline when provided with false cues: after the adaptation block ends, a
+baseline block ends but the cue still points to the adaptation block.
+
+Simulations from this can be seen in the function
+slow\_context\_inference\_badcues of /adaptation/code/simulations.py
+
+The agent parameters are as follows:
+
+\begin{center}
+\begin{tabularx}{0.8\textwidth}{ c | c | X}
+  Variable & Value & Description \\ \hline
+  prediction\_noise & 0 & Setting this to a nonzero value screws everything up. May be a bug. \\  
+  cue\_noise & 0.015 & Changing this by 0.001 creates slower or faster context inference \\
+  force\_sds & 0.0002 & Changing this makes the model learn more quickly or slowly. With this value, the agent barely finishes adaptation by the time the adaptation block is over. \\
+  context\_noise & 0 & Higher numbers make context inference more uncertain after evidence accumulation.
+\end{tabularx}
+\end{center}
+
+\subsubsection{Example: LR model with no cues}
+Using the LR models with only two contexts (baseline, red) creates slow
+adaptation (across 400 trials or so), with fast-but-not-immediate switch back
+to baseline when no cues are provided. After each block (A or O), a slow switch
+can be observed, which is based entirely on the internal models producing noisy
+predictions.
+
+Simulations from this can be seen in the function
+slow\_context\_inference\_badcues of /adaptation/code/simulations.py
+
+\begin{center}
+\begin{tabularx}{0.8\textwidth}{ c | c | X}
+  Variable & Value & Description \\ \hline
+  prediction\_noise & 3 & The bigger this is, the slower correct inference is. Only with no cues, though. \\  
+  cue\_noise & 0.5 & Creates uninformative cues.
+  force\_sds & 0.0002 & Changing this makes the model learn more quickly or slowly. With this value, the agent barely finishes adaptation by the time the adaptation block is over.
+  context\_noise & 0 & Higher numbers make context inference more uncertain after evidence accumulation.  
+\end{tabularx}
+\end{center}
+
 \subsection{Context inference requires a new likelihood}
@@
 
-\subsection{Parameter values to look into}
-By visual inspection of the results of simulations, these are the parameter values that seem promising:
-
-\begin{tabularx}{\textwidth}{
-  >{\hsize=.2\hsize\linewidth=\hsize}X|
-  >{\hsize=.1\hsize\linewidth=\hsize}X|
-  >{\hsize=.1\hsize\linewidth=\hsize}X|
-  l}
-% {l|{\hsize=.4\hsize}X|{\hsize=0.3\hsize}X|{\hsize=0.3\hsize}X}
-Agent & Obs noise & Cue noise & Notes \\ \hline \hline
-LR    & 1.2       & 0.04      & Slow fake-context inference. Maybe the priors coming from the previous stage are messing things up. \\ \hline
-1 & 1 & 1 & 1
-\end{tabularx}
->>>>>>> adc91ff996e45d6eaf8974cd33ea69b5d380ed07
-
-\subsection{Adaptation}
-Adaptation happens mostly very quickly for all agents.
-For the current $\Delta t = \dt$, adaptation happens within 20 to 30 time steps, although many times (depending on agent and parametrization) it happens much more quickly.
-For example, for the LR agent with prediction noise of 0.15, cue noise and context noise of 0, and assumed force variability of 0.25, within 15-20 time steps the adaptation is complete.
 \section{Conclusions}

### 2021-05-11 adaptation/code/model.py

@@ -1,7 +1,7 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/model.py
 
-# import ipdb
+import ipdb
 
 import numpy as np
 from scipy import stats
@@ -70,7 +70,7 @@ class LeftRightAgent(object):
 
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None, prediction_noise=None,
-                 reset_after_change=None, force_sds=None):
+                 reset_after_change=None, force_sds=None, max_force=None):
         """Initializes the known left and right contexts, as well as
         baseline.
 
@@ -83,6 +83,9 @@ class LeftRightAgent(object):
             self.cue_noise = cue_noise
         if angles:
             self.angles = np.array(angles)
+            self.num_contexts = len(angles)
+        else:
+            self.num_contexts = 3
         if context_noise:
             self.context_noise = context_noise
         if prediction_noise:
@@ -91,15 +94,18 @@ class LeftRightAgent(object):
             self.reset_after_change = reset_after_change
         if force_sds is not None:
             self.force_sds = force_sds
+        if max_force:
+            self.max_force = max_force
 
-        self.num_contexts = 3
         _, self.magnitudes = self.__init_contexts()
         self.magnitude_history = [self.magnitudes]
         self.decision_history = []
         self.hand_position_history = [0]
         self.hand_position = 0
         self.action_history = [0]
-        self.log_context = np.log([0.8, 0.1, 0.1])
+        exp_log_context = 0.2 / self.num_contexts * np.ones_like(self.angles)
+        exp_log_context[0] = 0.8
+        self.log_context = np.log(exp_log_context)
         self.log_context_history = [self.log_context]
         self.cue_history = [-1]
 
@@ -139,11 +145,6 @@ class LeftRightAgent(object):
         The decision is returned and also saved into self.decision_history.
 
         """
-        c_pos = self.hand_position
-        c_context = self.sample_context('mode')
-        c_force = self.magnitudes[c_context][0]
-        action_mu = -c_pos - c_force
-        # action = self.sample_norm() * self.action_sd + action_mu
         action = self._make_decision_mixed()
         if abs(action) > self.max_force:
             action = action / abs(action) * self.max_force
@@ -151,7 +152,7 @@ class LeftRightAgent(object):
         self.action_history.append(action)
         return action
 
[…]

### 2021-05-11 adaptation/code/pars.py

@@ -1,0 +1,191 @@
+# -*- coding: utf-8 -*-
+
+import numpy as np
+
+"""Parameters for the model and the task. See /notes/notes.pdf for an
+explanation of each one of them and how these values were found.
+
+All values are in the international system of units (m, g, s, N, ...).
+"""
+
+CLAMP_INDEX = 150
+
+
+def _populate(nonzeros, vector=None, size=None, element=1):
+    """Creates a vector of zeros of size --size-- and puts an --element-- in each
+    of the indices in --nonzeros--.
+
+    If --vector-- is provided, it is used instead of creating zeros.
+
+    Beware of automatic casting!
+
+    """
+    if vector is None:
+        vector = np.zeros(size, dtype=int)
+    for index in nonzeros:
+        vector[int(index)] = element
+    return vector
+
+
+def _define_contexts_base():
+    """Defines the context of each trial. This is meant to be edited by
+    hand. """
+    # baseline = np.arange(50)
+    first_adaptation = np.arange(50, 200)
+    second_adaptation = np.arange(200, 350)
+    deadaptation = np.arange(350, 400)
+    clamp = np.arange(400, 500)
+    vector = _populate(size=500, nonzeros=first_adaptation, element=1)
+    _populate(vector=vector, nonzeros=second_adaptation, element=2)
+    _populate(vector=vector, nonzeros=deadaptation, element=2)
+    _populate(vector=vector, nonzeros=clamp, element=CLAMP_INDEX)
+    return vector
+
+
+def _define_cues_base():
+    """Defines the cues of each trial. This is meant to be edited by hand. """
+    # baseline = np.arange(50)
+    first_adaptation = np.arange(50, 200)
+    second_adaptation = np.arange(200, 350)
+    deadaptation = np.arange(350, 400)
+    clamp = np.arange(400, 500)
+    vector = _populate(size=500, nonzeros=first_adaptation, element=1)
+    _populate(vector=vector, nonzeros=second_adaptation, element=2)
+    _populate(vector=vector, nonzeros=deadaptation, element=1)
+    _populate(vector=vector, nonzeros=clamp, element=1)
+    return vector
+
+
+def _define_breaks():
[…]

### 2021-05-11 adaptation/code/simulations.py

@@ -9,10 +9,12 @@ import glob
 import numpy as np
 from matplotlib import pyplot as plt
 import pandas as pd
+from scipy import stats
 
 import model
 import task_hold_hand as thh
 from pars import task as task_pars
+import pars
 
 
 def test_obsnoise(fignum=1):
@@ -80,10 +82,9 @@ def grid_sims():
     all_pars = product(*grid_elements)
     my_robots = mp.Pool()
 
-    for pars in all_pars:
+    for c_pars in all_pars:
         inputs = {label: value
-                  for label, value in zip(labels, pars)}
-        # _one_sim(inputs)
+                  for label, value in zip(labels, c_pars)}
         my_robots.apply_async(_one_sim, args=[inputs])
     my_robots.close()
     my_robots.join()
@@ -113,7 +114,8 @@ def _define_grid():
     labels = ['prediction_noise', 'cue_noise', 'context_noise', 'force_sds']
     try:  # Test labels on agent
         test_values = [value[0] for value in values]
-        agents[0](**{label: value for label, value in zip(labels, test_values)})
+        agents[0](**{label: value
+                     for label, value in zip(labels, test_values)})
     except TypeError:
         raise ValueError('Testing the labels on the agent failed. '
                          'Maybe the labels are misspelled?')
@@ -158,12 +160,7 @@ def read_data():
             else:
                 c_part = part
             panda[labels[ix_part]] = float(c_part) * np.ones(len_panda)
-        # cue_noise = float(split[3]) * np.ones(len_panda)
-        # con_noise = float(split[4]) * np.ones(len_panda)
         panda['agent'] = [split[-1][:-3]] * len_panda
-        # panda['cue_noise'] = cue_noise
-        # panda['context_noise'] = con_noise
-        # panda['agent'] = agent
         pandas.append(panda)
     pandatron = pd.concat(pandas, axis=0)
     pandatron.reset_index(inplace=True)
@@ -174,17 +171,17 @@ def read_data():
 
 def performance(pandata):
     """Calculates the performance of each agent and each parameter combination in
-    --pandata--. Performance is calculated as the sum of the absolute value of hand
-    position.
+    --pandata--. Performance is calculated as the sum of the absolute value of
+    hand position.
 
     Returns
[…]

### 2021-05-11 adaptation/code/task_hold_hand.py

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/task_hold_hand.py
-# import ipdb
+import ipdb
 from datetime import datetime
 
 import numpy as np
@@ -25,6 +25,8 @@ def run(agent=None, save=False, filename=None, pars=None):
     outs = []
     hand_position = 0
     for c_trial, do_break in enumerate(pars['breaks']):
+        # if c_trial > 520:
+        #     ipdb.set_trace()
         if do_break:
             agent.reset()
             hand_position = 0
@@ -104,6 +106,8 @@ def sample_observation(hand_position, pars):
     except TypeError:
         pass
     scale = pars['obs_noise']
+    if scale == 0:
+        return hand_position
     loc = hand_position
     distri = stats.norm(loc=loc, scale=scale)
     return distri.rvs()
@@ -125,7 +129,7 @@ def join_pandas(pandata, pandagent):
     into one big panda, aligned by trial number. Returns the panda.
 
     """
-    pandagent = pandagent.drop(['action', 'hand'], axis=1)
+    pandagent = pandagent.drop(['action'], axis=1)
     pandagent = pandagent.iloc[pandagent.index != -1]
     return pd.concat([pandata, pandagent], axis=1)
 

### 2021-05-11 code/adaptation/pars.py

@@ -1,100 +1,0 @@
-# -*- coding: utf-8 -*-
-
-import numpy as np
-
-"""Parameters for the model and the task. See /notes/notes.pdf for an
-explanation of each one of them and how these values were found.
-
-All values are in the international system of units (m, g, s, N, ...).
-"""
-
-CLAMP_INDEX = 150
-
-
-def _populate(nonzeros, vector=None, size=None, element=1):
-    """Creates a vector of zeros of size --size-- and puts an --element-- in each
-    of the indices in --nonzeros--.
-
-    If --vector-- is provided, it is used instead of creating zeros.
-
-    Beware of automatic casting!
-
-    """
-    if vector is None:
-        vector = np.zeros(size, dtype=int)
-    for index in nonzeros:
-        vector[int(index)] = element
-    return vector
-
-
-def _define_contexts_base():
-    """Defines the context of each trial. This is meant to be edited by
-    hand. """
-    # baseline = np.arange(50)
-    first_adaptation = np.arange(50, 200)
-    second_adaptation = np.arange(200, 350)
-    deadaptation = np.arange(350, 400)
-    clamp = np.arange(400, 500)
-    vector = _populate(size=500, nonzeros=first_adaptation, element=1)
-    _populate(vector=vector, nonzeros=second_adaptation, element=2)
-    _populate(vector=vector, nonzeros=deadaptation, element=2)
-    _populate(vector=vector, nonzeros=clamp, element=CLAMP_INDEX)
-    return vector
-
-
-def _define_cues_base():
-    """Defines the cues of each trial. This is meant to be edited by hand. """
-    # baseline = np.arange(50)
-    first_adaptation = np.arange(50, 200)
-    second_adaptation = np.arange(200, 350)
-    deadaptation = np.arange(350, 400)
-    clamp = np.arange(400, 500)
-    vector = _populate(size=500, nonzeros=first_adaptation, element=1)
-    _populate(vector=vector, nonzeros=second_adaptation, element=2)
-    _populate(vector=vector, nonzeros=deadaptation, element=1)
-    _populate(vector=vector, nonzeros=clamp, element=1)
-    return vector
-
-
-def _define_breaks():
-    """Defines the agent breaks for the entire session. This is meant to be edited
-    by hand."""
-    breaks = [50, 200, 350, 400]
[…]

### 2021-06-28 adaptation/code/model.py
Do not even know what this is...

@@ -68,6 +68,8 @@ class LeftRightAgent(object):
 
     sample_norm = stats.norm().rvs
 
+    all_learn = False
+
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None, prediction_noise=None,
                  reset_after_change=None, force_sds=None, max_force=None):
@@ -544,26 +546,40 @@ class LRMeanSD(LeftRightAgent):
         \alpha)$, where the hyperparameters are $\mu, \nu, \alpha, \beta$.
 
         """
-        ix_context = self.sample_context()
+        mag_hypers = self.mag_hypers.copy()
+        magnitudes = np.array(self.magnitudes)
+        if not self.all_learn:
+            ix_context = self.sample_context()
+            post_mag = self._update_magnitude(ix_context)
+            mag_hypers[ix_context] = post_mag
+            magnitudes[ix_context][0] = post_mag[0]
+            magnitudes[ix_context][1] = post_mag[3] / (post_mag[2] - 0.5) / post_mag[1]
+        else:
+            pos_con = np.exp(self.log_context)
+            for ix_context in range(self.num_contexts):
+                post_mag = self._update_magnitude(ix_context,
+                                                  pos_con[ix_context])
+                mag_hypers[ix_context] = post_mag
+                magnitudes[ix_context][0] = post_mag[0]
+                magnitudes[ix_context][1] = post_mag[3] / (post_mag[2] - 0.5) / post_mag[1]
+        self.mag_hypers = mag_hypers
+        self.mag_hypers_history.append(mag_hypers)
+        # post_mag[3] / post_mag[2] / post_mag[1]
+        self.magnitudes = magnitudes
+        self.magnitude_history.append(magnitudes)
+
+    def _update_magnitude(self, ix_context, data_weight=1):
+        """Updates the magnitude of the given context. """
         p_con = np.argmax(self.log_context_history[-2])  # called after infer_context        
         mu_l, sd_l = self.predict_outcome(previous_context=p_con)[1][ix_context]
         mag_pars = self.mag_hypers[ix_context]
         mu_l = mag_pars[0] + self.hand_position - mu_l
-        post_mag = [(mag_pars[0] * mag_pars[1] + mu_l) / (mag_pars[1] + 1),
-                    mag_pars[1] + 1,
-                    mag_pars[2] + 0.5,
[…]

### 2021-06-28 adaptation/code/simulations.py

@@ -433,6 +433,10 @@ def baseline_bias(fignum=9):
     """This reproduces the experiments in Davidson_Scaling_2004, plotting the
     results for all their groups and experiments in one big pot of lovin.
 
+    Note: for now, this does not work. The model would need some additions to
+    handle the creation of new models and it goes beyond the scope of the
+    paper.
+
     """
     fig, axes = plt.subplots(2, 4, gridspec_kw={'height_ratios': (3, 1)},
                              num=fignum, clear=True)
@@ -457,6 +461,7 @@ def baseline_bias(fignum=9):
                            cue_noise=1 / 3, context_noise=0.0,
                            force_sds=0.02 * np.ones(3),
                            max_force=1, hyper_sd=0.005)
+    agent.all_learn = True
     pars_task = pars.task_davidson_2_1
     sim_and_plot(agent, pars_task, axes=[axes[0, 2], axes[1, 2]])
 
@@ -471,31 +476,6 @@ def baseline_bias(fignum=9):
     axes[0, 1].get_legend().remove()
     axes[0, 2].get_legend().remove()
     axes[0, 3].get_legend().remove()
-
-
-def test():
-    agent = model.LRMeanSD(angles=[0, 0, 0],
-                           prediction_noise=1,
-                           cue_noise=1 / 3, context_noise=0.0,
-                           force_sds=0.02 * np.ones(3),
-                           max_force=1, hyper_sd=0.5)
-    pars_task = pars.task_davidson_2_1
-    pandota, agent = sim_and_plot(agent, pars_task, return_data=True)
-
-    manorm = stats.norm.logpdf
-    mag_hist = np.array(agent.magnitude_history)
-    probs = manorm(0.005, loc=mag_hist[..., 0], scale=mag_hist[:, :, 1])
-    probs = probs[1:, ...]
-    probs = np.exp((probs.T - probs.max(axis=1)).T)
-    pandota[['probs_0', 'probs_1', 'probs_2']] = probs
-    return pandota, agent
-
-
-def normal_gamma_logpdf(x, mu, nu, alpha, beta):
-    """Returns the logprobability of x given the pars.
-
-    """
-    constant = beta ** alpha * np.sqrt(nu)
 
 
 if __name__ == '__main__':

### 2021-06-28 notes/notes.tex
Added bullet points for results

@@
 \end{tabularx}
+
+\section{Results}
+
+\begin{enumerate}
+\item Quick de-adaptation/savings: gradual evidence accumulation during context inference will make the switching between internal models not be immediate.
+\item Learning: The speed in which the internal models are adjusted given errors depends on how certain the inference is. More uncertain context inference will lead to slower learning.
+\item Action selection: Action selection is sampled from the existing (relevant) internal models, weighted by their posterior probability during this trial.
+\item Cue reliability adjustments: As more bad-cue trials are observed, the agent will lower the estimated reliability of the cue.
+\item Error-clamp tappering off: Hundreds of error-clamp trials make adaptation seem to slowly disappear, but it never fully goes away.
+\item Error-clamp lag: For a while, the tappering off does not begin, depending on how clear it is that error-clamp trials have started.
+\end{enumerate}
+
 \end{chapter}
@@
 
-\cite{Smith_Interacting_2006} speak of error-clamp weirdness and presents a
-model to explain it. Their data is limited, however, and might not be specially
-useful for me.
-
-\cite{Ethier_Spontaneous_2008} have error-clamp weirdness and their
-within-block pauses might be useful for my interpretations.
-\end{chapter}
+\cite{Smith_Interacting_2006} present a typical mechanical arm task with a
+O-A-(-A)-E structure, with random E trials thrown in throughout all. Half of
+the trials were with outward movements and half with the inward movement, with
+the latter almost being E trials and not used for analysis. The experiment has
+no cues, but it was a design in which blocks of 60 trials were used, presumably
+with pauses in the middle, though they do not explicitly say this. The authors
+find that during the E block at the end, participants' responses show
+spontaneous recovery, where they seem to be moving towards A for a few trials,
+but never quite reaching it, and then taper off to
+zero. \cite{Smith_Interacting_2006} only show data aggregated over all
+participants, so it's difficult to extract too much information, but their
+figure 3D suggests that some participants assume -A during E, while most assume
+A.
+
+\cite{Ethier_Spontaneous_2008} present a saccade study (movements of 15° with
+adaptation of +-5°) in which participants go through E-A-(-A)-E blocks, each
+divided into sets of 60 trials (some blocks have different sets). Between each
+set (and block), a pause of 30s happened. There are no cues, although the size
+of the adaptations is big. It is important to note that the way the error clamp
+works in this paradigm probably gives no indication to the participant that an
+error has occured; contrast this with mechanical-arm paradigms, or
+visuorotation reaching paradigms, in which proprioceptive information gives
+feedback to the participant regardless of rewards.
+
+Figure 2 in \cite{Ethier_Spontaneous_2008} contains all the results relevant to
+this section. First of all, it is clear from the first E block that spontaneous
+recovery and slow tapering off do not exist if no adaptation has been learned;
+instead, behavior during these trials is the same as during baseline. Second,
+spontaneous recovery happens always in the direction of the first adaptation A,
+and never in the direction of -A, which is the block right before the last E
+block. The context-inference-based model explains these findings by noting that
+the -A block is very short, ending when participants reach 0 adaptation (during
+``de-adaptation'') or a little bit beyond zero (towards -A). This is important
+because, since adaptation was not finished, during the first trial of the E
+block the participant expected an error, but no error was observed. This leads
+the participant to think that a change in the context has occured, and context
+inference kicks in, taking the lack of an error as evidence in favor of either
+baseline or the previous adaptation A. Given that the action by the participant
+on this trial is more consistent with baseline than with A, why would A play
+any role in future responses? The answer lies in the structure of the
+experiment, as no baseline trials were ever performed in front of the screen,
+which makes baseline have very low priors.
 
@@
 \cite{Vaswani_Decay_2013} explores in detail the motor decay observed during EC
-trials after adaptation trials. They conclude, among other things, that the
-motor adaptation decays across as many as hundreds of trials, but never goes
-back to zero. Additionally, there is a lag before decay starts (all their
-experiments end in A-E, so no baseline before), and it depends on the
-participant, as well as how noticeable the transition to error-clamp trials
-is. They have some clever experiments.
+trials after adaptation trials in a viscous field paradigm. They conclude,
+among other things, that the motor adaptation decays across as many as hundreds
+of trials, but never goes back to zero. Additionally, there is a lag before
+decay starts (all their experiments end in A-E, so no baseline before), and it
+depends on the participant, as well as how noticeable the transition to
+error-clamp trials is. They have some clever experiments.
+
+More precisely, \cite{Vaswani_Decay_2013} have an experiment 1 in which
+participants learn an adaptation and then go through hundreds of trials of
+error clamp. Participants are divided into four groups: groups 1.1 and 1.4
+learn adaptation A and -A, respectively, and then go to EC trials. Group 1.2
+performs the task in baseline and then in A. Group 1.3 performs the task in
+-A/2, then A and then EC. In their results, all groups show ``adaptation'' that
+starts high (consistent with A) and, after some lag, goes down to an adaptation
+of almost zero, but decidedly positive. Group 1.3, in particular, decays faster
+than the other groups but does not seem to get closer to zero even at the end
+of the experiment. Group 1.4 is just like 1.1 and 1.2, but with its sign
+flipped.
+
+Our simulations can easily show these results to be a direct consequence of
+context inference, which starts by assuming that no change has happened after
+EC starts (thus staying with A), but eventually the evidence points towards a
+change. However, because no model explains EC fully, actions are taken by doing
+a weighted average of all available models, where the weights are given by the
+posterior over contexts. This explains both the lag, the fact that EC
+adaptation never reaches zero and that group 1.3 adapted faster, as the -A/2
+model pulls the average down faster than baseline. The fact that no difference
+is observed between 1.1 and 1.2 is due to the fact that the baseline model
+exists, regardless of whether participants trained on it or not. Just like the
+results from \cite{Davidson_Scaling_2004}, these results point toward motor
+output being a weighted average of all the available models, with weights being
+the posterior over contexts.
+
+The conclusions of experiments 2 and 3 by \cite{Vaswani_Decay_2013} reinforce
+the context-inference-based account: the more difficult it is to notice that EC
+has started, the less decay participants show, to the extreme of experiment 3
+in which most participants showed no decay at all.
+
+\cite{Pekny_Protection_2011} present a series of experiments in which they set
+out to determine whether A-B-E paradigms make the brain forget about A when
+learning B, or whether the brain protects A. The authors mention that there is
+a lot of literature that has used savings as a way to measure the protection of
+A and found that A is not being protected; however, the authors argue that
+classical conditioning experiments instead use spontaneous recovery during E
+trials to measure protection of memories during extinction, and the literature
+in that field has demonstrated that memories are indeed
+protected. \cite{Pekny_Protection_2011} present a series of experiments with
+mechanical arms to show the protection of memories. Their results, specifically
+those in their Figures 3 and 5, strongly support not only the
+context-inference-based motor learner, but also the idea that action selection
+happens through a weighted average, much like the results from
+\cite{Herzfeld_memory_2014}.
+
+In figure 3A, one group (red) went through O-A-b (where lowercase ``b'' means
+very few trials, used as extinction), and another group (black) went through
+O-B-A-b. When comparing the force compensation between those two groups, one
+can see that the introduction of B in the black group has the effect of pulling
+their response during E closer to baseline, as compared with the red
+group. This is because B has, for this group, been introduced as a potential
+context of importance equal that of A (although a longer time ago, as B is
+presented before A). This increased the prior probability of context B during
+the E trials in the minds of participants, which draws the weighted average
+closer to baseline, as $B = -A$. The same interpretation can be given to all of
+Figure 3 and Figure 5 as well.
+
+Additionally, \cite{Pekny_Protection_2011} present an experiment in which
+changes in context are presented gradually (see their figure 1 for
+details). They argue that because the changes are presented gradually,
+participants do not get a strong prediction error that can guide context
+inference towards the correct context. However, there is a caveat they do not
+discuss: in an experiment where context switches slowly, small changes become
+big evidence. For this reason, their Figure 5B has the same effect as that seen
+in Figure 3.
 
@@
 
+\end{chapter}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   prediction\_noise & 3 & The bigger this is, the slower correct inference is. Only with no cues, though. \\  
-  cue\_noise & 0.5 & Creates uninformative cues.
-  force\_sds & 0.0002 & Changing this makes the model learn more quickly or slowly. With this value, the agent barely finishes adaptation by the time the adaptation block is over.
+  cue\_noise & 0.5 & Creates uninformative cues. \\
+  force\_sds & 0.0002 & Changing this makes the model learn more quickly or slowly. With this value, the agent barely finishes adaptation by the time the adaptation block is over. \\
   context\_noise & 0 & Higher numbers make context inference more uncertain after evidence accumulation.  
@@
 
+\subsection{\cite{Davidson_Scaling_2004} results}
+Running simulations I realized that the model, as is, will not be able to
+reproduce the results from \cite{Davidson_Scaling_2004}. To do so, the model
+would need to be modified to create new models online, but only one at a
+time. A model would be created and kept at an adaptation of zero with very high
+standard deviation, and this model would very naturally learn any adaptation
+that is not explained by any other model.
+
+The biggest problem is in reproducing experiment 2, group 1 of
+\cite{Davidson_Scaling_2004}, in which the adaptations that will be presented
+are A and 3A. In this case, at the beginning there should be an empty model
+that learns A. Once this model is far away from baseline, a new empty model
+should be created (and kept ``on call''), which eventually would pick up on
+3A. If the two empty models are created at the same time, both will learn A,
+and later only one of them (picked randomly) will learn 3A, while the other
+stays at A; this is fine for simulations but difficult to explain from a
+mechanistic point of view: why would the brain train two models to do the same
+thing?
+
+Possible solutions to this would be: (1) as said before, that only one empty
+model exists at any time. (2) any model would have an empty counterpart that
+can always learn faster. (3) Context inference would be able to pick up on low
+model evidence (e.g. when no internal model explains the observations at all)
+and trigger the creation of a new model, either always as a copy of baseline or
+as a copy of the closest to the observations.
+
+This is work for future studies.
+
 \section{Conclusions}

### 2021-06-28 timetable.tex
Messssss

@@
 
-\section*{Time plan for the motor adaptation project.}
-
-The previous time table was interrupted by the possibility of the new project. Now that this new project has taken shape, this new time table applies:
-
-\noindent
-\begin{tabularx}{\textwidth}{
-  l|
-  >{\hsize=.4\hsize\linewidth=\hsize}X|
-  >{\hsize=.2\hsize\linewidth=\hsize}X|
-  >{\hsize=.2\hsize\linewidth=\hsize}X}
-End Date & Result & A priori comments & A posteriori comments \\ \hline \hline
-30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & \\ \hline
-07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
-14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
---- & Start writing the paper. & & \\ \hline
-\end{tabularx}
-
 \end{document}

### 2021-06-28 timetable.tex
Messy mergezz
Updated timetable.tex with results

@@
 \section*{Brief project description}
-The idea of the project is the following: We will create an agent that learns to control an arm in environment A (say, the real world carrying nothing), and learns how to control the arm in environment B (say, with high winds).
-The agent then learns to recognize environments A and B and to deploy the right controller for each environment depending on this recognition.
+In this chapter, I describe the everything related to the paper that should come
+out of this project, including modeling and experimental findings, as well as
+the direction of the paper and things needed to get there.
 
-We will argue that this is how humans do it and present evidence (based on literature research) for the predictions this model makes, as well as try to make new predictions.
+\section{Project description}
+The idea of the project is the following: We will create an agent that learns to
+control an arm in environment A (say, the real world carrying nothing), and
+learns how to control the arm in environment B (say, with high winds). The agent
+then learns to recognize environments A and B and to deploy the right controller
+for each environment depending on this recognition.
 
-After this methods paper is done, model validation can be carried out in future studies.
+We will argue that this is how humans do it and present evidence (based on
+literature research) for the predictions this model makes, as well as try to
+make new predictions.
 
-\section*{Time plan for the arm movement project.}
+After this methods paper is done, model validation can be carried out in future
+studies.
+
+\section{Planning}
+Here I enumerate the things I need to learn and do in order to get this project
+moving and to finish it.
 
@@
 
+The previous time table was interrupted by the possibility of the new project. Now that this new project has taken shape, this new time table applies:
+
+\begin{tabularx}{\textwidth}{
+  l|
+  >{\hsize=0.4\hsize\linewidth=\hsize}X|
+  >{\hsize=0.2\hsize\linewidth=\hsize}X|
+  >{\hsize=0.2\hsize\linewidth=\hsize}X}
+End Date & Result & A priori comments & A posteriori comments \\ \hline \hline
+30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & I found a lot of evidence. Moving on.\\ \hline
+07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
+14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
+--- & Start writing the paper. & & \\ \hline
+
+\end{tabularx}
+\section{Results}
+
+\begin{enumerate}
+\item Quick de-adaptation/savings: gradual evidence accumulation during context inference will make the switching between internal models not be immediate.
+\item Learning: The speed in which the internal models are adjusted given errors depends on how certain the inference is. More uncertain context inference will lead to slower learning.
+\item Action selection: Action selection is sampled from the existing (relevant) internal models, weighted by their posterior probability during this trial.
+\item Cue reliability adjustments: As more bad-cue trials are observed, the agent will lower the estimated reliability of the cue.
+\item Error-clamp tappering off: Hundreds of error-clamp trials make adaptation seem to slowly disappear, but it never fully goes away.
+\item Error-clamp lag: For a while, the tappering off does not begin, depending on how clear it is that error-clamp trials have started.
+\end{enumerate}
+
 \end{document}

### 2021-06-28 notes/notes.tex
Updated timetable.tex with new timings

@@
 30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & I found a lot of evidence. Moving on.\\ \hline
-07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
-14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
---- & Start writing the paper. & & \\ \hline
+07.05.2021 & Finding model pars to explain observations & This is mostly already done &  Parameters found\\\hline
+14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it.
+If there is, an extra week should be allocated to simulations.
+& I found a lot of evidence for context-inference-based motor adaptation.
+It will be incorporated into the paper.
+\\\hline
+23.06.2021 & Looking into single-cell recordings. & Perhaps I can find evidence here. & Evidence seems dubiuous, but might be worth including in Discussion \\
+--- & Start writing the paper. & & This was delayed by the defense and dysco. Restarted around 24.06.2021 \\ \hline
 
@@
 \section{Results}
-
+Possible results to include in the paper:
 \begin{enumerate}

### 2021-06-28 timetable.tex
Updated timetable.tex with new timings

@@
 \begin{document}
-\section*{Brief project description}
 In this chapter, I describe the everything related to the paper that should come
@@
 30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & I found a lot of evidence. Moving on.\\ \hline
-07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
-14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
---- & Start writing the paper. & & \\ \hline
+07.05.2021 & Finding model pars to explain observations & This is mostly already done &  Parameters found\\\hline
+14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it.
+If there is, an extra week should be allocated to simulations.
+& I found a lot of evidence for context-inference-based motor adaptation.
+It will be incorporated into the paper.
+\\\hline
+23.06.2021 & Looking into single-cell recordings. & Perhaps I can find evidence here. & Evidence seems dubiuous, but might be worth including in Discussion \\ \hline
+--- & Start writing the paper. & & This was delayed by the defense and dysco. Restarted around 24.06.2021 \\ \hline
 
 \end{tabularx}
+
 \section{Results}
-
+Possible results to include in the paper:
 \begin{enumerate}

### 2021-06-28 notes/notes.tex
Fixed typos

@@
 \\\hline
-23.06.2021 & Looking into single-cell recordings. & Perhaps I can find evidence here. & Evidence seems dubiuous, but might be worth including in Discussion \\
+23.06.2021 & Looking into single-cell recordings. & Perhaps I can find evidence here. & Evidence seems dubious, but might be worth including in Discussion \\
 --- & Start writing the paper. & & This was delayed by the defense and dysco. Restarted around 24.06.2021 \\ \hline
@@
 \item Cue reliability adjustments: As more bad-cue trials are observed, the agent will lower the estimated reliability of the cue.
-\item Error-clamp tappering off: Hundreds of error-clamp trials make adaptation seem to slowly disappear, but it never fully goes away.
-\item Error-clamp lag: For a while, the tappering off does not begin, depending on how clear it is that error-clamp trials have started.
+\item Error-clamp tapering off: Hundreds of error-clamp trials make adaptation seem to slowly disappear, but it never fully goes away.
+\item Error-clamp lag: For a while, the tapering off does not begin, depending on how clear it is that error-clamp trials have started.
 \end{enumerate}

### 2021-06-28 timetable.tex
Fixed typos

@@
 \\\hline
-23.06.2021 & Looking into single-cell recordings. & Perhaps I can find evidence here. & Evidence seems dubiuous, but might be worth including in Discussion \\ \hline
+23.06.2021 & Looking into single-cell recordings. & Perhaps I can find evidence here. & Evidence seems dubious, but might be worth including in Discussion \\ \hline
 --- & Start writing the paper. & & This was delayed by the defense and dysco. Restarted around 24.06.2021 \\ \hline

### 2021-07-01 article/article.tex
First commit of paper

\documentclass[a4paper,doc,floatsintext,natbib]{apa6}
% \documentclass{article, a4paper}
\usepackage[font=small]{caption}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage[utf8]{inputenc}
\usepackage{nameref}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{soul}

% Remember to start reftex-mode

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=black,
  urlcolor=black
  }

\setlength{\parskip}{1em}
\def \fref #1{Figure \ref{#1}}     % Reference figures
\def \tref #1{Table \ref{#1}}      % Reference tables
\def \eref #1{Equation \ref{#1}}   % Reference equations
\def \sref #1{Section '\nameref{#1}'}    % Reference sections
\def \supmat {the Supplementary Materials}

% For revision
\DeclareRobustCommand{\newcontent}[1]{#1}

\title{The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery}
\shorttitle{Context-inference-dependent motor learning}
\author[1]{Cuevas Rivera, Darío}
\author[1]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
\affiliation{~}

\begin{document}

\maketitle

\abstract{Abstract}
% \tableofcontents

\section{Introduction}
yeah

\section{Results}
In this work, we present a motor adaptation model in which the agent (e.g. a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation. Several models exist for error-based motor adaptation, but we expand on the existing models by adding an explicit component of context inference which guides the selection of the adequate internal model, its updating and the sampling of actions.

In what follows, we introduce the different parts of the model and explain how they work together. In addition, we will highlight experimental findings in motor adaptation from previous works for which our model provides explanations that are otherwise unavailable.

\subsection{The model}
The model we present has three main components: (1) a context-inference component; (2) a motor adaptation component; (3) an action selection component. Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.

Based on the Bayesian brain hypothesis, we chose to cast both context inference and motor adaptation as Bayesian inference, in effect using a unified mechanism for the preparation of motor commands in new environments.

We now describe the three components of the model separately. We begin by describing the motor adaptation part for clarity, as the definition of ``context'' (used in context inference) is better described in terms of motor adaptation.

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to learn from observed errors. For example, we can learn to perform a reaching movement in front of us --a movement that any person has been practicing all their lives-- underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals. While a number of models for motor adaptation exist, we chose to make use of an exact Bayesian learner. As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also does away with the need for explicit multiple time scales of learning. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, let us return to reaching movements. Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \beta)
\end{equation}
where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the reach movement), $s$ is the current state (e.g. the current position of the hand), $c$ is a motor command and $\beta$ are the parameters of the forward model $f(\cdot)$. Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements.

As a day progresses, and especially if the day included a lot of physical activity of the arm, muscle exhaustion leads to a different response of the muscles in the arm to motor commands. However, we are still able to produce accurate reach movements. This happens through adaptation, where the observed error signals of a moment are used to temporarily update the parameters $\beta$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that:
\begin{equation}
q(\beta | \vartheta, s, c) \propto p(\vartheta | s, c, \beta)p(\beta)
\end{equation}
where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$. $p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$.

In this account of motor adaptation, the prediction error signal takes the form of an observed outcome $\vartheta$ that differs from that of the most likely outcome predicted by the forward model. The amount of adaptation that happens after observing an error will depend not only on the size of the error (how far away from the most likely outcome) it is, but also of how precise the existing estimate of $\beta$ is. We will discuss this further below.

Not only is the internal model for everyday movements not fixed (i.e. adaptation is possible), it is also not unique. Multiple internal models are used and updated, depending on the task at hand. For example, a person might have an internal model for baseline reaching movements (e.g. with empty hands) $f_B$, and a different one for reaching movements while holding a heavy object, $f_H$. A key aspect of motor adaptation is to select the relevant model to update, should a prediction error arise.

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of hte environment (e.g. high winds when walking), as well as the task at hand. Previous models have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

In this work, we focus on the idea that the context can be inferred by integrating several sources of information, an idea with great intuitive appeal. Through this idea, we explain many previously-unexplained phenomena from different experiments with humans.

Identifying any possible context with a categorical variable $\zeta$, the identification of a context can be done through Bayesian inference:
\begin{equation}
q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
\end{equation}
where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from instant to instant, and (2) an overall estimation of observing any one context, if the context did change. Below, we discuss how these two components effect different experimentally observed phenomena.

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely intependent from them. For this section, the most important component of $\vartheta$ is contextual information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the contextual information can take the form of a visual cue [EGS], the place where the task must be carried out [EGS] or even the start of a new block of trials [EGS]. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.

\subsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model. In other words, a distribution is created as follows:
\begin{equation}
p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta)
\end{equation}
From this distribution, the current action (motor command $c_t$) can be sampled and carried out.

\section{Experimental results}
In this section, we go through a number of experimentally-observed phenomena and show how imperfect context inference, as done in our model, provides a parsimonious explanation for all of them.

In what follows, we show numerical results obtained the model outlined above. The full mathematical implementation of the model can be found in XXX.

Before discussing these phenomena, let us introduce terminology. As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force. Depending on the trial (and experiment), the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise. Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point (approximately) in opposite directions. Finally, we represent error-clamp trials with the letter E.

With this terminology, a typical experiment [EG] would have a block structure of O-A-B-E (or O-A-(-A)-E), which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces, and finally a block with error-clamp trials.

\subsection{Savings and quick de-adaptation}
Savings refers to the ability of humans and other animals to remember a previously-learned adaptation, and apply it without having to re-learn it. Savings is almost-universally observed in human participants [A THOUSAND CITATIONS; c.f. that one paper with single-internal-model], but animal experiments have shown mixed results [EG].

In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of an apparent adaptation rate might higher than that observed during the first A block.

On the other hand, quick de-adaptation refers to the observation that in an O-A-O paradigm, the adaptation from O to A is much slower than that from A back to O.

Importantly, neither savings nor quick de-adaptation are immediate, but instead manifest as a much increased adaptation speed. In this section, we will show that both phenomena are the two sides of a coin, and the speed at which this quick adaptation/de-adaptation happens is related to the availability and reliability of contextual cues, as well as to the quality of the prediction errors.

We can categorize experiments performed with human participants by looking at the amount of contextual information that is available. In some experiments [EGssss], the context is clearly cued to the participant, either in the form of visual cues on in changes to the location of the task. We call these cued-context experiments. In other experiments, partial information is available to participants [EGsss]; we refer to these as partially-cued experiments. While there are no experiments in which no contextual information is available, there are some experiments which contain blocks in which this is the case and participants have no way of inferring the current context. Abusing our notation, we call these experiments (or rather, these sections of the experiments) uncued experiments. In this section, we focus on the two former categories, while the latter is left for the next sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. Additionally, the sudden appearance of large motor errors can itself be a cue for contextual change.

To see the effects of context inference on savings and quick de-adaptation, we contrast the two former types of experiments.

In FIGURE XXX, data from representative experiments from each category are shown. It can be seen that in cued-context experiments, savings take the form of an immediate switch from the adaptation at trial $t$ to that in trial $t-1$, to the level that was learned before. In contrast, in partially-cued experiments, the switch is fast (compared to the initial adaptation), but not immediate. Our model explains this difference via context inference: in the cued-context experiments, context inference relies on the very reliable information of visual cues and the change in context is identified immediately; because of this, the model can re-activate the forward model adequate for the context from the first trial after the change occured. In contrast, in partially-cued experiments, the model needs to accumulate evidence in favor of the correct context throughout a number of trials, and the actions selected by the model are combinations of the actions that correspond to all the possible contexts (we discuss this in the following sections).

In FIGURE XXX, we show the same effects described above, but in the frame of quick de-adaptation. The explanation for these effects given by the model is the same for savings and quick de-adaptation, namely that the participant slowly accumulates evidence in favor of a switch in context.

\section{The effects on the rate of adaptation}
In our model, adaptation is gated by the uncertainty on the current context (see Methods). More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation, given a fixed size of prediction error.

The most direct evidence for the effects of context inference on the rate of adaptation come from the experiments by \cite{Herzfeld_memory_2014}. The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between O and A (and vice versa) change the speed at which adaptation occurs.

\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between two possibilities: no force (baseline O) and some force (context A). They divided participants into three groups for which the transition from one trial to the next were very uncommon, somewhat common and almost every trial. Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred when a motor error is observed.

\cite{Herzfeld_memory_2014} found that the volatility of the environment affects learning. The more volatile the environment, the less participants can adapt their motor responses. In FIGURE XXX, we show results adapted from \cite{Herzfeld_memory_2014} alongside results obtained with our model. In our model, learning is gated by context inference: the higher the uncertainty on the context in which an error occured, the lower the effect of this error on future trials (i.e. adaptation).

To account for volatility, the model includes a term in context inference which assumes a transition from the previous context to all other possible contexts (see Methods). Effectively, this transition term gives the model uncertainty on the current context, regardless of observations. The higher the volatility, the higher the uncertainty.

\section{Action selection}
As with learning, our model selects actions (motor commands) based on context inference. Is the identity of the current context is known, the forward model for this context is used to select the current action. However, if some uncertainty exists, the selected action is influenced by all the possible current contexts, with a weight directly related to how likely each one of those contexts is.

\section{Behavior in error-clamp blocks}

\section{Discussion}
\begin{enumerate}
\item The effect of pauses?
\end{enumerate}

\bibliography{../MyLibrary}

\end{document}

### 2021-07-01 adaptation/code/figures.py

@@ -1,0 +1,18 @@
+# -*- coding: utf-8 -*-
+# .adaptation/code/figures.py
+
+"""Figures for the context inference motor adaptation paper. The default
+parameter --fignum-- for each function defined here determines which figure it
+is in the paper. Those that do not have one are not paper figures, but
+auxiliary functions."""
+
+
+import numpy as np
+from matplotlib import pyplot as plt
+
+import simulations as sims
+import model
+import task_hold_hand as thh
+import pars
+
+

### 2021-07-01 article/article.tex
Added text on results. No figures yet

@@
 where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
-$p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$.
+$p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta$ is the previous estimate of $\beta$.
 
@@
 In experimental settings, the contextual information can take the form of a visual cue [EGS], the place where the task must be carried out [EGS] or even the start of a new block of trials [EGS].
-All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.
+All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1
+}, c_{t-1})$ and are integrated together to identify a context.
 
@@
 Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point (approximately) in opposite directions.
+We represent contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
+clockwise), but the second one has half the magnitude.
 Finally, we represent error-clamp trials with the letter E.
@@
 Savings is almost-universally observed in human participants [A THOUSAND CITATIONS; c.f.
-that one paper with single-internal-model], but animal experiments have shown mixed results [EG].
+that one paper with single-internal-model], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}.
 
@@
 
+While we focused on \cite{Herzfeld_memory_2014}, evidence for this can be seen in other studies, which we discuss in the Discussion section.
+
 \section{Action selection}
 As with learning, our model selects actions (motor commands) based on context inference.
-Is the identity of the current context is known, the forward model for this context is used to select the current action.
+If the identity of the current context is known, the forward model for this context is used to select the current action.
 However, if some uncertainty exists, the selected action is influenced by all the possible current contexts, with a weight directly related to how likely each one of those contexts is.
 
-\section{Behavior in error-clamp blocks}
+Experimental evidence supporting this view can be found in experiments with context switching.
+For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group (where -A refers to an adaptation in the opposite direction, but with the same magnitude), with a block sequence A-(-A)-A-(-A), keeping the size of the necessary adaptation equal across both groups.
+After A and 3A in one group, A and -A in the other, had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
+The authors interpreted this as evidence that a new adaptation happens faster if it is in the same direction as the current adaptation (e.g.
+both counter-clockwise), and slower if they are in the opposite direction (e.g.
+clockwise to counter-clockwise).
+
+Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context.
+As discussed above, action selection in our model is done via a weighted avarage:
+\begin{equation}
+p(a_t | s_t) \propto \displaystyle \sum_{m \in M} p(m | s_t, s_{t-1}, a_{t-1}) p(a_t | m)
+\end{equation}
+In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$.
+When a new block of trials starts (e.g.
+in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out.
+Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force).
+In the first group, this initial pull towards zero accelerates the transition towards A, as $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.
+
+We simulated data with our model fitting the experimental structure by \cite{Davidson_Scaling_2004}.
+We show the results in FIGURE XXX, alongside the experimental results from \cite{Davidson_Scaling_2004}.
+It can be seen that the model provides a good qualitative fit to the data.
+
+The model additionally predicts that the difference in observed switching speed between 3A-A and (-A)-A would disappear in an experiment in which the adaptations do not cross the baseline, i.e.
+all are positive or all are negative, or in an experiments in which the contexts are clearly cued.
+
+Further evidence can be found during error-clamp blocks in several experiments.
+We discuss these experiments in the following section.
+Additionally, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion section.
+
+\subsection{Action selection in error-clamp blocks}
+During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials.
+However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment.
+
+In this section, we show how our model can explain these different parameters of behavior by changing the way contextual cues mislead participants' context inference, which in turn influences action selection.
+
+\cite{Vaswani_Decay_2013} explored in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
+The authors found that during an E block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous trial.
+After that, the exerted force slowly dropped towards zero throughout tens of trials, but never reaching values around zero.
+Participants were divided into four groups, each of which going through a different block sequence: (1) A-E, (2) O-A-E, (3) (-A/2)-A-E, and (4) (-A)-E.
+No pauses were made during the experiment nor were there any contextual cues, so transitions between blocks were not signaled to participants.
+However, the mechanical arm used in the experiment provided force-feedback to participants.
+
+In FIGURE XXX, we show the data from \cite{Vaswani_Decay_2013}, where it can be seen that the responses during the E block trials starts to drop towards zero-compensation.
+Importantly, the observed movements of the groups display the same behavior as that discussed in the previous subsection: The group that only trained in A has the slowest drop towards zero, followed by the group that trained in O-A and finally the group that trained in (-A/2)-A.
+As before, the existance and training on the baseline context pulls the average adaptation towards zero during the E block, and the existance and training on the (-A/2) block in group 2 pulls the average adaptation even lower (as the adaptation is negative).
+Finally, group 4, who only saw a negative-adaptation block (-A) has an average adaptation that is negative.
+We show the simulations with our model in FIGURE XXX, which displays the same behavior as that in the data.
+
+As can be seen in FIGURE XXX, our model also reproduces the lag before the tapering off in the E block starts.
+In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
+
+During these error-clamp trials, the natural variability of movements leads to prediction error: during a normal (e.g.
+A) block, if the participant exerts less force than necessary by mistake, she observes a motor error which, if the adaptation is complete, produces no prediction error.
+However, in E trials, no error is observed, which leads to prediction error.
+It is this prediction error that lets the participant know that a transition has occurred.
+\cite{Vaswani_Decay_2013} performed an additional experiment in which they masked this lack of error by introducing artificial variability on the end position of the hand that mimicked that seen in the participant.
+Because this eliminated the only source of information regarding a switch of context, participants could no longer infer a switch.
+Consistent with our model, the authors found that no tapering off occured during the E block (see FIGURE XXX).
+
+We further discuss error-clamp trials in the Discussion section.
 

### 2021-07-07 adaptation/article/article.tex
Added model description

--- 2021-07-01 article/article.tex
@@
 \usepackage{soul}
+\usepackage{todonotes}
 
@@
 \def \supmat {the Supplementary Materials}
+
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 We can categorize experiments performed with human participants by looking at the amount of contextual information that is available.
-In some experiments [EGssss], the context is clearly cued to the participant, either in the form of visual cues on in changes to the location of the task.
+In some experiments [EGssss], the context is clearly \todo{Maybe not use the word 'cue', as many experiments don't have explicit sensory cues}cued to the participant, either in the form of visual cues on in changes to the location of the task.
 We call these cued-context experiments.
@@
 
+\section{Methods}
+\subsection{Model description}
+We first describe the general principle behind the model, followed, in the next
+subsection, by the full mathematical description of each one of its components.
+
+One of the strenghts of this model is that its general principle does not
+depend on the specifics of the experiment to which it is applied. However, for
+clarity, we will connect the different parts of the model to different
+experiments.
+
+This model is a Bayesian alternative to the N-state learners from
+\citep[e.g.][]{Lee_Dual_2009}. In this model, there are no explicit different
+timescales and multiple states; instead, the model relies on Bayesian learning
+for motor adaptation, combined with inference over the current context.
+
+Before proceding with the model description, let us introduce nomenclature:
+\begin{itemize}
+\item $t$: Time at the moment an observation is made, before inference is done.
+\item $\theta_t$: Observation. Note that this is assumed to be a noisy
+observation of the generalized state (see below) given by $\theta_t = h(z_t)$.
+\item $s_t$: Hidden state, e.g. the position of the hand in reaching
+experiments. It does not depend on the context.
+\item $a_t$: Action taken after observing $\theta_t$ and inferring the context.
+\item $c_t$: Context inferred after observing $\theta_t$.
+\item $\omega_t$: Contextual information. This term includes, e.g. visual cues,
+proprioceptive feedback, performance feedback.
+\item $z_t$: Generalized state, $z_t = \{s_t, c_t, \omega_t\}$. Observed noisily
+as $\theta_t$.
+\end{itemize}
+
+Context inference can work through context-specific (possibly unreliable) cues,
+or through the dynamics of the system (e.g. expected outputs of motor
+commands), using prediction error as the force behind inference. In this case,
+the generative model can is written as follows:
+\[
+p(z_t, z_{t-1}, a_{t-1}, \theta_t, \theta_{t-1}) = p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
+\]
+and inferring the current (hidden) generalized state is done with Bayes' theorem:
+\[
+  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
+\]
+
+Additionally, after having inferred the current state, motor adaptation is given by
+\[
+q(\gamma | z_t, z_{t-1}, a_{t-1}) \propto p(z_t | z_{t-1}, a_{t-1}, \gamma)p(z_{t-1})p(\gamma)
+\]
+
+where $\gamma$ are the parameters of the internal model for the dynamics,
+i.e. the function $f: (z_t, a_t) \rightarrow z_{t+1}$, which determines the beliefs of
+the agent regarding how the system evolves after having taken action $a_t$.
+
+Having inferred the generalized state (including context), a decision is made
+sampling from a distribution:
+\begin{equation}
+p(a_t | z_t, \beta) = \displaystyle\sum_{m \in M}q(C_t)p(a_t | z_t, \beta, m)
+\end{equation}
+where $M$ is the set of all relevant forward models (see below), and
+$p(a_t | ... m)$ is the probability distribution over actions prescribed by the
+forward model $m$.
+
+In the following sections, we present a specific mathematical implementation
+of the principles outlined above. This is the implementation used throughout the
+simulations in the main text.
+
+\subsubsection{Mathematical implementation}
+The generalized state is given by:
+\begin{equation}
+z_t = \{x, \theta_t, \gamma_t, \omega_t\}
+\end{equation}
+where $x$ is, e.g. in a reaching task, the current position of the hand in
+Cartesian coordinates, where the origin is the starting point. We assume that
+motor commands are issued every $\Delta t$ for simplicity. At the beginning of
+each time interval, the context is inferred combining the contextual
+information and the prediction error of the outcome of the previous motor
+command. This yields:
+\begin{align}
+  q(C_t) &= q(C_t | \omega_t)q(C_t | s_t, s_{t-1}, a_{t-1}) \\ \label{eqn:estimated-context}
+  q(C_t | \omega_t) &\propto p(\omega_t | C_t)p(C_t) \\
+  q(C_t | x_t, x_{t-1}, a_{t-1}) &\propto p(x_t | C_t, x_{t-1}, a_{t-1})p(x_t)p(x_{t-1})p(C_t)
+\end{align}
+The terms $p(C_t)$ refer to the prior probability of the context at the
+beginning of the trial: at the first trial, this refers to prior beliefs over
+which contexts are more common/likely, which we call $p_0(C_t)$. At each
+subsequent trial, it incorporates the belief that has so far been accumulated,
+given previous trials. Note that certain events, such as the start of a new
+block of trials (see main text) can return an agent to a state in which
+$p_0(C_t)$ becomes relevant.
+
+We do not directly model hand movements or eye position. Instead, we model
+systems in the motor error space: the position $s_t = 0$ represents a trial in
+which no motor error occurred. We make use of the fact that in most
+experiments, motor adaptation needs to happen in one of two directions
+(e.g. clockwise vs. counter-clockwise) to keep the error space
+two-dimensional. With this, $s_t > 0$ represents motor error in one direction
+(e.g. clockwise) and $s_t < 0$, error in the other direction. We made this
+choice because, in this work, we would model different experiments with
+different state spaces; but all these experiments have the commonality of
+two-dimensional error spaces. In order to model a specific experiment in its
+own physical space (e.g. hand positions in a reaching task), only the dynamical
+system of \eref{eqn:dynamical-system} would need to be adapted.
+
+We assume that the decision-making agent has an internal forward model that
+maps motor commands to outcomes:
+\begin{equation}
+p(s_{t+1} | s_t, a_t, m) = f_m(s_t, a_t)
+\end{equation}
+where $a_t$ are motor commands (actions). We further assume that participants
+have learned $f_m$ throughout their lives and use it to decide which motor
+command to issue. When a perturbation is introduced, to which the participant
+must adapt, we assume they learn a new forward model $f_m^{\phi}$, where $\phi$
+represents the parameters of the function that participants must learn. For
+example, in a saccade task, $\phi$ represents the new gain. Note that $\phi$
+depends on the context and participants need to use and update multiple
+forward adapted models $f_m^{\phi}$ throughout one experiment.
+
+Given a forward model $f_m$ (or $f_m^\phi$), action selection is done by
+building a probability distribution over possible actions $a_t$ based on how
+likely they are to produce zero motor error. For example, in saccade
+experiments, motor commands are issued based on how likely they are to move the
+fovea to the target. These motor commands are issued to counteract intrinsic
+sources of error, such as muscle variability \todo{expand this?}, and extrinsic
+sources of error, such as the experimental manipulations on saccade gain. Thus,
+the expected observed error can be written as:
+\begin{equation}
+p(\epsilon | a_t) \propto p(\epsilon_{\text{intrinsic}} | a_t) + p(\epsilon_{\text{extrinsic}} | a_t)
+\end{equation}
+and a motor command can be chosen to maximize the chance of observing zero error:
+\begin{equation}
+p(a_t) = argmin_{a_t} p(\epsilon | a_t)
+\end{equation}
+
+For simplicity, we assume that motor errors are Gaussian, such that:
+\begin{equation}
+p(\epsilon | a_t) = N(\mu_{\text{extrinsic}}, \sigma_\epsilon)
+\end{equation}
+where $\sigma_\epsilon$ incorporates both intrinsic and extrinsic sources of
+error. The mean of the error is assumed to have no component, as motor
+commands, in the absence of external errors, should produce an error centered
+at zero [CITATION MISSING].
+
+With this, a forward model will produce a motor command with outcomes centered
+at $-\mu_{\text{extrinsic}}$, which counteracts the externally-induced errors.
+
+To complete the action-selection picture, we return to context inference. As
+discussed above, action selection is made via a weighted sum of the different
+forward models:
+\begin{equation}
+p(a_t | ...) \propto \displaystyle \sum_{m \in M} q(C_i)p(a_t | m)
+\end{equation}
+From this distribution, an action is sampled at each trial.
+
+To update the agent's estimates on the parameters of the forward models, we
+make use of Bayes' theorem with conjugate priors. This approach has the
+advantage of not only greatly simplifying computational calculations, but also
+presenting update equations with great intuitive appeal.
+
+In the following equations, we drop the dependence on the context and
+observations for notational simplicity:
+\begin{equation}
+q(\phi | s_{1:t}) \propto p(s_{1:t} | \phi)p(\phi)
+\end{equation}
+where $s_{1:t}$ are the observed states up until time $t$. The likelihood will
+be a Gaussian:
+\begin{align}
+  p(s_t | \theta) &= N(\mu_x, \sigma_x)  \\
+  \mu_x &= \theta_{t-1} + x \\
+  \sigma_x &= \xi_0 1
+\end{align}
+This likelihood is the probability of observing $x$ given the previous estimate
+of the parameters of the force field $\phi_{t-1}$, and given the dynamics of
+the system (see below).
+
+To update $\phi$, we assumed that for each context, the agent's belief over
+the magnitude of the adaptation is given by a normal distribution:
+\begin{equation}
+p(\phi) = N(\mu_\phi, \sigma_\phi) \label{eqn:data-dist}
+\end{equation}
+where $\mu_\phi$ and $\sigma_\phi$ are parameters to be estimated at each
+trial. Before the experiment begins, the agent will have a prior distribution
+over these parameters. A standard Bayesian approach is to choose NormalGamma
+priors:
+\begin{equation}
+p(\mu_\phi, \sigma_\phi) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
+\end{equation}
+where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
+model. We assume the following hyperparameter values \todo{Justify}:
+\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
+  \nu_f^{(0)} &= 0.1 \\
+  \alpha_f^{(0)} &= 1 \\
+  \beta_f^{(0)} &= 0.5
+\end{align}
+Note that $\mu_f^{(0)}$ is different for each context and is thus provided as a
+vector $(0, -1, 1)$ representing the baseline and two opposing adaptations
+(e.g. clockwise and counter-clockwise visuomotor rotations), respectively. For
+all other hyperparameters, all contexts have the same value. These values
+ensure that the prior over the hyperparameters $(\mu_\phi, \sigma_\phi)$ has the mode
+at $(\mu_\phi^{(0)}, 1)$.
+
+The update equations for the magnitud parameters are given by:
+\begin{equation}
+q(\mu_f, \sigma_f | x_t, x_{t-1}, a_{t-1}) \propto p(x_t |
+x_{t-1}, a_{t-1}, C_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
+\end{equation}
+where $p(x_t | x_{t-1}, a_{t-1}, C_t)$ is a Gaussian distribution centered
+around $\mu_a$ with standard deviation $\sigma_a$, which is a free parameter of
+the model. It is assumed that the standard deviation over outcomes ($\sigma_a$)
+is related to that of the parameter $\mu_f$ as follows:
+\begin{equation}
+\sigma_f = \sigma_a / \nu_f
+\end{equation}
+which makes $\nu_f$ a parameter to estimate. Each one of the forward models can
+be updated at every trial, depending on context inference (see main text). The
+updated parameters for any one forward model $i$ are given by:
+\begin{align}
+  \mu_\phi^{(t)} &= \frac{\nu_\phi^{(t-1)} \mu_\phi^{(t-1)} + q(C_i)s_t}{\nu_\phi^{(t-1)} + q(C_i)} \\
+  \nu_\phi^{(t)} &= \nu_\phi^{(t-1)} + q(C_i) \\
+  \alpha_\phi^{(t)} &= \alpha_\phi^{(t-1)} + q(C_i) / 2 \\
+  \beta_\phi^{(t)} &= \beta_\phi^{(t-1)} + \frac{q(C_i)\nu_\phi^{(t-1)}}{\nu_\phi^{(t-1)} +
+                  q(C_i)}\frac{\left(s_t - \mu_\phi^{(t-1)}\right)^2}{2}
+\end{align}
+where $q(C_i)$ is the posterior probability of context $C_i$ from
+\eref{eqn:estimated-context}. Note that when $q(C_i) = 0$, the model $i$ is not
+updated.
+
+The effect of the hyperparameters of the priors are worth a note or two, as
+they are complex.
+
+Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
+units as the necessary adaptation. $\nu$ encodes how stable this hyperprior is:
+higher values (e.g. 10000) all but guarantee that the hypermean will not move
+in the face of evidence; In principle, enough evidence should still move it,
+but that won't happen during an experiment. Smaller values (e.g. 1 / force\_sd,
+as is the default on the code) make the hypermean (and thus the mean) follow
+evidence more freely. Note that as more observations are accumulated, $\nu$
+becames bigger and bigger, solidifying the value of the hypermean.
+
+The hyperparameters $\alpha$ and $\beta$ are a bit more complex. Note that the
+mean of a Gamma distribution is $\beta / (\alpha \nu)$; this mean is being used
+as the standard deviation of a Gaussian by the rest of the agent (cheating),
+which makes it an important measure of uncertainty. While setting the default
+hyperparameters, the values used are $\alpha = 0.5 / \sigma_0$ and
+$\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the
+standard deviation of the force exerted by the environment (force\_sds in the
+code), which controls the initial learning rate. This makes the initial
+standard deviation equal $\sigma_0$, which makes it consistent with the
+fixed-force model in its interface. The 0.5 values ensure that uncertainty is
+large at the beginning and during the experiment is greatly reduced, but never
+to a point where it's ``visually'' too small. Changing this 0.5 would make the
+standard deviation change more quickly, making the model more or less precise
+in its predictions, independently of the volatility of the mean of the
+adaptation (via the hypermean).
+
+The baseline model defaults to different values that make it a lot more
+stable. The hyperstd of the mean is set to 10,000, which makes the mean
+entirely stable during the duration of the experiment. The values of $\alpha$
+and $\beta$ are fixed regardless of $\sigma_0$ such that the standard deviation
+is 0.001 (compare that to the size of the adaptations, around 0.0125), and the
+hyperparameters of the standar deviation are stable during the experiment.
+
 \section{Discussion}
@@
 
-\bibliography{../MyLibrary}
+\bibliography{../../MyLibrary}
 

### 2021-07-09 adaptation/article/article.tex
Take home.

@@
 In FIGURE XXX, data from representative experiments from each category are shown.
-It can be seen that in cued-context experiments, savings take the form of an immediate switch from the adaptation at trial $t$ to that in trial $t-1$, to the level that was learned before.
+It can be seen that in cued-context experiments \cite[data from][]{Oh_Minimizing_2019}, savings take the form of an immediate switch from the adaptation at trial $t$ to that in trial $t-1$, to the level that was learned before.
 In contrast, in partially-cued experiments, the switch is fast (compared to the initial adaptation), but not immediate.

### 2021-07-19 adaptation/article/article.tex
Skeleton of results section

@@
 
-In what follows, we show numerical results obtained the model outlined above. The full mathematical implementation of the model can be found in XXX.
+In what follows, we show numerical results obtained the model outlined above. The full mathematical implementation of the model can be found in \sref{subsection:model-description}.
 
@@
 clockwise), but the second one has half the magnitude.
-Finally, we represent error-clamp trials with the letter E.
+Finally, we represent error-clamp trials with the letter E and washout trials with W.
 
-With this terminology, a typical experiment [EG] would have a block structure of O-A-B-E (or O-A-(-A)-E), which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces, and finally a block with error-clamp trials.
+With this terminology, a typical experiment [EGs] would have a block structure of O-A-B-E (or O-A-(-A)-E), which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).
 
 \subsection{Savings and quick de-adaptation}
-Savings refers to the ability of humans and other animals to remember a previously-learned adaptation, and apply it without having to re-learn it.
-Savings is almost-universally observed in human participants [A THOUSAND CITATIONS; c.f.
-that one paper with single-internal-model], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}.
-
-In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of an apparent adaptation rate might higher than that observed during the first A block.
+Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
+Savings is almost universally observed in human participants [A THOUSAND CITATIONS, c.f.
+papers with volatility or ramping up/down], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}.
+In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of an adaptation rate apparently higher than that observed during the first A block.
 
@@
 
-Importantly, neither savings nor quick de-adaptation are immediate, but instead manifest as a much increased adaptation speed.
-In this section, we will show that both phenomena are the two sides of a coin, and the speed at which this quick adaptation/de-adaptation happens is related to the availability and reliability of contextual cues, as well as to the quality of the prediction errors.
+In this section, we will show that both phenomena are the two sides of a coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.
 
-We can categorize experiments performed with human participants by looking at the amount of contextual information that is available.
-In some experiments [EGssss], the context is clearly \todo{Maybe not use the word 'cue', as many experiments don't have explicit sensory cues}cued to the participant, either in the form of visual cues on in changes to the location of the task.
+To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that is available.
+In some experiments [EGssss], the context is clearly \todo{Maybe not use the word 'cue', as many experiments don't have explicit sensory cues}cued to the participant, e.g.
+in the form of visual cues [EGs] or in changes to the location of the task [EGs].
 We call these cued-context experiments.
 In other experiments, partial information is available to participants [EGsss]; we refer to these as partially-cued experiments.
-While there are no experiments in which no contextual information is available, there are some experiments which contain blocks in which this is the case and participants have no way of inferring the current context.
+While there are no experiments in which no contextual information is available, there are some experiments which contain blocks of trials in which no contextual information is available and participants have no way of inferring the current context.
 Abusing our notation, we call these experiments (or rather, these sections of the experiments) uncued experiments.
@@
 For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
-Additionally, the sudden appearance of large motor errors can itself be a cue for contextual change.
+Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands].
+The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance.
 
-To see the effects of context inference on savings and quick de-adaptation, we contrast the two former types of experiments.
-
-In FIGURE XXX, data from representative experiments from each category are shown.
+In FIGURE XXX, we show data from representative experiments from the cued-context \citep{Kim_Neural_2015} and partially-cued \citep[][group 1c]{Oh_Minimizing_2019} categories are shown.
+\cite{Kim_Neural_2015} used a colored light as a visual cue at the beginning of each experiment, whereas \cite{Oh_Minimizing_2019} included a condition in which the necessary adaptation was in the order of the motor error, making feedback-based context inference slow.
 It can be seen that in cued-context experiments \cite[data from][]{Oh_Minimizing_2019}, savings take the form of an immediate switch from the adaptation at trial $t$ to that in trial $t-1$, to the level that was learned before.
@@
 
-In FIGURE XXX, we show the same effects described above, but in the frame of quick de-adaptation.
-The explanation for these effects given by the model is the same for savings and quick de-adaptation, namely that the participant slowly accumulates evidence in favor of a switch in context.
+Note that in the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
+\todo{Need to be careful here, as de-adaptation is sometimes discussed only in terms of experiments with no contextual cues or where the meaning of the cues is inverted}.
 
 \section{The effects on the rate of adaptation}
-In our model, adaptation is gated by the uncertainty on the current context (see Methods).
+In our model, adaptation is gated by the uncertainty on the current context.
 More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation, given a fixed size of prediction error.
@@
 They divided participants into three groups for which the transition from one trial to the next were very uncommon, somewhat common and almost every trial.
-Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred when a motor error is observed.
+Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.
 
 \cite{Herzfeld_memory_2014} found that the volatility of the environment affects learning.
-The more volatile the environment, the less participants can adapt their motor responses.
+The more volatile the environment, the less participants adapt their motor responses.
 In FIGURE XXX, we show results adapted from \cite{Herzfeld_memory_2014} alongside results obtained with our model.
@@
 
-While we focused on \cite{Herzfeld_memory_2014}, evidence for this can be seen in other studies, which we discuss in the Discussion section.
-
 \section{Action selection}
@@
 Experimental evidence supporting this view can be found in experiments with context switching.
-For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group (where -A refers to an adaptation in the opposite direction, but with the same magnitude), with a block sequence A-(-A)-A-(-A), keeping the size of the necessary adaptation equal across both groups.
+For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group (where -A refers to an adaptation in the opposite direction as A, but with the same magnitude), with a block sequence A-(-A)-A-(-A), keeping the size of the necessary adaptation equal across both groups.
 After A and 3A in one group, A and -A in the other, had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
-The authors interpreted this as evidence that a new adaptation happens faster if it is in the same direction as the current adaptation (e.g.
-both counter-clockwise), and slower if they are in the opposite direction (e.g.
+The authors interpreted this as evidence that a new adaptation happens more quickly if it is in the same direction as the current adaptation (e.g.
+both counter-clockwise), and more slowly if they are in the opposite direction (e.g.
 clockwise to counter-clockwise).
@@
 
+\section{Discussion}
+This is where the brunt of the paper will be.
+
 \section{Methods}
 \subsection{Model description}
+\label{subsection:model-description}
 We first describe the general principle behind the model, followed, in the next

### 2021-07-19 adaptation/article/article.tex
Update on Overleaf.

@@
 In our model, the idea of context is quite general.
-The context comprises the relevant elements of hte environment (e.g.
+The context comprises the relevant elements of the environment (e.g.
 high winds when walking), as well as the task at hand.
@@
 In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment.
-This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely intependent from them.
+This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely independent from them.
 For this section, the most important component of $\vartheta$ is contextual information, i.e.
@@
 
-Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use.
+Note that because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use.
 For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
@@
 
-Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context.
-As discussed above, action selection in our model is done via a weighted avarage:
+Our model offers a different explanation for the observed results: the asymmetry is due to the existence of the baseline context.
+As discussed above, action selection in our model is done via a weighted average:
 \begin{equation}

### 2021-07-22 adaptation/code/model.py
All results are ready, except error clamp trials

@@ -70,6 +70,8 @@ class LeftRightAgent(object):
 
     all_learn = False
 
+    threshold_learn = 0.1  # if p(context) lower than this, no learning happens
+
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None, prediction_noise=None,
                  reset_after_change=None, force_sds=None, max_force=None):
@@ -206,8 +208,7 @@ class LeftRightAgent(object):
         """
         ix_context = self.sample_context(t=-2)
         mu_p, sd_p = self.magnitudes[ix_context]
-        p_con = np.argmax(self.log_context_history[-2]) # called after infer_context
-        mu_l, sd_l = self.predict_outcome(previous_context=p_con)[1][ix_context]
+        mu_l, sd_l = self.predict_outcome()[1][ix_context]
         mu_l = mu_p + self.hand_position - mu_l
         mu_post = (mu_l * sd_p ** 2 + mu_p * sd_l ** 2) / \
             (sd_l ** 2 + sd_p ** 2)
@@ -217,8 +218,7 @@ class LeftRightAgent(object):
         self.magnitudes = c_con
         self.magnitude_history.append(self.magnitudes)
 
-    def predict_outcome(self, hand_position=None, action=None,
-                        previous_context=None):
+    def predict_outcome(self, hand_position=None, action=None):
         """Given an action, generates a probability distribution over all
         possible outcomes.
 
@@ -245,8 +245,6 @@ class LeftRightAgent(object):
         action_sd = self.action_sd
         if hand_position is None:
             hand_position = self.hand_position_history[-2]
-        if previous_context is None:
-            previous_context = np.argmax(self.log_context_history[-1])
         funs = []
         posterior_pars = []
         for ix_context in range(self.num_contexts):
@@ -268,9 +266,7 @@ class LeftRightAgent(object):
 
         """
         if not self.is_reset:
-            previous_context = np.argmax(self.log_context_history[-1])
-            lh_fun, lh_pars = self.predict_outcome(
-                previous_context=previous_context)
+            lh_fun, lh_pars = self.predict_outcome()
             # Hand position likelihood:
             log_li_hand = np.array([lh_fun(hand_position, ix_context)
                                     for ix_context in range(self.num_contexts)])
[…]

### 2021-07-22 adaptation/code/pars.py

@@ -64,6 +64,27 @@ def _define_breaks():
     return vector
 
 
+def define_contexts(structure):
+    """Given a list with elements [context_id, num_trials], it will piece together
+    the entire sequence of trial types for the experiment.
+
+    Example:
+    structure_dict = [[0, 10], [1, 20], [0, 5], [2, 100]]
+    This will create an experiment with 10 + 20 + 5 + 100 = 135 trials, the
+    first 10 being of context 0, the next 20 with context 1, etc.
+
+    Returns
+    -------
+    contexts : ndarray size=(total_trials, )
+    Context for each trial in the experiment.
+
+    """
+    context_list = []
+    for block in structure:
+        context_list.append(block[0] * np.ones(block[1], dtype=int))
+    return np.concatenate(context_list)
+
+
 # Parameters common to agent and task:
 delta_t = 0.05
 obs_noise = 0.0001

### 2021-07-22 adaptation/code/simulations.py

@@ -9,7 +9,6 @@ import glob
 import numpy as np
 from matplotlib import pyplot as plt
 import pandas as pd
-from scipy import stats
 
 import model
 import task_hold_hand as thh
@@ -372,7 +371,8 @@ def plot_adaptation(pandata, axis=None, fignum=5):
     plt.draw()
 
 
-def sim_and_plot(agent, pars_task, return_data=False, axes=None, fignum=6):
+def sim_and_plot(agent, pars_task, return_data=False,
+                 force_labels=None, axes=None, fignum=6):
     """Simulates the agent playing and makes a nice plot with context inference and
     adaptation and colors everywhere.
 
@@ -384,12 +384,12 @@ def sim_and_plot(agent, pars_task, return_data=False, axes=None, fignum=6):
     plot_adaptation(pandota, axis=axes[0])
     plot_contexts(pandota, axis=axes[1])
 
-    fake_forces = [-pars_task['forces'][1][1], 0, pars_task['forces'][1][1]]
-    real_forces = [-pars.forces[1], 0, pars.forces[1]]
+    fake_forces = np.array([-pars_task['forces'][1][1], 0, pars_task['forces'][1][1]])
 
     axes[0].set_ylabel('Adaptation (N)')
-    axes[0].set_yticks(ticks=fake_forces)
-    axes[0].set_yticklabels(real_forces)
+    # axes[0].set_yticks(ticks=fake_forces * 1.2)
+    if force_labels is not None:
+        axes[0].set_yticklabels(force_labels)
     axes[0].legend()
 
     axes[1].set_xlabel('Trial')
@@ -478,5 +478,216 @@ def baseline_bias(fignum=9):
     axes[0, 3].get_legend().remove()
 
 
+def oh_2019(plot=True, axes=None, fignum=10):
+    """Simulates the experiments from Oh and Schweighoffer 2019.
+
+    Parameters
+    ----------
+    plot : bool
+    Whether to simulate and plot. If True, will create the agent
+    with the parameters agent_pars and simulate with sim_and_plot().
+
+    Returns
+    -------
+    task_20 : dictionary
+    Parameters for the task with adaptation=20. To be used directly
+    as the input for tth.run().
+
+    task_10 : dictionary
+    Same as --task_20-- but for adaptation=10.
+
+    agent_pars : dictionary
+    Parameters for the agent for both tasks. To be used directly with
[…]

### 2021-07-22 adaptation/code/task_hold_hand.py

@@ -25,7 +25,7 @@ def run(agent=None, save=False, filename=None, pars=None):
     outs = []
     hand_position = 0
     for c_trial, do_break in enumerate(pars['breaks']):
-        # if c_trial > 520:
+        # if c_trial >= 111:
         #     ipdb.set_trace()
         if do_break:
             agent.reset()

### 2021-08-13 adaptation/article/article.tex
Added initial results, with plots and all.

--- 2021-07-19 adaptation/article/article.tex
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{./figures/figure_1.png}
+\caption{Savings and de-adaptation.
+Data from our simulations (A and C) compared to data from \cite{Oh_Minimizing_2019} (B and D).
+In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number.
+The black lines represent the optimal adaptation, i.e.
+the size of the visuomotor rotation during the task.
+(A) A simulated experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements.
+It can be seen that after context switches (e.g.
+at trial 120), the agent returns to the adaptation of the previous block, with one or two trials of delay.
+(B) Data adapted from \cite{Oh_Minimizing_2019}, figure 4a.
+The same immediate switch can be observed as that in A.
+(C) Simulated experiment with a 10 degree adaptation.
+In this case, savings are much slower.
+(D) Data adapted from \cite{Oh_Minimizing_2019}, figure 4a.} \label{fig:oh-2019}
+\end{figure}
+
 Note that in the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
@@
 The most direct evidence for the effects of context inference on the rate of adaptation come from the experiments by \cite{Herzfeld_memory_2014}.
-The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between O and A (and vice versa) change the speed at which adaptation occurs.
+The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between A and -A (and vice versa) change the speed at which adaptation occurs.
 
-\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between two possibilities: no force (baseline O) and some force (context A).
+\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between two possibilities: no force (baseline O) a clockwise force (context A) and a counter-clockwise force (context -A).
 They divided participants into three groups for which the transition from one trial to the next were very uncommon, somewhat common and almost every trial.
@@
 \cite{Herzfeld_memory_2014} found that the volatility of the environment affects learning.
-The more volatile the environment, the less participants adapt their motor responses.
-In FIGURE XXX, we show results adapted from \cite{Herzfeld_memory_2014} alongside results obtained with our model.
-In our model, learning is gated by context inference: the higher the uncertainty on the context in which an error occured, the lower the effect of this error on future trials (i.e.
-adaptation).
+The more volatile the environment, the less participants adapt their motor responses after observing an error.
+In \fref{fig:herzfeld-2014}, we show results obtained with our model in a paradigm matching that of \cite{Herzfeld_memory_2014}.
+We simulated the conditions of low volatility and maximum uncertainty ($z = 0.9$ and $z = 0.5$ in \cite{Herzfeld_memory_2014}) for thirty trials, repeating the experiment 300 times with random sequences of A and -A each time, following the transition probabilities from the experiment.
+In \fref{fig:herzfeld-2014}A we show the adaptation as a function of trial number, averaged across all 300 runs.
+It can be seen that the agent learns more slowly on the more volatile environment ($z = 0.5$) than on the less volatile one, matching the results from \cite{Herzfeld_memory_2014}.
+Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment.
+
+\begin{figure}
+\centering
+\includegraphics{./figures/figure_2.png}
+\caption{Learning under unstable and unpredictable contexts.
+30 trials were simulated in the task defined by \cite{Herzfeld_memory_2014}, repeated 300 times for better statistics.
+Two conditions were used, following the naming convention from \cite{Herzfeld_memory_2014}: $z=0.5$ is an unpredictable condition, where the context has a 0.5 probability of switching from trial to trial.
+$z=0.9$ is a stable condition, where the context has a 0.9 probability of staying the same from trial to trial.
+(A) Adaptation as a function of trial number, averaged across all runs.
+(B) Error sensitivity, binned into bins of 5 trials, averaged across all runs.
+The error is in arbitrary units, causing sensitivity to error (y axis) to also be in arbitrary units.
+Shaded areas are SEM.}
+\label{fig:herzfeld-2014}
+\end{figure}
+
+In our model, learning is gated by context inference: the higher the uncertainty on the context in which an error occured, the lower the effect of this error on future trials (i.e. adaptation).
 
@@
 We simulated data with our model fitting the experimental structure by \cite{Davidson_Scaling_2004}.
-We show the results in FIGURE XXX, alongside the experimental results from \cite{Davidson_Scaling_2004}.
-It can be seen that the model provides a good qualitative fit to the data.
+We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}.
+It can be seen that the agent exposed to the -8 context shifts back to 4 more slowly than the one exposed to +8, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{./figures/figure_3.png}
+\caption{Motor error when switching back to a previously learned adaptation.
+Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from +8 (yellow) or -8 (blue) to 4 is shown.
+(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}.
+On the y axis, the error is shown in arbitrary units, related to centimiters through a monotonically increasing transformation with the same origin.
+(B) Data adapted from \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
+\label{fig:davidson-2004}
+\end{figure}
 
@@
 all are positive or all are negative, or in an experiments in which the contexts are clearly cued.
+\todo{Maybe add simulations that show this, as well as simulations where the baseline context is not present.}
 
@@
 
-In FIGURE XXX, we show the data from \cite{Vaswani_Decay_2013}, where it can be seen that the responses during the E block trials starts to drop towards zero-compensation.
-Importantly, the observed movements of the groups display the same behavior as that discussed in the previous subsection: The group that only trained in A has the slowest drop towards zero, followed by the group that trained in O-A and finally the group that trained in (-A/2)-A.
-As before, the existance and training on the baseline context pulls the average adaptation towards zero during the E block, and the existance and training on the (-A/2) block in group 2 pulls the average adaptation even lower (as the adaptation is negative).
-Finally, group 4, who only saw a negative-adaptation block (-A) has an average adaptation that is negative.
-We show the simulations with our model in FIGURE XXX, which displays the same behavior as that in the data.
+In \fref{fig:vaswani-2013}A-E, we show the data simulated with our model for the task presented by \cite{Vaswani_Decay_2013}.
+The motor adaptation periods can be observed in \fref{fig:vaswani-2013}A-D for the four groups in the experiment, while on each bottom panel the inference over context is shown.
+Context inference works reliably until the error-clamp trials start, which do not correspond to any of the known contexts.
+This causes the agent to infer the combination of some of the known contexts that best fits the observations.
+Depending on the contexts learned by the agent (which change from group to group), inference during E trials will vary: groups 1.1 and 1.3 have exactly the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
+These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
+In contrast, group 1.3 has the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C, which has a non-zero posterior probility during E trials, pushing the agent's adaptation force more quickly towards zero.
+Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more slowly than groups 1.1 and 1.4.
+
+In \fref{fig:vaswani-2013}E, the same data is shown but only in the E trials.
+Here, the difference in the speed in which the force decays for the different groups is more readily visible.
+This is a strong qualitative match to the data from \cite{Vaswani_Decay_2013}, shown in \fref{fig:vaswani-2013}F.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{./figures/figure_4.png}
+\caption{Adaptation during error clamp trials.
+(A) On top, adaptation index shown by the agent belonging to group 1.1 at every trial, i.e.
+how much the agent adapted to the force that was present, where 1 is optimal adaptation.
+On bottom, inference over the current context, where contexts are color coded: black for baseline, orange for counter-clockwise force, purple for clockwise force and brown for clockwise force with half strenght.
+The lines represent the posterior probability of each ocontext at every trial, while the background color represents the real context.
+An olive-colored background represents error-clamp trials.
+(B) Same as A, for group 1.2.
+(C) Same as A, for group 1.3.
+(D) Same as A, for group 1.4.
+(E) Simulated data for all groups (same color coding as A-D), only during error-clamp trials.
+Perceptual noise (i.e.
+how well the agent can estimate the position of its own hand) was lowered to increase the clarity of the plotted lines; however, increasing it does not change the results.
+(F) Data adapted from Vaswani et al.
+2013, figure 2C, depicting the data equivalent to (E).}
+\label{fig:vaswani-2013}
+\end{figure}
 
@@
 In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
+\todo{Simulate the lag?
+May be worth it}
+
+\section{Discussion}
+Discuss that our model does not fully match vaswani in how it all decays. This is because our model does not have a mechanism to create new internal models.
 
@@
 A) block, if the participant exerts less force than necessary by mistake, she observes a motor error which, if the adaptation is complete, produces no prediction error.
-However, in E trials, no error is observed, which leads to prediction error.
+However, in E trials, no error is observed, which leads to a predicted error that is not observed.
 It is this prediction error that lets the participant know that a transition has occurred.
@@
 Because this eliminated the only source of information regarding a switch of context, participants could no longer infer a switch.
-Consistent with our model, the authors found that no tapering off occured during the E block (see FIGURE XXX).
-
-We further discuss error-clamp trials in the Discussion section.
-
-\section{Discussion}
-This is where the brunt of the paper will be.
+Consistent with our model, the authors found that no tapering off occured during the E block.
+\todo{Moved from Results}
 

### 2021-08-13 adaptation/code/figures.py

@@ -9,10 +9,325 @@ auxiliary functions."""
 
 import numpy as np
 from matplotlib import pyplot as plt
+import pandas as pd
+from matplotlib import gridspec as gs
+import seaborn as sns
 
 import simulations as sims
 import model
 import task_hold_hand as thh
 import pars
 
+FIGURE_FOLDER = '../article/figures/'
 
+
+def oh_2019(fignum=1, show=True):
+    """Reproduces the results from Oh_Minimizing_2019 and plots them
+    in a format similar to that of their figure 4.
+
+    An empty subplot is created to manually add the figure from their
+    paper.
+
+    """
+    fig, axes = plt.subplots(2, 2, clear=True, num=fignum,
+                             sharex=True, sharey=True)
+    task_20, task_10, agent_pars = sims.oh_2019(plot=False)
+
+    # a)
+    agent = model.LRMeanSD(**agent_pars)
+    agent.all_learn = True
+    pandata, pandagent, _ = thh.run(agent, pars=task_20)
+    pandota_20 = thh.join_pandas(pandata, pandagent)
+    axes[0, 0].plot(-pandota_20['pos(t)'] - pandota_20['action'])
+    contexts = pandota_20['ix_context'].values
+    contexts[contexts == pars.CLAMP_INDEX] = 0
+    contexts[contexts == 1] = 20
+    contexts_20 = contexts
+    axes[0, 0].plot(contexts, color='black')
+    axes[0, 0].set_ylabel('Adaptation')
+    axes[0, 0].set_title('Simulations')
+    axes[0, 0].text(x=-0.1, y=1, s='A',
+                    transform=axes[0, 0].transAxes,
+                    fontdict={'size': 14})
+
+    # c)
+    agent = model.LRMeanSD(**agent_pars)
+    agent.all_learn = True
+    pandata, pandagent, _ = thh.run(agent, pars=task_10)
+    pandota_10 = thh.join_pandas(pandata, pandagent)
+    axes[1, 0].plot(-pandota_10['pos(t)'] - pandota_10['action'])
+    contexts = pandota_10['ix_context'].values
+    contexts[contexts == pars.CLAMP_INDEX] = 0
+    contexts[contexts == 1] = 10
+    contexts_10 = contexts
+    axes[1, 0].plot(contexts, color='black')
+    axes[1, 0].set_xlabel('Trials')
+    axes[1, 0].set_ylabel('Adaptation')
[…]

### 2021-08-13 adaptation/code/model.py

@@ -504,7 +504,8 @@ class LRMeanSD(LeftRightAgent):
     """
     name = 'LRMS'
 
-    def __init__(self, hyper_sd=None, *args, **kwargs):
+    def __init__(self, hyper_sd=None, learn_rate=1, all_learn=False,
+                 threshold_learn=0.2, *args, **kwargs):
         """Initializes hyperparameters and calls LeftRightAgent.__init__
 
         Note: The hyperparameters are initialized such that the mean of the
@@ -517,7 +518,8 @@ class LRMeanSD(LeftRightAgent):
         if hyper_sd is None:
             hyper_sd = 0.5
         super().__init__(*args, **kwargs)
-        self.mag_hypers = [[angle, 1, hyper_sd / (sd + self.obs_sd), 2 * hyper_sd]
+        self.mag_hypers = [[angle, learn_rate, hyper_sd / (sd + self.obs_sd),
+                            2 * hyper_sd]
                            for angle, sd in zip(self.angles, self.force_sds)]
         self.mag_hypers[0][1] = 10000
         self.mag_hypers[0][2] = 100000 / (self.force_sds[0] + self.obs_sd)
@@ -529,6 +531,8 @@ class LRMeanSD(LeftRightAgent):
             prior_mags.append([c_hypers[0], sds])
         self.magnitudes = prior_mags
         self.magnitude_history[0] = prior_mags
+        self.all_learn = all_learn
+        self.threshold_learn = threshold_learn
 
     def update_magnitudes(self):
         r"""Updates all four hyperparameters of the force magnitudes, using

### 2021-08-13 adaptation/code/simulations.py

@@ -504,7 +504,7 @@ def oh_2019(plot=True, axes=None, fignum=10):
     task_20['force_noise'] = 1 * np.ones(2)
     task_20['forces'] = [[0, 0], [1, 20]]
     contexts_20 = [[0, 20], [1, 60], [0, 40], [1, 50], [0, 20],
-                   [1, 30], [0, 40], [1, 50], [pars.CLAMP_INDEX, 120]]
+                   [1, 30], [0, 40], [1, 50]]
     task_20['context_seq'] = pars.define_contexts(contexts_20)
     task_20['breaks'] = np.zeros(len(task_20['context_seq']))
     task_20['cues'] = np.zeros(len(task_20['context_seq']), dtype=int)
@@ -530,10 +530,12 @@ def oh_2019(plot=True, axes=None, fignum=10):
         agent.all_learn = True
         sim_and_plot(agent, task_20, axes=axes[:, 0])
         axes[0, 0].set_title('Adaptation: 20')
+        axes[0, 0].set_ylim((-25, 25))
         agent = model.LRMeanSD(**agent_pars)
         agent.all_learn = True
         sim_and_plot(agent, task_10, axes=axes[:, 1])
         axes[0, 1].set_title('Adaptation: 10')
+        axes[0, 1].set_ylim((-12, 12))
 
     return task_20, task_10, agent_pars
 
@@ -583,7 +585,7 @@ def kim_2015(plot=True, axis=None, fignum=11):
     return task, agent_pars
 
 
-def herzfeld_2014(plot=True, axes=None, fignum=12):
+def herzfeld_2014(num_trials=150, plot=True, axes=None, fignum=12):
     """Simulates something like the experiments from Herzfeld et al. 2014.
 
     This does two conditions: high uncertainty and low uncertainty, where
@@ -592,9 +594,8 @@ def herzfeld_2014(plot=True, axes=None, fignum=12):
     change. In the low uncertainty, there is a 0.1 chance of change.
 
     """
-    num_trials = 300
     task = task_pars.copy()
-    task['obs_noise'] = 2
+    task['obs_noise'] = 3
     task['force_noise'] = 0.01 * np.ones(3)
     task['forces'] = [[0, 0], [-1, 13], [1, 13]]
     task['breaks'] = np.zeros(num_trials, dtype=int)
@@ -607,16 +608,18 @@ def herzfeld_2014(plot=True, axes=None, fignum=12):
     switches = np.random.choice([0, 1], p=(0.99, 0.01), size=num_trials)
[…]

### 2021-08-24 adaptation/article/article.tex
80% of first draft done. Abstract missing, intro incomplete.
Changed figure 1 to now include the results from kim 2015.

@@
 \abstract{Abstract}
-% \tableofcontents
+Very interesting abstract. Much wow.
 
 \section{Introduction}
-yeah
+It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997,}, although interference can occur between tasks \cite[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
+
+In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement.
+For example, in experiments with a mechanical arm \cite[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
+Other examples include measuring the force profile in error-clamp trials \cite{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \cite{Kim_Neural_2015} and saccade adaptation \cite{Catz_Cerebellardependent_2008}.
+
+By introducing blocks of trials in which a manipulation is present (e.g.
+the mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error.
+A number of phenomena have been observed that persist across participants and across many different motor adaptation experiments: savings, quick de-adaptation, anterograde interference and spontaneous recovery.
+
+In order to explain these phenomena, a number of computational models have been introduced.
+The least complex of these models are linear learners [CITATION NEEDED], in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
+By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, explaining these distinct experimental phenomena.
+
+Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via prediction error generated by these internal models.
+These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses, as well as increases the stability of the estimate.
+
+While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of the scope of these models. (1)
+
+[unexplained phenomena. The truth is out there. I want to believe]
 
@@
 \subsection{The model}
-The model we present has three main components: (1) a context-inference component; (2) a motor adaptation component; (3) an action selection component.
+The model we present has three main components: (1) context-inference, (2) motor adaptation and (3) action selection.
 Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.
 
-Based on the Bayesian brain hypothesis, we chose to cast both context inference and motor adaptation as Bayesian inference, in effect using a unified mechanism for the preparation of motor commands in new environments.
-
-We now describe the three components of the model separately.
-We begin by describing the motor adaptation part for clarity, as the definition of ``context'' (used in context inference) is better described in terms of motor adaptation.
+We now briefly describe the three components of the model separately.
+For clarity, we start with the second component (i.e.
+motor adaptation), in order to introduce the terminology that we use throughout this work.
+For a complete mathematical description of the model, see the Methods section.
 
 \subsubsection{Motor adaptation}
-Motor adaptation refers to the ability of humans and other animals to learn from observed errors.
-For example, we can learn to perform a reaching movement in front of us --a movement that any person has been practicing all their lives-- underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
+Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors.
+For example, we can learn to perform a reaching movement in front of us underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
 
-It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals.
-While a number of models for motor adaptation exist, we chose to make use of an exact Bayesian learner.
-As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also does away with the need for explicit multiple time scales of learning.
+It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \cite{Wolpert_Multiple_1998a} \todo{Is this the original one?
+Include more}.
+While a number of models for motor adaptation exist [EGs], we chose to make use of an exact Bayesian learner.
+As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning.
 We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.
@@
 the current position of the hand), $c$ is a motor command and $\beta$ are the parameters of the forward model $f(\cdot)$.
-Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements.
+Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which can be updated during motor adaptation.
 
-As a day progresses, and especially if the day included a lot of physical activity of the arm, muscle exhaustion leads to a different response of the muscles in the arm to motor commands.
-However, we are still able to produce accurate reach movements.
-This happens through adaptation, where the observed error signals of a moment are used to temporarily update the parameters $\beta$ of the forward model.
+When an observed outcome does not match the predictions produced by $f$, the brain uses the prediction error to update the parameters $\beta$ of the forward model.
 To describe this adaptation process, our model makes uses of Bayesian inference, such that:
@@
 where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
-$p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta$ is the previous estimate of $\beta$.
-
-In this account of motor adaptation, the prediction error signal takes the form of an observed outcome $\vartheta$ that differs from that of the most likely outcome predicted by the forward model.
-The amount of adaptation that happens after observing an error will depend not only on the size of the error (how far away from the most likely outcome) it is, but also of how precise the existing estimate of $\beta$ is.
-We will discuss this further below.
-
-Not only is the internal model for everyday movements not fixed (i.e.
-adaptation is possible), it is also not unique.
-Multiple internal models are used and updated, depending on the task at hand.
-For example, a person might have an internal model for baseline reaching movements (e.g.
-with empty hands) $f_B$, and a different one for reaching movements while holding a heavy object, $f_H$.
-A key aspect of motor adaptation is to select the relevant model to update, should a prediction error arise.
+$p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta)$ is the previous estimate of $\beta$.
 
@@
 In our model, the idea of context is quite general.
-The context comprises the relevant elements of hte environment (e.g.
-high winds when walking), as well as the task at hand.
-Previous models have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
+The context comprises the relevant elements of the environment (e.g.
+the strength and direction of the wind when walking) and of the task at hand (immediate and long-term goals, reward structure, etc).
+Previous models \cite[e.g.][]{Wolpert_Multiple_1998a,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
 $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e.
@@
 
-In this work, we focus on the idea that the context can be inferred by integrating several sources of information, an idea with great intuitive appeal.
-Through this idea, we explain many previously-unexplained phenomena from different experiments with humans.
+In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.
+We show how the incorporation of all available contextual information explains many previously-unexplained phenomena from experiments with human subjects, without invoking \textit{ad hoc} model components.
 
-Identifying any possible context with a categorical variable $\zeta$, the identification of a context can be done through Bayesian inference:
+Identifying any possible context with a categorical variable $\zeta$, the identification of a context is done by our model through Bayesian inference:
 \begin{equation}
@@
 \end{equation}
-where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from instant to instant, and (2) an overall estimation of observing any one context, if the context did change.
+where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall estimation of observing any one context, if the context did change.
 Below, we discuss how these two components effect different experimentally observed phenomena.
@@
 In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment.
-This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely intependent from them.
-For this section, the most important component of $\vartheta$ is contextual information, i.e.
+This includes, as before, the outcomes of actions (e.g.
+the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely intependent from them.
+For this section, the most important component of $\vartheta$ is sensory contextual information, i.e.
 information that might directly help infer the identity of the current context.
 
-In experimental settings, the contextual information can take the form of a visual cue [EGS], the place where the task must be carried out [EGS] or even the start of a new block of trials [EGS].
-All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1
-}, c_{t-1})$ and are integrated together to identify a context.
+In experimental settings, the sensory contextual information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}.
+All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.
+
+By making use of NormalGamma priors on context, the model performs exact Bayesian inference at every trial. The exact update equations can be found in the Methods section.
 
@@
 \section{Experimental results}
-In this section, we go through a number of experimentally-observed phenomena and show how imperfect context inference, as done in our model, provides a parsimonious explanation for all of them.
+In this section, we go through a number of experimentally-observed phenomena and show that context inference, as done in our model, provides a parsimonious explanation for all of them.
 
-In what follows, we show numerical results obtained the model outlined above. The full mathematical implementation of the model can be found in \sref{subsection:model-description}.
-
-Before discussing these phenomena, let us introduce terminology.
+For clarity, we introduce some terminology before discussing these phenomena.
 As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force.
-Depending on the trial (and experiment), the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
+Depending on the trial, the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
 Let us define the baseline context O as that in which the mechanical arm exerts no force.
 Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise.
-Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point (approximately) in opposite directions.
-We represent contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
+Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point in opposite directions.
+We can also talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
 clockwise), but the second one has half the magnitude.
-Finally, we represent error-clamp trials with the letter E and washout trials with W.
+Finally, error-clamp trials, in which the mechanical arm forces the participant to make straight-line movements, are represented with the letter E.
 
-With this terminology, a typical experiment [EGs] would have a block structure of O-A-B-E (or O-A-(-A)-E), which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).
+With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of O-A-B-E, or O-A-(-A)-E, which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).
 
@@
 papers with volatility or ramping up/down], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}.
-In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of an adaptation rate apparently higher than that observed during the first A block.
+In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
 
@@
 
-In this section, we will show that both phenomena are the two sides of a coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.
+In this section, we will show that both phenomena are the two sides of the same coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.
 
-To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that is available.
-In some experiments [EGssss], the context is clearly \todo{Maybe not use the word 'cue', as many experiments don't have explicit sensory cues}cued to the participant, e.g.
-in the form of visual cues [EGs] or in changes to the location of the task [EGs].
+To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that made available to participants.
+In some experiments [EGssss], the context is clearly revealed to the participant.
 We call these cued-context experiments.
-In other experiments, partial information is available to participants [EGsss]; we refer to these as partially-cued experiments.
+In other experiments, partial information is available to participants [EGsss] in the form of prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.
 While there are no experiments in which no contextual information is available, there are some experiments which contain blocks of trials in which no contextual information is available and participants have no way of inferring the current context.
@@
 For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
+In fact, the mere act of holding the handle informs participants that something is different from baseline.
 Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands].
-The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance.
+The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.
 
-In FIGURE XXX, we show data from representative experiments from the cued-context \citep{Kim_Neural_2015} and partially-cued \citep[][group 1c]{Oh_Minimizing_2019} categories are shown.
-\cite{Kim_Neural_2015} used a colored light as a visual cue at the beginning of each experiment, whereas \cite{Oh_Minimizing_2019} included a condition in which the necessary adaptation was in the order of the motor error, making feedback-based context inference slow.
-It can be seen that in cued-context experiments \cite[data from][]{Oh_Minimizing_2019}, savings take the form of an immediate switch from the adaptation at trial $t$ to that in trial $t-1$, to the level that was learned before.
-In contrast, in partially-cued experiments, the switch is fast (compared to the initial adaptation), but not immediate.
-Our model explains this difference via context inference: in the cued-context experiments, context inference relies on the very reliable information of visual cues and the change in context is identified immediately; because of this, the model can re-activate the forward model adequate for the context from the first trial after the change occured.
-In contrast, in partially-cued experiments, the model needs to accumulate evidence in favor of the correct context throughout a number of trials, and the actions selected by the model are combinations of the actions that correspond to all the possible contexts (we discuss this in the following sections).
+To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies which differ in the amount of contextual information.
+\cite{Kim_Neural_2015} performed a visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
+Participants performed shooting movements in blocks of trials with the same rotation.
+Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment.
+Consistent with our model, the authors found that switching was immediate and accurate.
+In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
+The correspondence between simulations and experimental results is quite evident, both in the immediacy of the switches as well as in the fact that learning continues from the state it was when the context was last observed.
 
@@
 \centering
-\includegraphics[width=\textwidth]{./figures/figure_1.png}
+\includegraphics{./figures/figure_1.png}
 \caption{Savings and de-adaptation.
-Data from our simulations (A and C) compared to data from \cite{Oh_Minimizing_2019} (B and D).
+Data from our simulations (left column) compared to data from \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} (right column).
 In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number.
@@
 the size of the visuomotor rotation during the task.
-(A) A simulated experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements.
-It can be seen that after context switches (e.g.
+(A) Experiment by \cite{Kim_Neural_2015} with two rotations (-40 and 40 degrees), in addition to baseline.
+The switches between contexts were visually cued and therefore switches are immediate, as can be seen in both simulations (left) and the experimental data from \cite{Kim_Neural_2015} (right).
+(B) An experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements.
+It can be seen in both the simulations (left) and experimental data from \cite{Oh_Minimizing_2019} (right) that after context switches (e.g.
 at trial 120), the agent returns to the adaptation of the previous block, with one or two trials of delay.
-(B) Data adapted from \cite{Oh_Minimizing_2019}, figure 4a.
-The same immediate switch can be observed as that in A.
-(C) Simulated experiment with a 10 degree adaptation.
-In this case, savings are much slower.
-(D) Data adapted from \cite{Oh_Minimizing_2019}, figure 4a.} \label{fig:oh-2019}
+(C) Same experiment as (B) but  with a 10 degree adaptation.
+In this case, savings are much slower.} \label{fig:oh-2019}
 \end{figure}
 
-Note that in the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
+\cite{Oh_Minimizing_2019} performed a similar experiment in which contexts were not visually cued.
+In addition, the authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10.
+The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
+\cite{Oh_Minimizing_2019} immediate switching in the first experiment, while switching was much slower in the second experiment.
+As can be seen on the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low.
+This difference is explained by our model in terms of the size of the adaptation: the smaller the adaptation is, the more difficult it is for the agent to notice when the context has changed.
+Because of this, the model requires more evidence (i.e.
+more trials) to infer a switch in contexts.
+
+Note that the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
 \todo{Need to be careful here, as de-adaptation is sometimes discussed only in terms of experiments with no contextual cues or where the meaning of the cues is inverted}.
@@
 In our model, adaptation is gated by the uncertainty on the current context.
-More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation, given a fixed size of prediction error.
+More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
 
@@
 
-\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between two possibilities: no force (baseline O) a clockwise force (context A) and a counter-clockwise force (context -A).
-They divided participants into three groups for which the transition from one trial to the next were very uncommon, somewhat common and almost every trial.
+\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between O, A and -A.
+They divided participants into three groups for which the context transition from one trial to the next was very uncommon, somewhat common and almost every trial.
 Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.
@@
 It can be seen that the agent learns more slowly on the more volatile environment ($z = 0.5$) than on the less volatile one, matching the results from \cite{Herzfeld_memory_2014}.
-Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment.
+Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment, a difference that is consistent across all trials.
 
@@
 
-In our model, learning is gated by context inference: the higher the uncertainty on the context in which an error occured, the lower the effect of this error on future trials (i.e. adaptation).
-
-To account for volatility, the model includes a term in context inference which assumes a transition from the previous context to all other possible contexts (see Methods).
+To account for volatility, our model includes a term in context inference which assumes a transition from the previous context to all other possible contexts (see Methods).
 Effectively, this transition term gives the model uncertainty on the current context, regardless of observations.
 The higher the volatility, the higher the uncertainty.
+This higher uncertainty is what lowers the observed learning rate across all trials.
 
@@
 If the identity of the current context is known, the forward model for this context is used to select the current action.
-However, if some uncertainty exists, the selected action is influenced by all the possible current contexts, with a weight directly related to how likely each one of those contexts is.
+However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is.
 
 Experimental evidence supporting this view can be found in experiments with context switching.
-For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group (where -A refers to an adaptation in the opposite direction as A, but with the same magnitude), with a block sequence A-(-A)-A-(-A), keeping the size of the necessary adaptation equal across both groups.
-After A and 3A in one group, A and -A in the other, had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
-The authors interpreted this as evidence that a new adaptation happens more quickly if it is in the same direction as the current adaptation (e.g.
+For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A).
+After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
+The authors interpreted this as evidence that switching between adaptations happens more quickly if it is in the same direction as the current adaptation (e.g.
 both counter-clockwise), and more slowly if they are in the opposite direction (e.g.
@@
 Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force).
-In the first group, this initial pull towards zero accelerates the transition towards A, as $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.
+In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.
 
-We simulated data with our model fitting the experimental structure by \cite{Davidson_Scaling_2004}.
+We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}.
 We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}.
@@
 \centering
-\includegraphics[width=\textwidth]{./figures/figure_3.png}
+\includegraphics{./figures/figure_3.png}
 \caption{Motor error when switching back to a previously learned adaptation.
@@
 
-The model additionally predicts that the difference in observed switching speed between 3A-A and (-A)-A would disappear in an experiment in which the adaptations do not cross the baseline, i.e.
-all are positive or all are negative, or in an experiments in which the contexts are clearly cued.
-\todo{Maybe add simulations that show this, as well as simulations where the baseline context is not present.}
-
 Further evidence can be found during error-clamp blocks in several experiments.
@@
 \subsection{Action selection in error-clamp blocks}
-During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials.
-However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment.
+During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery (i.e.
+displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials.
+However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment [EGs].
 
@@
 \cite{Vaswani_Decay_2013} explored in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
-The authors found that during an E block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous trial.
+The authors found that during an E block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous block.
 After that, the exerted force slowly dropped towards zero throughout tens of trials, but never reaching values around zero.
@@
 
-In \fref{fig:vaswani-2013}A-E, we show the data simulated with our model for the task presented by \cite{Vaswani_Decay_2013}.
-The motor adaptation periods can be observed in \fref{fig:vaswani-2013}A-D for the four groups in the experiment, while on each bottom panel the inference over context is shown.
-Context inference works reliably until the error-clamp trials start, which do not correspond to any of the known contexts.
-This causes the agent to infer the combination of some of the known contexts that best fits the observations.
-Depending on the contexts learned by the agent (which change from group to group), inference during E trials will vary: groups 1.1 and 1.3 have exactly the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
-These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
-In contrast, group 1.3 has the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C, which has a non-zero posterior probility during E trials, pushing the agent's adaptation force more quickly towards zero.
-Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more slowly than groups 1.1 and 1.4.
-
-In \fref{fig:vaswani-2013}E, the same data is shown but only in the E trials.
-Here, the difference in the speed in which the force decays for the different groups is more readily visible.
-This is a strong qualitative match to the data from \cite{Vaswani_Decay_2013}, shown in \fref{fig:vaswani-2013}F.
+In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in In \fref{fig:vaswani-2013}B we show the experimental data adapted from \cite{Vaswani_Decay_2013}.
+The displayed adaptation is shown during the E trials for the three experimental groups in the experiment.
+It can be seen that group 1.3 (i.e.
+the participants who had learned in the -A/2 context in addition to A) more quickly recognizes that a change in context has occurred and lowers the force applied on the mechanical handle, as can be seen in the experimental data.
 
@@
 \caption{Adaptation during error clamp trials.
-(A) On top, adaptation index shown by the agent belonging to group 1.1 at every trial, i.e.
-how much the agent adapted to the force that was present, where 1 is optimal adaptation.
-On bottom, inference over the current context, where contexts are color coded: black for baseline, orange for counter-clockwise force, purple for clockwise force and brown for clockwise force with half strenght.
+(A) Simulated adaptation during the error-clamp trials for the three groups of participants in \cite{Vaswani_Decay_2013}, using the same colors.
+The solid line is the average across 10 runs (i.e.
+10 simulated participants) and the shaded area represents the standard deviation.
+The vertical dashed line is the start of the error-clamp trials.(B) Corresponding experimental data adapted from \cite{Vaswani_Decay_2013}.
+(C) Inference over the current context, where contexts are color coded: black for baseline, orange for the counter-clockwise force, purple for the clockwise force and brown for clockwise force with half strenght.
 The lines represent the posterior probability of each ocontext at every trial, while the background color represents the real context.
 An olive-colored background represents error-clamp trials.
-(B) Same as A, for group 1.2.
-(C) Same as A, for group 1.3.
-(D) Same as A, for group 1.4.
-(E) Simulated data for all groups (same color coding as A-D), only during error-clamp trials.
-Perceptual noise (i.e.
-how well the agent can estimate the position of its own hand) was lowered to increase the clarity of the plotted lines; however, increasing it does not change the results.
-(F) Data adapted from Vaswani et al.
-2013, figure 2C, depicting the data equivalent to (E).}
+As in (A), solid lines represent the average across all runs and shaded areas represent the standard deviation.
+(D) Visualization of the lag before error-clamp trials are detected.
+Each line represents one run (groups 1.1, 1.2 and 1.3, 10 runs per group).}
 \label{fig:vaswani-2013}
@@
 
-As can be seen in FIGURE XXX, our model also reproduces the lag before the tapering off in the E block starts.
+This effect is explained in our model in terms of context inference.
+In \fref{fig:vaswani-2013}C, the inference over context is shown for each group separately.
+Context inference works reliably until the error-clamp trials start, which do not correspond to any of the known contexts.
+This causes the agent to infer the combination of some of the known contexts that best fits the observations.
+Depending on the contexts previously learned by the agent (which change from group to group), inference during E trials will vary: groups 1.1 and 1.4 have exactly the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
+These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
+In contrast, group 1.3 has learned the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C), which has a non-zero posterior probility during E trials, pushing the agent's adaptation force more quickly towards zero.
+Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more slowly than groups 1.1 and 1.4.
+
+Our model also reproduces the variability in the lag before adaptation starts to drop to zero after the E block starts.
 In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
-\todo{Simulate the lag?
-May be worth it}
+To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context (A for groups 1.1, 1.2 and 1.3, -A for 1.4) for each simulated run (one line per run, 10 runs per group, all groups).
+It can be seen that the moment in which the drop starts, as well as the speed of the drops, depends on the run.
 
 \section{Discussion}
-Discuss that our model does not fully match vaswani in how it all decays. This is because our model does not have a mechanism to create new internal models.
+We presented a hierarchical model for motor adaptation based on Bayesian inference, which not only learns to adapt its motor outputs by observing errors, but also infers the context in which these arrors were commited and updates the forward model associated with that context.
 
-During these error-clamp trials, the natural variability of movements leads to prediction error: during a normal (e.g.
-A) block, if the participant exerts less force than necessary by mistake, she observes a motor error which, if the adaptation is complete, produces no prediction error.
-However, in E trials, no error is observed, which leads to a predicted error that is not observed.
-It is this prediction error that lets the participant know that a transition has occurred.
-\cite{Vaswani_Decay_2013} performed an additional experiment in which they masked this lack of error by introducing artificial variability on the end position of the hand that mimicked that seen in the participant.
-Because this eliminated the only source of information regarding a switch of context, participants could no longer infer a switch.
-Consistent with our model, the authors found that no tapering off occured during the E block.
-\todo{Moved from Results}
+Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998a,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
+We built on this studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
+
+We selected representative experimental studies that show savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
+Using our model, we showed how each of these effects can be explained by the specifics of context inference, which integrates all the available information (e.g.
+sensory cues, workspace location, reward and endpoint feedback, etc), in some cases throughout many trials.
+
+It is important to note that the most essential part of the model we presented is the concept of the context and the inclusion of a Bayesian inference scheme over the context, which integrates all the information that is available to the participant during the experiment.
+The mathematical details of the model are not strictly necessary to reproduce the results.
+Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations.
+This choice greatly simplifies our simulations and provides very intuitive intepretations for the results.
+However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998a,Kording_Bayesian_2004}.
+A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.
+
+In what follows, we dive deeper into the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model.
+In addition, we discuss several predictions that our model makes, which could be verified experimentally in future studies.
+
+\subsection{Extended explanatory power}
+Context is often directly observable, either via environmental cues or via feedback observed after actions.
+In many behavioral experiments, for example, the context of the current block of trials is easily identifiable by the color of a cue light \citep[e.g.][]{Ethier_Spontaneous_2008}, the posture of the hand \citep[e.g.][]{Gandolfo_Motor_1996}, the target of the movement \citep[e.g.][]{Lee_Dual_2009} or by the workspace in which the task is being performed \citep[e.g.][]{Shadmehr_Adaptive_1994}.
+The context can also be selected by the agent itself.
+For example, a person can decide the level of care devoted to moving an egg around, which determines the way in which the person would move their hands.
+In these cases, the context is clearly determined and known to the agent and switches between different adaptations are immediate.
+
+In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context, including infinite.
+It is in these cases where the effects of context inference are most noticeable.
+Laboratory experiments with this property are common.
+While many experiments exist that give probabilistic cues [EGs], or cues that only reveal part of the context [EGs], evidence accumulation is not limited to these explicit cases.
+Indeed, as we argued in the Results section, many experiments inadvertently include partial contextual information used by participants.
+
+The most direct secondary contextual information comes in the form of reward and end-point feedback.
+Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
+When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be wrong.
+This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified.; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
+
+The same rationale explains the results from \cite{Davidson_Scaling_2004}: motor learning, which in our model is gated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
+This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
+It is important to note that they did not calculate a learning of exactly zero in these near-maximal-error trials, which suggests that participants are doing \textit{post hoc} learning: after observing the outcome of their actions, they identify that they had inferred the context incorrectly, and can then adapt the movements for the correct context.
+
+A subtler source of information can be found in long pauses between blocks of adaptation trials.
+\cite{Ethier_Spontaneous_2008} performed an experiment where the adaptation blocks were divided into sub-blocks of 60 trials each, with a long 30s pause between sub-blocks.
+The authors observed that at the beginning of each sub-block, participants showed an average dip in their learned adaptation, suggesting that they were partially returning to baseline (O) after each pause.
+This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occured.
+Because this was inferred without any information that points at the identity of the new context, prior beliefs take over.
+This means that participants rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context A, but now includes a component of the baseline O, as it is the most common in their lives.
+
+Error-clamp (E) trials present another insight.
+During E trials, error is artificially kept at zero.
+This takes different forms, depending on the type of experiment.
+If error is kept at zero, one could assume that participants would continue doing what they were doing before, as there is no reason (no observed error) to inferr a change in context.
+However, this is almost never the case \cite[e.g.][]{Smith_Interacting_2006,Ethier_Spontaneous_2008,Forano_Timescales_2020,Vaswani_Decay_2013,Scheidt_Persistence_2000}.
+Instead, participants slowly reduce their adaptation, often displaying spontaneous recovery \cite[e.g.][]{Smith_Interacting_2006}, i.e.
+an initial response in the direction of one of the learned adaptations.
+Our model provides a principled account of this behavior: while the observed error (i.e.
+the cursor on the screen) is kept at zero, participants still receive proprioceptive feedback which tells them that something has changed.
+The natural variability in a participant's behavior will lead her to expect observing errors, which clashes with the observed zero error.
+This prompts the participant to re-evaluate their inferred context, which can partially activate a previously-learned adaptation, as we showed in \fref{fig:vaswani-2013}.
+This can be seen in \cite{Criscimagna-Hemminger_Consolidation_2008}, where the authors showed that introducing long periods before the E block begins lowers the initial force that participants exerted on the mechanical arm during the E block; longer periods of time make context inference revert to the prior expectation that a new baseline block begins, because participants are free to move their arm about during the pause \todo{check this}.
+
+In our account, if all information indicating a change in context is removed from the experiment, participants would continue to behave as they were in the previous block.
+Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where they showed random errors to participants during E trials, with a variance matching that of previously observed motor commands.
+The authors showed that by matching the errors expected by participants, they eliminated the slow tappering off observed in most E blocks.
+As we showed in \fref{fig:vaswani-2013}D, this effect need not be all or none: in some experimental sessions, participants might fail to notice that a change has occurred due to the stochasticity of their own actions.
+This would lead to an observed delay before a participant changes her behavior after the E block has started.
+
+[Maybe: talk about generalization in experiments with different workspaces]
+
+\subsection{Model predictions and future work}
+The basic principle behind the results we presented is that context inference is a process that develops over time and that carries with it uncertainty.
+This uncertainty affects learning and behavior, effecting phenomena that are directly observable during behavioral experiments.
+
+In addition to providing a unifying explanation for phenomena such as switching, savings, spontaneous recovery and slow return to baseline, our model makes unique predictions that could be tested in future experiments.
+
+As we already discussed above, the inclusion of reliable sensory contextual cues (e.g.
+lights whose color uniquely identify a context) makes switching immediate, as in the experiments by \cite{Kim_Neural_2015}.
+We expect that the same effect would be observed in error-clamp trials.
+If the E block is learned by participants during training, it might still be difficult for them to identify that an E block has started, which would create delays similar to those in \fref{fig:vaswani-2013}.
+However, the model predicts that if a visual cue is introduced that identifies the E block, participants would immediately switch to their baseline behavior, no longer displaying spontaneous recovery, lag, nor the slow return to baseline.
+This immediate switch in the presence of contextual cues would persist even if endpoint feedback is manipulated as \cite{Vaswani_Decay_2013} did.
+
+As we discussed in the Results section, the volatility of the environment (i.e.
+unpredictable switching between contexts) lowers the measured learning rate.
+Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can identify the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
+This would manifest in \fref{fig:herzfeld-2014} in the form of no discernible differences between the two groups of participants.
+
+If this prediction is confirmed, our model would additionally suggest that human participants do not revisit the learned adaptation of the last trial when a new observation comes in.
+To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}: at trial $t$, the real context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
+When the outcome is observed, it becomes clear to the participant that the context was B.
+Does the participant update the internal model of A or of B?
+According to our model, the participant incorrectly updates A and, upon learning of her mistake at trial $t+1$, does not revert this update.
+If this were not the case, the volatility of the environment would have no effect, as the context at trial $t$ can almost always be identified at trial $t+1$.
+
+The model also predicts an effect reminescent of multisensory integration: in order to integrate contextual information from conflicting sources (e.g.
+probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its precision (the inverse of its variance).
+Such integration would manifest itself in experiments in which observations are noisy, as in the experiments by \cite{Kording_Bayesian_2004}, in which the position of the finger was obscured and instead participants are shown a blurry cursor which was some times shifted from its real position.
+If the added observation noise gives evidence for a particular context (the real underlying context or another one) and a visual cue gave partial information for another context, the participants' behavior would be more consistent with the most reliable source of contextual information.
 
@@
 own physical space (e.g. hand positions in a reaching task), only the dynamical
-system of \eref{eqn:dynamical-system} would need to be adapted.
+system would need to be adapted.
 
@@
 
-\section{Discussion}
-\begin{enumerate}
-\item The effect of pauses?
-\end{enumerate}
-
 \bibliography{../../MyLibrary}

### 2021-08-24 adaptation/code/figures.py

@@ -29,7 +29,8 @@ def oh_2019_kim_2015(fignum=1, show=True):
     paper.
 
     """
-    fig, axes = plt.subplots(3, 2, clear=True, num=fignum)
+    figsize = (5, 6)
+    fig, axes = plt.subplots(3, 2, clear=True, num=fignum, figsize=figsize)
 
     # a)
     task_kim, agent_pars = sims.kim_2015(plot=False)
@@ -104,6 +105,8 @@ def oh_2019_kim_2015(fignum=1, show=True):
                     transform=axes[2, 0].transAxes,
                     fontdict={'size': 14})
 
+    fig.tight_layout()
+
     plt.savefig(FIGURE_FOLDER + 'figure_{}.png'.format(fignum), dpi=600)
     plt.savefig(FIGURE_FOLDER + 'figure_{}.svg'.format(fignum), format='svg')
     if show:
@@ -115,12 +118,13 @@ def herzfeld_2014(fignum=2, show=True):
     """Reproduces the results from Herzfeld et al 2014.
 
     """
+    figsize = (3, 4)
     colors = {'high': np.array((95, 109, 212)) / 256,
               'low': np.array((212, 198, 95)) / 256}
     optimal = -13
     repeats = 300  # participants * (number of blocks)
     num_trials = 30
-    fig = plt.figure(clear=True, num=fignum)
+    fig = plt.figure(clear=True, num=fignum, figsize=figsize)
     fig, axes = plt.subplots(2, 1, squeeze=True, clear=True,
                              num=fignum, sharex=True)
 
@@ -188,7 +192,7 @@ def herzfeld_2014(fignum=2, show=True):
     sns.lineplot(data=sensitive_panda, y='sensi', x='bin_trial',
                  hue='z_label', palette=colors, ax=axes[1])
     axes[1].set_xlabel('Trials')
-    axes[1].set_ylabel('Sensitivity to error (a.u.)')
+    axes[1].set_ylabel('Sensitivity\nto error (a.u.)')
     yticks = axes[1].get_yticks()
     axes[1].set_yticks(yticks[[0, -1]])
     axes[1].set_yticklabels([0, ''])
@@ -197,6 +201,8 @@ def herzfeld_2014(fignum=2, show=True):
     axes[1].text(x=-0.1, y=1, s='B',
                  transform=axes[1].transAxes,
                  fontdict={'size': 14})
+
+    fig.tight_layout()
 
[…]

### 2021-08-24 adaptation/code/model.py

@@ -61,6 +61,7 @@ class LeftRightAgent(object):
 
     sample_context_mode = 'mode'
     sample_force_mode = 'mode'
+    sample_action = False  # Whether actions are sampled or deterministic
 
     max_force = pars['max_force']  # Maximum force the agent can exert
 
@@ -74,7 +75,8 @@ class LeftRightAgent(object):
 
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None, prediction_noise=None,
-                 reset_after_change=None, force_sds=None, max_force=None):
+                 reset_after_change=None, force_sds=None, max_force=None,
+                 sample_action=None):
         """Initializes the known left and right contexts, as well as
         baseline.
 
@@ -100,6 +102,8 @@ class LeftRightAgent(object):
             self.force_sds = force_sds
         if max_force:
             self.max_force = max_force
+        if sample_action:
+            self.sample_action = sample_action
 
         _, self.magnitudes = self.__init_contexts()
         self.magnitude_history = [self.magnitudes]
@@ -149,7 +153,7 @@ class LeftRightAgent(object):
         The decision is returned and also saved into self.decision_history.
 
         """
-        action = self._make_decision_mixed()
+        action = self._make_decision_mixed(sampled=self.sample_action)
         if abs(action) > self.max_force:
             action = action / abs(action) * self.max_force
         self.action = action
@@ -177,9 +181,10 @@ class LeftRightAgent(object):
         probs /= probs.sum()
         if not sampled:
             return (bests * probs).sum() / probs.sum()
-        sampled_context = np.random.choice(np.arange(len(probs)), p=probs)
-        sampled_decision = bests[sampled_context] + \
-            self.action_sd * self.sample_norm()
+        # sampled_context = np.random.choice(np.arange(len(probs)), p=probs)
+        # sampled_decision = bests[sampled_context] + \
+        #     self.action_sd * self.sample_norm()
+        sampled_decision = bests.dot(probs) + self.action_sd * self.sample_norm()
         return sampled_decision
 
     def _best(self, context):

### 2021-08-24 adaptation/code/simulations.py

@@ -512,11 +512,14 @@ def oh_2019(plot=True, axes=None, fignum=10):
     agent_pars = {'angles': [0, 0],
                   'cue_noise': 1 / 2,
                   'max_force': 50,
-                  'hyper_sd': 1,
-                  'obs_sd': 2.5,
-                  'context_noise': 0.05,
+                  'hyper_sd': 10,
+                  'obs_sd': 3.5,
+                  'context_noise': 0.01,
                   'force_sds': np.ones(2),
-                  'prediction_noise': 1}
+                  'prediction_noise': 0.1,
+                  'action_sd': 2,
+                  'all_learn': True,
+                  'sample_action': True}
 
     task_10 = task_20.copy()
     task_10['forces'] = [[0, 0], [1, 10]]
@@ -527,15 +530,13 @@ def oh_2019(plot=True, axes=None, fignum=10):
 
         agent = model.LRMeanSD(**agent_pars)
 
-        agent.all_learn = True
         sim_and_plot(agent, task_20, axes=axes[:, 0])
         axes[0, 0].set_title('Adaptation: 20')
         axes[0, 0].set_ylim((-25, 25))
         agent = model.LRMeanSD(**agent_pars)
-        agent.all_learn = True
         sim_and_plot(agent, task_10, axes=axes[:, 1])
         axes[0, 1].set_title('Adaptation: 10')
-        axes[0, 1].set_ylim((-12, 12))
+        axes[0, 1].set_ylim((-25, 25))
 
     return task_20, task_10, agent_pars
 
@@ -577,7 +578,10 @@ def kim_2015(plot=True, axis=None, fignum=11):
                   'force_sds': np.ones(3),
                   'max_force': 60,
                   'hyper_sd': 1,
-                  'obs_sd': 3}
+                  'obs_sd': 3,
+                  'all_learn': True,
+                  'learn_rate': 10,
+                  'sample_action': True}
     if plot:
         agent = model.LRMeanSD(**agent_pars)
         sim_and_plot(agent, task, force_labels=[-40, 0, 40],
@@ -620,7 +624,8 @@ def herzfeld_2014(num_trials=150, plot=True, axes=None, fignum=12):
                        'max_force': 20,
                        'hyper_sd': 1000,
                        'learn_rate': 10,
-                       'obs_sd': 2}
[…]

### 2021-08-30 adaptation/article/article.tex
Added abstract. Edited discussion a bit

@@
 \abstract{Abstract}
-Very interesting abstract. Much wow.
+Humans are able to adapt motor commands to changing dynamics of the environment and of the body itself.
+While we must adapt to slow changes in our body dynamics, e.g.
+because of muscle exhaustion throughout the day, we can also adapt to sudden changes in the environment.
+For example, when picking up a heavy objects, we must adapt our motor commands to be able to move the object around.
+The size and material of an object provide us with contextual information which we can use to prepare the correct response to moving the object, and our ability to switch between clear contexts (e.g.
+between lifting a big rock and small piece of paper) has been well studied.
+However, contextual information is not always clear or realiable (e.g.
+a big object of unknown material).
+In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
+We show how our model explains the effects of sensory cues and propriosceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.
 
@@
 The context can also be selected by the agent itself.
-For example, a person can decide the level of care devoted to moving an egg around, which determines the way in which the person would move their hands.
+For example, a person can decide the level of care devoted to moving an egg or a lead ball around, which determines the way in which the person would move their hands.
 In these cases, the context is clearly determined and known to the agent and switches between different adaptations are immediate.
 
-In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context, including infinite.
+In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
 It is in these cases where the effects of context inference are most noticeable.
-Laboratory experiments with this property are common.
-While many experiments exist that give probabilistic cues [EGs], or cues that only reveal part of the context [EGs], evidence accumulation is not limited to these explicit cases.
+While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases.
 Indeed, as we argued in the Results section, many experiments inadvertently include partial contextual information used by participants.
 
-The most direct secondary contextual information comes in the form of reward and end-point feedback.
+Contextual information can come in the form of reward and end-point feedback.
 Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
 When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be wrong.
-This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified.; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
+This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C, as well as those by \cite{Zarahn_Explaining_2008}: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
 
@@
 This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
-It is important to note that they did not calculate a learning of exactly zero in these near-maximal-error trials, which suggests that participants are doing \textit{post hoc} learning: after observing the outcome of their actions, they identify that they had inferred the context incorrectly, and can then adapt the movements for the correct context.
 
@@
 Because this was inferred without any information that points at the identity of the new context, prior beliefs take over.
-This means that participants rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context A, but now includes a component of the baseline O, as it is the most common in their lives.
+This means that participants rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context A, but now includes a component of the baseline O, as it is the most common one in everyday life.
 
@@
 If error is kept at zero, one could assume that participants would continue doing what they were doing before, as there is no reason (no observed error) to inferr a change in context.
-However, this is almost never the case \cite[e.g.][]{Smith_Interacting_2006,Ethier_Spontaneous_2008,Forano_Timescales_2020,Vaswani_Decay_2013,Scheidt_Persistence_2000}.
+However, this is almost never the case \cite[e.g.][]{Smith_Interacting_2006,Ethier_Spontaneous_2008,Forano_Timescales_2020,Vaswani_Decay_2013,Scheidt_Persistence_2000,Pekny_Protection_2011}.
 Instead, participants slowly reduce their adaptation, often displaying spontaneous recovery \cite[e.g.][]{Smith_Interacting_2006}, i.e.
@@
 This prompts the participant to re-evaluate their inferred context, which can partially activate a previously-learned adaptation, as we showed in \fref{fig:vaswani-2013}.
-This can be seen in \cite{Criscimagna-Hemminger_Consolidation_2008}, where the authors showed that introducing long periods before the E block begins lowers the initial force that participants exerted on the mechanical arm during the E block; longer periods of time make context inference revert to the prior expectation that a new baseline block begins, because participants are free to move their arm about during the pause \todo{check this}.
+\cite{Pekny_Protection_2011} found similar results, demonstrating that the duration of the previously-observed adaptation block also affects behavior in the E block.
+Additionally, \cite{Criscimagna-Hemminger_Consolidation_2008} showed that introducing long periods before the E block begins lowers the initial force that participants exerted on the mechanical arm during the E block; longer periods of time make context inference revert to the prior expectation that a new baseline block begins, because participants are free to move their arm about during the pause.
 
@@
 
-[Maybe: talk about generalization in experiments with different workspaces]
-
 \subsection{Model predictions and future work}
@@
 
-As we already discussed above, the inclusion of reliable sensory contextual cues (e.g.
+\subsubsection{Error-clamp as a context}
+The inclusion of reliable sensory contextual cues (e.g.
 lights whose color uniquely identify a context) makes switching immediate, as in the experiments by \cite{Kim_Neural_2015}.
@@
 
-As we discussed in the Results section, the volatility of the environment (i.e.
+\subsubsection{Volatility has no effect on learning}
+The volatility of the environment (i.e.
 unpredictable switching between contexts) lowers the measured learning rate.
@@
 
+\subsubsection{Multi-source integration}
 The model also predicts an effect reminescent of multisensory integration: in order to integrate contextual information from conflicting sources (e.g.

### 2021-08-31 adaptation/article/article.tex
Finished intro.

--- 2021-08-24 adaptation/article/article.tex
@@
 \section{Introduction}
-It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997,}, although interference can occur between tasks \cite[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
+It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
 
 In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement.
-For example, in experiments with a mechanical arm \cite[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
-Other examples include measuring the force profile in error-clamp trials \cite{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \cite{Kim_Neural_2015} and saccade adaptation \cite{Catz_Cerebellardependent_2008}.
+For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
+Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.
 
@@
 In order to explain these phenomena, a number of computational models have been introduced.
-The least complex of these models are linear learners [CITATION NEEDED], in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
+The least complex of these models are linear learners \citep[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
 By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, explaining these distinct experimental phenomena.
 
-Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via prediction error generated by these internal models.
+Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
 These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses, as well as increases the stability of the estimate.
 
-While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of the scope of these models. (1)
+While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of their scope.
+For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}.
+Learning new adaptations has been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013}, as well as the way in which those adaptations were presented [CITATION NEEDED].
 
-[unexplained phenomena. The truth is out there. I want to believe]
+In this work, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
+We expand on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
+In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions.
+To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.
+Having inferred the context, the model can choose the best-fitting internal model with which to make decisions.
+
+We show that the dynamics of context inference provide a unifying explanation for all the experimental phenomena outlined above, providing a single account for motor adaptation under changing contexts, while relying on known mechanisms for adaptation.
 
@@
 
-In what follows, we introduce the different parts of the model and explain how they work together.
-In addition, we will highlight experimental findings in motor adaptation from previous works for which our model provides explanations that are otherwise unavailable.
+We used our model to explain several experimental phenomena found in previous studies.
+We simulated representative experiments from these studies to show how our model explains their findings using the dynamics of context inference.
+In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss how context inference explains the data.
+
+Before presenting our simulations, we will briefly explain the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.
 
@@
 motor adaptation), in order to introduce the terminology that we use throughout this work.
-For a complete mathematical description of the model, see the Methods section.
 
@@
 
-It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \cite{Wolpert_Multiple_1998a} \todo{Is this the original one?
-Include more}.
-While a number of models for motor adaptation exist [EGs], we chose to make use of an exact Bayesian learner.
+It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}.
+We chose to make use of an exact Bayesian learner.
 As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning.
@@
 
-To describe this component of the model, let us return to reaching movements.
+To describe this component of the model, we make use of the example of reaching movements.
 Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model.
@@
 \begin{equation}
-p(\vartheta | s, c) = f(s, c; \beta)
+p(\vartheta | s, c) = f(s, c; \phi) \label{eqn:forward-model}
 \end{equation}
-where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the reach movement), $s$ is the current state (e.g.
-the current position of the hand), $c$ is a motor command and $\beta$ are the parameters of the forward model $f(\cdot)$.
-Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which can be updated during motor adaptation.
+where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the hand during the reaching movement), $s$ is the current state (e.g.
+the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$.
+Throughout our lives, the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.
 
-When an observed outcome does not match the predictions produced by $f$, the brain uses the prediction error to update the parameters $\beta$ of the forward model.
+When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model.
 To describe this adaptation process, our model makes uses of Bayesian inference, such that:
 \begin{equation}
-q(\beta | \vartheta, s, c) \propto p(\vartheta | s, c, \beta)p(\beta)
+q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi)
 \end{equation}
-where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
-$p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta)$ is the previous estimate of $\beta$.
+where $q(\phi | ...)$ represents the new estimate of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
+The likelihood $p(\vartheta | s, c, \phi)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.
 
@@
 the strength and direction of the wind when walking) and of the task at hand (immediate and long-term goals, reward structure, etc).
-Previous models \cite[e.g.][]{Wolpert_Multiple_1998a,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
-$f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e.
+Previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
+two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e.
 the model that best represents the context.
@@
 In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.
-We show how the incorporation of all available contextual information explains many previously-unexplained phenomena from experiments with human subjects, without invoking \textit{ad hoc} model components.
 
@@
 This includes, as before, the outcomes of actions (e.g.
-the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely intependent from them.
-For this section, the most important component of $\vartheta$ is sensory contextual information, i.e.
+the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them.
+For this section, the most important component of $\vartheta$ is contextual sensory information, i.e.
 information that might directly help infer the identity of the current context.
 
-In experimental settings, the sensory contextual information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}.
-All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.
+In experimental settings, the contextual sensory information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}.
+All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.
 
@@
 \end{equation}
-From this distribution, the current action (motor command $c_t$) can be sampled and carried out.
+where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.
 
@@
 
-\subsection{Savings and quick de-adaptation}
+\subsection{Savings}
 Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
-Savings is almost universally observed in human participants [A THOUSAND CITATIONS, c.f.
-papers with volatility or ramping up/down], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}.
+Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
 In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
 
-On the other hand, quick de-adaptation refers to the observation that in an O-A-O paradigm, the adaptation from O to A is much slower than that from A back to O.
+In this section, we discuss savings, and the related phenomenon of deadaptation, in terms of the switching between forward models.
+This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019}, but we show that expanding switching into context inference explains facets of savings that these previous models do not.
 
-In this section, we will show that both phenomena are the two sides of the same coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.
+To show this, we categorized experiments by looking at the amount of contextual information made available to participants.
+In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
+We call these cued-context experiments.
+In other experiments, partial information is available to participants \citep[e.g.][]{Davidson_Scaling_2004,Zarahn_Explaining_2008} in the form of large prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.
 
-To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that made available to participants.
-In some experiments [EGssss], the context is clearly revealed to the participant.
-We call these cued-context experiments.
-In other experiments, partial information is available to participants [EGsss] in the form of prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.
-While there are no experiments in which no contextual information is available, there are some experiments which contain blocks of trials in which no contextual information is available and participants have no way of inferring the current context.
-Abusing our notation, we call these experiments (or rather, these sections of the experiments) uncued experiments.
-In this section, we focus on the two former categories, while the latter is left for the next sections.
+To our knowledge, there are no experiments in which no contextual information is available.
+However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants have no way of identifying the current context.
+These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is still available to participants.
+
+In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.
 
@@
 In fact, the mere act of holding the handle informs participants that something is different from baseline.
-Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands].
+Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
 The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.
@@
 Note that the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
-\todo{Need to be careful here, as de-adaptation is sometimes discussed only in terms of experiments with no contextual cues or where the meaning of the cues is inverted}.
 
@@
 During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery (i.e.
-displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials.
-However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment [EGs].
+displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials \citep{Brennan_Decay_2015}.
+However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
 
@@
 
-Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998a,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
+Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
 We built on this studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
@@
 This choice greatly simplifies our simulations and provides very intuitive intepretations for the results.
-However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998a,Kording_Bayesian_2004}.
+However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
 A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.
@@
 
-The most direct secondary contextual information comes in the form of reward and end-point feedback.
+The most direct secondary contextual information comes in the form of reward and endpoint feedback.
 Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
@@
 
-This model is a Bayesian alternative to the N-state learners from
-\citep[e.g.][]{Lee_Dual_2009}. In this model, there are no explicit different
-timescales and multiple states; instead, the model relies on Bayesian learning
-for motor adaptation, combined with inference over the current context.
-
 Before proceding with the model description, let us introduce nomenclature:
@@
 
-Context inference can work through context-specific (possibly unreliable) cues,
+Context inference can work through context-specific (possibly unreliable) sensory cues,
 or through the dynamics of the system (e.g. expected outputs of motor
@@
 of the principles outlined above. This is the implementation used throughout the
-simulations in the main text.
+simulations in the Results section.
 
@@
 (e.g. clockwise) and $s_t < 0$, error in the other direction. We made this
-choice because, in this work, we would model different experiments with
-different state spaces; but all these experiments have the commonality of
+choice because, in this work modeled different experiments with
+different state spaces, but with the commonality of (at most)
 two-dimensional error spaces. In order to model a specific experiment in its
@@
 
-We assume that the decision-making agent has an internal forward model that
-maps motor commands to outcomes:
+We assume that the decision-making agent ha a baseline internal forward models that
+map motor commands to outcomes:
 \begin{equation}
-p(s_{t+1} | s_t, a_t, m) = f_m(s_t, a_t)
+p(s_{t+1} | s_t, a_t, m) = f_O(s_t, a_t)
 \end{equation}
 where $a_t$ are motor commands (actions). We further assume that participants
-have learned $f_m$ throughout their lives and use it to decide which motor
-command to issue. When a perturbation is introduced, to which the participant
-must adapt, we assume they learn a new forward model $f_m^{\phi}$, where $\phi$
-represents the parameters of the function that participants must learn. For
-example, in a saccade task, $\phi$ represents the new gain. Note that $\phi$
-depends on the context and participants need to use and update multiple
-forward adapted models $f_m^{\phi}$ throughout one experiment.
+have learned $f_O$ throughout their lives and use it to decide which motor
+command to issue under normal circumstances (i.e. no adaptation). When a
+perturbation is introduced, to which the participant must adapt, we assume they
+learn a new forward model $f_m^{\phi}$, where $\phi$ represents the parameters
+of the function that participants must learn. For example, in a saccade task,
+$\phi$ represents the new gain. Note that $\phi$ depends on the context and
+participants may need to use and update multiple forward adapted models
+$f_m^{\phi}$ throughout one experiment.
 
@@
 fovea to the target. These motor commands are issued to counteract intrinsic
-sources of error, such as muscle variability \todo{expand this?}, and extrinsic
+sources of error, such as muscle variability and exhaustion, and extrinsic
 sources of error, such as the experimental manipulations on saccade gain. Thus,
@@
 where $\sigma_\epsilon$ incorporates both intrinsic and extrinsic sources of
-error. The mean of the error is assumed to have no component, as motor
-commands, in the absence of external errors, should produce an error centered
-at zero [CITATION MISSING].
+error and is a free parameter of the model. The mean of the error is assumed to
+have no intrinsic component, as motor commands, in the absence of external
+errors, should produce an error centered at zero [CITATION MISSING].
 
@@
 \end{equation}
-where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
-model. We assume the following hyperparameter values \todo{Justify}:
-\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
-  \nu_f^{(0)} &= 0.1 \\
-  \alpha_f^{(0)} &= 1 \\
-  \beta_f^{(0)} &= 0.5
-\end{align}
-Note that $\mu_f^{(0)}$ is different for each context and is thus provided as a
-vector $(0, -1, 1)$ representing the baseline and two opposing adaptations
-(e.g. clockwise and counter-clockwise visuomotor rotations), respectively. For
-all other hyperparameters, all contexts have the same value. These values
-ensure that the prior over the hyperparameters $(\mu_\phi, \sigma_\phi)$ has the mode
-at $(\mu_\phi^{(0)}, 1)$.
+where $\mu_{\phi}^0, \nu_{\phi}^0, \alpha_{\phi}^0$ and $\beta_{\phi}^0$ are free hyperparameters of the
+model.
 
@@
 
-The effect of the hyperparameters of the priors are worth a note or two, as
-they are complex.
-
+\subsubsection{Interpreting the hyperparameters}
 Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
@@
 
-The hyperparameters $\alpha$ and $\beta$ are a bit more complex. Note that the
-mean of a Gamma distribution is $\beta / (\alpha \nu)$; this mean is being used
+The hyperparameters $\alpha$ and $\beta$ have a more complex effect. Note that the
+mean of a Gamma distribution is $1 / (\alpha \nu)$; this mean is being used
 as the standard deviation of a Gaussian by the rest of the agent (cheating),
@@
 is 0.001 (compare that to the size of the adaptations, around 0.0125), and the
-hyperparameters of the standar deviation are stable during the experiment.
+hyperparameters of the standard deviation are stable during the experiment.
 

### 2021-08-31 adaptation/article/article.tex

@@
 The context can also be selected by the agent itself.
-For example, a person can decide the level of care devoted to moving an egg around, which determines the way in which the person would move their hands.
+For example, a person can decide the level of care devoted to moving an egg or a lead ball around, which determines the way in which the person would move their hands.
 In these cases, the context is clearly determined and known to the agent and switches between different adaptations are immediate.
 
-In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context, including infinite.
+In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
 It is in these cases where the effects of context inference are most noticeable.
-Laboratory experiments with this property are common.
-While many experiments exist that give probabilistic cues [EGs], or cues that only reveal part of the context [EGs], evidence accumulation is not limited to these explicit cases.
+While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases.
 Indeed, as we argued in the Results section, many experiments inadvertently include partial contextual information used by participants.

### 2021-08-31 adaptation/article/article.tex
First draft. Before letting it sit

@@
 For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}.
-Learning new adaptations has been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013}, as well as the way in which those adaptations were presented [CITATION NEEDED].
+Learning new adaptations has also been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013}.
 
@@
 Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.
+A schematic representation of the model can be seen in \fref{fig:model}.
+
+\begin{figure}
+\centering
+\includegraphics[]{./figures/figure_1.png}
+\caption{Description of the model.
+(A) Definition of a context in terms of its components.
+(B) Update equations for context inference, motor adaptation and action selection.
+$\zeta$ represents the identity of the context as a categorical variable, $\vartheta$ are observations, $c_t$ are motor commands at trial $t$, $\phi$ are the parameters of the forward model and $s$ are the hidden states of the environment (e.g.
+the position of the hand).
+Note that because action selection is not performed via Bayes' theorem, the notation $p(c_t)$ is used.
+(C) Schematic representation of the model.
+At trial $t$, an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$).
+After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process repeated.
+Note that the posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
+\label{fig:model}
+\end{figure}
 
@@
 motor adaptation), in order to introduce the terminology that we use throughout this work.
+To see how these three components affect each other, see \fref{fig:model}B and C.
 
@@
 \subsection{Context inference}
-In our model, the idea of context is quite general.
+In our model, the idea of context is quite general (see \fref{fig:model}A).
 The context comprises the relevant elements of the environment (e.g.
-the strength and direction of the wind when walking) and of the task at hand (immediate and long-term goals, reward structure, etc).
+the strength and direction of the wind when walking), of the task at hand (immediate and long-term goals, reward structure, etc) and task-related information internal to the decision-making agent (e.g.
+forward model, decision-making algorithms, motor adaptation equations, expected observations).
 Previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
@@
 \begin{equation}
-p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta)
+p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta) \label{eq:action-selection}
 \end{equation}
@@
 \section{Experimental results}
-In this section, we go through a number of experimentally-observed phenomena and show that context inference, as done in our model, provides a parsimonious explanation for all of them.
+In this section, we go through a number of experimentally-observed phenomena and show that context inference, as done in our model, provides a unifying explanation for all of them.
 
-For clarity, we introduce some terminology before discussing these phenomena.
+For clarity, we first introduce necessary terminology.
 As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force.
@@
 Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise.
-Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point in opposite directions.
+Additionally, abusing notation, we can say that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
 We can also talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
@@
 
-\subsection{Savings}
+\subsection{Savings and slow/fast switching}
 Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
@@
 In this section, we discuss savings, and the related phenomenon of deadaptation, in terms of the switching between forward models.
-This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019}, but we show that expanding switching into context inference explains facets of savings that these previous models do not.
+This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019}, but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.
 
@@
 To our knowledge, there are no experiments in which no contextual information is available.
-However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants have no way of identifying the current context.
+However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information.
 These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is still available to participants.
-
 In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.
@@
 \centering
-\includegraphics{./figures/figure_1.png}
+\includegraphics{./figures/figure_2.png}
 \caption{Savings and de-adaptation.
@@
 The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
-\cite{Oh_Minimizing_2019} immediate switching in the first experiment, while switching was much slower in the second experiment.
+Participants in \cite{Oh_Minimizing_2019} showed immediate switching in the first experiment, while switching was much slower in the second experiment.
 As can be seen on the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low.
-This difference is explained by our model in terms of the size of the adaptation: the smaller the adaptation is, the more difficult it is for the agent to notice when the context has changed.
+This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly identifying the context from the noise due to trial-to-trial variation in motor output.
 Because of this, the model requires more evidence (i.e.
@@
 
-\section{The effects on the rate of adaptation}
-In our model, adaptation is gated by the uncertainty on the current context.
-More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
+\section{Slower adaptation in volatile and unpredictable environments}
+In our model, adaptation is gated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
 
 The most direct evidence for the effects of context inference on the rate of adaptation come from the experiments by \cite{Herzfeld_memory_2014}.
-The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between A and -A (and vice versa) change the speed at which adaptation occurs.
+The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between A and -A (and vice versa), change the speed at which adaptation occurs.
 
 \cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between O, A and -A.
-They divided participants into three groups for which the context transition from one trial to the next was very uncommon, somewhat common and almost every trial.
+They divided participants into three groups for which the context transition from one trial to the next had differing probabilities, depending on the group: for one group transitions were very uncommon, for the second group they were somewhat common and for the third group they were almost every trial.
 Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.
@@
 \centering
-\includegraphics{./figures/figure_2.png}
+\includegraphics{./figures/figure_3.png}
 \caption{Learning under unstable and unpredictable contexts.
@@
 
-To account for volatility, our model includes a term in context inference which assumes a transition from the previous context to all other possible contexts (see Methods).
+To account for volatility, our model includes a term in context inference which assumes a transition can occur from the previous context to all other possible contexts (see Methods).
 Effectively, this transition term gives the model uncertainty on the current context, regardless of observations.
@@
 
-\section{Action selection}
+\section{Uncertainty over contexts affects action selection}
 As with learning, our model selects actions (motor commands) based on context inference.
@@
 Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context.
-As discussed above, action selection in our model is done via a weighted avarage:
-\begin{equation}
-p(a_t | s_t) \propto \displaystyle \sum_{m \in M} p(m | s_t, s_{t-1}, a_{t-1}) p(a_t | m)
-\end{equation}
+As discussed above, action selection in our model is done via a weighted avarage (see \eref{eq:action-selection}).
 In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$.
@@
 in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out.
-Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force).
+Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force).
 In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.
@@
 \centering
-\includegraphics{./figures/figure_3.png}
+\includegraphics{./figures/figure_4.png}
 \caption{Motor error when switching back to a previously learned adaptation.
@@
 No pauses were made during the experiment nor were there any contextual cues, so transitions between blocks were not signaled to participants.
-However, the mechanical arm used in the experiment provided force-feedback to participants.
 
@@
 \centering
-\includegraphics[width=\textwidth]{./figures/figure_4.png}
+\includegraphics[width=\textwidth]{./figures/figure_5.png}
 \caption{Adaptation during error clamp trials.
@@
 Depending on the contexts previously learned by the agent (which change from group to group), inference during E trials will vary: groups 1.1 and 1.4 have exactly the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
-These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
+These agents will slowly lower the force appl ied until eventually reaching zero (after hundreds of trials; not shown).
 In contrast, group 1.3 has learned the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C), which has a non-zero posterior probility during E trials, pushing the agent's adaptation force more quickly towards zero.
@@
 To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context (A for groups 1.1, 1.2 and 1.3, -A for 1.4) for each simulated run (one line per run, 10 runs per group, all groups).
-It can be seen that the moment in which the drop starts, as well as the speed of the drops, depends on the run.
+It can be seen that the moment in which the drop starts, as well as the speed of the drops, depends on the run, reproducing the different lags observed in experimental data.
+This can also be seen directly in context inference, displayed as the shaded areas in \fref{fig:vaswani-2013}C.
 
@@
 
-\subsection{Extended explanatory power}
+\subsection{Alternative explanations}
+For each of the representative studies presented in the Results section, the authors put forward explanations for the phenomena they observed in their data.
+While our model provides a single explanation for all these phenomena, it is not necessarily at odds with those presented by the respective authors and might instead act on top of the explanations the authors presented.
+
+\cite{Oh_Minimizing_2019} posited that the difference between their two groups (~\fref{fig:oh-2019}B and C) can be explained by considering that the smaller visual rotation of 10 degrees caused participants to update their baseline forward model instead of creating a new forward model for the context.
+If such an account were true, it would predict that when participants walk away from the experiment they would have trouble adapting to the no-rotation reality, having to re-learn it.
+Our model instead would suggest that participants would immediatly switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.
+More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
+While we believe our model provides a better explanation for the observations by \cite{Oh_Minimizing_2019}, the effect the authors describe has been documented before in chimpanzees \citep{Kojima_Memory_2004}; however, to our knowledge such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
+
+The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed at which participants adapted to curl forces.
+While it has been shown that sensitivity to error can be modulated \cite{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}, to our knowledge there are no other experiments suggesting that it depends on the history of errors.
+As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data.
+The same can be said of the experiments by \cite{Davidson_Scaling_2004}, in which different contexts involve movement in different directions, which requires the use of different muscles.
+This alone could cause delays in switching unrelated to context inference [CITATION NEEDED on the delay in switching muscles?], but we expect the effects to be insufficient to explain the observed data.
+
+\subsection{Further experimental evidence}
 Context is often directly observable, either via environmental cues or via feedback observed after actions.
@@
 
-One of the strenghts of this model is that its general principle does not
-depend on the specifics of the experiment to which it is applied. However, for
-clarity, we will connect the different parts of the model to different
-experiments.
-
 Before proceding with the model description, let us introduce nomenclature:
@@
 \item $t$: Time at the moment an observation is made, before inference is done.
-\item $\theta_t$: Observation. Note that this is assumed to be a noisy
-observation of the generalized state (see below) given by $\theta_t = h(z_t)$.
+\item $\vartheta_t$: Observation. Note that this is assumed to be a noisy
+observation of the generalized state (see below) given by $\vartheta_t = h(z_t)$.
 \item $s_t$: Hidden state, e.g. the position of the hand in reaching
 experiments. It does not depend on the context.
-\item $a_t$: Action taken after observing $\theta_t$ and inferring the context.
-\item $c_t$: Context inferred after observing $\theta_t$.
+\item $c_t$: Motor command (action) taken after observing $\vartheta_t$ and inferring the context.
+\item $\zeta_t$: Context inferred after observing $\vartheta_t$.
 \item $\omega_t$: Contextual information. This term includes, e.g. visual cues,
 proprioceptive feedback, performance feedback.
-\item $z_t$: Generalized state, $z_t = \{s_t, c_t, \omega_t\}$. Observed noisily
-as $\theta_t$.
+\item $z_t$: Generalized state, $z_t = \{s_t, \zeta_t, \omega_t\}$. Observed noisily
+as $\vartheta_t$.
 \end{itemize}
@@
 \[
-p(z_t, z_{t-1}, a_{t-1}, \theta_t, \theta_{t-1}) = p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
+p(z_t, z_{t-1}, c_{t-1}, \vartheta_t, \vartheta_{t-1}) = p(\vartheta_t | z_t)p(z_t|\vartheta_{t-1}, c_{t-1})
 \]
@@
 \[
-  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
+  q(z_t | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | z_t)p(z_t|\vartheta_{t-1}, c_{t-1})
 \]
@@
 \[
-q(\gamma | z_t, z_{t-1}, a_{t-1}) \propto p(z_t | z_{t-1}, a_{t-1}, \gamma)p(z_{t-1})p(\gamma)
+q(\gamma | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)p(z_{t-1})p(\phi)
 \]
 
-where $\gamma$ are the parameters of the internal model for the dynamics,
-i.e. the function $f: (z_t, a_t) \rightarrow z_{t+1}$, which determines the beliefs of
-the agent regarding how the system evolves after having taken action $a_t$.
+where $\phi$ are the parameters of the internal model for the dynamics,
+i.e. the function $f: (z_t, c_t) \rightarrow z_{t+1}$, which determines the beliefs of
+the agent regarding how the system evolves after having taken action $c_t$.
 
-Having inferred the generalized state (including context), a decision is made
-sampling from a distribution:
+Having inferred the generalized state (including context), a motor command is
+sampled from a distribution:
 \begin{equation}
-p(a_t | z_t, \beta) = \displaystyle\sum_{m \in M}q(C_t)p(a_t | z_t, \beta, m)
+p(c_t | z_t) = \displaystyle\sum_{m \in M}q(\zeta_t)p(c_t | z_t, m)
 \end{equation}
@@
 \begin{equation}
-z_t = \{x, \theta_t, \gamma_t, \omega_t\}
+z_t = \{s, \vartheta_t, \phi_t, \omega_t\}
 \end{equation}
-where $x$ is, e.g. in a reaching task, the current position of the hand in
+where $s$ is, e.g. in a reaching task, the current position of the hand in
 Cartesian coordinates, where the origin is the starting point. We assume that
@@
 \begin{align}
-  q(C_t) &= q(C_t | \omega_t)q(C_t | s_t, s_{t-1}, a_{t-1}) \\ \label{eqn:estimated-context}
-  q(C_t | \omega_t) &\propto p(\omega_t | C_t)p(C_t) \\
-  q(C_t | x_t, x_{t-1}, a_{t-1}) &\propto p(x_t | C_t, x_{t-1}, a_{t-1})p(x_t)p(x_{t-1})p(C_t)
+  q(\zeta_t) &= q(\zeta_t | \omega_t)q(\zeta_t | s_t, s_{t-1}, c_{t-1}) \\ \label{eqn:estimated-context}
+  q(\zeta_t | \omega_t) &\propto p(\omega_t | \zeta_t)p(\zeta_t) \\
+  q(\zeta_t | s_t, s_{t-1}, c_{t-1}) &\propto p(s_t | \zeta_t, s_{t-1}, c_{t-1})p(x_t)p(s_{t-1})p(\zeta_t)
 \end{align}
-The terms $p(C_t)$ refer to the prior probability of the context at the
+The terms $p(\zeta_t)$ refer to the prior probability of the context at the
 beginning of the trial: at the first trial, this refers to prior beliefs over
-which contexts are more common/likely, which we call $p_0(C_t)$. At each
+which contexts are more common/likely, which we call $p_0(\zeta_t)$. At each
 subsequent trial, it incorporates the belief that has so far been accumulated,
@@
 block of trials (see main text) can return an agent to a state in which
-$p_0(C_t)$ becomes relevant.
+$p_0(\zeta_t)$ becomes relevant.
 
@@
 \begin{equation}
-p(s_{t+1} | s_t, a_t, m) = f_O(s_t, a_t)
+p(s_{t+1} | s_t, c_t, m) = f_O(s_t, c_t)
 \end{equation}
-where $a_t$ are motor commands (actions). We further assume that participants
+where $c_t$ are motor commands (actions). We further assume that participants
 have learned $f_O$ throughout their lives and use it to decide which motor
@@
 Given a forward model $f_m$ (or $f_m^\phi$), action selection is done by
-building a probability distribution over possible actions $a_t$ based on how
+building a probability distribution over possible actions $c_t$ based on how
 likely they are to produce zero motor error. For example, in saccade
@@
 \begin{equation}
-p(\epsilon | a_t) \propto p(\epsilon_{\text{intrinsic}} | a_t) + p(\epsilon_{\text{extrinsic}} | a_t)
+p(\epsilon | c_t) \propto p(\epsilon_{\text{intrinsic}} | c_t) + p(\epsilon_{\text{extrinsic}} | c_t)
 \end{equation}
@@
 \begin{equation}
-p(a_t) = argmin_{a_t} p(\epsilon | a_t)
+p(c_t) = argmin_{c_t} p(\epsilon | c_t)
 \end{equation}
@@
 \begin{equation}
-p(\epsilon | a_t) = N(\mu_{\text{extrinsic}}, \sigma_\epsilon)
+p(\epsilon | c_t) = N(\mu_{\text{extrinsic}}, \sigma_\epsilon)
 \end{equation}
@@
 error and is a free parameter of the model. The mean of the error is assumed to
-have no intrinsic component, as motor commands, in the absence of external
-errors, should produce an error centered at zero [CITATION MISSING].
+have no intrinsic component.
 
@@
 \begin{equation}
-p(a_t | ...) \propto \displaystyle \sum_{m \in M} q(C_i)p(a_t | m)
+p(c_t | ...) \propto \displaystyle \sum_{m \in M} q(\zeta_i)p(c_t | m)
 \end{equation}
@@
 \end{equation}
-where $s_{1:t}$ are the observed states up until time $t$. The likelihood will
-be a Gaussian:
+where $s_{1:t}$ are the observed states up until time $t$. The likelihood is a
+Gaussian:
 \begin{align}
-  p(s_t | \theta) &= N(\mu_x, \sigma_x)  \\
-  \mu_x &= \theta_{t-1} + x \\
-  \sigma_x &= \xi_0 1
+  p(s_t | \vartheta) &= N(\mu_s, \sigma_s)  \\
+  \mu_s &= \vartheta_{t-1} + s \\
+  \sigma_s &= \xi_0 1
 \end{align}
-This likelihood is the probability of observing $x$ given the previous estimate
+This likelihood is the probability of observing $s$ given the previous estimate
 of the parameters of the force field $\phi_{t-1}$, and given the dynamics of
@@
 \begin{equation}
-q(\mu_f, \sigma_f | x_t, x_{t-1}, a_{t-1}) \propto p(x_t |
-x_{t-1}, a_{t-1}, C_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
+q(\mu_f, \sigma_f | s_t, s_{t-1}, c_{t-1}) \propto p(s_t |
+s_{t-1}, c_{t-1}, \zeta_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
 \end{equation}
-where $p(x_t | x_{t-1}, a_{t-1}, C_t)$ is a Gaussian distribution centered
-around $\mu_a$ with standard deviation $\sigma_a$, which is a free parameter of
-the model. It is assumed that the standard deviation over outcomes ($\sigma_a$)
+where $p(s_t | s_{t-1}, c_{t-1}, \zeta_t)$ is a Gaussian distribution centered
+around $\mu_c$ with standard deviation $\sigma_c$, which is a free parameter of
+the model. It is assumed that the standard deviation over outcomes ($\sigma_c$)
 is related to that of the parameter $\mu_f$ as follows:
@@
 \begin{align}
-  \mu_\phi^{(t)} &= \frac{\nu_\phi^{(t-1)} \mu_\phi^{(t-1)} + q(C_i)s_t}{\nu_\phi^{(t-1)} + q(C_i)} \\
-  \nu_\phi^{(t)} &= \nu_\phi^{(t-1)} + q(C_i) \\
-  \alpha_\phi^{(t)} &= \alpha_\phi^{(t-1)} + q(C_i) / 2 \\
-  \beta_\phi^{(t)} &= \beta_\phi^{(t-1)} + \frac{q(C_i)\nu_\phi^{(t-1)}}{\nu_\phi^{(t-1)} +
-                  q(C_i)}\frac{\left(s_t - \mu_\phi^{(t-1)}\right)^2}{2}
+  \mu_\phi^{(t)} &= \frac{\nu_\phi^{(t-1)} \mu_\phi^{(t-1)} + q(\zeta_i)s_t}{\nu_\phi^{(t-1)} + q(\zeta_i)} \\
+  \nu_\phi^{(t)} &= \nu_\phi^{(t-1)} + q(\zeta_i) \\
+  \alpha_\phi^{(t)} &= \alpha_\phi^{(t-1)} + q(\zeta_i) / 2 \\
+  \beta_\phi^{(t)} &= \beta_\phi^{(t-1)} + \frac{q(\zeta_i)\nu_\phi^{(t-1)}}{\nu_\phi^{(t-1)} +
+                  q(\zeta_i)}\frac{\left(s_t - \mu_\phi^{(t-1)}\right)^2}{2}
 \end{align}
-where $q(C_i)$ is the posterior probability of context $C_i$ from
-\eref{eqn:estimated-context}. Note that when $q(C_i) = 0$, the model $i$ is not
+where $q(\zeta_i)$ is the posterior probability of context $\zeta_i$ from
+\eref{eqn:estimated-context}. Note that when $q(\zeta_i) = 0$, the model $i$ is not
 updated.
@@
 \subsubsection{Interpreting the hyperparameters}
+\label{sec:interpreting-hyperparameters}
 Naturally, $\mu$ affects the initial estimate of the adaptation, in the same

### 2021-08-31 adaptation/code/figures.py

@@ -21,7 +21,9 @@ import pars
 FIGURE_FOLDER = '../article/figures/'
 
 
-def oh_2019_kim_2015(fignum=1, show=True):
+# Note: Figure 1 is made by hand in inkscape
+
+def oh_2019_kim_2015(fignum=2, show=True):
     """Reproduces the results from Oh_Minimizing_2019 and kim_neural_2015 and plots
     them in a format similar to their plots.
 
@@ -114,7 +116,7 @@ def oh_2019_kim_2015(fignum=1, show=True):
         plt.show(block=False)
 
 
-def herzfeld_2014(fignum=2, show=True):
+def herzfeld_2014(fignum=3, show=True):
     """Reproduces the results from Herzfeld et al 2014.
 
     """
@@ -211,7 +213,7 @@ def herzfeld_2014(fignum=2, show=True):
         plt.show(block=False)
 
 
-def davidson_2004(fignum=3, show=True):
+def davidson_2004(fignum=4, show=True):
     """Reproduces the results from Davidson_Scaling_2004, leaving an empty
     subplot to put in the results from their paper.
 
@@ -256,7 +258,7 @@ def davidson_2004(fignum=3, show=True):
         plt.show(block=False)
 
 
-def vaswani_2013(fignum=4, show=True, pandota=None):
+def vaswani_2013(fignum=5, show=True, pandota=None):
     """Reproduces the results from Vaswani_Decay_2013, especifically their
     figures 2a-c.
 

### 2021-09-09 adaptation/article/article.tex
First draft.

@@
 It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
+Additionally, humans have been shown to dynamically switch between different learned tasks.
+What prompts these switches is still a matter of research.
 
@@
 By introducing blocks of trials in which a manipulation is present (e.g.
-the mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error.
+a mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error.
 A number of phenomena have been observed that persist across participants and across many different motor adaptation experiments: savings, quick de-adaptation, anterograde interference and spontaneous recovery.
 
-In order to explain these phenomena, a number of computational models have been introduced.
+To explain these phenomena, a number of computational models have been introduced.
 The least complex of these models are linear learners \citep[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
-By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, explaining these distinct experimental phenomena.
+By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, partially explaining these distinct experimental phenomena.
 
@@
 
-In this work, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
+In this work, we show that these unexplained phenomena, alongside savings, de-adaptation, anterograde interference and spontaneous recovery, can be explained in terms of the concept of context, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
+
+To this end, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
 We expand on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
@@
 
-\subsection{The model}
+\subsection{Modeling context-dependant adaptation}
 The model we present has three main components: (1) context-inference, (2) motor adaptation and (3) action selection.
@@
 More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
-While we believe our model provides a better explanation for the observations by \cite{Oh_Minimizing_2019}, the effect the authors describe has been documented before in chimpanzees \citep{Kojima_Memory_2004}; however, to our knowledge such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
+While we believe our model provides a better explanation for the observations by \cite{Oh_Minimizing_2019}, the effect the authors describe has been documented before in chimpanzees \citep{Kojima_Memory_2004}\todo{@Stefan: you mentioned once you knew of papers talking about this.
+Do you have any citations I could add?}; however, to our knowledge such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
 
 The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed at which participants adapted to curl forces.
-While it has been shown that sensitivity to error can be modulated \cite{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}, to our knowledge there are no other experiments suggesting that it depends on the history of errors.
+While it has been shown that sensitivity to error can be modulated \cite{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}), to our knowledge there are no other experiments suggesting that it depends on the history of errors.
 As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data.
 The same can be said of the experiments by \cite{Davidson_Scaling_2004}, in which different contexts involve movement in different directions, which requires the use of different muscles.
-This alone could cause delays in switching unrelated to context inference [CITATION NEEDED on the delay in switching muscles?], but we expect the effects to be insufficient to explain the observed data.
+This alone could cause delays in switching unrelated to context inference\todo{Need citation for this}, but we expect the effects to be insufficient to explain the observed data.
 

### 2021-09-09 adaptation/article/article.tex
Update on Overleaf.

@@
 between lifting a big rock and small piece of paper) has been well studied.
-However, contextual information is not always clear or realiable (e.g.
+However, contextual information is not always clear or reliable (e.g.
 a big object of unknown material).
 In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
-We show how our model explains the effects of sensory cues and propriosceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.
+We show how our model explains the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.
 
@@
 
-Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context.
-As discussed above, action selection in our model is done via a weighted avarage (see \eref{eq:action-selection}).
+Our model offers a different explanation for the observed results: the asymmetry is due to the existence of the baseline context.
+As discussed above, action selection in our model is done via a weighted average (see \eref{eq:action-selection}).
 In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$.
@@
 (A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}.
-On the y axis, the error is shown in arbitrary units, related to centimiters through a monotonically increasing transformation with the same origin.
+On the y axis, the error is shown in arbitrary units, related to centimeters through a monotonically increasing transformation with the same origin.
 (B) Data adapted from \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
@@
 The authors observed that at the beginning of each sub-block, participants showed an average dip in their learned adaptation, suggesting that they were partially returning to baseline (O) after each pause.
-This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occured.
+This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occurred.
 Because this was inferred without any information that points at the identity of the new context, prior beliefs take over.
@@
 This takes different forms, depending on the type of experiment.
-If error is kept at zero, one could assume that participants would continue doing what they were doing before, as there is no reason (no observed error) to inferr a change in context.
+If error is kept at zero, one could assume that participants would continue doing what they were doing before, as there is no reason (no observed error) to infer a change in context.
 However, this is almost never the case \cite[e.g.][]{Smith_Interacting_2006,Ethier_Spontaneous_2008,Forano_Timescales_2020,Vaswani_Decay_2013,Scheidt_Persistence_2000,Pekny_Protection_2011}.
@@
 Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where they showed random errors to participants during E trials, with a variance matching that of previously observed motor commands.
-The authors showed that by matching the errors expected by participants, they eliminated the slow tappering off observed in most E blocks.
+The authors showed that by matching the errors expected by participants, they eliminated the slow tapering off observed in most E blocks.
 As we showed in \fref{fig:vaswani-2013}D, this effect need not be all or none: in some experimental sessions, participants might fail to notice that a change has occurred due to the stochasticity of their own actions.
@@
 \subsubsection{Multi-source integration}
-The model also predicts an effect reminescent of multisensory integration: in order to integrate contextual information from conflicting sources (e.g.
+The model also predicts an effect reminiscent of multisensory integration: in order to integrate contextual information from conflicting sources (e.g.
 probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its precision (the inverse of its variance).
@@
 
-Before proceding with the model description, let us introduce nomenclature:
+Before proceeding with the model description, let us introduce nomenclature:
 \begin{itemize}
@@
 
-The update equations for the magnitud parameters are given by:
+The update equations for the magnitude parameters are given by:
 \begin{equation}
@@
 in the face of evidence; In principle, enough evidence should still move it,
-but that won't happen during an experiment. Smaller values (e.g. 1 / force\_sd,
+but that will not happen during an experiment. Smaller values (e.g. 1 / force\_sd,
 as is the default on the code) make the hypermean (and thus the mean) follow
 evidence more freely. Note that as more observations are accumulated, $\nu$
-becames bigger and bigger, solidifying the value of the hypermean.
+becomes bigger and bigger, solidifying the value of the hypermean.
 
@@
 mean of a Gamma distribution is $1 / (\alpha \nu)$; this mean is being used
-as the standard deviation of a Gaussian by the rest of the agent (cheating),
+as the standard deviation of a Gaussian by the rest of the agent,
 which makes it an important measure of uncertainty. While setting the default
@@
 large at the beginning and during the experiment is greatly reduced, but never
-to a point where it's ``visually'' too small. Changing this 0.5 would make the
+to a point where it is too small. Changing this 0.5 would make the
 standard deviation change more quickly, making the model more or less precise
@@
 
-\bibliography{../../MyLibrary}
+\bibliography{MyLibrary}
 

### 2021-09-09 adaptation/article/article.tex
Added repetitions to simulations in fig4

@@
 It can be seen that the agent exposed to the -8 context shifts back to 4 more slowly than the one exposed to +8, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.
+In our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data.
+However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.
 
@@
 Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from +8 (yellow) or -8 (blue) to 4 is shown.
-(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}.
+(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}, running eight simulations per group, each simulation representing one participant.
 On the y axis, the error is shown in arbitrary units, related to centimeters through a monotonically increasing transformation with the same origin.
+The shaded area is the SEM across simulated runs (participants).
 (B) Data adapted from \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
@@
 
-\bibliography{MyLibrary}
+\bibliography{../../MyLibrary}
 

### 2021-09-09 adaptation/code/figures.py

@@ -218,10 +218,11 @@ def davidson_2004(fignum=4, show=True):
     subplot to put in the results from their paper.
 
     """
+    repeats = 8  # No. of participants per group
     figsize = (4, 3)
-    colors = [np.array((95, 109, 212)) / 256,
-              np.array((212, 198, 95)) / 256]
-    ran = [161, 200]
+    colors = {'-8': np.array((95, 109, 212)) / 256,
+              '+8': np.array((212, 198, 95)) / 256}
+    ran = [161, 201]
 
     fig, axes = plt.subplots(1, 2, num=fignum, clear=True, squeeze=True,
                              figsize=figsize, sharex=True, sharey=True)
@@ -233,13 +234,24 @@ def davidson_2004(fignum=4, show=True):
     tasks = [task_m8, task_p8]
     agents = [agent_m8, agent_p8]
     names = ['-8', '+8']
+    data = {name: [] for name in names}
     for idx, (task, agent, name) in enumerate(zip(tasks, agents, names)):
-        pandata, pandagent, _ = thh.run(agent, pars=task)
-        pandota = thh.join_pandas(pandata, pandagent)
-        error = np.abs(pandota['pos(t)'])[ran[0]:ran[1]]
-        axes[0].plot(np.arange(ran[1] - ran[0] + 1), error, color=colors[idx],
-                     label=name)
-    axes[0].legend()
+        for idx in range(repeats):
+            pandata, pandagent, _ = thh.run(agent, pars=task)
+            pandota = thh.join_pandas(pandata, pandagent)
+            data[name].append(pandota)
+        data[name] = pd.concat(data[name])
+        data[name]['pos(t)'] = np.abs(data[name]['pos(t)'])
+    data = pd.concat(data, axis=0, names=['Group', 'trial'])
+    data.reset_index('Group', inplace=True)
+    sns.lineplot(data=data, x='trial', y='pos(t)', ax=axes[0],
+                 hue='Group', palette=colors, ci='sd')
+    axes[0].set_xlim(ran)
+    ticks = np.array(axes[0].get_xticks(), dtype=int) - ran[0] + 1
+    axes[0].set_xticklabels(ticks)
+    # axes[0].plot(np.arange(ran[1] - ran[0] + 1), error[name], color=colors[idx],
+    #              label=name)
+    # axes[0].legend()
     axes[0].set_xlabel('Trials after switch')
     axes[1].set_xlabel('Trials after switch')
     axes[0].set_ylabel('Error (a.u.)')

### 2021-09-20 adaptation/article/article.tex
Update on Overleaf.

@@
 \shorttitle{Context-inference-dependent motor learning}
-\author[1]{Cuevas Rivera, Darío}
-\author[1]{Kiebel, Stefan J.}
+\author[1,2]{Cuevas Rivera, Darío}
+\author[1,2]{Kiebel, Stefan J.}
 \affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
+\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
 \affiliation{~}
@@
 \abstract{Abstract}
-Humans are able to adapt motor commands to changing dynamics of the environment and of the body itself.
+ Humans are able to adapt their movements to changes in the dynamics of the environment and of their own bodies, a phenomenon known as motor adaptation.
+Motor adaptation has been shown to be contextual in experiments in which participants must learn to adapt to different altered dynamics, and must be able to switch between the learned adaptations dynamically.
+In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
+We show how our model provides a unifying explanation for the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.
+
+\section{Introduction}
+Humans are able to adapt movements to changing dynamics of the environment and of the body itself.
 While we must adapt to slow changes in our body dynamics, e.g.
 because of muscle exhaustion throughout the day, we can also adapt to sudden changes in the environment.
-For example, when picking up a heavy objects, we must adapt our motor commands to be able to move the object around.
-The size and material of an object provide us with contextual information which we can use to prepare the correct response to moving the object, and our ability to switch between clear contexts (e.g.
-between lifting a big rock and small piece of paper) has been well studied.
-However, contextual information is not always clear or reliable (e.g.
-a big object of unknown material).
-In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
-We show how our model explains the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.
+For example, when picking up a heavy object, we must adapt our motor commands to be able to move the object around.
+Even before picking it up, the size and material of an object provide us with contextual information which we can use to prepare the correct posture for moving the object.
+However, contextual information cannot always be relied on, e.g.
+a big object of unknown material.
 
-\section{Introduction}
-It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
-Additionally, humans have been shown to dynamically switch between different learned tasks.
-What prompts these switches is still a matter of research.
+It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}, which slows down learning.
+Additionally, humans have been shown to dynamically switch between different learned tasks \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
+The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the way in which switching can be impeded or improved, are still an active topic of research.
 
-In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement.
+In order to study motor adaptation, motor error is defined in experiments as the deviation between the observed movements of a participant and a baseline movement.
 For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
@@
 a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation.
-Several models exist for error-based motor adaptation, but we expand on the existing models by adding an explicit component of context inference which guides the selection of the adequate internal model, its updating and the sampling of actions.
+Several models exist for such error-based motor adaptation, but we expand on existing models by adding an explicit component of context inference.
+As we will show, this context inference guides the selection of the adequate internal model, its updating and the sampling of actions.
 
-We used our model to explain several experimental phenomena found in previous studies.
-We simulated representative experiments from these studies to show how our model explains their findings using the dynamics of context inference.
-In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss how context inference explains the data.
+In what follows, we show that the proposed model explains several experimental phenomena found in previous studies.
+We simulated representative experiments from these studies to illustrate how our model explains these experimental findings using the dynamics of context inference.
+In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss in detail how context inference explains the experimental phenomena.
 
-Before presenting our simulations, we will briefly explain the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.
+Before presenting our simulations, we will briefly go through the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.
 
-\subsection{Modeling context-dependant adaptation}
-The model we present has three main components: (1) context-inference, (2) motor adaptation and (3) action selection.
-Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.
-A schematic representation of the model can be seen in \fref{fig:model}.
+\subsection{Modeling context-dependent adaptation}
+The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection.
+The processes defined by these component happen in this order, and each component informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.
+
+Central to the model is the concept of context.
+In our model, contexts are defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the variables of the environment.
+Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to identify the context.
+For example, when lifting an object of unknown weight, an agent might have learned one context for heavy objects and one for light objects.
+When observing an object to be lifted, the agent can use its size and texture to estimate the weight of the object, which in turn allows the agent to infer the appropriate context and, with it, decide how to lift the object.
+These contextual cues (size and texture) do not necessarily identify the context (e.g.
+it could be a movie prop that looks like a rock).
 
@@
 the position of the hand).
-Note that because action selection is not performed via Bayes' theorem, the notation $p(c_t)$ is used.
 (C) Schematic representation of the model.
-At trial $t$, an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$).
-After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process repeated.
-Note that the posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
+The arrows show flow of information, i.e.
+the dependencies in the update equations, as well as the flow of a single trial.
+The elements (e.g.
+$q(\zeta)$) are those shown in B.
+Trial $t$ starts at the top when an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$).
+After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process is repeated.
+The posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
 \label{fig:model}
@@
 motor adaptation), in order to introduce the terminology that we use throughout this work.
-To see how these three components affect each other, see \fref{fig:model}B and C.
+For an overview, \fref{fig:model}B and C shows how these three components affect each other.
 
@@
 Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors.
-For example, we can learn to perform a reaching movement in front of us underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
+For example, we can learn to perform a reaching movement even when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
 
 It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}.
-We chose to make use of an exact Bayesian learner.
-As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning.
+Here, we use this principle and chose to make use of an exact Bayesian learner.
+As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials.
+As we will show below this feature will be useful to explain specific key phenomena parsimoniously.
 We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.
@@
 To describe this component of the model, we make use of the example of reaching movements.
-Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model.
+It is now widely assumed that humans learn associations between motor commands and outcomes in different environments \cite{Wolpert_Multiple_1998}.
+These associations, called forward models, are used by humans for action selection.
 We can write this forward model as follows:
@@
 \end{equation}
-where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the hand during the reaching movement), $s$ is the current state (e.g.
-the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$.
-Throughout our lives, the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.
+where $\vartheta$ are the outcomes of an action (e.g.
+the final position of the hand in a reaching movement), $s$ is the current state (the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$.
+We assume that during learning the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.
 
 When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model.
-To describe this adaptation process, our model makes uses of Bayesian inference, such that:
+To describe this adaptation process, our model makes uses of Bayesian inference, such that
 \begin{equation}
@@
 \end{equation}
-where $q(\phi | ...)$ represents the new estimate of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
-The likelihood $p(\vartheta | s, c, \phi)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.
+where $q(\phi | ...)$ represents the new estimated probability distribution over the possible values of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
+The likelihood $p(\vartheta | s, c, \phi)$ is the probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.
 
-\subsection{Context inference}
+\subsubsection{Context inference}
 In our model, the idea of context is quite general (see \fref{fig:model}A).
 The context comprises the relevant elements of the environment (e.g.
-the strength and direction of the wind when walking), of the task at hand (immediate and long-term goals, reward structure, etc) and task-related information internal to the decision-making agent (e.g.
+the strength and direction of the wind when walking), of the task at hand (e.g.
+immediate and long-term goals and reward structure) and task-related information internal to the decision-making agent (e.g.
 forward model, decision-making algorithms, motor adaptation equations, expected observations).
-Previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
+Most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
 two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e.
 the model that best represents the context.
+In addition, \cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.
 
-In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.
+Following the these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.
 
-Identifying any possible context with a categorical variable $\zeta$, the identification of a context is done by our model through Bayesian inference:
+Context is a categorical variable $\zeta$, where the identification of a context is done through Bayesian inference:
 \begin{equation}
@@
 \end{equation}
-where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall estimation of observing any one context, if the context did change.
-Below, we discuss how these two components effect different experimentally observed phenomena.
+where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall expectation of observing any one context, if the context did change.
+Below, we discuss how these two sources of information effect different experimentally observed phenomena.
 
@@
 This includes, as before, the outcomes of actions (e.g.
-the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them.
+the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them, as is the case for visual cues in experiments \cite[e.g.][]{Addou_Colored_2011}.
 For this section, the most important component of $\vartheta$ is contextual sensory information, i.e.
@@
 
-In experimental settings, the contextual sensory information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}.
-All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.
+In experimental settings, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}.
+All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to infer the identity of a context.
 
-By making use of NormalGamma priors on context, the model performs exact Bayesian inference at every trial. The exact update equations can be found in the Methods section.
-
-\subsection{Action selection}
+\subsubsection{Action selection}
 Action selection is affected by the current context via the context-dependent forward model that is used.
@@
 where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.
+\todo{Maybe add a section with the progression of a trial}
+\section{Experimental results}
+In this section, we present experimentally-observed phenomena in four sections, and show that context inference, as done in our model, provides a unifying explanation for all of them.
+In the first section, we discuss switches between contexts, and how slow context inference affects these switches.
+In the second section, we focus on the effects of volatility in the environment on motor adaptation.
+In the third section, we discuss the effects of slow context inference on action selection.
+Finally, in the fourth section we discuss context inference during error-clamp trials, and its effect on behavior.
 
-\section{Experimental results}
-In this section, we go through a number of experimentally-observed phenomena and show that context inference, as done in our model, provides a unifying explanation for all of them.
-
-For clarity, we first introduce necessary terminology.
-As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force.
+For clarity, we first introduce necessary terminology that is typically used in experimental studies.
+As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force (i.e.
+a force that pushes the participant's hand in the direction perpendicular to movement).
 Depending on the trial, the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
@@
 Additionally, abusing notation, we can say that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
-We can also talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
+Studies often talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
 clockwise), but the second one has half the magnitude.
-Finally, error-clamp trials, in which the mechanical arm forces the participant to make straight-line movements, are represented with the letter E.
+Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter E.
 
@@
 
+Additionally, we use the term "learned adaptation" and similar to refer to motor outputs (e.g.
+reaching movements) that have already been adapted by the participant to a change in the dynamics of the environment.
+For example, the learned adaptation during a task with a curl force refers to the movements of the participant that already counteract the force exerted by the mechanical arm.
+
 \subsection{Savings and slow/fast switching}
-Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
+The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
 Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
-In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
+In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
 
-In this section, we discuss savings, and the related phenomenon of deadaptation, in terms of the switching between forward models.
-This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019}, but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.
+In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models.
+This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019} but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.
 
-To show this, we categorized experiments by looking at the amount of contextual information made available to participants.
+To show this, we examined multiple experimental studies in which savings is displayed.
+We categorized these studies based on the amount of contextual information made available to participants.
 In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
@@
 However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information.
-These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is still available to participants.
+These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is nevertheless available to participants.
 In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.
 
-Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use.
+Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
 For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
@@
 Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
-The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.
+The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.
 
-To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies which differ in the amount of contextual information.
+To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \cite{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information.
 \cite{Kim_Neural_2015} performed a visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
@@
 In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
-The correspondence between simulations and experimental results is quite evident, both in the immediacy of the switches as well as in the fact that learning continues from the state it was when the context was last observed.
+The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e.
+when the solid black line (representing the real context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.
+However, adaptation is incomplete (i.e.
+the blue line never reaches 40 or -40) in both experimental data and simulations until late in the experiment (around trial 500).
 
@@
 Data from our simulations (left column) compared to data from \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} (right column).
+Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown.
 In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number.
@@
 the size of the visuomotor rotation during the task.
-(A) Experiment by \cite{Kim_Neural_2015} with two rotations (-40 and 40 degrees), in addition to baseline.
-The switches between contexts were visually cued and therefore switches are immediate, as can be seen in both simulations (left) and the experimental data from \cite{Kim_Neural_2015} (right).
-(B) An experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements.
-It can be seen in both the simulations (left) and experimental data from \cite{Oh_Minimizing_2019} (right) that after context switches (e.g.
-at trial 120), the agent returns to the adaptation of the previous block, with one or two trials of delay.
-(C) Same experiment as (B) but  with a 10 degree adaptation.
-In this case, savings are much slower.} \label{fig:oh-2019}
+(A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O.
+Participants must adapt to the rotations during shooting movements.
+The y-axis is in degrees; the black line represents the real visuomotor rotation, at either zero or 20 degrees, and the blue line shows the participants' adaptation (averaged across all participants), which can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation.
+When black and blue lines match, the participant perfectly counteracted the visuomotor rotation and reached the target.
+(B) An experiment by \cite{Oh_Minimizing_2019} similar to A, but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced.
+O-A and A-O transitions (i.e.
+adaptation and de-adaptation) can be observed.
+Blue and black lines are as in A.
+(C) Same experiment as (B) but with a 10 degree adaptation.} \label{fig:oh-2019}
 \end{figure}
 
-\cite{Oh_Minimizing_2019} performed a similar experiment in which contexts were not visually cued.
-In addition, the authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10.
+\cite{Oh_Minimizing_2019} performed partially-cued experiment similar to \cite{Kim_Neural_2015}.
+The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10.
 The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
-Participants in \cite{Oh_Minimizing_2019} showed immediate switching in the first experiment, while switching was much slower in the second experiment.
-As can be seen on the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low.
+Participants in \cite{Oh_Minimizing_2019} showed immediate switching in the first experiment, which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (real rotation).
+In the second experiment, switching happens more slowly, with adaptation (blue line) lagging behind the switches in real context (black line), and slowly catching up.
+As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low.
 This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly identifying the context from the noise due to trial-to-trial variation in motor output.
@@
 
-\section{Slower adaptation in volatile and unpredictable environments}
-In our model, adaptation is gated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
+\subsection{Slower adaptation in volatile and unpredictable environments}
+One central element in our model is that adaptation is gated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
 
-The most direct evidence for the effects of context inference on the rate of adaptation come from the experiments by \cite{Herzfeld_memory_2014}.
+Direct experimental evidence for the effects of context inference on the rate of adaptation is provided by \cite{Herzfeld_memory_2014}.
 The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between A and -A (and vice versa), change the speed at which adaptation occurs.
@@
 \cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between O, A and -A.
-They divided participants into three groups for which the context transition from one trial to the next had differing probabilities, depending on the group: for one group transitions were very uncommon, for the second group they were somewhat common and for the third group they were almost every trial.
+Participants were divided into three groups for which the context transition from one trial to the next had differing probabilities, depending on the group: for the first group, transitions were very uncommon, for the second group they were somewhat common and for the third group they were occurred at almost every trial.
 Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.
@@
 In \fref{fig:herzfeld-2014}A we show the adaptation as a function of trial number, averaged across all 300 runs.
-It can be seen that the agent learns more slowly on the more volatile environment ($z = 0.5$) than on the less volatile one, matching the results from \cite{Herzfeld_memory_2014}.
+It can be seen that the agent matches the findings from \cite{Herzfeld_memory_2014} where adaptation is slower in the more volatile environment ($z = 0.5$) than in the less volatile one.
 Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment, a difference that is consistent across all trials.
@@
 \includegraphics{./figures/figure_3.png}
-\caption{Learning under unstable and unpredictable contexts.
-30 trials were simulated in the task defined by \cite{Herzfeld_memory_2014}, repeated 300 times for better statistics.
+\caption{Learning under uncertainty about contexts.
+30 trials were simulated in the task defined by \cite{Herzfeld_memory_2014} and repeated 300 times to estimate the mean effects.
 Two conditions were used, following the naming convention from \cite{Herzfeld_memory_2014}: $z=0.5$ is an unpredictable condition, where the context has a 0.5 probability of switching from trial to trial.
@@
 (B) Error sensitivity, binned into bins of 5 trials, averaged across all runs.
-The error is in arbitrary units, causing sensitivity to error (y axis) to also be in arbitrary units.
+The error and consequently sensitivity to error (y axis) is in arbitrary units.
 Shaded areas are SEM.}
@@
 
-\section{Uncertainty over contexts affects action selection}
+\subsection{Uncertainty over contexts affects action selection}
 As with learning, our model selects actions (motor commands) based on context inference.
 If the identity of the current context is known, the forward model for this context is used to select the current action.
-However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is.
+However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eq:action-selection}).
 
@@
 
-Our model offers a different explanation for the observed results: the asymmetry is due to the existence of the baseline context.
+Our model offers a different and more parsimonious\todo{Is this the right word?} explanation for the observed results: the asymmetry is due to the existence of the baseline context.
 As discussed above, action selection in our model is done via a weighted average (see \eref{eq:action-selection}).
-In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$.
+In experiments without cued contexts, the baseline context O and its associated forward model $m_O$, have a non-zero probability $p(\zeta | s_t ...)$.
 When a new block of trials starts (e.g.
@@
 However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.
+While we focused here on a specific, well-known experiment, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion section.
 
@@
 
-Further evidence can be found during error-clamp blocks in several experiments.
+Further evidence for this effect of uncertainty about contexts can be found during error-clamp blocks in several experiments.
 We discuss these experiments in the following section.
@@
 \subsection{Action selection in error-clamp blocks}
-During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery (i.e.
-displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials \citep{Brennan_Decay_2015}.
-However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
+During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Spontaneous recovery, during which participants' behavior is consistent with a previously-encountered context (e.g.
+in an O-A-E experiment, consistent with A); this phase, when present, is seen during the early trials of the E block.
+(2) A slow return to baseline (i.e.
+no observable adaptation, consistent with the baseline context), which can last as long as hundreds of trials \citep{Brennan_Decay_2015}.
+However, the direction of the spontaneous recovery, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
 
@@
 
-\cite{Vaswani_Decay_2013} explored in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
+\cite{Vaswani_Decay_2013} studied in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
 The authors found that during an E block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous block.
@@
 
-In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in In \fref{fig:vaswani-2013}B we show the experimental data adapted from \cite{Vaswani_Decay_2013}.
+In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
 The displayed adaptation is shown during the E trials for the three experimental groups in the experiment.
 It can be seen that group 1.3 (i.e.
-the participants who had learned in the -A/2 context in addition to A) more quickly recognizes that a change in context has occurred and lowers the force applied on the mechanical handle, as can be seen in the experimental data.
+the participants who had learned in the -A/2 context in addition to A) more quickly recognizes a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.
 
@@
 (A) Simulated adaptation during the error-clamp trials for the three groups of participants in \cite{Vaswani_Decay_2013}, using the same colors.
+The groups differ in the sequence of adaptations: 1.1 performed an A-E sequence; 1.2, O-A-E; 1.3, (-A/2)-A-E; and 1.4, (-A)-E.
+Following \cite{Vaswani_Decay_2013}, group 1.4 is not shown in A and B, as their behavior is identical to group 1.1.
 The solid line is the average across 10 runs (i.e.
-10 simulated participants) and the shaded area represents the standard deviation.
-The vertical dashed line is the start of the error-clamp trials.(B) Corresponding experimental data adapted from \cite{Vaswani_Decay_2013}.
-(C) Inference over the current context, where contexts are color coded: black for baseline, orange for the counter-clockwise force, purple for the clockwise force and brown for clockwise force with half strenght.
-The lines represent the posterior probability of each ocontext at every trial, while the background color represents the real context.
+a group of 10 simulated participants) and the shaded area represents the standard deviation.
+The vertical dashed line is the start of the error-clamp trials.
+(B) Corresponding experimental data adapted from \cite{Vaswani_Decay_2013}, figure 2.
+(C) Simulations: Inference over the current context, where contexts are color coded: black for baseline, orange for the counter-clockwise force, purple for the clockwise force and brown for clockwise force with half strength.
+The lines represent the posterior probability of each context in every trial, while the background color represents the true context.
 An olive-colored background represents error-clamp trials.
 As in (A), solid lines represent the average across all runs and shaded areas represent the standard deviation.
-(D) Visualization of the lag before error-clamp trials are detected.
+(D) Simulations: Visualization of the lag before a change in context is detected by the agent during the E trials.
 Each line represents one run (groups 1.1, 1.2 and 1.3, 10 runs per group).}
@@
 Depending on the contexts previously learned by the agent (which change from group to group), inference during E trials will vary: groups 1.1 and 1.4 have exactly the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
-These agents will slowly lower the force appl ied until eventually reaching zero (after hundreds of trials; not shown).
+These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
 In contrast, group 1.3 has learned the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C), which has a non-zero posterior probility during E trials, pushing the agent's adaptation force more quickly towards zero.
@@
 \section{Discussion}
-We presented a hierarchical model for motor adaptation based on Bayesian inference, which not only learns to adapt its motor outputs by observing errors, but also infers the context in which these arrors were commited and updates the forward model associated with that context.
+We presented a hierarchical model for motor adaptation based on Bayesian inference, which not only learns to adapt its motor outputs by observing errors, but also infers the context in which these errors were committed and updates the forward model associated with that context.
 
 Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
-We built on this studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
+We built on these studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
 
@@
 Using our model, we showed how each of these effects can be explained by the specifics of context inference, which integrates all the available information (e.g.
-sensory cues, workspace location, reward and endpoint feedback, etc), in some cases throughout many trials.
+sensory cues, workspace location, reward and endpoint feedback), in some cases throughout many trials.
 
 It is important to note that the most essential part of the model we presented is the concept of the context and the inclusion of a Bayesian inference scheme over the context, which integrates all the information that is available to the participant during the experiment.
-The mathematical details of the model are not strictly necessary to reproduce the results.
 Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations.
-This choice greatly simplifies our simulations and provides very intuitive intepretations for the results.
+This choice greatly simplifies our simulations and provides intuitive interpretations of experimental results.
 However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
@@
 
-In what follows, we dive deeper into the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model.
+In what follows, we dive deeper\todo{Rephrase} into the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model.
 In addition, we discuss several predictions that our model makes, which could be verified experimentally in future studies.
@@
 If such an account were true, it would predict that when participants walk away from the experiment they would have trouble adapting to the no-rotation reality, having to re-learn it.
-Our model instead would suggest that participants would immediatly switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.
-More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
-While we believe our model provides a better explanation for the observations by \cite{Oh_Minimizing_2019}, the effect the authors describe has been documented before in chimpanzees \citep{Kojima_Memory_2004}\todo{@Stefan: you mentioned once you knew of papers talking about this.
-Do you have any citations I could add?}; however, to our knowledge such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
+Our model instead would suggest that participants would immediately switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.\todo{@Stefan: You: It's reasonable to speak of adaptable learning rates, but their model (and, in general, the many-state linear models) have a fixed learning rate, which does not change when going back to baseline.
+It has been tested before they came up with the many-state models.} More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
+While we believe our model provides a more general explanation for the observations by \cite{Oh_Minimizing_2019}, the effect the authors describe has been documented before in chimpanzees \citep{Kojima_Memory_2004}\todo{Maybe find more citations}; however, to our knowledge such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
 
-The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed at which participants adapted to curl forces.
+The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed with which participants adapted to curl forces.
 While it has been shown that sensitivity to error can be modulated \cite{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}), to our knowledge there are no other experiments suggesting that it depends on the history of errors.
@@
 The same can be said of the experiments by \cite{Davidson_Scaling_2004}, in which different contexts involve movement in different directions, which requires the use of different muscles.
-This alone could cause delays in switching unrelated to context inference\todo{Need citation for this}, but we expect the effects to be insufficient to explain the observed data.
+This alone could cause delays in switching unrelated to context inference\todo{Need citation for this, but maybe use task switching literature instead?}, but we expect the effects to be insufficient to explain the observed data.
 
@@
 In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
-It is in these cases where the effects of context inference are most noticeable.
+It is in these cases where the effects of context inference are most noticeable \todo{Maybe move all text until here to the introduction}.
 While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases.
@@
 Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
-When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be wrong.
-This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified.; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
+When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be incorrect.
+This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
 
@@
 In our account, if all information indicating a change in context is removed from the experiment, participants would continue to behave as they were in the previous block.
-Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where they showed random errors to participants during E trials, with a variance matching that of previously observed motor commands.
+Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where participants were shown random errors during E trials, with a variance matching that of previously observed motor commands.
 The authors showed that by matching the errors expected by participants, they eliminated the slow tapering off observed in most E blocks.
@@
 The basic principle behind the results we presented is that context inference is a process that develops over time and that carries with it uncertainty.
-This uncertainty affects learning and behavior, effecting phenomena that are directly observable during behavioral experiments.
+This uncertainty affects learning and behavior during motor adaptation, effecting phenomena that are directly observable during behavioral experiments.
 
@@
 If this prediction is confirmed, our model would additionally suggest that human participants do not revisit the learned adaptation of the last trial when a new observation comes in.
-To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}: at trial $t$, the real context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
+To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}\todo{Talk about what Herzfeld showed, as evidence of what follows}: at trial $t$, the true context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
 When the outcome is observed, it becomes clear to the participant that the context was B.
@@
 Such integration would manifest itself in experiments in which observations are noisy, as in the experiments by \cite{Kording_Bayesian_2004}, in which the position of the finger was obscured and instead participants are shown a blurry cursor which was some times shifted from its real position.
-If the added observation noise gives evidence for a particular context (the real underlying context or another one) and a visual cue gave partial information for another context, the participants' behavior would be more consistent with the most reliable source of contextual information.
+If the added observation noise gives evidence for a particular context (the true underlying context or another one) and a visual cue gave partial information for another context, the participants' behavior would be more consistent with the most reliable source of contextual information.
 
@@
 commands), using prediction error as the force behind inference. In this case,
-the generative model can is written as follows:
+the generative model can be written as follows:
 \[
@@
 advantage of not only greatly simplifying computational calculations, but also
-presenting update equations with great intuitive appeal.
+presenting update equations with intuitive appeal.
 
@@
 \label{sec:interpreting-hyperparameters}
+\todo{Remove references to code}
 Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
 units as the necessary adaptation. $\nu$ encodes how stable this hyperprior is:
-higher values (e.g. 10000) all but guarantee that the hypermean will not move
+higher values (e.g. 10,000) all but guarantee that the hypermean will not move
 in the face of evidence; In principle, enough evidence should still move it,
@@
 
-\bibliography{../../MyLibrary}
+\section{Aknowledgement}
+Funded by the German Research Foundation (DFG, Deutsche Forschungsgemeinschaft) as part of Germany’s Excellence Strategy – EXC 2050/1 – Project ID 390696704 – Cluster of Excellence “Centre for Tactile Internet with Human-in-the-Loop” (CeTI) of Technische Universität Dresden.
+
+\bibliography{MyLibrary}
 

### 2021-09-20 adaptation/article/article.tex
First modification after Stefan's comments. Small things all addressed.

@@
 In this work, we show that these unexplained phenomena, alongside savings, de-adaptation, anterograde interference and spontaneous recovery, can be explained in terms of the concept of context, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
+In mechanical arm experiments, we can think of one context in which the arm exerts no force and one in which it does.
+By identifying the current context, a participant can know how to counteract the mechanical arm and succeed in a reaching movement.
+In some experiments, the context is clearly identifiable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}.
+In many other experiments, the context is not directly observable, and context inference takes the form of an evidence-accumulating process.
+In this work, we propose that this process can be observed in behavioral data and can be used to explain the phenomena outlined above.
 
@@
 \begin{equation}
-q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi)
+q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi) \label{eqn:update-mini}
 \end{equation}
@@
 q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
+\label{eqn:context-inference-mini}
 \end{equation}
@@
 where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.
-\todo{Maybe add a section with the progression of a trial}
+
+\subsection{From observation to action selection}
+Each trial starts with a new observation $\vartheta_t$, which comprises contextual sensory information (when available), as well as the outcome of the previously-selected action.
+
+The first step in the model is to perform context inference using the current observation $\vartheta_t$ according to \eref{eqn:context-inference-mini} and \eref{eqn:estimated-context}.
+Once a probability distribution over contexts has been created, all models are updated according to \eref{eqn:update-mini} and \eref{eqn:update-full}, based on the distribution over contexts.
+Finally, action selection is performed using the distribution over contexts as well as the updated forward models, following \eref{eq:action-selection}.
+
 \section{Experimental results}
@@
 
-Our model offers a different and more parsimonious\todo{Is this the right word?} explanation for the observed results: the asymmetry is due to the existence of the baseline context.
+Our model offers a different explanation of the observed results, based on context inference: the asymmetry is due to the existence of the baseline context.
 As discussed above, action selection in our model is done via a weighted average (see \eref{eq:action-selection}).
@@
 
-In what follows, we dive deeper\todo{Rephrase} into the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model.
+In what follows, we review the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model.
 In addition, we discuss several predictions that our model makes, which could be verified experimentally in future studies.
@@
 If such an account were true, it would predict that when participants walk away from the experiment they would have trouble adapting to the no-rotation reality, having to re-learn it.
-Our model instead would suggest that participants would immediately switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.\todo{@Stefan: You: It's reasonable to speak of adaptable learning rates, but their model (and, in general, the many-state linear models) have a fixed learning rate, which does not change when going back to baseline.
+Our model instead would suggest that participants would immediately switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.\todo{@Stefan: It's reasonable to speak of adaptable learning rates, but their model (and, in general, the many-state linear models) have a fixed learning rate, which does not change when going back to baseline.
 It has been tested before they came up with the many-state models.} More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
-While we believe our model provides a more general explanation for the observations by \cite{Oh_Minimizing_2019}, the effect the authors describe has been documented before in chimpanzees \citep{Kojima_Memory_2004}\todo{Maybe find more citations}; however, to our knowledge such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
+We believe our model provides a more general and principled explanation for the observations by \cite{Oh_Minimizing_2019} and, to our knowledge, such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
 
@@
 As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data.
-The same can be said of the experiments by \cite{Davidson_Scaling_2004}, in which different contexts involve movement in different directions, which requires the use of different muscles.
-This alone could cause delays in switching unrelated to context inference\todo{Need citation for this, but maybe use task switching literature instead?}, but we expect the effects to be insufficient to explain the observed data.
+The same can be said of the experiments by \cite{Davidson_Scaling_2004}.
 
@@
 \begin{align}
+  \begin{split}
   q(\zeta_t) &= q(\zeta_t | \omega_t)q(\zeta_t | s_t, s_{t-1}, c_{t-1}) \\ \label{eqn:estimated-context}
@@
   q(\zeta_t | s_t, s_{t-1}, c_{t-1}) &\propto p(s_t | \zeta_t, s_{t-1}, c_{t-1})p(x_t)p(s_{t-1})p(\zeta_t)
+  \end{split}
 \end{align}
@@
 \begin{align}
+  \begin{split}
   \mu_\phi^{(t)} &= \frac{\nu_\phi^{(t-1)} \mu_\phi^{(t-1)} + q(\zeta_i)s_t}{\nu_\phi^{(t-1)} + q(\zeta_i)} \\
@@
   \beta_\phi^{(t)} &= \beta_\phi^{(t-1)} + \frac{q(\zeta_i)\nu_\phi^{(t-1)}}{\nu_\phi^{(t-1)} +
-                  q(\zeta_i)}\frac{\left(s_t - \mu_\phi^{(t-1)}\right)^2}{2}
+    q(\zeta_i)}\frac{\left(s_t - \mu_\phi^{(t-1)}\right)^2}{2}  \label{eqn:update-full}
+  \end{split}
 \end{align}
@@
 \label{sec:interpreting-hyperparameters}
-\todo{Remove references to code}
-Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
-units as the necessary adaptation. $\nu$ encodes how stable this hyperprior is:
-higher values (e.g. 10,000) all but guarantee that the hypermean will not move
-in the face of evidence; In principle, enough evidence should still move it,
-but that will not happen during an experiment. Smaller values (e.g. 1 / force\_sd,
-as is the default on the code) make the hypermean (and thus the mean) follow
-evidence more freely. Note that as more observations are accumulated, $\nu$
-becomes bigger and bigger, solidifying the value of the hypermean.
+$\mu$ determines the initial estimate of the adaptation, in the same units as
+the necessary adaptation. $\nu$ encodes how stable this hyperprior is: higher
+values (e.g. 10,000) all but guarantee that the hypermean will not change its
+value after observations; In principle, enough evidence should still modify it,
+but that would not happen during an experiment. Smaller values (i.e. $\sim 1$)
+make the hypermean follow evidence more freely. Note that as more observations
+are accumulated, $\nu$ becomes bigger and bigger, stabilizing the value of the
+hypermean.
 
-The hyperparameters $\alpha$ and $\beta$ have a more complex effect. Note that the
-mean of a Gamma distribution is $1 / (\alpha \nu)$; this mean is being used
-as the standard deviation of a Gaussian by the rest of the agent,
-which makes it an important measure of uncertainty. While setting the default
-hyperparameters, the values used are $\alpha = 0.5 / \sigma_0$ and
-$\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the
-standard deviation of the force exerted by the environment (force\_sds in the
-code), which controls the initial learning rate. This makes the initial
-standard deviation equal $\sigma_0$, which makes it consistent with the
-fixed-force model in its interface. The 0.5 values ensure that uncertainty is
-large at the beginning and during the experiment is greatly reduced, but never
-to a point where it is too small. Changing this 0.5 would make the
-standard deviation change more quickly, making the model more or less precise
-in its predictions, independently of the volatility of the mean of the
-adaptation (via the hypermean).
+The hyperparameters $\alpha$ and $\beta$ have a more complex effect. Note that
+the mean of a gamma distribution is $1 / (\alpha \nu)$; this mean is being used
+as the standard deviation of a Gaussian by the rest of the agent, which makes it
+an important measure of uncertainty. While setting the default hyperparameters,
+the values used are $\alpha = 0.5 / \sigma_0$ and $\beta = 0.5$, where
+$\sigma_0$ is the \textit{a priori} estimate of the standard deviation of the
+force exerted by the environment, which controls the initial learning rate. This
+makes the initial standard deviation equal $\sigma_0$, which makes it consistent
+with the fixed-force model. The 0.5 values ensure that uncertainty is large at
+the beginning and is greatly reduced during the experiment, but never to a point
+where it is so small that it makes trial-to-trial variation in the environment
+surprising. Changing this 0.5 would make the standard deviation change more
+quickly, making the model more or less precise in its predictions, independently
+of the volatility of the mean of the adaptation (via the hypermean).
 
 The baseline model defaults to different values that make it a lot more
-stable. The hyperstd of the mean is set to 10,000, which makes the mean
-entirely stable during the duration of the experiment. The values of $\alpha$
-and $\beta$ are fixed regardless of $\sigma_0$ such that the standard deviation
-is 0.001 (compare that to the size of the adaptations, around 0.0125), and the
-hyperparameters of the standard deviation are stable during the experiment.
+stable. The hyper-standard deviation of the mean is set to 10,000, which makes
+the mean entirely stable during the duration of the experiment. The values of
+$\alpha$ and $\beta$ are fixed regardless of $\sigma_0$ such that the standard
+deviation is 0.001 (compared that to the size of the adaptations in mechanical
+arm experiments, around 0.0125), and the hyperparameters of the standard
+deviation are stable during the experiment.
 

### 2021-09-22 adaptation/article/article.tex
Second draft

@@
  Humans are able to adapt their movements to changes in the dynamics of the environment and of their own bodies, a phenomenon known as motor adaptation.
-Motor adaptation has been shown to be contextual in experiments in which participants must learn to adapt to different altered dynamics, and must be able to switch between the learned adaptations dynamically.
+Motor adaptation has been shown to be contextual in experiments in which participants must learn to adapt to different modified dynamics, and must be able to switch between the learned adaptations dynamically.
+In real life, the information that we use to determine how to perform a task is often incomplete or misleading, which affects our ability to switch between tasks.
 In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
@@
 Humans are able to adapt movements to changing dynamics of the environment and of the body itself.
-While we must adapt to slow changes in our body dynamics, e.g.
-because of muscle exhaustion throughout the day, we can also adapt to sudden changes in the environment.
-For example, when picking up a heavy object, we must adapt our motor commands to be able to move the object around.
+We adapt to slow changes in our body dynamics, e.g.
+because of muscle exhaustion throughout the day, but we can also adapt to sudden changes in the environment.
+For example, when picking up a heavy object, we must adapt our motor commands to be able to move the extra weight safely.
 Even before picking it up, the size and material of an object provide us with contextual information which we can use to prepare the correct posture for moving the object.
@@
 Additionally, humans have been shown to dynamically switch between different learned tasks \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
-The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the way in which switching can be impeded or improved, are still an active topic of research.
+The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are still an active topic of research.
 
 In order to study motor adaptation, motor error is defined in experiments as the deviation between the observed movements of a participant and a baseline movement.
-For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
+For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient can be defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
 Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.
@@
 
-Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
-These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses, as well as increases the stability of the estimate.
+Bayesian accounts of motor adaptation have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
+These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses and increases the stability of the estimate.
 
-While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of their scope.
+While these general models of adaptation explain the most common phenomena observed in experiments, other known phenomena remain outside of their scope.
 For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}.
@@
 In mechanical arm experiments, we can think of one context in which the arm exerts no force and one in which it does.
-By identifying the current context, a participant can know how to counteract the mechanical arm and succeed in a reaching movement.
-In some experiments, the context is clearly identifiable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}.
+By inferring the current context, a participant can know how to counteract the mechanical arm and succeed in a reaching movement.
+In some experiments, the context is clearly inferable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}.
 In many other experiments, the context is not directly observable, and context inference takes the form of an evidence-accumulating process.
-In this work, we propose that this process can be observed in behavioral data and can be used to explain the phenomena outlined above.
+In this work, we propose that this process can be observed in behavior and can be used to explain the phenomena outlined above.
 
@@
 In our model, contexts are defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the variables of the environment.
-Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to identify the context.
+Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to infer the context.
 For example, when lifting an object of unknown weight, an agent might have learned one context for heavy objects and one for light objects.
@@
 These contextual cues (size and texture) do not necessarily identify the context (e.g.
-it could be a movie prop that looks like a rock).
+it could be a movie prop that looks like a rock) and integration over time might be necessary.
 
@@
 motor adaptation), in order to introduce the terminology that we use throughout this work.
-For an overview, \fref{fig:model}B and C shows how these three components affect each other.
+For an overview, we show in \fref{fig:model}B and C how these three components affect each other.
 
@@
 Here, we use this principle and chose to make use of an exact Bayesian learner.
-As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials.
+As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials.
 As we will show below this feature will be useful to explain specific key phenomena parsimoniously.
@@
 
-To describe this component of the model, we make use of the example of reaching movements.
-It is now widely assumed that humans learn associations between motor commands and outcomes in different environments \cite{Wolpert_Multiple_1998}.
-These associations, called forward models, are used by humans for action selection.
-We can write this forward model as follows:
+To describe this component of the model, we make use of the example of reaching movements. We define a forward model as follows:
 \begin{equation}
@@
 
-Additionally, we use the term "learned adaptation" and similar to refer to motor outputs (e.g.
+In some experiments, adaptation blocks are repeated throughout the experiments.
+For example, \cite{Oh_Minimizing_2019} used an O-A-O-A-O-A-O-A sequence.
+To differentiate the repeating blocks, the authors named them $O_i$ and $A_i$, where $i = 1, 2, ...$.
+Their experiment can be described as $O_1-A_1-O_2-...$.
+We adopt this naming convention.
+
+We use the term "learned adaptation" and similar to refer to motor outputs (e.g.
 reaching movements) that have already been adapted by the participant to a change in the dynamics of the environment.
@@
 In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models.
-This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019} but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.
+This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007} but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.
 
-To show this, we examined multiple experimental studies in which savings is displayed.
+To show this, we examined multiple experimental studies in which savings are observed.
 We categorized these studies based on the amount of contextual information made available to participants.
@@
 
-Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
+Because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
 For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
@@
 Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
-The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.
+The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.
 
 To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \cite{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information.
-\cite{Kim_Neural_2015} performed a visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
+\cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
 Participants performed shooting movements in blocks of trials with the same rotation.
@@
 when the solid black line (representing the real context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.
-However, adaptation is incomplete (i.e.
-the blue line never reaches 40 or -40) in both experimental data and simulations until late in the experiment (around trial 500).
 
@@
 \caption{Savings and de-adaptation.
-Data from our simulations (left column) compared to data from \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} (right column).
+Data from our simulations (left column) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (right column).
 Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown.
@@
 The black lines represent the optimal adaptation, i.e.
-the size of the visuomotor rotation during the task.
+the size of the true visuomotor rotation during the task.
 (A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O.
-Participants must adapt to the rotations during shooting movements.
-The y-axis is in degrees; the black line represents the real visuomotor rotation, at either zero or 20 degrees, and the blue line shows the participants' adaptation (averaged across all participants), which can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation.
+Participants must adapt to the rotations during shooting movements, and the blue line can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation.
 When black and blue lines match, the participant perfectly counteracted the visuomotor rotation and reached the target.
@@
 Blue and black lines are as in A.
+Different blocks with the same adaptation are differentiated by the subscript (e.g.
+$A_1$ and $A_2$ have the same 20 degree visuomotor rotation).
 (C) Same experiment as (B) but with a 10 degree adaptation.} \label{fig:oh-2019}
@@
 
-\cite{Oh_Minimizing_2019} performed partially-cued experiment similar to \cite{Kim_Neural_2015}.
+Participants did not completely adapt to the rotation during the first block, as evidenced by the blue line not reaching as low or high as the black line during the first blocks.
+Switching between opposite adaptations was nevertheless immediate in simulations and experimental data.
+After around 500 trials, participants had adapted completely, as seen by the blue line reaching the heights of the black line, and switches continued to be immediate.
+
+These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches.
+To expand on these results, we now turn to performance feedback and its effects on switching behavior.
+
+\cite{Oh_Minimizing_2019} performed partially-cued experiments similar to \cite{Kim_Neural_2015}.
 The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10.
 The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
-Participants in \cite{Oh_Minimizing_2019} showed immediate switching in the first experiment, which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (real rotation).
+Participants in the first experiment first learned the adaptation in $A_1$.
+From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (real rotation).
 In the second experiment, switching happens more slowly, with adaptation (blue line) lagging behind the switches in real context (black line), and slowly catching up.
 As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low.
-This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly identifying the context from the noise due to trial-to-trial variation in motor output.
+This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output.
 Because of this, the model requires more evidence (i.e.
@@
 \subsection{Slower adaptation in volatile and unpredictable environments}
-One central element in our model is that adaptation is gated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
+One central element in our model is that adaptation is modulated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
 
@@
 The shaded area is the SEM across simulated runs (participants).
-(B) Data adapted from \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
+(B) Data adapted from figure 4 by \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
 \label{fig:davidson-2004}
@@
 The vertical dashed line is the start of the error-clamp trials.
-(B) Corresponding experimental data adapted from \cite{Vaswani_Decay_2013}, figure 2.
+(B) Corresponding experimental data adapted from figure 2C by \cite{Vaswani_Decay_2013}.
 (C) Simulations: Inference over the current context, where contexts are color coded: black for baseline, orange for the counter-clockwise force, purple for the clockwise force and brown for clockwise force with half strength.
@@
 (D) Simulations: Visualization of the lag before a change in context is detected by the agent during the E trials.
-Each line represents one run (groups 1.1, 1.2 and 1.3, 10 runs per group).}
+Each line represents one run (10 runs per group).}
 \label{fig:vaswani-2013}
@@
 These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
-In contrast, group 1.3 has learned the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C), which has a non-zero posterior probility during E trials, pushing the agent's adaptation force more quickly towards zero.
+In contrast, group 1.3 has learned the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C), which has a non-zero posterior probability during E trials, pushing the agent's adaptation force more quickly towards zero.
 Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more slowly than groups 1.1 and 1.4.
@@
 In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
-To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context (A for groups 1.1, 1.2 and 1.3, -A for 1.4) for each simulated run (one line per run, 10 runs per group, all groups).
+To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context for each simulated run (one line per run, 10 runs per group, all groups).
 It can be seen that the moment in which the drop starts, as well as the speed of the drops, depends on the run, reproducing the different lags observed in experimental data.
@@
 
-Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
+Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003,Imamizu_Explicit_2007}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
 We built on these studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
@@
 In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
-It is in these cases where the effects of context inference are most noticeable \todo{Maybe move all text until here to the introduction}.
+It is in these cases where the effects of context inference are most noticeable.
 While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases.
@@
 
-The same rationale explains the results from \cite{Davidson_Scaling_2004}: motor learning, which in our model is gated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
+The same rationale explains the results from \cite{Davidson_Scaling_2004}: motor learning, which in our model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
 This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
@@
 We expect that the same effect would be observed in error-clamp trials.
-If the E block is learned by participants during training, it might still be difficult for them to identify that an E block has started, which would create delays similar to those in \fref{fig:vaswani-2013}.
+If the E block is learned by participants during training, it might still be difficult for them to infer that an E block has started, which would create delays similar to those in \fref{fig:vaswani-2013}.
 However, the model predicts that if a visual cue is introduced that identifies the E block, participants would immediately switch to their baseline behavior, no longer displaying spontaneous recovery, lag, nor the slow return to baseline.
@@
 unpredictable switching between contexts) lowers the measured learning rate.
-Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can identify the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
+Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
 This would manifest in \fref{fig:herzfeld-2014} in the form of no discernible differences between the two groups of participants.
@@
 If this prediction is confirmed, our model would additionally suggest that human participants do not revisit the learned adaptation of the last trial when a new observation comes in.
-To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}\todo{Talk about what Herzfeld showed, as evidence of what follows}: at trial $t$, the true context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
+To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}: at trial $t$, the true context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
 When the outcome is observed, it becomes clear to the participant that the context was B.
@@
 
-\section{Aknowledgement}
+\section{Acknowledgment}
 Funded by the German Research Foundation (DFG, Deutsche Forschungsgemeinschaft) as part of Germany’s Excellence Strategy – EXC 2050/1 – Project ID 390696704 – Cluster of Excellence “Centre for Tactile Internet with Human-in-the-Loop” (CeTI) of Technische Universität Dresden.

### 2021-09-22 adaptation/code/figures.py

@@ -31,8 +31,19 @@ def oh_2019_kim_2015(fignum=2, show=True):
     paper.
 
     """
-    figsize = (5, 6)
+    context_color = np.ones(3) * 0.5
+    figsize = (6, 6)
     fig, axes = plt.subplots(3, 2, clear=True, num=fignum, figsize=figsize)
+    # fig = plt.figure(figsize=figsize, clear=True, num=fignum)
+    # axes_00 = fig.add_subplot(3, 2, 1)
+    # axes_01 = fig.add_subplot(3, 2, 2, sharey=axes_00)
+    # axes_10 = fig.add_subplot(3, 2, 3)
+    # axes_11 = fig.add_subplot(3, 2, 4, sharex=axes_10, sharey=axes_10)
+    # axes_20 = fig.add_subplot(3, 2, 5, sharex=axes_10)
+    # axes_21 = fig.add_subplot(3, 2, 6, sharex=axes_10, sharey=axes_20)
+    # axes = np.array([[axes_00, axes_01],
+    #                  [axes_10, axes_11],
+    #                  [axes_20, axes_21]])
 
     # a)
     task_kim, agent_pars = sims.kim_2015(plot=False)
@@ -43,8 +54,8 @@ def oh_2019_kim_2015(fignum=2, show=True):
     contexts[contexts == pars.CLAMP_INDEX] = 0
     contexts[contexts == 1] = -40
     contexts[contexts == 2] = 40
-    axes[0, 0].plot(contexts, color='black')
-    axes[0, 1].plot(contexts, color='black')
+    axes[0, 0].plot(contexts, color=context_color)
+    axes[0, 1].plot(contexts, color=context_color)
     axes[0, 0].plot(-pandota_kim['pos(t)'] - pandota_kim['action'])
     axes[0, 0].set_yticks([-40, 0, 40])
     axes[0, 0].set_ylabel('Adaptation')
@@ -54,6 +65,7 @@ def oh_2019_kim_2015(fignum=2, show=True):
     ylim = axes[0, 0].get_ylim()
     axes[0, 1].set_xlim(xlim)
     axes[0, 1].set_ylim(ylim)
+
     # b)
     task_20, task_10, agent_pars = sims.oh_2019(plot=False)
 
@@ -65,13 +77,17 @@ def oh_2019_kim_2015(fignum=2, show=True):
     contexts[contexts == pars.CLAMP_INDEX] = 0
     contexts[contexts == 1] = 20
[…]

### 2021-09-23 adaptation/article/article.tex
Update on Overleaf.
Minor changes to abstract

@@
 \abstract{Abstract}
- Humans are able to adapt their movements to changes in the dynamics of the environment and of their own bodies, a phenomenon known as motor adaptation.
-Motor adaptation has been shown to be contextual in experiments in which participants must learn to adapt to different modified dynamics, and must be able to switch between the learned adaptations dynamically.
-In real life, the information that we use to determine how to perform a task is often incomplete or misleading, which affects our ability to switch between tasks.
+Humans have been shown to adapt their movements when a sudden change to the dynamics of the environment is introduced, a phenomenon called motor adaptation.
+If the change is reverted, the adaptation is also quickly reverted.
+Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly.
+Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
 In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
@@
 
-It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}, which slows down learning.
-Additionally, humans have been shown to dynamically switch between different learned tasks \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
+It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
+Additionally, humans have been shown to dynamically switch between different learned adaptation \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
 The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are still an active topic of research.

### 2021-09-30 adaptation/article/article.tex
Update on Overleaf.

@@
 \def \sref #1{Section '\nameref{#1}'}    % Reference sections
-\def \supmat {the Supplementary Materials}
 
@@
 Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
-We show how our model provides a unifying explanation for the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.
+In this work, we use a computational mechanistic model to explain the behavioral phenomena effected by imperfect contextual information.
+Specifically, we present a hierarchical computational model for motor adaptation based on exact Bayesian inference.
+This model explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
+We show how the proposed model provides a unifying explanation for four different experimentally established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of environmental volatility and predictability on learning, (iii) the effects of previously-learned adaptations on switching between tasks, in addition to  (iv) the well studied savings, de-adaptation and spontaneous recovery.
 
 \section{Introduction}
-Humans are able to adapt movements to changing dynamics of the environment and of the body itself.
+Humans are able to adapt their movements to changing dynamics of the environment and of the body itself.
 We adapt to slow changes in our body dynamics, e.g.
@@
 However, contextual information cannot always be relied on, e.g.
-a big object of unknown material.
+a large parcel with unknown content.
 
-It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
-Additionally, humans have been shown to dynamically switch between different learned adaptation \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
-The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are still an active topic of research.
+It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
+Additionally, humans have been shown to dynamically switch between different learned adaptation \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
+The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are an ongoing, active topic of research.
 
@@
 a mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error.
-A number of phenomena have been observed that persist across participants and across many different motor adaptation experiments: savings, quick de-adaptation, anterograde interference and spontaneous recovery.
+Across many different motor adaptation experiments, now well-established phenomena have been observed, for example: (i) savings (the ability to recall previously-learned skills), (ii) quick de-adaptation (the ability to return to unmodified dynamics), (iii) anterograde interference (the interference in motor learning between opposing manipulations in dynamics) and (iv) spontaneous recovery (during trials where errors are forced to be zero, the display of behavior consistent with the previously-learned dynamics).
 
 To explain these phenomena, a number of computational models have been introduced.
-The least complex of these models are linear learners \citep[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
+The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
 By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, partially explaining these distinct experimental phenomena.
@@
 Bayesian accounts of motor adaptation have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
-These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses and increases the stability of the estimate.
+These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the the same task is practiced, the uncertainty on an estimate of the underlying modification is reduced, which lowers variability in responses and increases the stability of the estimate.
 
@@
 In this work, we show that these unexplained phenomena, alongside savings, de-adaptation, anterograde interference and spontaneous recovery, can be explained in terms of the concept of context, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
-In mechanical arm experiments, we can think of one context in which the arm exerts no force and one in which it does.
+Intuitively, in mechanical arm experiments, one can think of one context in which the arm exerts no force and one in which it does.
 By inferring the current context, a participant can know how to counteract the mechanical arm and succeed in a reaching movement.
 In some experiments, the context is clearly inferable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}.
-In many other experiments, the context is not directly observable, and context inference takes the form of an evidence-accumulating process.
-In this work, we propose that this process can be observed in behavior and can be used to explain the phenomena outlined above.
+In many other experiments, the context is not directly observable, and context inference is probabilistic and takes the form of an evidence-accumulating process.
+In this work, we propose that the behavioural consequences of this slow process of evidence accumulation can be observed in behavior and can be used to explain the phenomena outlined above.
 
 To this end, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
-We expand on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
+We build on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
 In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions.
-To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.
+To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment (e.g.
+whether the task was successfully completed).
 Having inferred the context, the model can choose the best-fitting internal model with which to make decisions.
 
-We show that the dynamics of context inference provide a unifying explanation for all the experimental phenomena outlined above, providing a single account for motor adaptation under changing contexts, while relying on known mechanisms for adaptation.
+We show that the dynamics of context inference provide a principled mechanistic explanation for the experimental phenomena outlined above, providing a single account for motor adaptation under changing contexts, while relying on known mechanisms for adaptation.
 
@@
 a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation.
-Several models exist for such error-based motor adaptation, but we expand on existing models by adding an explicit component of context inference.
+Several models exist for such error-based motor adaptation \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019}.
+We expand on these existing models by adding an explicit component of context inference.
 As we will show, this context inference guides the selection of the adequate internal model, its updating and the sampling of actions.
@@
 We simulated representative experiments from these studies to illustrate how our model explains these experimental findings using the dynamics of context inference.
-In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss in detail how context inference explains the experimental phenomena.
+We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.
 
@@
 When observing an object to be lifted, the agent can use its size and texture to estimate the weight of the object, which in turn allows the agent to infer the appropriate context and, with it, decide how to lift the object.
-These contextual cues (size and texture) do not necessarily identify the context (e.g.
+These contextual cues (size and texture) do not necessarily identify the context with certainty (e.g.
 it could be a movie prop that looks like a rock) and integration over time might be necessary.
@@
 As we will show below this feature will be useful to explain specific key phenomena parsimoniously.
-We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.
 
@@
 
-Following the these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.
+Following these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.
 
-Context is a categorical variable $\zeta$, where the identification of a context is done through Bayesian inference:
+In the model, context is a categorical variable $\zeta$, where the identification of a context is done through Bayesian inference:
 \begin{equation}
@@
 In the third section, we discuss the effects of slow context inference on action selection.
-Finally, in the fourth section we discuss context inference during error-clamp trials, and its effect on behavior.
+Finally, in the fourth section we discuss context inference during so-called error-clamp trials, and its effect on behavior.
 
-For clarity, we first introduce necessary terminology that is typically used in experimental studies.
+For clarity, we first introduce necessary terminology that is typically used in experimental studies and will be useful for linking experimental to computational observations.
 As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force (i.e.
 a force that pushes the participant's hand in the direction perpendicular to movement).
-Depending on the trial, the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
+Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
 Let us define the baseline context O as that in which the mechanical arm exerts no force.
 Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise.
-Additionally, abusing notation, we can say that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
+Additionally, abusing notation, the usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
 Studies often talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
@@
 In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models.
-This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007} but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.
+This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007} where we show in addition that expanding forward model switching into context inference explains facets of savings that these previous models do not.
 
 To show this, we examined multiple experimental studies in which savings are observed.
-We categorized these studies based on the amount of contextual information made available to participants.
-In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
+We categorized these studies based on the amount of contextual information made available to participants into cued-context and partially-cued experiments: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
 We call these cued-context experiments.
@@
 
-Because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
+Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
 For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
@@
 The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e.
-when the solid black line (representing the real context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.
+when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.
 
@@
 
-Participants did not completely adapt to the rotation during the first block, as evidenced by the blue line not reaching as low or high as the black line during the first blocks.
+Critically, in \fref{fig:oh-2019}A it can be seen until trial 250 and especially for the first four context switches,  that participants did not completely adapt to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks.
 Switching between opposite adaptations was nevertheless immediate in simulations and experimental data.
@@
 The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
-Participants in the first experiment first learned the adaptation in $A_1$.
-From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (real rotation).
-In the second experiment, switching happens more slowly, with adaptation (blue line) lagging behind the switches in real context (black line), and slowly catching up.
-As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low.
+Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A_1$.
+From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (true rotation).
+In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation (blue line) lagging behind the switches in the real context (black line), and slowly catching up.
+As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C).
 This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output.
@@
 
-Note that the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
+Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
 
 \subsection{Slower adaptation in volatile and unpredictable environments}
-One central element in our model is that adaptation is modulated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
+One central element in the proposed model is that adaptation is modulated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
 
@@
 \cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between O, A and -A.
-Participants were divided into three groups for which the context transition from one trial to the next had differing probabilities, depending on the group: for the first group, transitions were very uncommon, for the second group they were somewhat common and for the third group they were occurred at almost every trial.
+Participants were divided into three groups for which the context transition from one trial to the next had different probabilities, depending on the group: for the first group, transitions were very uncommon, for the second group they were somewhat common and for the third group they occurred at almost every trial.
 Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.
@@
 When a new block of trials starts (e.g.
-in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out.
+in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out).
 Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force).
-In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.
+In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$.
 
@@
 It can be seen that the agent exposed to the -8 context shifts back to 4 more slowly than the one exposed to +8, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.
-In our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data.
+Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data.
 However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.
-While we focused here on a specific, well-known experiment, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion section.
+While we focused here on a specific, well-known experiment, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion.
 
@@
 (B) Corresponding experimental data adapted from figure 2C by \cite{Vaswani_Decay_2013}.
-(C) Simulations: Inference over the current context, where contexts are color coded: black for baseline, orange for the counter-clockwise force, purple for the clockwise force and brown for clockwise force with half strength.
+(C) Simulations: Inference over the current context, where contexts are color coded: black for baseline, orange for the counter-clockwise force, purple for the clockwise force and brown for counter-clockwise force with half strength.
 The lines represent the posterior probability of each context in every trial, while the background color represents the true context.
@@
 Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003,Imamizu_Explicit_2007}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
-We built on these studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
+We built on these studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
 
@@
 Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations.
-This choice greatly simplifies our simulations and provides intuitive interpretations of experimental results.
+This choice greatly simplified our simulations and provides intuitive interpretations of experimental results.
 However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
@@
 If such an account were true, it would predict that when participants walk away from the experiment they would have trouble adapting to the no-rotation reality, having to re-learn it.
-Our model instead would suggest that participants would immediately switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.\todo{@Stefan: It's reasonable to speak of adaptable learning rates, but their model (and, in general, the many-state linear models) have a fixed learning rate, which does not change when going back to baseline.
-It has been tested before they came up with the many-state models.} More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
+Our model instead would suggest that participants would immediately switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.
+More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
 We believe our model provides a more general and principled explanation for the observations by \cite{Oh_Minimizing_2019} and, to our knowledge, such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
@@
 The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed with which participants adapted to curl forces.
-While it has been shown that sensitivity to error can be modulated \cite{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}), to our knowledge there are no other experiments suggesting that it depends on the history of errors.
+While it has been shown that sensitivity to error can be modulated \citep{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}), to our knowledge there are no other experiments suggesting that it depends on the history of errors.
 As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data.
-The same can be said of the experiments by \cite{Davidson_Scaling_2004}.
+The same applies to the experiments by \cite{Davidson_Scaling_2004}.
 
@@
 While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases.
-Indeed, as we argued in the Results section, many experiments inadvertently include partial contextual information used by participants.
+Indeed, as we noted in the Results section, many experiments inadvertently include partial contextual information used by participants.
 
@@
 
-The same rationale explains the results from \cite{Davidson_Scaling_2004}: motor learning, which in our model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
+The same rationale explains the results by \cite{Davidson_Scaling_2004}: motor learning, which in our model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
 This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
@@
 \subsubsection{Multi-source integration}
-The model also predicts an effect reminiscent of multisensory integration: in order to integrate contextual information from conflicting sources (e.g.
+The model also predicts an effect reminiscent of multisensory integration \citep[e.g.][]{Ernst_Humans_2002}: in order to integrate contextual information from conflicting sources (e.g.
 probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its precision (the inverse of its variance).
@@
 If the added observation noise gives evidence for a particular context (the true underlying context or another one) and a visual cue gave partial information for another context, the participants' behavior would be more consistent with the most reliable source of contextual information.
+
+\subsubsection{Adaptation history affects lag in E blocks}
+As \cite{Vaswani_Decay_2013} discussed, participants had different delays before their behavior is modified after entering the E block.
+Our model presents the same run-dependent delay, as can be seen in \fref{fig:vaswani-2013}D.
+In addition, the model predicts that this delay depends on the history of adaptation, as evidenced by the red group having very small delays (one or two trials) compared to the other groups.
+Importantly, \cite{Vaswani_Decay_2013} found no significant difference between groups in their fitted parameter for delay, contrary to our results.
+This discrepancy could be due to the models used by \cite{Vaswani_Decay_2013} to determine the lag, but could also point to an explicit component of inertia that is missing from our model.
+Future experimental and modelling work is needed to understand this.
 

### 2021-10-01 adaptation/article/article.tex
Third draft. Got rid of the section on volatility

@@
 Specifically, we present a hierarchical computational model for motor adaptation based on exact Bayesian inference.
-This model explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection.
-We show how the proposed model provides a unifying explanation for four different experimentally established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of environmental volatility and predictability on learning, (iii) the effects of previously-learned adaptations on switching between tasks, in addition to  (iv) the well studied savings, de-adaptation and spontaneous recovery.
+This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection.
+We show how the proposed model provides a unifying explanation for four different experimentally established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, in addition to  (iii) the well-studied savings, de-adaptation and spontaneous recovery.
 
@@
 For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}.
-Learning new adaptations has also been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013}.
+Action selection has also been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013,Davidson_Scaling_2004}.
 
@@
 In many other experiments, the context is not directly observable, and context inference is probabilistic and takes the form of an evidence-accumulating process.
-In this work, we propose that the behavioural consequences of this slow process of evidence accumulation can be observed in behavior and can be used to explain the phenomena outlined above.
+In this work, we propose that the behavioral consequences of this slow process of evidence accumulation can be observed in behavior and can be used to explain the phenomena outlined above.
 
@@
 
-In the model, context is a categorical variable $\zeta$, where the identification of a context is done through Bayesian inference:
+In the model, context is a categorical variable $\zeta$ which is inferred using Bayes' theorem:
 \begin{equation}
@@
 To differentiate the repeating blocks, the authors named them $O_i$ and $A_i$, where $i = 1, 2, ...$.
-Their experiment can be described as $O_1-A_1-O_2-...$.
+Their experiment can be described as $O_1$-$A_1$-$O_2$-....
 We adopt this naming convention.
@@
 As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C).
-This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output.
+This difference is explained by our model in terms of the size of the adaptation: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output.
 Because of this, the model requires more evidence (i.e.
@@
 Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
-
-\subsection{Slower adaptation in volatile and unpredictable environments}
-One central element in the proposed model is that adaptation is modulated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
-
-Direct experimental evidence for the effects of context inference on the rate of adaptation is provided by \cite{Herzfeld_memory_2014}.
-The authors showed that the volatility of the environment, expressed in terms of unpredictable, stochastic transitions between A and -A (and vice versa), change the speed at which adaptation occurs.
-
-\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between O, A and -A.
-Participants were divided into three groups for which the context transition from one trial to the next had different probabilities, depending on the group: for the first group, transitions were very uncommon, for the second group they were somewhat common and for the third group they occurred at almost every trial.
-Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.
-
-\cite{Herzfeld_memory_2014} found that the volatility of the environment affects learning.
-The more volatile the environment, the less participants adapt their motor responses after observing an error.
-In \fref{fig:herzfeld-2014}, we show results obtained with our model in a paradigm matching that of \cite{Herzfeld_memory_2014}.
-We simulated the conditions of low volatility and maximum uncertainty ($z = 0.9$ and $z = 0.5$ in \cite{Herzfeld_memory_2014}) for thirty trials, repeating the experiment 300 times with random sequences of A and -A each time, following the transition probabilities from the experiment.
-In \fref{fig:herzfeld-2014}A we show the adaptation as a function of trial number, averaged across all 300 runs.
-It can be seen that the agent matches the findings from \cite{Herzfeld_memory_2014} where adaptation is slower in the more volatile environment ($z = 0.5$) than in the less volatile one.
-Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment, a difference that is consistent across all trials.
-
-\begin{figure}
-\centering
-\includegraphics{./figures/figure_3.png}
-\caption{Learning under uncertainty about contexts.
-30 trials were simulated in the task defined by \cite{Herzfeld_memory_2014} and repeated 300 times to estimate the mean effects.
-Two conditions were used, following the naming convention from \cite{Herzfeld_memory_2014}: $z=0.5$ is an unpredictable condition, where the context has a 0.5 probability of switching from trial to trial.
-$z=0.9$ is a stable condition, where the context has a 0.9 probability of staying the same from trial to trial.
-(A) Adaptation as a function of trial number, averaged across all runs.
-(B) Error sensitivity, binned into bins of 5 trials, averaged across all runs.
-The error and consequently sensitivity to error (y axis) is in arbitrary units.
-Shaded areas are SEM.}
-\label{fig:herzfeld-2014}
-\end{figure}
-
-To account for volatility, our model includes a term in context inference which assumes a transition can occur from the previous context to all other possible contexts (see Methods).
-Effectively, this transition term gives the model uncertainty on the current context, regardless of observations.
-The higher the volatility, the higher the uncertainty.
-This higher uncertainty is what lowers the observed learning rate across all trials.
 
@@
 Experimental evidence supporting this view can be found in experiments with context switching.
-For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A).
+For example, \cite{Davidson_Scaling_2004} reported an curl-force experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A).
 After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
@@
 We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}.
-It can be seen that the agent exposed to the -8 context shifts back to 4 more slowly than the one exposed to +8, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.
+It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.
 Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data.
@@
 \centering
-\includegraphics{./figures/figure_4.png}
+\includegraphics{./figures/figure_3.png}
 \caption{Motor error when switching back to a previously learned adaptation.
-Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from +8 (yellow) or -8 (blue) to 4 is shown.
-(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}, running eight simulations per group, each simulation representing one participant.
+Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from 3A (yellow) or -A (blue) to A is shown(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}, running eight simulations per group, each simulation representing one participant.
 On the y axis, the error is shown in arbitrary units, related to centimeters through a monotonically increasing transformation with the same origin.
@@
 We discuss these experiments in the following section.
-Additionally, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion section.
+Additionally, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion.
 
@@
 \centering
-\includegraphics[width=\textwidth]{./figures/figure_5.png}
+\includegraphics[width=\textwidth]{./figures/figure_4.png}
 \caption{Adaptation during error clamp trials.
@@
 To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context for each simulated run (one line per run, 10 runs per group, all groups).
-It can be seen that the moment in which the drop starts, as well as the speed of the drops, depends on the run, reproducing the different lags observed in experimental data.
+The start and speed of the drop dependd on the run (i.e.
+the participant), reproducing the different lags observed in experimental data.
 This can also be seen directly in context inference, displayed as the shaded areas in \fref{fig:vaswani-2013}C.
@@
 More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
-We believe our model provides a more general and principled explanation for the observations by \cite{Oh_Minimizing_2019} and, to our knowledge, such an effect has not been observed in humans in tasks like that by \cite{Oh_Minimizing_2019}.
 

### 2021-10-01 adaptation/code/figures.py

@@ -34,16 +34,6 @@ def oh_2019_kim_2015(fignum=2, show=True):
     context_color = np.ones(3) * 0.5
     figsize = (6, 6)
     fig, axes = plt.subplots(3, 2, clear=True, num=fignum, figsize=figsize)
-    # fig = plt.figure(figsize=figsize, clear=True, num=fignum)
-    # axes_00 = fig.add_subplot(3, 2, 1)
-    # axes_01 = fig.add_subplot(3, 2, 2, sharey=axes_00)
-    # axes_10 = fig.add_subplot(3, 2, 3)
-    # axes_11 = fig.add_subplot(3, 2, 4, sharex=axes_10, sharey=axes_10)
-    # axes_20 = fig.add_subplot(3, 2, 5, sharex=axes_10)
-    # axes_21 = fig.add_subplot(3, 2, 6, sharex=axes_10, sharey=axes_20)
-    # axes = np.array([[axes_00, axes_01],
-    #                  [axes_10, axes_11],
-    #                  [axes_20, axes_21]])
 
     # a)
     task_kim, agent_pars = sims.kim_2015(plot=False)
@@ -148,110 +138,15 @@ def oh_2019_kim_2015(fignum=2, show=True):
         plt.show(block=False)
 
 
-def herzfeld_2014(fignum=3, show=True):
-    """Reproduces the results from Herzfeld et al 2014.
-
-    """
-    figsize = (4, 5)
-    colors = {'high': np.array((95, 109, 212)) / 256,
-              'low': np.array((212, 198, 95)) / 256}
-    optimal = -13
-    repeats = 300  # participants * (number of blocks)
-    num_trials = 30
-    fig = plt.figure(clear=True, num=fignum, figsize=figsize)
-    fig, axes = plt.subplots(2, 1, squeeze=True, clear=True,
-                             num=fignum, sharex=True)
-
-    agent = {}
-    data = {'high': [], 'low': []}
-    for idx in range(repeats):
-        outs = sims.herzfeld_2014(num_trials=num_trials, plot=False)
-        tasks = outs[:2]
-        agent_pars = outs[2:]
-
-        for name, t_pars, a_pars in zip(['high', 'low'], tasks, agent_pars):
-            agent[name] = model.LRMeanSD(**a_pars)
-            agent[name].all_learn = True
-            agent[name].threshold_learn = 0.2
-        for name, c_pars in zip(['high', 'low'], tasks):
[…]

### 2021-10-01 adaptation/code/simulations.py

@@ -663,14 +663,15 @@ def davidson_2004(plot=True, axes=None, fignum=13):
                      'cue_noise': 1 / 3,
                      'max_force': 20,
                      'hyper_sd': 100,
-                     'obs_sd': 2.2,
+                     'obs_sd': 2.5,
                      'context_noise': 0.01,
                      'force_sds': 0.5 * np.ones(3),
                      'prediction_noise': 0.5,
                      'learn_rate': 0.5,
                      'all_learn': True,
                      'threshold_learn': 0.2,
-                     'sample_action': True}
+                     'sample_action': True,
+                     'action_sd': 2}
     agent_pars_p8 = agent_pars_m8.copy()
     agent_pars_p8['angles'] = [0, 4, 12]
     if plot:

### 2021-10-01 prediction/notes/notes.tex
Restructuring and adding new project

\documentclass{report}
\usepackage{graphicx}
\usepackage{breakcites}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tabularx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{todonotes}
\bibliographystyle{apalike}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

% Defining parameters
\def \fref #1{Figure \ref{#1}}     % Reference figures
\def \tref #1{Table \ref{#1}}      % Reference tables
\def \eref #1{Equation \ref{#1}}   % Reference equations
\def \sref #1{Section '\nameref{#1}'}    % Reference sections

% Remember to start reftex-mode

\begin{document}

\begin{chapter}{Introduction}

\end{chapter}

\begin{chapter}{Model}
The full model consists of four components, all interconnected in a way that the agent is capable of single movements, as well as the planning and execution of squences of movements, all without the need for fast changes in the connectivity between the involved neurons.

To describe the model, I will refer to the example of controlling a pointer (e.g. a hand) in a 2d space. However, the model can be extended to three physical dimmensions and, perhaps, to any number of joints. All in the future.

\section{First layer: movement layer}
To begin the description, we look at the outermost layer of the model, which will have a direct correspondence with the physical world. For simplicity, let us think of this layer as two populations of neurons. We refer to the average firing rate of each population as $x_1$ and $x_2$ respectively. To simplify, we normalize the firing rate such that $x_i \in [0, 1]$, where $i = {1, 2}$; 0 represents a population that is completely inactive, 1 represents the maximum possible activity (firing rate) of the population. We assume that the hand movements are made within a square space (e.g. a table in front of a participant) and the current position of the hand in this space is represented by $x$.

The dynamics of $x$ will be given by the following equation:
\begin{equation}
\dot x = x - f(y) \label{eqn:first-layer}
\end{equation}
\todo{It might be better to use a different system of equations to match the experimental observation that the speed of movement peaks halfway there.}
where $y$ represents another population of neurons (described below) and $f:R^{N_y}\rightarrow R^2$ is a function whose properties are yet to be determined. As will be discussed below, $y$ does not depend on $x$, and therefore the system \ref{eqn:first-layer} has an instantaneous stable equilibrium point located at $f(y)$

\section{Second layer: control layer}
The second layer is comprised of a large population of neurons, divided into subpopulations represented by $y_i$. As with $x$, $y_{ij}$ represents the average firing rate of population $i,j$. The subpopulations $y_{ij}$ can be thought of as being arranged in a lattice of size N by N, where $N^2$ equals the number of subpopulations. When a single population $y_{i^+j^*}$ is active, $f(y)$ creates a single equilibrium point in the $x$ space at the location with coordinates corresponding to the center of the $i,j$ lattice. With this mechanism, the hand can be moved to a desired location by activating the population $y_{ij}$ that corresponds to this location.

The neurons of these layer do not have inherent dynamics. Their activity depends entirely on the third layer. We discuss their connections in the following section.

In what follows, we will simplify the notation to $y_i$ unless the explicit reference to lattice positions is necessary, in which case it will be clearly stated.

\section{Third layer: sequential planning}
This layer is used for the learning and execution of sequences of actions. This is done by having a large number of subpopulations $z_i$, each one of them connected to one $y_j$. These connections are many-to-one: many $z_i$ could be connected to a single $y_j$. For coding efficiency, many-to-many connectivity could also be used, but we do not describe this possibility in detail.

The dynamics of these subpopulations are such that when the population receives an external input that activates a set of subpopulations, a sequence of sets of subpopulations will ensue. To create these dynamics, we use a system similar to that in \citep{CuevasRivera_Modelling_2015}, to which we refer as cluster heteroclinic sequences, or CSHS.

The CSHS is based on the generalized Lotka-Volterra (LV) equations, which can be written as follows:
\begin{equation}
\dot z_i = z_i(\alpha_i - \rho_{ij}z_j) + \xi \label{eqn:third-level}
\end{equation}
where $\alpha$ and $\rho$ have to be set as in \citep{CuevasRivera_Modelling_2015} to create CSHS dynamics. The term $\xi$, which was not present in the original model, is a signal given by the task monitoring layer (discussed below), which is zero except when a motion has ended (e.g. the hand has reached the next desired position). This feedback signal affects all neurons in the third layer equally, and its magnitude is constant. It serves as a jolt that will counteract the all-to-all inhibition (given by $\rho$) and allow the next cluster in the sequence to activate.

By setting $\rho$ as in \cite{CuevasRivera_Modelling_2015}, we can create sequences of clusters, where each one of the clusters in a sequence activates one $y_i$, essentially creating sequences in the second layer. In general, the number of clusters, $N_\zeta$, is much larger than the number of subpopulations in the second layer $N_y$. This allows us to include a particular $y_i$ in as many different sequences as we want, without any interference between sequences. Note, however, that the number of clusters could vastly exceed the number of subpopulations on the third level, if we allow any one subpopulation to belong to more than one clusters, as was done in \citep{CuevasRivera_Modelling_2015}.

We define $S_i$ as a sequence of clusters $\zeta_j$, where $j = i_1, i_2, ..., i_{K_i}$ and $K_i$ is the number of clusters in sequence $S_i$. The third layer can be embedded with as many sequences $S$ as desired, within the limits explored by \cite{CuevasRivera_Modelling_2015}. This is done by modifying $\rho$ from \eref{eqn:third-level}. We do not explore in this model the learning mechanisms that can lead to the modification of $\rho$. In order to execute a learned sequence of actions, the first cluster of the sequence $\zeta_{i_1}$ must be excited, which will excite the corresponding subpopulation $y_{i_1}$ and create a stable equilibrium point in the $x$ space, initiating movement. Due to the high all-to-all inhibition of the $z$ space, the system will remain there until a completion signal comes in, such that $\xi \ne 0$, which will jolt the entire system $z$, and the connectivity $\rho$ will make it go towards $\zeta_{i_2}$, and so on, executing the entire sequence of movements.

Let us define $\lambda_{ij}$ as the connection strength between cluster $i$ in the third layer and subpopulation $y_j$ in the second cluster. With this, we can define the activity of subpopulation $y_i$ as follows:
\begin{align}
  \begin{split}
  y_i &= \displaystyle\sum_{i = 1}^{N_\zeta} \lambda_{ij}\zeta_j \\ \label{eqn:second-level}
  \zeta_j &= \frac{1}{M_j}\displaystyle\sum_{k=1}^{M_j}z_k
  \end{split}
\end{align}
where $\zeta_j$ represents the average activation of all neurons belonging to cluster $j$, and $M_i$ is the number of subpopulations in cluster $i$.

\section{Fourth layer: goals}
This layer represents the overall goal of an agent, which is then carried out as a sequence of actions. For example, a typist might have the overall goal of typing a word. In order to do this, the agent will activate the sequence of clusters $S$ which will then carry out each action (e.g. typing each letter) until the goal is achieved. In this example, the third layer would be embedded with a sequence for each word that the agent can type, and in order to carry them out, only the initial cluster would need to be activated.

The function of this layer can also be thought of as a context switcher. Let us present a thought experiment with human participants: the participant sits in front of a screen and a keyboard and must, at each trial, type a word. Which word she must type depends on the current context. The participant has learned to type five words, thus defining five contexts. After learning, the trials are groupped into blocks of ten, during each trial of which the same word must be typed, and correct/incorrect feedback is given after each trial. To inform participants of the current context, a visual cue is included, which is an image of the word to type (or symbol representing this word). When the image is perfectly clear, the current context is identified and the corresponding word can be typed successfuly. If the image is blurred or partially obscured, context inference can take place during a number of trials, in which the visual stimulus and the feedback after each trial are integrated over time.

In this layer, each context is represented by a single population of neurons $w_i$, where $i = 1, 2, ..., N$, where $N$ is the number of contexts. If context $i$ is identified, $w_i$ becomes activated (average activity of 1) and stays active until the task has been completed (e.g. the word has been typed). $w_i$ will send constant activity to the first cluster of the adequate sequence $S_i$, and the CSHS will take care of the rest.

Furthermore, we think of the whole $w$ population as representing a probability distribution over contexts. This means that if the context is unclear, $w_i \ne 0$ for more than one $i$. This will create behavior that could be a mix of contexts, representing the participant's uncertainty over the context, or the model's uncertainty over the actions of the participant.

It might prove useful to extend this paradigm and make $w_i$ also partially activate all the neurons that are involved in the sequence of actions associated with $w_i$, while keeping the first element with a higher activation, which is what triggers the sequence. This might have the effect of enabling interference between sequences when the context is unclear.

\end{chapter}

\bibliography{../../MyLibrary}

\end{document}

### 2021-10-13 adaptation/article/article.tex
Update on Overleaf.

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{authblk}
@@
 Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-In this work, we use a computational mechanistic model to explain the behavioral phenomena effected by imperfect contextual information.
+In this work, we use a model to explain the behavioral phenomena effected by imperfect contextual information.
 Specifically, we present a hierarchical computational model for motor adaptation based on exact Bayesian inference.
@@
 a mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error.
-Across many different motor adaptation experiments, now well-established phenomena have been observed, for example: (i) savings (the ability to recall previously-learned skills), (ii) quick de-adaptation (the ability to return to unmodified dynamics), (iii) anterograde interference (the interference in motor learning between opposing manipulations in dynamics) and (iv) spontaneous recovery (during trials where errors are forced to be zero, the display of behavior consistent with the previously-learned dynamics).
+Across many different motor adaptation experiments, now well-established phenomena have been observed, for example: (i) savings (the ability to recall previously-learned skills), (ii) quick de-adaptation (the ability to return to unmodified dynamics), (iii) anterograde interference (the interference in motor learning between opposing manipulations in dynamics) and (iv) spontaneous recovery (during trials where errors are forced to be zero, the display of behavior consistent with the previously-learned dynamics) \citep[e.g.][]{Davidson_Scaling_2004}.
 
@@
 In some experiments, the context is clearly inferable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}.
-In many other experiments, the context is not directly observable, and context inference is probabilistic and takes the form of an evidence-accumulating process.
+In many other experiments, the context is not directly observable, and context inference is probabilistic and takes the form of an evidence-accumulating process across multiple trials.
 In this work, we propose that the behavioral consequences of this slow process of evidence accumulation can be observed in behavior and can be used to explain the phenomena outlined above.
@@
 it could be a movie prop that looks like a rock) and integration over time might be necessary.
+All the variables of the model are fully described in the Methods section.
 
@@
 Here, we use this principle and chose to make use of an exact Bayesian learner.
-As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials.
+As has been discussed before \citep{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for multiple explicit time scales of learning as the learner can deal with adaptation after any number of trials.
 As we will show below this feature will be useful to explain specific key phenomena parsimoniously.
@@
 \begin{equation}
-p(\vartheta | s, c) = f(s, c; \phi) \label{eqn:forward-model}
+p(\vartheta | s, c; \phi) = f(s, c; \phi) \label{eqn:forward-model}
 \end{equation}
-where $\vartheta$ are the outcomes of an action (e.g.
+where $\vartheta$ are the predicted outcomes of an action (e.g.
 the final position of the hand in a reaching movement), $s$ is the current state (the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$.
@@
 
-\subsection{From observation to action selection}
+\subsubsection{From observation to action selection}
 Each trial starts with a new observation $\vartheta_t$, which comprises contextual sensory information (when available), as well as the outcome of the previously-selected action.
@@
 
-\section{Experimental results}
+\subsection{Experimental results}
 In this section, we present experimentally-observed phenomena in four sections, and show that context inference, as done in our model, provides a unifying explanation for all of them.
@@
 Finally, in the fourth section we discuss context inference during so-called error-clamp trials, and its effect on behavior.
+For each of the four sections, we selected one or two studies which are representative of the phenomenon being discussed; a more extensive review of the literature can be found in the Discussion.
 
@@
 
-\subsection{Savings and slow/fast switching}
+\subsubsection{Savings and slow/fast switching}
 The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
-Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
+Savings is almost universally observed in human participants \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
 In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
@@
 
-To our knowledge, there are no experiments in which no contextual information is available.
-However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information.
-These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is nevertheless available to participants.
-In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.
-
 Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
@@
 
-To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \cite{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information.
+To our knowledge, there are no experiments in which no contextual information is available.
+However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information.
+These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is nevertheless available to participants.
+In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.
+
+To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies: \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} which differ in the amount of contextual information.
 \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
@@
 \cite{Oh_Minimizing_2019} performed partially-cued experiments similar to \cite{Kim_Neural_2015}.
-The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10.
+The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10 degrees.
 The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
@@
 
-\subsection{Uncertainty over contexts affects action selection}
+\subsubsection{Uncertainty over contexts affects action selection}
 As with learning, our model selects actions (motor commands) based on context inference.
@@
 
-\subsection{Action selection in error-clamp blocks}
+\subsubsection{Action selection in error-clamp blocks}
 During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Spontaneous recovery, during which participants' behavior is consistent with a previously-encountered context (e.g.
@@
 To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context for each simulated run (one line per run, 10 runs per group, all groups).
-The start and speed of the drop dependd on the run (i.e.
+The start and speed of the drop depend on the run (i.e.
 the participant), reproducing the different lags observed in experimental data.
@@
 \subsubsection{Volatility has no effect on learning}
-The volatility of the environment (i.e.
-unpredictable switching between contexts) lowers the measured learning rate.
+Experiments \citep{Herzfeld_memory_2014} have shown an effect of the volatility of the environment (i.e.
+unpredictable switching between contexts) on measured learning rate.
 Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
-This would manifest in \fref{fig:herzfeld-2014} in the form of no discernible differences between the two groups of participants.
+In the experiments by \cite{Herzfeld_memory_2014}, participants had to adapt to two contexts, and the context could change from trial to trial, with a fix probability.
+This probability was different for two groups of participants.
+Our model predicts that no discernible differences would be observed between the participants that performed the task in a highly volatile environment, and those that performed in a very stable environment.
 
@@
 The model also predicts an effect reminiscent of multisensory integration \citep[e.g.][]{Ernst_Humans_2002}: in order to integrate contextual information from conflicting sources (e.g.
-probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its precision (the inverse of its variance).
+probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its reliability (e.g.
+the inverse of its variance).
 Such integration would manifest itself in experiments in which observations are noisy, as in the experiments by \cite{Kording_Bayesian_2004}, in which the position of the finger was obscured and instead participants are shown a blurry cursor which was some times shifted from its real position.
@@
 \begin{itemize}
-\item $t$: Time at the moment an observation is made, before inference is done.
+\item $t$: Discrete time at the moment an observation is made, before inference is done. 
+Here taken as the trial number.
 \item $\vartheta_t$: Observation. Note that this is assumed to be a noisy
-observation of the generalized state (see below) given by $\vartheta_t = h(z_t)$.
+observation of the generalized state (see below)
 \item $s_t$: Hidden state, e.g. the position of the hand in reaching
@@
 \item $c_t$: Motor command (action) taken after observing $\vartheta_t$ and inferring the context.
-\item $\zeta_t$: Context inferred after observing $\vartheta_t$.
+\item $\zeta_t$: Categorical variable representing the context inferred after observing $\vartheta_t$.
 \item $\omega_t$: Contextual information. This term includes, e.g. visual cues,
 proprioceptive feedback, performance feedback.
-\item $z_t$: Generalized state, $z_t = \{s_t, \zeta_t, \omega_t\}$. Observed noisily
+\item $z_t$: Generalized state, $z_t = \{s_t, \zeta_t\}$. Observed noisily
 as $\vartheta_t$.
+\item $m_i$: Forward model for context $\zeta_i$.
+\item $\phi_i$: Parametrization of the forward model $m_i$. In adaptation experiments, this
+can be interpreted as the inferred size of the adaptation.
 \end{itemize}
@@
 \[
-p(z_t, z_{t-1}, c_{t-1}, \vartheta_t, \vartheta_{t-1}) = p(\vartheta_t | z_t)p(z_t|\vartheta_{t-1}, c_{t-1})
+p(z_t, \vartheta_t | z_{t-1}, c_{t-1}) = p(\vartheta_t | z_t)p(z_t|\z_{t-1}, c_{t-1})
 \]
@@
 \[
-q(\gamma | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)p(z_{t-1})p(\phi)
+q(\phi | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)q(z_{t-1}|\ldots)p(\phi)
 \]
-
-where $\phi$ are the parameters of the internal model for the dynamics,
+where $\phi$ are the parameters of the internal model for the dynamics, also called the forward model,
 i.e. the function $f: (z_t, c_t) \rightarrow z_{t+1}$, which determines the beliefs of
-the agent regarding how the system evolves after having taken action $c_t$.
+the agent regarding how the system evolves after having taken action $c_t$. $q(z_{t-1}|\ldots)$ is the
+posterior probability calculated in the previous trial.
 
@@
 \begin{equation}
-p(c_t | z_t) = \displaystyle\sum_{m \in M}q(\zeta_t)p(c_t | z_t, m)
+p(c_t | s_t, \zeta_t) = \displaystyle\sum_{i}q(\zeta_i)p(c_t | s_t, \zeta_i)
 \end{equation}
-where $M$ is the set of all relevant forward models (see below), and
-$p(a_t | ... m)$ is the probability distribution over actions prescribed by the
-forward model $m$.
+where $i$ runs over all contexts and $p(a_t | ... m)$ is the probability distribution over actions
+prescribed by the forward model $m_i$ for context $\zeta_i$.
 
@@
 \begin{equation}
-z_t = \{s, \vartheta_t, \phi_t, \omega_t\}
+z_t = \{s_t, \zeta_t\}
 \end{equation}
-where $s$ is, e.g. in a reaching task, the current position of the hand in
+where $s_t$ is, e.g. in a reaching task, the current position of the hand in
 Cartesian coordinates, where the origin is the starting point. We assume that
@@
   q(\zeta_t | \omega_t) &\propto p(\omega_t | \zeta_t)p(\zeta_t) \\
-  q(\zeta_t | s_t, s_{t-1}, c_{t-1}) &\propto p(s_t | \zeta_t, s_{t-1}, c_{t-1})p(x_t)p(s_{t-1})p(\zeta_t)
+  q(\zeta_t | s_t, s_{t-1}, c_{t-1}) &\propto p(s_t | \zeta_t, s_{t-1}, c_{t-1})p(\zeta_t)
   \end{split}
@@
 (e.g. clockwise vs. counter-clockwise) to keep the error space
-two-dimensional. With this, $s_t > 0$ represents motor error in one direction
+one-dimensional. With this, $s_t > 0$ represents motor error in one direction
 (e.g. clockwise) and $s_t < 0$, error in the other direction. We made this
-choice because, in this work modeled different experiments with
+choice because in this work we modeled different experiments with
 different state spaces, but with the commonality of (at most)
@@
 own physical space (e.g. hand positions in a reaching task), only the dynamical
-system would need to be adapted.
+system  $p(s_t | \zeta_t, s_{t-1}, c_{t-1})$ would need to be adapted.
 
-We assume that the decision-making agent ha a baseline internal forward models that
-map motor commands to outcomes:
+We assume that the decision-making agent has a baseline internal forward model $m_O$ that
+maps motor commands to outcomes:
 \begin{equation}
-p(s_{t+1} | s_t, c_t, m) = f_O(s_t, c_t)
+p(s_{t+1} | s_t, c_t, m=m_O) = f_O(s_t, c_t)
 \end{equation}
@@
 To complete the action-selection picture, we return to context inference. As
-discussed above, action selection is made via a weighted sum of the different
-forward models:
+discussed above, action selection is made via a weighted sum of the motor commands
+produced by each forward model:
 \begin{equation}
-p(c_t | ...) \propto \displaystyle \sum_{m \in M} q(\zeta_i)p(c_t | m)
+p(c_t | ...) \propto \displaystyle \sum_{i} q(\zeta_i)p(c_t | \zeta_i, ...)
 \end{equation}
@@
 \begin{align}
-  p(s_t | \vartheta) &= N(\mu_s, \sigma_s)  \\
-  \mu_s &= \vartheta_{t-1} + s \\
+  p(s_t | \phi) &= N(\mu_s, \sigma_s)  \\
+  \mu_s &= \phi + s_t \\
   \sigma_s &= \xi_0 1
@@
 This likelihood is the probability of observing $s$ given the previous estimate
-of the parameters of the force field $\phi_{t-1}$, and given the dynamics of
+of the parameters of the forward model, $\phi_{t-1}$, and given the dynamics of
 the system (see below).
@@
 \end{equation}
-where $\mu_{\phi}^0, \nu_{\phi}^0, \alpha_{\phi}^0$ and $\beta_{\phi}^0$ are free hyperparameters of the
-model.
+where $\mu_{\phi}^0, \nu_{\phi}^0, \alpha_{\phi}^0$ and $\beta_{\phi}^0$ are free
+hyperparameters of the model.
 
@@
 \begin{equation}
-q(\mu_f, \sigma_f | s_t, s_{t-1}, c_{t-1}) \propto p(s_t |
-s_{t-1}, c_{t-1}, \zeta_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
+q(\mu_\phi, \sigma_\phi | s_t, s_{t-1}, c_{t-1}) \propto p(s_t |
+s_{t-1}, c_{t-1}, \zeta_t)p(\mu_\phi, \sigma_\phi) \label{eqn:context-from-x}
 \end{equation}
@@
 the model. It is assumed that the standard deviation over outcomes ($\sigma_c$)
-is related to that of the parameter $\mu_f$ as follows:
+is related to that of the parameter $\mu_\phi$ as follows:
 \begin{equation}
-\sigma_f = \sigma_a / \nu_f
+\sigma_\phi = \sigma_c / \nu_\phi
 \end{equation}
-which makes $\nu_f$ a parameter to estimate. Each one of the forward models can
+which makes $\nu_\phi$ a parameter to estimate. Each one of the forward models can
 be updated at every trial, depending on context inference (see main text). The

### 2021-10-20 adaptation/article/article.tex
Small changes on article

--- 2021-10-01 adaptation/article/article.tex
@@
 Here, we use this principle and chose to make use of an exact Bayesian learner.
-As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials.
+As has been discussed before \citep{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials.
 As we will show below this feature will be useful to explain specific key phenomena parsimoniously.
@@
 The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
-Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
+Savings is almost universally observed in human participants \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
 In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
@@
 In fact, the mere act of holding the handle informs participants that something is different from baseline.
-Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
-The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.
+Proprioceptive signals can also provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
+The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \citep{Ethier_Spontaneous_2008}.
 
-To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \cite{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information.
+To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \citep{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information.
 \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
@@
 Consistent with our model, the authors found that switching was immediate and accurate.
-In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
+In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results by \cite{Kim_Neural_2015}.
 The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e.

### 2021-10-20 adaptation/article/article.tex
Removed redundancy. 1.5k words to go

--- 2021-10-13 adaptation/article/article.tex
@@
 \section{Introduction}
-Humans are able to adapt their movements to changing dynamics of the environment and of the body itself.
-We adapt to slow changes in our body dynamics, e.g.
-because of muscle exhaustion throughout the day, but we can also adapt to sudden changes in the environment.
-For example, when picking up a heavy object, we must adapt our motor commands to be able to move the extra weight safely.
-Even before picking it up, the size and material of an object provide us with contextual information which we can use to prepare the correct posture for moving the object.
-However, contextual information cannot always be relied on, e.g.
-a large parcel with unknown content.
-
 It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
 Additionally, humans have been shown to dynamically switch between different learned adaptation \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
-The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are an ongoing, active topic of research.
 
@@
 To explain these phenomena, a number of computational models have been introduced.
-The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors.
-By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, partially explaining these distinct experimental phenomena.
-
-Bayesian accounts of motor adaptation have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
-These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the the same task is practiced, the uncertainty on an estimate of the underlying modification is reduced, which lowers variability in responses and increases the stability of the estimate.
+The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}.
+Bayesian accountshave also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
 
 While these general models of adaptation explain the most common phenomena observed in experiments, other known phenomena remain outside of their scope.
-For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}.
+For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} or adaptations slowly introduced \citep{Huang_Persistence_2009}.
 Action selection has also been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013,Davidson_Scaling_2004}.
@@
 whether the task was successfully completed).
-Having inferred the context, the model can choose the best-fitting internal model with which to make decisions.
 
@@
 Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors.
-For example, we can learn to perform a reaching movement even when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
+For example, we can learn to perform a reaching movement when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
 
@@
 Here, we use this principle and chose to make use of an exact Bayesian learner.
-As has been discussed before \citep{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for multiple explicit time scales of learning as the learner can deal with adaptation after any number of trials.
-As we will show below this feature will be useful to explain specific key phenomena parsimoniously.
-
-To describe this component of the model, we make use of the example of reaching movements. We define a forward model as follows:
-\begin{equation}
-p(\vartheta | s, c; \phi) = f(s, c; \phi) \label{eqn:forward-model}
-\end{equation}
-where $\vartheta$ are the predicted outcomes of an action (e.g.
-the final position of the hand in a reaching movement), $s$ is the current state (the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$.
-We assume that during learning the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.
-
-When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model.
-To describe this adaptation process, our model makes uses of Bayesian inference, such that
-\begin{equation}
-q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi) \label{eqn:update-mini}
-\end{equation}
-where $q(\phi | ...)$ represents the new estimated probability distribution over the possible values of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$.
-The likelihood $p(\vartheta | s, c, \phi)$ is the probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.
+To adapt to altered dynamics (e.g.
+a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model.
+These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation.
+With this, a posterior estimate on $\phi$, $q(\phi)$, is calculated after each trial.
 
@@
 Following these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.
+As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.
 
-In the model, context is a categorical variable $\zeta$ which is inferred using Bayes' theorem:
-\begin{equation}
-q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
-\label{eqn:context-inference-mini}
-\end{equation}
-where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall expectation of observing any one context, if the context did change.
-Below, we discuss how these two sources of information effect different experimentally observed phenomena.
-
-In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions.
-In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment.
-This includes, as before, the outcomes of actions (e.g.
-the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them, as is the case for visual cues in experiments \cite[e.g.][]{Addou_Colored_2011}.
-For this section, the most important component of $\vartheta$ is contextual sensory information, i.e.
-information that might directly help infer the identity of the current context.
-
-In experimental settings, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}.
-All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to infer the identity of a context.
+This component of the model can be used to include contextual sensory information from many sources.
+In experimental settings, for example, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}.
+All these sources of information, alongside the outcomes of previous actions and the prior beliefs over context $p(\vartheta)$ are integrated to infer the identity of a context.
 
@@
 In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model.
-In other words, a distribution is created as follows:
-\begin{equation}
-p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta) \label{eq:action-selection}
-\end{equation}
-where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.
 
-\subsubsection{From observation to action selection}
-Each trial starts with a new observation $\vartheta_t$, which comprises contextual sensory information (when available), as well as the outcome of the previously-selected action.
-
-The first step in the model is to perform context inference using the current observation $\vartheta_t$ according to \eref{eqn:context-inference-mini} and \eref{eqn:estimated-context}.
-Once a probability distribution over contexts has been created, all models are updated according to \eref{eqn:update-mini} and \eref{eqn:update-full}, based on the distribution over contexts.
-Finally, action selection is performed using the distribution over contexts as well as the updated forward models, following \eref{eq:action-selection}.
+In other words, a distribution $p(c_t)$ over all possible motor commands is created following \eref{eqn:dist-comm}, from which the next action (motor command $c_t$) can be sampled and carried out.
 
@@
 In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models.
-This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007} where we show in addition that expanding forward model switching into context inference explains facets of savings that these previous models do not.
+This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}.
+Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not.
 
 To show this, we examined multiple experimental studies in which savings are observed.
-We categorized these studies based on the amount of contextual information made available to participants into cued-context and partially-cued experiments: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
+We categorized these studies based on the amount of contextual information made available to participants: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
 We call these cued-context experiments.
@@
 In fact, the mere act of holding the handle informs participants that something is different from baseline.
-Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
+Proprioceptive signals can also provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
 The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.
@@
 However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.
-While we focused here on a specific, well-known experiment, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion.
 
@@
 \end{figure}
-
-Further evidence for this effect of uncertainty about contexts can be found during error-clamp blocks in several experiments.
-We discuss these experiments in the following section.
-Additionally, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion.
 
@@
 It can be seen that group 1.3 (i.e.
-the participants who had learned in the -A/2 context in addition to A) more quickly recognizes a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.
+the participants who had learned in the -A/2 context in addition to A) more quickly recognized a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.
 
@@
 
-Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003,Imamizu_Explicit_2007}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command.
-We built on these studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
+We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task.
+With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
 
@@
 
-It is important to note that the most essential part of the model we presented is the concept of the context and the inclusion of a Bayesian inference scheme over the context, which integrates all the information that is available to the participant during the experiment.
-Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations.
-This choice greatly simplified our simulations and provides intuitive interpretations of experimental results.
-However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
+Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
 A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.
-
-In what follows, we review the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model.
-In addition, we discuss several predictions that our model makes, which could be verified experimentally in future studies.
 
@@
 The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed with which participants adapted to curl forces.
-While it has been shown that sensitivity to error can be modulated \citep{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}), to our knowledge there are no other experiments suggesting that it depends on the history of errors.
+While it has been shown that sensitivity to error can be modulated \citep{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience, to our knowledge there are no other experiments suggesting that it depends on the history of errors.
 As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data.
@@
 Context is often directly observable, either via environmental cues or via feedback observed after actions.
-In many behavioral experiments, for example, the context of the current block of trials is easily identifiable by the color of a cue light \citep[e.g.][]{Ethier_Spontaneous_2008}, the posture of the hand \citep[e.g.][]{Gandolfo_Motor_1996}, the target of the movement \citep[e.g.][]{Lee_Dual_2009} or by the workspace in which the task is being performed \citep[e.g.][]{Shadmehr_Adaptive_1994}.
-The context can also be selected by the agent itself.
-For example, a person can decide the level of care devoted to moving an egg or a lead ball around, which determines the way in which the person would move their hands.
-In these cases, the context is clearly determined and known to the agent and switches between different adaptations are immediate.
-
-In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
+It can also be selected by the agent itself.
+For example, a person can decide the level of care devoted to moving an egg or a lead ball around.
+In many other cases, however, context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
 It is in these cases where the effects of context inference are most noticeable.
@@
 The authors showed that by matching the errors expected by participants, they eliminated the slow tapering off observed in most E blocks.
-As we showed in \fref{fig:vaswani-2013}D, this effect need not be all or none: in some experimental sessions, participants might fail to notice that a change has occurred due to the stochasticity of their own actions.
-This would lead to an observed delay before a participant changes her behavior after the E block has started.
 
@@
 This uncertainty affects learning and behavior during motor adaptation, effecting phenomena that are directly observable during behavioral experiments.
-
-In addition to providing a unifying explanation for phenomena such as switching, savings, spontaneous recovery and slow return to baseline, our model makes unique predictions that could be tested in future experiments.
 
@@
 \[
-p(z_t, \vartheta_t | z_{t-1}, c_{t-1}) = p(\vartheta_t | z_t)p(z_t|\z_{t-1}, c_{t-1})
+p(z_t, \vartheta_t | z_{t-1}, c_{t-1}) = p(\vartheta_t | z_t)p(z_t|z_{t-1}, c_{t-1})
 \]
@@
 Additionally, after having inferred the current state, motor adaptation is given by
-\[
-q(\phi | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)q(z_{t-1}|\ldots)p(\phi)
-\]
+\begin{equation}
+q(\phi | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)q(z_{t-1}|\ldots)p(\phi) \label{eqn:motor-adaptation}
+\end{equation}
 where $\phi$ are the parameters of the internal model for the dynamics, also called the forward model,
@@
 \begin{equation}
-p(c_t | ...) \propto \displaystyle \sum_{i} q(\zeta_i)p(c_t | \zeta_i, ...)
+p(c_t | ...) \propto \displaystyle \sum_{i} q(\zeta_i)p(c_t | \zeta_i, ...) \label{eqn:dist-comm}
 \end{equation}

### 2021-10-22 adaptation/article/article.tex
5k words reached. All good

@@
-\documentclass[a4paper,doc,floatsintext,natbib]{apa6}
-% \documentclass{article, a4paper}
+\documentclass[a4paper,doc,floatsintext,natbib]{article}
+% \documentclass{article}
 \usepackage[font=small]{caption}
-\usepackage{lscape}
+\usepackage{natbib}
 \usepackage{graphicx}
@@
 \usepackage{amsfonts}
-\usepackage{authblk}
 \usepackage[utf8]{inputenc}
 \usepackage{nameref}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{soul}
 \usepackage{todonotes}
+\usepackage{authblk}
+\usepackage[nomarkers,figuresonly]{endfloat}
 
 % Remember to start reftex-mode
-
-\hypersetup{
-  colorlinks=true,
-  linkcolor=black,
-  citecolor=black,
-  urlcolor=black
-  }
+\bibliographystyle{ieeetr}
 
@@
 
-\title{The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery}
-\shorttitle{Context-inference-dependent motor learning}
+\title{The effects of ongoing context inference on motor learning}
+% \shorttitle{Context-inference-dependent motor learning}
 \author[1,2]{Cuevas Rivera, Darío}
@@
 \affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
-\affiliation{~}
+% \affiliation{~}
 
@@
 \section{Introduction}
-It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.
-Additionally, humans have been shown to dynamically switch between different learned adaptation \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
+It has been shown that humans can adapt motor commands to counteract changes in the dynamics of the environment and their own bodies, such as performing reaching movements with a weight attached to the wrist.
+This is known as motor adaptation.
+Moreover, human participants have been shown to adapt to different, even opposing, such changes during the course of a single experiment \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}.
+Additionally, humans have been shown to dynamically switch between different learned adaptations \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
 
-In order to study motor adaptation, motor error is defined in experiments as the deviation between the observed movements of a participant and a baseline movement.
-For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient can be defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error.
-Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.
+By introducing blocks of trials in which body dynamics are altered (e.g.
+a mechanical arm exerts a force on the participant's hand), experimenters are able to observe motor adaptation through the lens of motor error.
+Across many different motor adaptation experiments \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Davidson_Scaling_2004}, well-established phenomena have been observed: (i) the ability to recall previously-learned skills, called savings; (ii) the ability to return to unmodified dynamics, termed de-adaptation; (iii) the interference in motor learning between opposing manipulations in dynamics, called anterograde interference; (iv) spontaneous display of behavior consistent with a previously-learned adaptation, during trials where errors are forced to be zero, called spontaneous recovery.
 
-By introducing blocks of trials in which a manipulation is present (e.g.
-a mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error.
-Across many different motor adaptation experiments, now well-established phenomena have been observed, for example: (i) savings (the ability to recall previously-learned skills), (ii) quick de-adaptation (the ability to return to unmodified dynamics), (iii) anterograde interference (the interference in motor learning between opposing manipulations in dynamics) and (iv) spontaneous recovery (during trials where errors are forced to be zero, the display of behavior consistent with the previously-learned dynamics) \citep[e.g.][]{Davidson_Scaling_2004}.
-
-To explain these phenomena, a number of computational models have been introduced.
-The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}.
-Bayesian accountshave also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
+To explain these phenomena, a number of computational models have been introduced, which adapt their motor commands after observing motor errors.
+The most studied are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, but Bayesian accounts have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between forward models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
 
@@
 
-In this work, we show that these unexplained phenomena, alongside savings, de-adaptation, anterograde interference and spontaneous recovery, can be explained in terms of the concept of context, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
-Intuitively, in mechanical arm experiments, one can think of one context in which the arm exerts no force and one in which it does.
-By inferring the current context, a participant can know how to counteract the mechanical arm and succeed in a reaching movement.
-In some experiments, the context is clearly inferable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}.
-In many other experiments, the context is not directly observable, and context inference is probabilistic and takes the form of an evidence-accumulating process across multiple trials.
-In this work, we propose that the behavioral consequences of this slow process of evidence accumulation can be observed in behavior and can be used to explain the phenomena outlined above.
-
-To this end, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
-We build on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
-In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions.
-To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment (e.g.
-whether the task was successfully completed).
+In this work, we show that these unexplained phenomena can be explained in terms context inference, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
+We build on models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019} by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
+In the proposed model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions.
+To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.
 
@@
 \section{Results}
-In this work, we present a motor adaptation model in which the agent (e.g.
-a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation.
-Several models exist for such error-based motor adaptation \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019}.
-We expand on these existing models by adding an explicit component of context inference.
-As we will show, this context inference guides the selection of the adequate internal model, its updating and the sampling of actions.
+We present a motor adaptation model in which an existing model of the environment is updated based on error signals produced by the need of adaptation.
+We expanded on existing models \citep{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019} by adding an explicit component of context inference.
+As we will show, context inference guides the selection of the adequate forward model, its updating and the sampling of actions.
 
-In what follows, we show that the proposed model explains several experimental phenomena found in previous studies.
-We simulated representative experiments from these studies to illustrate how our model explains these experimental findings using the dynamics of context inference.
+We simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference.
 We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.
@@
 The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection.
-The processes defined by these component happen in this order, and each component informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.
+The processes defined by these component occur in this order, and each component informs the ones that follow.
 
-Central to the model is the concept of context.
-In our model, contexts are defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the variables of the environment.
+Central to the model is the concept of context, defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the changing environment.
 Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to infer the context.
@@
 When observing an object to be lifted, the agent can use its size and texture to estimate the weight of the object, which in turn allows the agent to infer the appropriate context and, with it, decide how to lift the object.
-These contextual cues (size and texture) do not necessarily identify the context with certainty (e.g.
-it could be a movie prop that looks like a rock) and integration over time might be necessary.
-All the variables of the model are fully described in the Methods section.
 
@@
 \subsubsection{Motor adaptation}
-Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors.
-For example, we can learn to perform a reaching movement when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.
-
 It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}.
-Here, we use this principle and chose to make use of an exact Bayesian learner.
+We used this principle and chose to make use of an exact Bayesian learner.
 To adapt to altered dynamics (e.g.
@@
 These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation.
-With this, a posterior estimate on $\phi$, $q(\phi)$, is calculated after each trial.
+With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}B and C, is calculated after each trial.
 
 \subsubsection{Context inference}
-In our model, the idea of context is quite general (see \fref{fig:model}A).
-The context comprises the relevant elements of the environment (e.g.
-the strength and direction of the wind when walking), of the task at hand (e.g.
-immediate and long-term goals and reward structure) and task-related information internal to the decision-making agent (e.g.
-forward model, decision-making algorithms, motor adaptation equations, expected observations).
-Most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g.
-two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e.
-the model that best represents the context.
-In addition, \cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.
+To infer the adequate forward model to deploy, most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the forward models: the model that best predicts the current observation given the past is deemed to be the ``correct'' model, i.e.
+the model that best represents the current context.
+\cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.
 
-Following these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.
+Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors.
 As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.
-
-This component of the model can be used to include contextual sensory information from many sources.
-In experimental settings, for example, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}.
-All these sources of information, alongside the outcomes of previous actions and the prior beliefs over context $p(\vartheta)$ are integrated to infer the identity of a context.
 
@@
 \subsection{Experimental results}
-In this section, we present experimentally-observed phenomena in four sections, and show that context inference, as done in our model, provides a unifying explanation for all of them.
+In this section, we present experimentally-observed phenomena in three sections, and show that the dynamics of context inference provide a unifying explanation for all of them.
 In the first section, we discuss switches between contexts, and how slow context inference affects these switches.
 In the second section, we focus on the effects of volatility in the environment on motor adaptation.
-In the third section, we discuss the effects of slow context inference on action selection.
-Finally, in the fourth section we discuss context inference during so-called error-clamp trials, and its effect on behavior.
-For each of the four sections, we selected one or two studies which are representative of the phenomenon being discussed; a more extensive review of the literature can be found in the Discussion.
+Finally, in the third section we discuss context inference during so-called error-clamp trials, and its effect on behavior.
+For each of the three sections, we selected one or two studies which are representative of the phenomenon being discussed.
 
-For clarity, we first introduce necessary terminology that is typically used in experimental studies and will be useful for linking experimental to computational observations.
-As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force (i.e.
-a force that pushes the participant's hand in the direction perpendicular to movement).
+For clarity, we first introduce necessary terminology that is typically used in experimental studies.
+As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force on the participant's hand.
 Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
 Let us define the baseline context O as that in which the mechanical arm exerts no force.
-Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise.
-Additionally, abusing notation, the usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
-Studies often talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g.
-clockwise), but the second one has half the magnitude.
+Contexts A and B can be defined as those with clockwise or counter-clockwise, respectively.
+Abusing notation, a usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
+Similarly, one can define context A/2, with the same direction of adaptation as A, but half the magnitude.
 Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter E.
@@
 With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of O-A-B-E, or O-A-(-A)-E, which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).
-
-In some experiments, adaptation blocks are repeated throughout the experiments.
-For example, \cite{Oh_Minimizing_2019} used an O-A-O-A-O-A-O-A sequence.
-To differentiate the repeating blocks, the authors named them $O_i$ and $A_i$, where $i = 1, 2, ...$.
-Their experiment can be described as $O_1$-$A_1$-$O_2$-....
-We adopt this naming convention.
-
-We use the term "learned adaptation" and similar to refer to motor outputs (e.g.
-reaching movements) that have already been adapted by the participant to a change in the dynamics of the environment.
-For example, the learned adaptation during a task with a curl force refers to the movements of the participant that already counteract the force exerted by the mechanical arm.
+With repeated contexts \citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1$-$A_1$-$O_2$-\ldots .
 
 \subsubsection{Savings and slow/fast switching}
-The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it.
-Savings is almost universally observed in human participants \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
+The term 'savings' refer to the ability to remember a previously-learned adaptation and apply it without having to re-learn it.
+Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
 In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
+The related concept of quick de-adaptation occurs in A-O transitions, where participants switch back to baseline without having to re-learn it.
 
-In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models.
+In this section, we discuss savings and de-adaptation in terms of switching between forward models.
 This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}.
-Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not.
+Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not, namely that savings are not immediate, but a relatively fast process that reflects context inference.
 
@@
 
-Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
-For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone.
-In fact, the mere act of holding the handle informs participants that something is different from baseline.
-Proprioceptive signals can also provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
-The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.
-
-To our knowledge, there are no experiments in which no contextual information is available.
-However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information.
-These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is nevertheless available to participants.
-In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.
-
-To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies: \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} which differ in the amount of contextual information.
+We selected three representative experiments from two studies \citep{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information available to participants.
 \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees.
 Participants performed shooting movements in blocks of trials with the same rotation.
-Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment.
+Importantly, a colored light identified the current context, making this a cued-context experiment.
 Consistent with our model, the authors found that switching was immediate and accurate.
@@
 
-Critically, in \fref{fig:oh-2019}A it can be seen until trial 250 and especially for the first four context switches,  that participants did not completely adapt to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks.
+Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks.
 Switching between opposite adaptations was nevertheless immediate in simulations and experimental data.
 After around 500 trials, participants had adapted completely, as seen by the blue line reaching the heights of the black line, and switches continued to be immediate.
-
 These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches.
@@
 
-\cite{Oh_Minimizing_2019} performed partially-cued experiments similar to \cite{Kim_Neural_2015}.
-The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10 degrees.
-The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model.
+\cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visual rotation of 20 and 10 degrees, respectively.
+The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with our model.
 Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A_1$.
 From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (true rotation).
-In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation (blue line) lagging behind the switches in the real context (black line), and slowly catching up.
+In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up.
 As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C).
@@
 
-Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.
+Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism.
 
@@
 If the identity of the current context is known, the forward model for this context is used to select the current action.
-However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eq:action-selection}).
+However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eqn:dist-comm}).
 
 Experimental evidence supporting this view can be found in experiments with context switching.
-For example, \cite{Davidson_Scaling_2004} reported an curl-force experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A).
+For example, \cite{Davidson_Scaling_2004} reported a curl-force experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A).
 After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
@@
 
-Our model offers a different explanation of the observed results, based on context inference: the asymmetry is due to the existence of the baseline context.
-As discussed above, action selection in our model is done via a weighted average (see \eref{eq:action-selection}).
-In experiments without cued contexts, the baseline context O and its associated forward model $m_O$, have a non-zero probability $p(\zeta | s_t ...)$.
+In our context-inference account, the asymmetry is caused by the baseline context, which have a non-zero probability $p(\zeta_O | s_t ...)$.
 When a new block of trials starts (e.g.
-in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out).
+in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $\zeta_O$ becomes more likely (given that $\zeta_{3A}$ has been ruled out).
 Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force).
@@
 We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}.
-It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.
+It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}.
 Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data.
@@
 in an O-A-E experiment, consistent with A); this phase, when present, is seen during the early trials of the E block.
-(2) A slow return to baseline (i.e.
-no observable adaptation, consistent with the baseline context), which can last as long as hundreds of trials \citep{Brennan_Decay_2015}.
-However, the direction of the spontaneous recovery, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
+(2) A slow return to baseline, which can last as long as hundreds of trials \citep{Brennan_Decay_2015}.
+However, the direction of the spontaneous recovery, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
 
-In this section, we show how our model can explain these different parameters of behavior by changing the way contextual cues mislead participants' context inference, which in turn influences action selection.
+Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
+For example, proprioceptive signals  provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
+The sudden appearance of motor errors can itself be a cue for contextual change \citep{Herzfeld_memory_2014} and even a pause between two trials could suggest a change in context \cite{Ethier_Spontaneous_2008}.
+
+In this section, we show how context inference can explain these different parameters of behavior by changing the way contextual cues mislead participants' context inference, which in turn influences action selection.
 
@@
 
-This effect is explained in our model in terms of context inference.
 In \fref{fig:vaswani-2013}C, the inference over context is shown for each group separately.
@@
 This causes the agent to infer the combination of some of the known contexts that best fits the observations.
-Depending on the contexts previously learned by the agent (which change from group to group), inference during E trials will vary: groups 1.1 and 1.4 have exactly the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
-These agents will slowly lower the force applied until eventually reaching zero (after hundreds of trials; not shown).
-In contrast, group 1.3 has learned the additional -A/2 context (represented by brown in \fref{fig:vaswani-2013}C), which has a non-zero posterior probability during E trials, pushing the agent's adaptation force more quickly towards zero.
-Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more slowly than groups 1.1 and 1.4.
+Depending on the contexts previously learned by the agent: groups 1.1 and 1.4 display the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
+These agents will slowly lower the force applied.
+In contrast, group 1.3 has learned the additional -A/2 context, which has a non-zero posterior probability during E trials, pushing the agent's adaptation force more quickly towards zero.
+Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more quickly than groups 1.1 and 1.4.
 
-Our model also reproduces the variability in the lag before adaptation starts to drop to zero after the E block starts.
+Context inference also explains the variability in the lag before adaptation starts to drop to zero after the E block starts.
 In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
@@
 the participant), reproducing the different lags observed in experimental data.
-This can also be seen directly in context inference, displayed as the shaded areas in \fref{fig:vaswani-2013}C.
 
@@
 Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
-A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.
-
-\subsection{Alternative explanations}
-For each of the representative studies presented in the Results section, the authors put forward explanations for the phenomena they observed in their data.
-While our model provides a single explanation for all these phenomena, it is not necessarily at odds with those presented by the respective authors and might instead act on top of the explanations the authors presented.
-
-\cite{Oh_Minimizing_2019} posited that the difference between their two groups (~\fref{fig:oh-2019}B and C) can be explained by considering that the smaller visual rotation of 10 degrees caused participants to update their baseline forward model instead of creating a new forward model for the context.
-If such an account were true, it would predict that when participants walk away from the experiment they would have trouble adapting to the no-rotation reality, having to re-learn it.
-Our model instead would suggest that participants would immediately switch back to their baseline movements the moment they walk away from the computer screen, as that is a clear indication that the context has switched back to baseline.
-More generally, our model predicts that the observed difference between the two groups would disappear if clear sensory contextual cues were added to the experiment.
-
-The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed with which participants adapted to curl forces.
-While it has been shown that sensitivity to error can be modulated \citep{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience, to our knowledge there are no other experiments suggesting that it depends on the history of errors.
-As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data.
-The same applies to the experiments by \cite{Davidson_Scaling_2004}.
+An exception is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, as these models do not incorporate uncertainty in the estimate of the parameters of the forward models.
 
 \subsection{Further experimental evidence}
-Context is often directly observable, either via environmental cues or via feedback observed after actions.
-It can also be selected by the agent itself.
-For example, a person can decide the level of care devoted to moving an egg or a lead ball around.
-In many other cases, however, context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
+In many cases, the context is not directly observable and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context.
 It is in these cases where the effects of context inference are most noticeable.
@@
 Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
-When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be incorrect.
+When a participants observe an unexpectedly large error, they can infer that the inferred context might be incorrect.
 This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
@@
 
-A subtler source of information can be found in long pauses between blocks of adaptation trials.
-\cite{Ethier_Spontaneous_2008} performed an experiment where the adaptation blocks were divided into sub-blocks of 60 trials each, with a long 30s pause between sub-blocks.
-The authors observed that at the beginning of each sub-block, participants showed an average dip in their learned adaptation, suggesting that they were partially returning to baseline (O) after each pause.
-This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occurred.
-Because this was inferred without any information that points at the identity of the new context, prior beliefs take over.
-This means that participants rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context A, but now includes a component of the baseline O, as it is the most common one in everyday life.
+A subtler source of information can be found in long pauses between blocks of adaptation trials, after which an unprompted partial return to baseline has been observed \cite{Ethier_Spontaneous_2008}.
+This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occurred, prompting participants to rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context A, but now includes a component of the baseline O, as it is the most common one in everyday life.
 
 Error-clamp (E) trials present another insight.
-During E trials, error is artificially kept at zero.
-This takes different forms, depending on the type of experiment.
 If error is kept at zero, one could assume that participants would continue doing what they were doing before, as there is no reason (no observed error) to infer a change in context.
 However, this is almost never the case \cite[e.g.][]{Smith_Interacting_2006,Ethier_Spontaneous_2008,Forano_Timescales_2020,Vaswani_Decay_2013,Scheidt_Persistence_2000,Pekny_Protection_2011}.
-Instead, participants slowly reduce their adaptation, often displaying spontaneous recovery \cite[e.g.][]{Smith_Interacting_2006}, i.e.
-an initial response in the direction of one of the learned adaptations.
-Our model provides a principled account of this behavior: while the observed error (i.e.
-the cursor on the screen) is kept at zero, participants still receive proprioceptive feedback which tells them that something has changed.
-The natural variability in a participant's behavior will lead her to expect observing errors, which clashes with the observed zero error.
-This prompts the participant to re-evaluate their inferred context, which can partially activate a previously-learned adaptation, as we showed in \fref{fig:vaswani-2013}.
+Instead, participants slowly reduce their adaptation, often displaying spontaneous recovery \cite[e.g.][]{Smith_Interacting_2006}.
+Our model provides a principled account of this behavior: the natural variability in participants' behavior lead them to expect errors, which clashes with the observed zero error.
+This prompts participants to re-evaluate their inferred context, which can partially activate a previously-observed context, as we showed in \fref{fig:vaswani-2013}.
 \cite{Pekny_Protection_2011} found similar results, demonstrating that the duration of the previously-observed adaptation block also affects behavior in the E block.
@@
 
-\subsubsection{Error-clamp as a context}
+\subsubsection{Error-clamp as a known context}
 The inclusion of reliable sensory contextual cues (e.g.
@@
 
-\subsubsection{Volatility has no effect on learning}
+\subsubsection{Reducing the effect of volatility on learning}
 Experiments \citep{Herzfeld_memory_2014} have shown an effect of the volatility of the environment (i.e.
@@
 Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
-In the experiments by \cite{Herzfeld_memory_2014}, participants had to adapt to two contexts, and the context could change from trial to trial, with a fix probability.
-This probability was different for two groups of participants.
-Our model predicts that no discernible differences would be observed between the participants that performed the task in a highly volatile environment, and those that performed in a very stable environment.
 
-If this prediction is confirmed, our model would additionally suggest that human participants do not revisit the learned adaptation of the last trial when a new observation comes in.
+If this prediction is confirmed, our model would additionally suggest that human participants do not revisit the learned adaptation of the previous trial when a new observation comes in.
 To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}: at trial $t$, the true context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
@@
 Does the participant update the internal model of A or of B?
-According to our model, the participant incorrectly updates A and, upon learning of her mistake at trial $t+1$, does not revert this update.
-If this were not the case, the volatility of the environment would have no effect, as the context at trial $t$ can almost always be identified at trial $t+1$.
+According to our model, participants incorrectly update A and, upon learning of their mistake at trial $t+1$, do not revert this update.
+If this were not the case, the volatility of the environment would have no effect even without contextual cues, as the context at trial $t$ can almost always be identified at trial $t+1$.
 
 \subsubsection{Multi-source integration}
-The model also predicts an effect reminiscent of multisensory integration \citep[e.g.][]{Ernst_Humans_2002}: in order to integrate contextual information from conflicting sources (e.g.
-probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its reliability (e.g.
-the inverse of its variance).
+The model also predicts an effect reminiscent of multisensory integration \citep{Ernst_Humans_2002}: in order to integrate contextual information from conflicting sources (e.g.
+probabilistic visual cues and noisy endpoint feedback), the weight placed on a source increases with its reliability.
 Such integration would manifest itself in experiments in which observations are noisy, as in the experiments by \cite{Kording_Bayesian_2004}, in which the position of the finger was obscured and instead participants are shown a blurry cursor which was some times shifted from its real position.
 If the added observation noise gives evidence for a particular context (the true underlying context or another one) and a visual cue gave partial information for another context, the participants' behavior would be more consistent with the most reliable source of contextual information.
-
-\subsubsection{Adaptation history affects lag in E blocks}
-As \cite{Vaswani_Decay_2013} discussed, participants had different delays before their behavior is modified after entering the E block.
-Our model presents the same run-dependent delay, as can be seen in \fref{fig:vaswani-2013}D.
-In addition, the model predicts that this delay depends on the history of adaptation, as evidenced by the red group having very small delays (one or two trials) compared to the other groups.
-Importantly, \cite{Vaswani_Decay_2013} found no significant difference between groups in their fitted parameter for delay, contrary to our results.
-This discrepancy could be due to the models used by \cite{Vaswani_Decay_2013} to determine the lag, but could also point to an explicit component of inertia that is missing from our model.
-Future experimental and modelling work is needed to understand this.
 

### 2021-11-10 adaptation/article/article.tex

@@
-\documentclass[a4paper,doc,floatsintext,natbib]{article}
+\documentclass[a4paper,doc,floatsintext,natbib]{apa6}%article}
 % \documentclass{article}
@@
 \usepackage{authblk}
-\usepackage[nomarkers,figuresonly]{endfloat}
+% \usepackage[nomarkers,figuresonly]{endfloat}
 
@@
 \title{The effects of ongoing context inference on motor learning}
-% \shorttitle{Context-inference-dependent motor learning}
+\shorttitle{Context-inference-dependent motor learning}
 \author[1,2]{Cuevas Rivera, Darío}
@@
 \affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
-% \affiliation{~}
+\affiliation{~}
 
@@
 
-Central to the model is the concept of context, defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the changing environment.
+Central to the model is the concept of context, defined in terms of the task to be performed, the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the changing environment.
 Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to infer the context.
@@
 
+We now briefly describe the three components of the model separately.
+For clarity, we start with the second component (i.e.
+motor adaptation), in order to introduce the terminology that we use throughout this work.
+For an overview, we show in \fref{fig:model} how these three components affect each other.
+
 \begin{figure}
@@
 \includegraphics[]{./figures/figure_1.png}
-\caption{Description of the model.
-(A) Definition of a context in terms of its components.
-(B) Update equations for context inference, motor adaptation and action selection.
+\caption{Schematic representation of the model.
 $\zeta$ represents the identity of the context as a categorical variable, $\vartheta$ are observations, $c_t$ are motor commands at trial $t$, $\phi$ are the parameters of the forward model and $s$ are the hidden states of the environment (e.g.
 the position of the hand).
-(C) Schematic representation of the model.
 The arrows show flow of information, i.e.
 the dependencies in the update equations, as well as the flow of a single trial.
-The elements (e.g.
-$q(\zeta)$) are those shown in B.
 Trial $t$ starts at the top when an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$).
+The update equations can be found in the Methods section.
 After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process is repeated.
@@
 \end{figure}
-
-We now briefly describe the three components of the model separately.
-For clarity, we start with the second component (i.e.
-motor adaptation), in order to introduce the terminology that we use throughout this work.
-For an overview, we show in \fref{fig:model}B and C how these three components affect each other.
 
@@
 These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation.
-With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}B and C, is calculated after each trial.
+With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}, is calculated after each trial.
 

### 2021-11-11 adaptation/article/article.tex
Update on Overleaf.

@@
 \author[1,2]{Kiebel, Stefan J.}
-\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
+\affil[1]{Chair of Neuroimaging, Faculty of Psychology, Technische Universität Dresden, 01062 Dresden, Germany.}
 \affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
@@
 Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-In this work, we use a model to explain the behavioral phenomena effected by imperfect contextual information.
-Specifically, we present a hierarchical computational model for motor adaptation based on exact Bayesian inference.
+In this work, we introduce a computational model to explain the behavioral phenomena effected by imperfect contextual information.
+Specifically, we present a hierarchical model for motor adaptation based on exact Bayesian inference.
 This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection.
@@
 For an overview, we show in \fref{fig:model} how these three components affect each other.
+The full mathematical description of the model can be found in the Methods.
 
@@
 Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors.
-As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.
+As with motor adaptation, the context is inferred as a probability distribution $q(\zeta)$ with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.
 
@@
 We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task.
-With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.
+With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had assumed the existence of multiple mechanisms.
 
-We selected representative experimental studies that show savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
+We selected representative experimental studies that show the well-established effects of savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
 Using our model, we showed how each of these effects can be explained by the specifics of context inference, which integrates all the available information (e.g.

### 2021-11-11 adaptation/article/article.tex
Finalized. Submitted to Nat.Comm.

@@
 % Remember to start reftex-mode
-\bibliographystyle{ieeetr}
 
@@
 \shorttitle{Context-inference-dependent motor learning}
-\author[1,2]{Cuevas Rivera, Darío}
-\author[1,2]{Kiebel, Stefan J.}
-\affil[1]{Chair of Neuroimaging, Faculty of Psychology, Technische Universität Dresden, 01062 Dresden, Germany.}
-\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
+% \author[1,2]{Cuevas Rivera, Darío}
+% \author[1,2]{Kiebel, Stefan J.}
+% \affil[1]{Chair of Neuroimaging, Faculty of Psychology, Technische Universität Dresden, 01062 Dresden, Germany.}
+% \affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
+\author[1]{Author list}
+\affil[1]{Affiliation list}
 \affiliation{~}
@@
 
-\section{Acknowledgment}
-Funded by the German Research Foundation (DFG, Deutsche Forschungsgemeinschaft) as part of Germany’s Excellence Strategy – EXC 2050/1 – Project ID 390696704 – Cluster of Excellence “Centre for Tactile Internet with Human-in-the-Loop” (CeTI) of Technische Universität Dresden.
+% \section{Acknowledgment}
+% Funded by the German Research Foundation (DFG, Deutsche Forschungsgemeinschaft) as part of Germany’s Excellence Strategy – EXC 2050/1 – Project ID 390696704 – Cluster of Excellence “Centre for Tactile Internet with Human-in-the-Loop” (CeTI) of Technische Universität Dresden.
 
-\bibliography{MyLibrary}
+\bibliography{../../MyLibrary}
 

### 2022-01-12 adaptation/article/article.tex
Added Heald 2021. Modified for plos

--- 2021-11-11 adaptation/article/article.tex
@@
 
-\title{The effects of ongoing context inference on motor learning}
-\shorttitle{Context-inference-dependent motor learning}
+\title{The effects of uncertain context inference on motor adaptation}
+\shorttitle{Context-inference-dependent motor adaptation}
 \author[1,2]{Cuevas Rivera, Darío}
@@
 \affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
+% \author[1]{Author list}
+% \affil[1]{Affiliation list}
 \affiliation{~}
@@
 Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-In this work, we introduce a computational model to explain the behavioral phenomena effected by imperfect contextual information.
+In this work, we introduce a computational model to explain the behavioral phenomena effected by uncertain contextual information.
 Specifically, we present a hierarchical model for motor adaptation based on exact Bayesian inference.
 This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection.
-We show how the proposed model provides a unifying explanation for four different experimentally established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, in addition to (iii) the well-studied savings, de-adaptation and spontaneous recovery.
+We show how the proposed model provides a unifying explanation for four different experimentally-established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, (iii) the effects of training history on behavior in new contexts, in addition to (iv) the well-studied savings, de-adaptation and spontaneous recovery.
 
@@
 
-In this work, we show that these unexplained phenomena can be explained in terms context inference, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
+In this work, we show that these unexplained phenomena can be explained in terms context inference, an idea that has recently been gaining traction in motor adaptation \citep{Heald_Contextual_2021}, as well as in other fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
 We build on models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019} by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
@@
 To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.
+
+This work expands on the work by \cite{Heald_Contextual_2021}, who showed that context inference modulates learning rate in cases of unstable contexts.
+To do this, we focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
 
@@
 
-Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
+Together with the results by \cite{Heald_Contextual_2021}, our results show that many behavioral phenomena observed across different experiments can be explained by the uncertainty in context inference and its effects on learning and action selection.
+
+Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}, as well as the recently-introduced COIN model \citep{Heald_Contextual_2021}.
 An exception is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, as these models do not incorporate uncertainty in the estimate of the parameters of the forward models.
@@
 
-The same rationale explains the results by \cite{Davidson_Scaling_2004}: motor learning, which in our model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
+The same rationale explains the results by \cite{Herzfeld_memory_2014}: motor learning, which in our model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
 This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
+Additionally, as was shown by \cite{Heald_Contextual_2021}, context inference explains the modulation of learning rate by the volatility of the environment observed by \cite{Herzfeld_memory_2014}.
 

### 2022-01-18 adaptation/article/article.tex
Update on Overleaf.

@@
 We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task.
-With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had assumed the existence of multiple mechanisms.
+With this, we showed how this seemingly-simple generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had assumed the existence of multiple mechanisms.
 
@@
 sensory cues, workspace location, reward and endpoint feedback), in some cases throughout many trials.
-
-Together with the results by \cite{Heald_Contextual_2021}, our results show that many behavioral phenomena observed across different experiments can be explained by the uncertainty in context inference and its effects on learning and action selection.
 
@@
 If the added observation noise gives evidence for a particular context (the true underlying context or another one) and a visual cue gave partial information for another context, the participants' behavior would be more consistent with the most reliable source of contextual information.
+
+\subsection{Conclusions}
+The results we presented in this work show that many behavioral phenomena observed across different experiments can be explained by the uncertainty in context inference and its effects on learning and action selection.
+Together with the results by \cite{Heald_Contextual_2021}, these results suggest new venues of investigation for future works in motor adaptation and context-dependent behavior.
 

### 2022-03-31 adaptation/article/article.tex
Finished COIN-based model description

@@
 % \usepackage[nomarkers,figuresonly]{endfloat}
+\usepackage{multirow}
+\usepackage{tabularx}
 
@@
 \section{Methods}
-\subsection{Model description}
-\label{subsection:model-description}
-We first describe the general principle behind the model, followed, in the next
-subsection, by the full mathematical description of each one of its components.
+\subsection{The COIN model}
+In this work, we used the recently-introduced COIN model \citep{Heald_Contextual_2021}, adapted to the experiments that we covered in our simulations.
+In this section, we give a brief introduction to the COIN model and, in the subsequent subsection, describe how we adapted the model to the experimental tasks.
+For a full description of the model, refer to \cite{Heald_Contextual_2021}.
 
-Before proceeding with the model description, let us introduce nomenclature:
-\begin{itemize}
-\item $t$: Discrete time at the moment an observation is made, before inference is done. 
-Here taken as the trial number.
-\item $\vartheta_t$: Observation. Note that this is assumed to be a noisy
-observation of the generalized state (see below)
-\item $s_t$: Hidden state, e.g. the position of the hand in reaching
-experiments. It does not depend on the context.
-\item $c_t$: Motor command (action) taken after observing $\vartheta_t$ and inferring the context.
-\item $\zeta_t$: Categorical variable representing the context inferred after observing $\vartheta_t$.
-\item $\omega_t$: Contextual information. This term includes, e.g. visual cues,
-proprioceptive feedback, performance feedback.
-\item $z_t$: Generalized state, $z_t = \{s_t, \zeta_t\}$. Observed noisily
-as $\vartheta_t$.
-\item $m_i$: Forward model for context $\zeta_i$.
-\item $\phi_i$: Parametrization of the forward model $m_i$. In adaptation experiments, this
-can be interpreted as the inferred size of the adaptation.
-\end{itemize}
+\subsubsection{Generative model}
+At each trial $t$, the agent infers both the context and the context-dependent adaptation (e.g.
+the parameters of the force field in mechanical-arm experiments).
+The context is represented by a latent, categorical variable $\zeta_t$:
+\begin{equation}
+p(\zeta_t | \zeta_{t-1}, \pi_{\zeta_{t-1}}) = \text{Discrete}\left(\pi_{\zeta_{t-1}}\right)
+\end{equation}
+where $\pi_{\zeta_{t-1}}$ is the transition probability vector from context $\zeta_{t-1}$ to all other contexts.
+MARKOV CHAIN.
+The contextual cues (when present in an experiment) are assumed to be drawn depending on the context following:
+\begin{equation}
+p(q_t | c_t, \Phi) = \text{Discrete}(\Phi_{\zeta_t})
+\end{equation}
+where $Q_{\zeta_t}$ is the probability vector with which the contextual cue $q_t$ is shown to the agent in context $\zeta_t$.
+As pointed out by \cite{Heald_Contextual_2021}, both $\Phi$ and $\pi$ are in principle infinite, but a task-relevant finite set can be used instead.
 
-Context inference can work through context-specific (possibly unreliable) sensory cues,
-or through the dynamics of the system (e.g. expected outputs of motor
-commands), using prediction error as the force behind inference. In this case,
-the generative model can be written as follows:
-\[
-p(z_t, \vartheta_t | z_{t-1}, c_{t-1}) = p(\vartheta_t | z_t)p(z_t|z_{t-1}, c_{t-1})
-\]
-and inferring the current (hidden) generalized state is done with Bayes' theorem:
-\[
-  q(z_t | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | z_t)p(z_t|\vartheta_{t-1}, c_{t-1})
-\]
+The context-dependent adaptation is represented by the latent variable $x_{\zeta,t}$ and assumed to arise from an autoregressive process AR(1), with a stationary Gaussian distribution of unknown mean and variance:
+\begin{equation}
+p(x_{\zeta,t}) = \mathcal{N}(\mu_{\zeta,x}, \sigma_{\zeta,x})
+\end{equation}
+where $\sigma$ is assumed to be shared across all contexts.
+Note that $\mu$ and $\sigma$ are parametrized by the parameters of the AR(1) process, namely $\mu_{\zeta,x} = d_\zeta / (1 - a_\zeta)$ and $\sigma_{\zeta, x} = \sigma_q / 1 - a_\zeta^2$, where $\sigma_q$ is a free parameter of the model.
 
-Additionally, after having inferred the current state, motor adaptation is given by
+Observations take the form of state feedback (e.g. the position of the cursor on the screen in visuomotor rotation tasks), given by:
 \begin{equation}
-q(\phi | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)q(z_{t-1}|\ldots)p(\phi) \label{eqn:motor-adaptation}
+y_t = x_{\zeta_t, t} + \nu_t
 \end{equation}
-where $\phi$ are the parameters of the internal model for the dynamics, also called the forward model,
-i.e. the function $f: (z_t, c_t) \rightarrow z_{t+1}$, which determines the beliefs of
-the agent regarding how the system evolves after having taken action $c_t$. $q(z_{t-1}|\ldots)$ is the
-posterior probability calculated in the previous trial.
+where $\nu_t$ is a zero-mean Gaussian noise term with unknown standard deviation $\sigma_r$, which is a free parameter of the model.
 
-Having inferred the generalized state (including context), a motor command is
-sampled from a distribution:
+Action selection (i.e. motor output $u_t$) is done via the weighted mean of $x_{j,t}$:
 \begin{equation}
-p(c_t | s_t, \zeta_t) = \displaystyle\sum_{i}q(\zeta_i)p(c_t | s_t, \zeta_i)
+u_t  = \displaystyle\sum_{j}p(\zeta_{j,t} | ...) x_{j,t}
 \end{equation}
-where $i$ runs over all contexts and $p(a_t | ... m)$ is the probability distribution over actions
-prescribed by the forward model $m_i$ for context $\zeta_i$.
+To include motor noise (independent from estimation uncertainty), as well as carry over the uncertainty over $x_{j,t}$, we instead sample motor commands from a Gaussian centered on this mean, with a standard deviation $\sigma_u$, which is a free parameter of the model.
 
-In the following sections, we present a specific mathematical implementation
-of the principles outlined above. This is the implementation used throughout the
-simulations in the Results section.
+\subsubsection{Using the model}
+The free parameters of this model can be fit to participants' data, as was done by \cite{Heald_Contextual_2021}.
+In this work, we instead chose values for these parameters to show that the model is capable of explaining the experimental phenomena in the Results section.
+Additionally, by fixing these parameters the agent is able to perform exact Bayesian inference at each trial using conjugate priors, replacing the MCMC approach used by \cite{Heald_Contextual_2021} due to the mathematical intractability of the full formulation.
+This, however, does not change the model and was done purely for computational efficiency.
+In this section, we describe how we fixed parameters and the procedure for Bayesian inference.
 
-\subsubsection{Mathematical implementation}
-The generalized state is given by:
+As explained above, context is assumed to be a discrete variable which evolves as a Markov process.
+The transition matrices $\pi$ were generated via a Dirichlet process, with parameters that can be inferred from participant data ($\alpha$ and $\kappa$ in \cite{Heald_Contextual_2021}).
+For a fixed value of these parameters, the transition matrices also become fixed.
+In particular, setting the parameters to CALCULATE THIS, the transition matrices are such that contexts are assumed to be sticky, i.e.
+a transition from a context to itself is more likely than to any other context.
+In our simulations, we set the probability of self-transitioning (denoted $p_\zeta$) depending on the experiments (see below), to numbers that match the information given to the participants of each study.
+
+Contextual cues are assumed by the agent to be sampled from a distribution that depends on the current context.
+This is done through a set of cue probability vectors that are generated via a parametric distribution, whose parameters are fitted to participants' data.
+Because the experiments we simulate do not include probabilistic or deceiving cues, contextual cues, when present, unequivocally reflect the current context, i.e.
+$p(q_t = i | c_t = j) = d_{ij}$, where $\delta_{ij}$ is the Kronecker delta, equaling zero unless $i = j$.
+
+Finally, we chose an equivalent parametrization for the priors over the hidden variables $x_{j,t}$, using the mean $\mu_x$ and standard deviation $\sigma_x$ directly instead of the AR(1) $a$ and $d$ parameters used by \cite{Heald_Contextual_2021}.
+Note that we dropped the $j$ dependency for clarity.
+This parametrization, in conjunction with the fixed parameters outlined above,  allows us to set priors over $\mu_x$ and $\sigma_x$ that enable exact inference over the latent variables.
+The priors are given by a Normal-Gamma distribution, typically used for priors of the mean and standard deviation of a Gaussian distribution [CITATION NEEDED]:
 \begin{equation}
-z_t = \{s_t, \zeta_t\}
+\mu_x, \sigma_x \sim \mathcal{NG}(\mu_0, \nu_0, \alpha_0, \beta_0)
 \end{equation}
-where $s_t$ is, e.g. in a reaching task, the current position of the hand in
-Cartesian coordinates, where the origin is the starting point. We assume that
-motor commands are issued every $\Delta t$ for simplicity. At the beginning of
-each time interval, the context is inferred combining the contextual
-information and the prediction error of the outcome of the previous motor
-command. This yields:
-\begin{align}
-  \begin{split}
-  q(\zeta_t) &= q(\zeta_t | \omega_t)q(\zeta_t | s_t, s_{t-1}, c_{t-1}) \\ \label{eqn:estimated-context}
-  q(\zeta_t | \omega_t) &\propto p(\omega_t | \zeta_t)p(\zeta_t) \\
-  q(\zeta_t | s_t, s_{t-1}, c_{t-1}) &\propto p(s_t | \zeta_t, s_{t-1}, c_{t-1})p(\zeta_t)
-  \end{split}
-\end{align}
-The terms $p(\zeta_t)$ refer to the prior probability of the context at the
-beginning of the trial: at the first trial, this refers to prior beliefs over
-which contexts are more common/likely, which we call $p_0(\zeta_t)$. At each
-subsequent trial, it incorporates the belief that has so far been accumulated,
-given previous trials. Note that certain events, such as the start of a new
-block of trials (see main text) can return an agent to a state in which
-$p_0(\zeta_t)$ becomes relevant.
+with free parameters $\mu_0$, $\nu_0$, $\alpha_0$ and $\beta_0$, which we fixed for each experiment separately.
+Because $x$ is context-specific, so are these parameters.
+This formulation comes with four free parameters (i.e.
+the hyper-priors $\mu_0, \nu_0, \alpha_0, \beta_0$), in accordance with the original formulation (note that \cite{Heald_Contextual_2021} fixed the mean of the priors for $b$ to zero).
+While the two formulations are not mathematically identical, the effects of the hyper-priors for both are the same; we discuss these effects in the next section.
+Additionally, we found numerically that these two formulations could provide almost identical samples for given parametrizations (see SUPPLEMENTARY MATERIALS).
+\todo{Do this}
 
-We do not directly model hand movements or eye position. Instead, we model
-systems in the motor error space: the position $s_t = 0$ represents a trial in
-which no motor error occurred. We make use of the fact that in most
-experiments, motor adaptation needs to happen in one of two directions
-(e.g. clockwise vs. counter-clockwise) to keep the error space
-one-dimensional. With this, $s_t > 0$ represents motor error in one direction
-(e.g. clockwise) and $s_t < 0$, error in the other direction. We made this
-choice because in this work we modeled different experiments with
-different state spaces, but with the commonality of (at most)
-two-dimensional error spaces. In order to model a specific experiment in its
-own physical space (e.g. hand positions in a reaching task), only the dynamical
-system  $p(s_t | \zeta_t, s_{t-1}, c_{t-1})$ would need to be adapted.
-
-We assume that the decision-making agent has a baseline internal forward model $m_O$ that
-maps motor commands to outcomes:
-\begin{equation}
-p(s_{t+1} | s_t, c_t, m=m_O) = f_O(s_t, c_t)
-\end{equation}
-where $c_t$ are motor commands (actions). We further assume that participants
-have learned $f_O$ throughout their lives and use it to decide which motor
-command to issue under normal circumstances (i.e. no adaptation). When a
-perturbation is introduced, to which the participant must adapt, we assume they
-learn a new forward model $f_m^{\phi}$, where $\phi$ represents the parameters
-of the function that participants must learn. For example, in a saccade task,
-$\phi$ represents the new gain. Note that $\phi$ depends on the context and
-participants may need to use and update multiple forward adapted models
-$f_m^{\phi}$ throughout one experiment.
-
-Given a forward model $f_m$ (or $f_m^\phi$), action selection is done by
-building a probability distribution over possible actions $c_t$ based on how
-likely they are to produce zero motor error. For example, in saccade
-experiments, motor commands are issued based on how likely they are to move the
-fovea to the target. These motor commands are issued to counteract intrinsic
-sources of error, such as muscle variability and exhaustion, and extrinsic
-sources of error, such as the experimental manipulations on saccade gain. Thus,
-the expected observed error can be written as:
-\begin{equation}
-p(\epsilon | c_t) \propto p(\epsilon_{\text{intrinsic}} | c_t) + p(\epsilon_{\text{extrinsic}} | c_t)
-\end{equation}
-and a motor command can be chosen to maximize the chance of observing zero error:
-\begin{equation}
-p(c_t) = argmin_{c_t} p(\epsilon | c_t)
-\end{equation}
-
-For simplicity, we assume that motor errors are Gaussian, such that:
-\begin{equation}
-p(\epsilon | c_t) = N(\mu_{\text{extrinsic}}, \sigma_\epsilon)
-\end{equation}
-where $\sigma_\epsilon$ incorporates both intrinsic and extrinsic sources of
-error and is a free parameter of the model. The mean of the error is assumed to
-have no intrinsic component.
-
-With this, a forward model will produce a motor command with outcomes centered
-at $-\mu_{\text{extrinsic}}$, which counteracts the externally-induced errors.
-
-To complete the action-selection picture, we return to context inference. As
-discussed above, action selection is made via a weighted sum of the motor commands
-produced by each forward model:
-\begin{equation}
-p(c_t | ...) \propto \displaystyle \sum_{i} q(\zeta_i)p(c_t | \zeta_i, ...) \label{eqn:dist-comm}
-\end{equation}
-From this distribution, an action is sampled at each trial.
-
-To update the agent's estimates on the parameters of the forward models, we
-make use of Bayes' theorem with conjugate priors. This approach has the
-advantage of not only greatly simplifying computational calculations, but also
-presenting update equations with intuitive appeal.
-
-In the following equations, we drop the dependence on the context and
-observations for notational simplicity:
-\begin{equation}
-q(\phi | s_{1:t}) \propto p(s_{1:t} | \phi)p(\phi)
-\end{equation}
-where $s_{1:t}$ are the observed states up until time $t$. The likelihood is a
-Gaussian:
-\begin{align}
-  p(s_t | \phi) &= N(\mu_s, \sigma_s)  \\
-  \mu_s &= \phi + s_t \\
-  \sigma_s &= \xi_0 1
-\end{align}
-This likelihood is the probability of observing $s$ given the previous estimate
-of the parameters of the forward model, $\phi_{t-1}$, and given the dynamics of
-the system (see below).
-
-To update $\phi$, we assumed that for each context, the agent's belief over
-the magnitude of the adaptation is given by a normal distribution:
-\begin{equation}
-p(\phi) = N(\mu_\phi, \sigma_\phi) \label{eqn:data-dist}
-\end{equation}
-where $\mu_\phi$ and $\sigma_\phi$ are parameters to be estimated at each
-trial. Before the experiment begins, the agent will have a prior distribution
-over these parameters. A standard Bayesian approach is to choose NormalGamma
-priors:
-\begin{equation}
-p(\mu_\phi, \sigma_\phi) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
-\end{equation}
-where $\mu_{\phi}^0, \nu_{\phi}^0, \alpha_{\phi}^0$ and $\beta_{\phi}^0$ are free
-hyperparameters of the model.
-
-The update equations for the magnitude parameters are given by:
-\begin{equation}
-q(\mu_\phi, \sigma_\phi | s_t, s_{t-1}, c_{t-1}) \propto p(s_t |
-s_{t-1}, c_{t-1}, \zeta_t)p(\mu_\phi, \sigma_\phi) \label{eqn:context-from-x}
-\end{equation}
-where $p(s_t | s_{t-1}, c_{t-1}, \zeta_t)$ is a Gaussian distribution centered
-around $\mu_c$ with standard deviation $\sigma_c$, which is a free parameter of
-the model. It is assumed that the standard deviation over outcomes ($\sigma_c$)
-is related to that of the parameter $\mu_\phi$ as follows:
-\begin{equation}
-\sigma_\phi = \sigma_c / \nu_\phi
-\end{equation}
-which makes $\nu_\phi$ a parameter to estimate. Each one of the forward models can
-be updated at every trial, depending on context inference (see main text). The
-updated parameters for any one forward model $i$ are given by:
+Because the likelihood function $p(y_t | x_t, ...)$ is Gaussian, this choice of priors allows us to calculate the update equations as follows:
 \begin{align}
@@
 \end{align}
-where $q(\zeta_i)$ is the posterior probability of context $\zeta_i$ from
-\eref{eqn:estimated-context}. Note that when $q(\zeta_i) = 0$, the model $i$ is not
-updated.
+Note that the effect of the evidence (i.e.
+observations) on the inference over the context-dependent hidden states is gated by the probability of each context $p(\zeta_i)$, as in \cite[][supplementary materials]{Heald_Contextual_2021}.
+
+\subsubsection{Model parameters}
+\tref{tab:parameters} lists all the parameter values that we used during our simulations.
+The parameters are divided into two categories: (1) task parameters, which encode the way we simulated the experimental design; (2) agent parameters, which correspond to the free parameters listed in the previous section.
+The variable names for the model parameters are given in the ``Var'' column, corresponding to the variables in the previous section.
+The values are divided into experiments and, within experiments, into the different groups or conditions that we simulated.
+\begin{table}[h!]
+\resizebox{\textwidth}{!}{%
+  \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
+  \cline{1-11}
+  & & & Kim (2015) & \multicolumn{2}{c|}{Oh (2009)} & \multicolumn{2}{c|}{Davidson (2004)} & \multicolumn{3}{c|}{Vaswani (2013)} \\
+  \cline{2-11}
+  & Var & Description & & Exp. 1 & Exp. 2 & Grp. 3A & Grp. -A & Grp. 1 & Grp. 2 & Grp. 3 \\
+  \cline{1-11}\cline{1-11}
+  \multirow{4}{*}{Task pars.} & & Contextual cues & Yes & \multicolumn{7}{c|}{No} \\
+  \cline{2-11}
+  & $x_j^*$ & Adaptation sizes & 0, 40, -40 & 0, 20 & 0, 10 & 0, 4, -4 & 0, 4, 12 & 1 & 0, 1 & -0.5, 1 \\
+  \cline{2-11}
+  & & Adaptation noise & 0.01 & \multicolumn{2}{c|}{1} & \multicolumn{2}{c|}{0.5} & \multicolumn{3}{c|}{0.01} \\
+  \cline{2-11}
+  & $\sigma_r^*$ & Obs. noise & 3 & \multicolumn{2}{c|}{1.5} & \multicolumn{2}{c|}{0.1} & \multicolumn{3}{c|}{0.01} \\
+  \cline{1-11}
+  \cline{1-11}
+  \multirow{7}{*}{Agent pars.} & $p_\zeta$& Context self-transition & 0.9 & \multicolumn{2}{c|}{0.98} &  \multicolumn{2}{c|}{0.98} & \multicolumn{2}{c|}{0.9} & 0.8 \\
+  \cline{2-11}
+  & $\mu_0$ & \multirow{4}{*}{Hyper priors} & 0, -1, 1 & \multicolumn{2}{c|}{0, 0} & 0, 4, -4 & 0, 4, 12 & \multicolumn{2}{c|}{0, 1} & 0, 1, -0.5 \\
+  \cline{2-2} \cline{4-11}
+  & $\nu_0$ & & 1e4, 10, 10 & \multicolumn{2}{c|}{0, 0} & \multicolumn{2}{c|}{1e4, 0.5, 0.5} & \multicolumn{2}{c|}{1e4, 1} & 1e4, 1, 1 \\
+  \cline{2-2} \cline{4-11}
+  & $\alpha_0$ & & 25e4, 0.25, 0.25 & \multicolumn{2}{c|}{22e4, 2.2} & \multicolumn{2}{c|}{33e4, 33, 33} & \multicolumn{2}{c|}{5e5, 5} & 15e5, 5, 5 \\
+  \cline{2-2} \cline{4-11}
+  & $\beta_0$ & & 1e5, 2, 2 & \multicolumn{2}{c|}{1e5, 20} & \multicolumn{2}{c|}{1e5, 200, 200} & \multicolumn{2}{c|}{1e5, 2} & 1e5, 2, 2 \\
+  \cline{2-11}
+  & $\sigma_u$ & Motor noise & 0.001 & \multicolumn{4}{c|}{2} & \multicolumn{3}{c|}{0.17} \\
+  \cline{2-11}
+  & $\sigma_r$& Obs. noise & 3 & \multicolumn{2}{c|}{3.5} & \multicolumn{2}{c|}{2.5} & \multicolumn{3}{c|}{0.1} \\
+  \cline{1-11}
+  \end{tabular}}
+\caption{Model and simulation parameters.}
+\label{tab:parameters}
+\end{table}
+
+We estimated the task parameters from the information provided in their respective publications; when direct information was not provided, we estimated it from the reported results; these estimations are not exact, but function as a proof of context.
+Agent parameter values are held constant for the different conditions or groups for each experiment, except those parameters that are expected to vary across conditions.
+
+The context-dependent hyper-priors $\{\mu_0, \nu_0, \alpha_0, \beta_0\}$ effect a learning rate and, as such, were set such that the learning rate for the baseline context is nearly zero, following the assumption that participants will not change the way they perform these movements outside of the lab.
+The values for contexts outside of baseline were set identically within each experiment.
+
+The agent has direct knowledge on the adaptation and observation noises of the task; we do not expect this to be the case in an experiment (instead participants would estimate it during their early trials).
+\todo{check simulations to see if I can make them identical; if not, change this}
+
+The star notation (e.g. $x_j^*$) denotes the real value used in the simulation of the task, which may be different from that assumed by the agent.
 

### 2022-04-12 adaptation/article/article.tex
New figures ready. Started rewriting paper

@@
 We show how the proposed model provides a unifying explanation for four different experimentally-established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, (iii) the effects of training history on behavior in new contexts, in addition to (iv) the well-studied savings, de-adaptation and spontaneous recovery.
+
+Humans have been shown to adapt their movements when sudden or gradual changes to the dynamics of the environment are introduced, a phenomenon called motor adaptation.
+If the change is reverted, the adaptation is also quickly reverted.
+Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly.
+Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
+Recently, the COIN computational model for motor adaptation and context inference was introduced, which contains explicit components of context inference and Bayesian motor adaptation.
+This model was used to show the effects of context inference on learning rates across different experiments.
+We expanded on that work by using the COIN model to show that the effects of context inference on motor adaptation and control go further than previously shown.
+Here, we used the COIN model to simulate classical motor adaptation experiments from previous works and showed that context inference, and how it is affected by the presence and reliability of feedback, effect a host of behavioral phenomena that had so far required multiple hypothesized mechanisms, lacking a unified explanation.
+Concretely, we show that the reliability of direct contextual information, as well as noisy sensory feedback, typical of many experiments, effect measurable changes in switching-task behavior, as well as in action selection, that stem directly from imprecise context inference.
 
@@
 a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model.
-These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation.
+These updates are done according to \eref{eqn:update-full}, using the difference between predictions made with the current parameters of the forward model, and the current observation.
 With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}, is calculated after each trial.

### 2022-04-12 adaptation/code/figures.py

@@ -6,12 +6,13 @@ parameter --fignum-- for each function defined here determines which figure it
 is in the paper. Those that do not have one are not paper figures, but
 auxiliary functions."""
 
-
+import ipdb
 import numpy as np
 from matplotlib import pyplot as plt
 import pandas as pd
 from matplotlib import gridspec as gs
 import seaborn as sns
+import matplotlib as mpl
 
 import simulations as sims
 import model
@@ -20,45 +21,209 @@ import pars
 
 FIGURE_FOLDER = '../article/figures/'
 
+mpl.rcParams['font.size'] = 8
+mpl.rcParams['figure.titlesize'] = 'small'
 
-# Note: Figure 1 is made by hand in inkscape
+
+def model_showoff(fignum=1, show=True):
+    """Plots a bunch of arbitrary simulations that show how the parameters of
+    the model work and affect inference. Also leaves an empty space on top
+    to insert the diagram of the generative model by hand.
+    """
+    runs = 50
+    colors = ['black', 'tab:orange', 'tab:blue']
+    cycler = plt.cycler('color', colors)
+    tasks, agents_pars = sims.model_showoff(plot=False)
+    numbers = tasks.shape
+    figsize = (6, 8)
+    fig = plt.figure(fignum, clear=True, figsize=figsize)
+    height_ratios = np.ones(numbers[0] + 2)
+    height_ratios[0] = 3
+    height_ratios[1] = 0.3
+    gsbig = gs.GridSpec(numbers[0] + 2, numbers[1], figure=fig,
+                        height_ratios=height_ratios, hspace=0.1, wspace=0.1)
+    axis_diagram = fig.add_subplot(gsbig[0, :])
+    ctx_switch = np.diff(tasks.reshape(-1)[0]['context_seq']).nonzero()[0][0] + 1
+    num_trials = len(tasks.reshape(-1)[0]['context_seq'])
+    axes = np.empty((*numbers, 2), dtype=object)
+    small_hr = [2, 1]  # height ratios for adaptation/p(ctx)
+    for ix_row in np.arange(2, numbers[0] + 2):
+        for ix_col in np.arange(numbers[1]):
+            gssmall = gsbig[ix_row, ix_col].subgridspec(2, 1, hspace=0.05,
+                                                        height_ratios=small_hr)
+            axes[ix_row - 2, ix_col, 0] = fig.add_subplot(gssmall[0])
[…]

### 2022-04-12 adaptation/code/model.py

@@ -76,34 +76,38 @@ class LeftRightAgent(object):
     def __init__(self, obs_sd=None, action_sd=None, cue_noise=None,
                  angles=None, context_noise=None, prediction_noise=None,
                  reset_after_change=None, force_sds=None, max_force=None,
-                 sample_action=None):
+                 sample_action=None, prior_over_contexts=None):
         """Initializes the known left and right contexts, as well as
         baseline.
 
         """
-        if obs_sd:
+        if obs_sd is not None:
             self.obs_sd = obs_sd
-        if action_sd:
+        if action_sd is not None:
             self.action_sd = action_sd
-        if cue_noise:
+        if cue_noise is not None:
             self.cue_noise = cue_noise
-        if angles:
+        if angles is not None:
             self.angles = np.array(angles)
             self.num_contexts = len(angles)
         else:
             self.num_contexts = 3
-        if context_noise:
+        if context_noise is not None:
             self.context_noise = context_noise
-        if prediction_noise:
+        if prediction_noise is not None:
             self.prediction_noise = prediction_noise
-        if reset_after_change:
+        if reset_after_change is not None:
             self.reset_after_change = reset_after_change
         if force_sds is not None:
             self.force_sds = force_sds
-        if max_force:
+        if max_force is not None:
             self.max_force = max_force
-        if sample_action:
+        if sample_action is not None:
             self.sample_action = sample_action
+        if prior_over_contexts is None:
+            self.prior_over_contexts = np.ones(self.num_contexts) / self.num_contexts
+        else:
+            self.prior_over_contexts = prior_over_contexts
 
         _, self.magnitudes = self.__init_contexts()
         self.magnitude_history = [self.magnitudes]
@@ -290,12 +294,13 @@ class LeftRightAgent(object):
             cue_li = self.cue_noise * np.ones(self.num_contexts)
             cue_li[cue] = 1 - (self.num_contexts - 1) * self.cue_noise
             log_cue = np.log(cue_li)
-        # if np.argmax(log_cue) != np.argmax(log_li_hand):
-        #     ipdb.set_trace()
+    
+        # Now all together!
         prior = np.exp(self.log_context - self.log_context.max()) + \
             self.context_noise
         log_prior = np.log(prior / prior.sum())
-        full_logli = log_li_hand + log_cue + log_prior
+        full_logli = log_li_hand + log_cue + log_prior +\
+            np.log(self.prior_over_contexts)
[…]

### 2022-04-12 adaptation/code/simulations.py

@@ -373,8 +373,8 @@ def plot_adaptation(pandata, axis=None, fignum=5):
 
 def sim_and_plot(agent, pars_task, return_data=False,
                  force_labels=None, axes=None, fignum=6):
-    """Simulates the agent playing and makes a nice plot with context inference and
-    adaptation and colors everywhere.
+    """Simulates the agent playing and makes a nice plot with context inference
+    and adaptation and colors everywhere.
 
     """
     pandata, pandagent, agent = thh.run(agent, pars=pars_task)
@@ -476,6 +476,110 @@ def baseline_bias(fignum=9):
     axes[0, 3].get_legend().remove()
 
 
+def herzfeld_2014(num_trials=150, plot=True, axes=None, fignum=12):
+    """Simulates something like the experiments from Herzfeld et al. 2014.
+
+    This does two conditions: high uncertainty and low uncertainty, where
+    the uncertainty is over the current context given the previous one. In
+    the high uncertainty condition, there's a 50/50 chance of a context
+    change. In the low uncertainty, there is a 0.1 chance of change.
+
+    """
+    task = task_pars.copy()
+    task['obs_noise'] = 3
+    task['force_noise'] = 0.01 * np.ones(3)
+    task['forces'] = [[0, 0], [-1, 13], [1, 13]]
+    task['breaks'] = np.zeros(num_trials, dtype=int)
+    task['cues'] = np.zeros(num_trials, dtype=int)
+
+    task_high = task.copy()
+    contexts = np.random.choice([1, 2], size=num_trials)
+    task_high['context_seq'] = contexts
+    task_low = task.copy()
+    switches = np.random.choice([0, 1], p=(0.99, 0.01), size=num_trials)
+    switches_ct = np.concatenate([[0], np.nonzero(switches)[0], [num_trials]])
+    switches_delta = np.diff(switches_ct)
+    contexts = [[(idx % 2) + 1, switch]
+                for idx, switch in enumerate(switches_delta)]
+    task_low['context_seq'] = pars.define_contexts(contexts)
+
+    agent_pars_high = {'angles': [0, -1, 1],
+                       'prediction_noise': 0.2,
+                       'cue_noise': 1 / 3,
+                       'context_noise': 0.25,
+                       'force_sds': 0.1 * np.ones(3),
+                       'max_force': 20,
[…]

### 2022-04-26 adaptation/article/article.tex
First draft for post-Heald version ready

@@
 \abstract{Abstract}
-Humans have been shown to adapt their movements when a sudden change to the dynamics of the environment is introduced, a phenomenon called motor adaptation.
-If the change is reverted, the adaptation is also quickly reverted.
-Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly.
-Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-In this work, we introduce a computational model to explain the behavioral phenomena effected by uncertain contextual information.
-Specifically, we present a hierarchical model for motor adaptation based on exact Bayesian inference.
-This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection.
-We show how the proposed model provides a unifying explanation for four different experimentally-established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, (iii) the effects of training history on behavior in new contexts, in addition to (iv) the well-studied savings, de-adaptation and spontaneous recovery.
-
 Humans have been shown to adapt their movements when sudden or gradual changes to the dynamics of the environment are introduced, a phenomenon called motor adaptation.
@@
 
-In this work, we show that these unexplained phenomena can be explained in terms context inference, an idea that has recently been gaining traction in motor adaptation \citep{Heald_Contextual_2021}, as well as in other fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}.
-We build on models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019} by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference.
-In the proposed model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions.
-To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.
+Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced \cite{Heald_Contextual_2021}, called COIN (for context inference).
+\cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, but is informed by it, establishing a loop whereby context inference also informs motor learning.
+With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning un previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.
 
-This work expands on the work by \cite{Heald_Contextual_2021}, who showed that context inference modulates learning rate in cases of unstable contexts.
-To do this, we focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
+In this work, we use the COIN model to show that the process of context inference underlies more behavioral phenomena than previously shown.
+we focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
+More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured.
+We show that through context inference, switching behavior can display three main effects that have been previously attributed to hypothesized ad-hoc mechanisms: (1) The size of an adaptation dictates how quick and reliable switching between tasks is \cite{Oh_Minimizing_2019,Kim_Neural_2015}, which we explain in terms of the effects of perceptual noise on context inference.
+(2) Previously-learned adaptations can interfere with switching behavior \citep{Davidson_Scaling_2004}, which we explain in terms of uncertain context inference.
+(3) Training history (i.e.
+which adaptations have been learned and for how long) affects switching during error-clamp trials \citep{Vaswani_Decay_2013}, which we also attribute to uncertain context inference.
+To do this, we used the COIN model to simulate the experimental setups and the decision-making agents (i.e.
+participants) during those experiments.
 
@@
 \section{Results}
-We present a motor adaptation model in which an existing model of the environment is updated based on error signals produced by the need of adaptation.
-We expanded on existing models \citep{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019} by adding an explicit component of context inference.
-As we will show, context inference guides the selection of the adequate forward model, its updating and the sampling of actions.
-
-We simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference.
+Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference.
 We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.
 
-Before presenting our simulations, we will briefly go through the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.
+Before presenting these results, we make a brief description of the COIN model, leaving a more thorough explanation for the methods section.
+Additionally, we present simulations using the model that show the effects of contextual cues and perceptual noise on context inference, which pave the way for the simulated experiments that we show later on.
 
 \subsection{Modeling context-dependent adaptation}
-The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection.
+We focused on three main components of the COIN model: (1) context inference, (2) motor adaptation and (3) action selection.
 The processes defined by these component occur in this order, and each component informs the ones that follow.
@@
 
-We now briefly describe the three components of the model separately.
-For clarity, we start with the second component (i.e.
-motor adaptation), in order to introduce the terminology that we use throughout this work.
-For an overview, we show in \fref{fig:model} how these three components affect each other.
-The full mathematical description of the model can be found in the Methods.
+The COIN model contains, additionally to these three main components, components for learning new contexts (i.e.
+inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities.
+In this work, we focus on switching behavior between known contexts, as well as in the perceptual aspect of context inference.
+Additionally, we do not fit the COIN model to participant data, instead focusing on showing the way the COIN model explains known experimental phenomena.
+For these reasons, we chose not to use these components of the COIN model, instead fixing both the transition probabilities between contexts and the number of contexts based on the real values during the experiments, which were made available to participants \todo{Check this}.
+Fixing these values, in turn, allowed us to simplify the generative model, internal to the agent, using conjugate priors, with which the agent can do exact Bayesian inference.
+The generative model we used here can be seen in \fref{fig:model}A, including the priors for both context inference and motor adaptation.
+For more details on these choices, see the Methods section.
 
@@
 \includegraphics[]{./figures/figure_1.png}
-\caption{Schematic representation of the model.
-$\zeta$ represents the identity of the context as a categorical variable, $\vartheta$ are observations, $c_t$ are motor commands at trial $t$, $\phi$ are the parameters of the forward model and $s$ are the hidden states of the environment (e.g.
-the position of the hand).
-The arrows show flow of information, i.e.
-the dependencies in the update equations, as well as the flow of a single trial.
-Trial $t$ starts at the top when an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$).
-The update equations can be found in the Methods section.
-After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process is repeated.
-The posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
+\caption{Schematic representation of the model and simulations.
+(A) Generative model.
+Clear circles represent the observables, i.e.
+motor commands $u_t$ and direct observations $y_t$ (e.g.
+cursor position).
+The real context $\zeta_t$ is not directly observable, but influences $y_t$.
+The dark rectangles represent the prior distributions for the inferred adaptation level $x_t$ (Normal-Gamma distribution) and the current context $\zeta_t$ (discrete distribution with known $\pi$; see Methods).
+At every trial, the context is inferred, then motor adaptation is carried out and finally a motor command is issued; the flow of this process is indicated by the gray arrow in the background, while black arrows show the direction of information flow.
+(B) Simulations obtained with the model in (A), using an experimental setup similar to that by \cite{Davidson_Scaling_2004}.
+The plots are organized as follows: each pair of plots represent context inference (bottom; p(ctx)) and response/adaptation (top).
+In the context inference plot, the y-axis represents the posterior probability of each context $p(\zeta_t = j$; the black line represents the baseline context (i.e.
+no adaptation), the green line represents the only adaptation to be learned during the simulated experiment.
+In the top plot (adaptation), the y-axis is the adaptation to be learned (e.g.
+the angle in a visuomotor rotation experiment), using the same color scheme (black for baseline).
+The purple line is the agent's response for the trial.
+The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.
+Note, however, that the top plots are shown for a single run, for clarity (see main text).
+A total of 3x3 experiments were simulated, with low, medium and high levels of both cue uncertainty and observation noise.
+All plots share the same x-axis, which represents the trial number.}
 \label{fig:model}
 \end{figure}
+\todo{Probably remove the top plots, as they are hard to explain and add very little to the paper}
+While these priors are not mathematically equivalent to those used by \cite{Heald_Contextual_2021}, the prior samples they produce can be shown to be similar.
+For more details, see the supplementary materials.
+\todo{Do this}
 
-\subsubsection{Motor adaptation}
-It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}.
-We used this principle and chose to make use of an exact Bayesian learner.
-To adapt to altered dynamics (e.g.
-a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model.
-These updates are done according to \eref{eqn:update-full}, using the difference between predictions made with the current parameters of the forward model, and the current observation.
-With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}, is calculated after each trial.
+\subsection{Contextual cues and feedback}
+The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment.
+To show these effects, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to a visuomotor rotation of the cursor on a screen in relation to the movement of a pointing device.
 
-\subsubsection{Context inference}
-To infer the adequate forward model to deploy, most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the forward models: the model that best predicts the current observation given the past is deemed to be the ``correct'' model, i.e.
-the model that best represents the current context.
-\cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.
+To show the effects that will be the base for the results presented below, we manipulated the presence and reliability of contextual cues, as well as the perceptual noise added to the cursor.
+In \fref{fig:model}B, a 3x3 grid of results can be seen, where each simulation in this grid is a combination of low, medium or high contextual cue reliability (where low reliability is equivalent to presenting no cues), and low, medium or high perceptual noise.
 
-Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors.
-As with motor adaptation, the context is inferred as a probability distribution $q(\zeta)$ with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.
+The first column of \fref{fig:model} shows that, in the presence of reliable contextual cues, context inference is correct, certain and fast to switch.
+However, as contextual cues become less reliable, switching between known contexts becomes slower, as seen in the posterior probabilities over contexts at and after trial 20 in the second and third columns.
+Furthermore, as perceptual noise increases, switching becomes not only slower, but also more uncertain, with individual agents incorrectly missing the switch entirely.
 
-\subsubsection{Action selection}
-Action selection is affected by the current context via the context-dependent forward model that is used.
-In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model.
+Additionally, the motor responses (shown in purple) can be seen to depend not only on the current estimate of the rotation angle (shown as the green line), but also the uncertainty over the current context.
+This is most evident in the high cue uncertainty and observation noise panel, in which the response varies substantially from the estimate of the angle (i.e.
+the purple line deviates from the green line), even when the real context has the highest estimated probability.
 
-In other words, a distribution $p(c_t)$ over all possible motor commands is created following \eref{eqn:dist-comm}, from which the next action (motor command $c_t$) can be sampled and carried out.
+As we show below, these effects are at the heart of the behavioral phenomena observed in the experiments by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004}, and \cite{Vaswani_Decay_2013}, which we directly simulate in this work, alongside others that we discuss in the Discussion section.
 
@@
 
-In this section, we discuss savings and de-adaptation in terms of switching between forward models.
-This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}.
-Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not, namely that savings are not immediate, but a relatively fast process that reflects context inference.
+In this section, we discuss savings and de-adaptation in terms of switching between contexts.
+We show that through context inference and how its affected by contextual cues and obsevation noise, savings are not immediate, but a relatively fast process that reflects context inference.
 
@@
 Consistent with our model, the authors found that switching was immediate and accurate.
+
 In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
-The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e.
-when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.
+The correspondence between simulations and experimental results can be seen in the switches between contexts in the center panel, i.e.
+when the solid gray line (representing the true context) switches between 40, 0 and -40; in both panels (second and third), the thick black line, representing the agent's adaptation, quickly follows these switches.
 
@@
 \caption{Savings and de-adaptation.
-Data from our simulations (left column) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (right column).
+Data from our simulations (first and second columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column).
 Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown.
-In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number.
-The black lines represent the optimal adaptation, i.e.
+In the first column, the simulated context over inference is represented by the posterior probabilities over all available contexts.
+Each color (black, blue, green) represents a different context, with black always representing the baseline (i.e.
+no adaptation).
+Vertical, dashed lines represent switches in the real context.
+In the second and third columns, the black line represents the adaptation (i.e.
+response) displayed by the agent as a function of trial number.
+The thin gray lines represent the optimal adaptation, i.e.
 the size of the true visuomotor rotation during the task.
+In all panels, the x-axis represents trial number, from 0 to 300.
 (A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O.
-Participants must adapt to the rotations during shooting movements, and the blue line can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation.
-When black and blue lines match, the participant perfectly counteracted the visuomotor rotation and reached the target.
+Participants must adapt to the rotations during shooting movements, and the participant's response can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation.
+For clarity, both experimental and simulated results are shown only up to trial 300, of the original 600.
 (B) An experiment by \cite{Oh_Minimizing_2019} similar to A, but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced.
@@
 Blue and black lines are as in A.
-Different blocks with the same adaptation are differentiated by the subscript (e.g.
-$A_1$ and $A_2$ have the same 20 degree visuomotor rotation).
-(C) Same experiment as (B) but with a 10 degree adaptation.} \label{fig:oh-2019}
+(C) Same experiment as (B) but with a 10 degree adaptation.}
+\label{fig:oh-2019}
 \end{figure}
 
-Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks.
-Switching between opposite adaptations was nevertheless immediate in simulations and experimental data.
-After around 500 trials, participants had adapted completely, as seen by the blue line reaching the heights of the black line, and switches continued to be immediate.
+Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by their responses not being on par with the true rotation angle.
+This undershooting of responses (i.e.
+not doing the 40 degree rotation) happens despite participants being able to immediately and with high certainty identify the real context, as shown in the first-column panel.
 These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches.
-To expand on these results, we now turn to performance feedback and its effects on switching behavior.
+To expand on these results, we now turn to feedback and its effects on switching behavior.
 
@@
 The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with our model.
-Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A_1$.
-From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (true rotation).
+Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A$.
+In subsequent $A - O$ transitions, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in their responses (black line in the last two columns of \fref{fig:oh-2019}B) closely following the switches in the black line (true rotation).
 In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up.
 As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C).
-This difference is explained by our model in terms of the size of the adaptation: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output.
+This difference is explained in our simulations in terms of the size of the adaptation in relation to observation noise: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output.
 Because of this, the model requires more evidence (i.e.
@@
 \subsubsection{Uncertainty over contexts affects action selection}
-As with learning, our model selects actions (motor commands) based on context inference.
+As with learning, action selection is affected by context inference.
 If the identity of the current context is known, the forward model for this context is used to select the current action.
@@
 
-In our context-inference account, the asymmetry is caused by the baseline context, which have a non-zero probability $p(\zeta_O | s_t ...)$.
+In the context-inference account, the asymmetry is caused by the existance of the baseline context, which has a non-zero probability $p(\zeta_O | s_t ...)$.
 When a new block of trials starts (e.g.
@@
 We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}.
-We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}.
+We show the results in \fref{fig:davidson-2004}A, alongside the experimental results from \cite{Davidson_Scaling_2004}.
 It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}.
-Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data.
-However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.
 
@@
 \includegraphics{./figures/figure_3.png}
-\caption{Motor error when switching back to a previously learned adaptation.
-Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from 3A (yellow) or -A (blue) to A is shown(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}, running eight simulations per group, each simulation representing one participant.
-On the y axis, the error is shown in arbitrary units, related to centimeters through a monotonically increasing transformation with the same origin.
-The shaded area is the SEM across simulated runs (participants).
-(B) Data adapted from figure 4 by \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
+\caption{Motor error when switching back to a previously-learned adaptation.
+(A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the COIN model are shown.
+In the first column, each panel represents context inference for one group of participants (top: group A; bottom: group 3A), with each line representing the posterior probability of a context (black for the baseline O).
+The second column represents the error made by our simulated agent after returning to the previously-learned context, with blue and yellow representing groups -A and 3A, respectively.
+The last panel represents the same data, from the \cite{Davidson_Scaling_2004} experiment.
+All panels share the x-axis, representing the number of trials elapsed since the switch to the new context.
+All simulations were executed eight times per group, following the number of participants in the experiments; shaded areas represent the standard deviation across these eight simulations.
+(B) Simulated results for an experiment similar to \cite{Davidson_Scaling_2004}, but changing the contexts seen by the two groups from -A to -2A for the first group, and from 3A to 4A for the second group.
+The panels follow the same structure as (A), without the last panel for experimental results.
+(C) Simulations for an experiment in which the baseline context has been removed altogether.}
 \label{fig:davidson-2004}
 \end{figure}
+
+To confirm our hypothesis, we simulated variants of the experiments in which the context-inference account predicts that the difference between groups disappears.
+First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments.
+To do this, one group adapts in a $O-A-(-2A)$ paradigm, while the other group adapts in a $0-A-4A$ paradigm.
+As in the original experiment, the second contexts ($-2A$ for one group, $4A$ for the other) are equally spaced from the first context.
+However, given the larger distance from baseline, the baseline context has a much lower attribution when the switch occurs compared to the original experiment.
+This change makes both simulated groups infer the correct context in much more similar times, making the difference between their errors be much smaller compared to the original experiment.
+Furthermore, we simulated an experiment with an identical structure to that of \cite{Davidson_Scaling_2004}, but eliminated the baseline context from the agent.
+The results can be seen in \fref{fig:davidson-2004}C, where the switches between contexts are made identically by the two groups.
+We discuss possible implementations of such an experiment in the Discussion section, alongside with other predictions made by the model.
+\todo{Promises, promises}
 
@@
 
-In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
+In \fref{fig:vaswani-2013}A, we show data simulated with the COIN, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
 The displayed adaptation is shown during the E trials for the three experimental groups in the experiment.
@@
 Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more quickly than groups 1.1 and 1.4.
+\todo{Danger!
+Danger!
+A keen observer might point out that the model might want to learn a new context, as COIN does.
+What to do?
+What to do?}
 
@@
 \section{Discussion}
-We presented a hierarchical model for motor adaptation based on Bayesian inference, which not only learns to adapt its motor outputs by observing errors, but also infers the context in which these errors were committed and updates the forward model associated with that context.
-
-We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task.
-With this, we showed how this seemingly-simple generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had assumed the existence of multiple mechanisms.
+We expand on previous works \cite{Heald_Contextual_2021,Oh_Minimizing_2019} that introduced the idea that context inference is a process that informs and is informed by motor adaptation by showing that it explains behavioral phenomena that had previously required different specific, ad-hoc mechanisms outside of contextual motor adaptation.
 
@@
 
-Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}, as well as the recently-introduced COIN model \citep{Heald_Contextual_2021}.
+We used the recently-introduced COIN model \cite{Heald_Contextual_2021}, with a different choice of prior distributions that greatly simplified our simulations without sacrificing relevant parts of the original specification.
+While we focused on the Bayesian motor adaptation of the COIN model, we expect the results we presented here to hold using alternative models for motor adaptation, such as other Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
 An exception is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, as these models do not incorporate uncertainty in the estimate of the parameters of the forward models.
@@
 Instead, participants slowly reduce their adaptation, often displaying spontaneous recovery \cite[e.g.][]{Smith_Interacting_2006}.
-Our model provides a principled account of this behavior: the natural variability in participants' behavior lead them to expect errors, which clashes with the observed zero error.
+Context inference provides a principled account of this behavior: the natural variability in participants' behavior lead them to expect errors, which clashes with the observed zero error.
 This prompts participants to re-evaluate their inferred context, which can partially activate a previously-observed context, as we showed in \fref{fig:vaswani-2013}.
@@
 
+Note that the original COIN specification includes a component to learn new contexts.
+However, this component works exclusively by creating new contexts in which the forward models take the same form but have different parameter values.
+It is unclear whether the model can be adapted to create contexts in an online fashion that operate in an essentially different manner, as is the case of error-clamp trials, in which participants' responses do not affect the outcome and motor commands are issued based on criteria not directly related to the goal of the task (e.g.
+energy minimization or comfort maximization).
+
 \subsubsection{Reducing the effect of volatility on learning}
@@
 Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
+\cite{Heald_Contextual_2021} showed that the COIN model explains this change in terms of the Kalman gain and the posterior probability over contexts (see their equation 14) \todo{Check this in more detail}.
 

### 2022-04-26 adaptation/code/figures.py

@@ -23,15 +23,24 @@ FIGURE_FOLDER = '../article/figures/'
 
 mpl.rcParams['font.size'] = 8
 mpl.rcParams['figure.titlesize'] = 'small'
+mpl.rcParams['lines.linewidth'] = 1
 
 
-def model_showoff(fignum=1, show=True):
-    """Plots a bunch of arbitrary simulations that show how the parameters of
+def model_showoff(fignum=1, show=True, do_a=True):
+    """Plots a bunch of arbitrary simulationsthat show how the parameters of
     the model work and affect inference. Also leaves an empty space on top
     to insert the diagram of the generative model by hand.
+
+    Parameters
+    ----------
+    do_a : bool
+    Whether to add the model diagram to the figure. If True, the model diagram
+    is assumed to be in FIGURE_FOLDER/generative.png. If False, the space will
+    be left empty at the top.
+
     """
     runs = 50
-    colors = ['black', 'tab:orange', 'tab:blue']
+    colors = ['black', 'tab:green', 'tab:blue']
     cycler = plt.cycler('color', colors)
     tasks, agents_pars = sims.model_showoff(plot=False)
     numbers = tasks.shape
@@ -111,9 +120,10 @@ def model_showoff(fignum=1, show=True):
     axes[-1, 0, 0].set_ylabel(r'Adaptation')
 
     # Import diagram of the model from png
-    diagram = plt.imread(FIGURE_FOLDER + '/generative.png')
-    axis_diagram.axis('off')
-    axis_diagram.imshow(diagram)
+    if do_a:
+        diagram = plt.imread(FIGURE_FOLDER + '/generative.png')
+        axis_diagram.axis('off')
+        axis_diagram.imshow(diagram)
 
     # subplot labels
     offset_multiplier = 0.03
@@ -122,7 +132,7 @@ def model_showoff(fignum=1, show=True):
     anchor_b = np.array(axes[0, 0, 0].get_position())[[0, 1], [0, 1]] + offset
     fig.text(s='A', x=anchor_a[0], y=anchor_a[1], fontdict={'size': 12})
     fig.text(s='B', x=anchor_b[0], y=anchor_b[1], fontdict={'size': 12})
-    
+
     plt.savefig(FIGURE_FOLDER + 'figure_{}.png'.format(fignum), dpi=600)
     plt.savefig(FIGURE_FOLDER + 'figure_{}.svg'.format(fignum), format='svg')
     if show:
@@ -155,7 +165,7 @@ def plot_adaptation(pandata, axis, colors):
[…]

### 2022-05-11 adaptation/article/article.tex
First COINed draft sent to Stefan. Small things remain to be finished.

@@
-\documentclass[a4paper,doc,floatsintext,natbib]{apa6}%article}
-% \documentclass{article}
-\usepackage[font=small]{caption}
+\documentclass[a4paper,doc,floatsintext,natbib]{apa6}
+%% \documentclass{article}
+\usepackage[font=large]{caption}
 \usepackage{natbib}
@@
 \section{Results}
-Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference.
+Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how this model explains these experimental findings using the dynamics of context inference.
 We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.
@@
 \centering
-\includegraphics[]{./figures/figure_1.png}
+\includegraphics[width=0.7\textwidth]{./figures/figure_1.png}
 \caption{Schematic representation of the model and simulations.
@@
 \todo{Probably remove the top plots, as they are hard to explain and add very little to the paper}
-While these priors are not mathematically equivalent to those used by \cite{Heald_Contextual_2021}, the prior samples they produce can be shown to be similar.
+These priors are not mathematically equivalent to those used by \cite{Heald_Contextual_2021}, but the prior samples they produce can be shown to be similar.
 For more details, see the supplementary materials.
-\todo{Do this}
+\todo{Do this}.
 
@@
 
-The first column of \fref{fig:model} shows that, in the presence of reliable contextual cues, context inference is correct, certain and fast to switch.
+The first column of \fref{fig:model} shows that, in the presence of reliable contextual cues, context inference is accurate, certain and fast to switch.
 However, as contextual cues become less reliable, switching between known contexts becomes slower, as seen in the posterior probabilities over contexts at and after trial 20 in the second and third columns.
@@
 In the first section, we discuss switches between contexts, and how slow context inference affects these switches.
-In the second section, we focus on the effects of volatility in the environment on motor adaptation.
-Finally, in the third section we discuss context inference during so-called error-clamp trials, and its effect on behavior.
+In the second section, we focus on interference between learned adaptations.
+Finally, in the third section we discuss context inference during error-clamp trials, and its effect on behavior.
 For each of the three sections, we selected one or two studies which are representative of the phenomenon being discussed.
@@
 Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
-In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.
+In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate than that observed during the first A block.
 The related concept of quick de-adaptation occurs in A-O transitions, where participants switch back to baseline without having to re-learn it.
@@
 To show this, we examined multiple experimental studies in which savings are observed.
-We categorized these studies based on the amount of contextual information made available to participants: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant.
+We categorized these studies based on the amount of contextual information made available to participants: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant using sensory cues.
 We call these cued-context experiments.
@@
 Importantly, a colored light identified the current context, making this a cued-context experiment.
-Consistent with our model, the authors found that switching was immediate and accurate.
+Consistent with the context-inference account, the authors found that switching was immediate and accurate.
 
-In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
+In \fref{fig:oh-2019}A we show the results of simulations with the model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
 The correspondence between simulations and experimental results can be seen in the switches between contexts in the center panel, i.e.
@@
 These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches.
+
 To expand on these results, we now turn to feedback and its effects on switching behavior.
-
 \cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visual rotation of 20 and 10 degrees, respectively.
-The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with our model.
+The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with the COIN model.
 Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A$.
@@
 
-Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism.
+Note that the results from our simulations from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism.
 
 \subsubsection{Uncertainty over contexts affects action selection}
-As with learning, action selection is affected by context inference.
+As with learning, action selection  affected by context inference.
 If the identity of the current context is known, the forward model for this context is used to select the current action.
@@
 
-We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}.
+We simulated data with the model fitting the experimental structure in \cite{Davidson_Scaling_2004}.
 We show the results in \fref{fig:davidson-2004}A, alongside the experimental results from \cite{Davidson_Scaling_2004}.
@@
 However, given the larger distance from baseline, the baseline context has a much lower attribution when the switch occurs compared to the original experiment.
-This change makes both simulated groups infer the correct context in much more similar times, making the difference between their errors be much smaller compared to the original experiment.
+This change makes both simulated groups infer the correct context almost equally quickly, making the difference between their errors be much smaller compared to the original experiment.
 Furthermore, we simulated an experiment with an identical structure to that of \cite{Davidson_Scaling_2004}, but eliminated the baseline context from the agent.
@@
 
-Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
-For example, proprioceptive signals  provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
-The sudden appearance of motor errors can itself be a cue for contextual change \citep{Herzfeld_memory_2014} and even a pause between two trials could suggest a change in context \cite{Ethier_Spontaneous_2008}.
-
 In this section, we show how context inference can explain these different parameters of behavior by changing the way contextual cues mislead participants' context inference, which in turn influences action selection.
@@
 No pauses were made during the experiment nor were there any contextual cues, so transitions between blocks were not signaled to participants.
+However, because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context.
+For example, proprioceptive signals  provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}.
+The sudden appearance of motor errors can itself be a cue for contextual change \citep{Herzfeld_memory_2014} and even a pause between two trials could suggest a change in context \citep{Ethier_Spontaneous_2008}.
 
-In \fref{fig:vaswani-2013}A, we show data simulated with the COIN, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
+In \fref{fig:vaswani-2013}A, we show data simulated with the COIN model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
 The displayed adaptation is shown during the E trials for the three experimental groups in the experiment.
@@
 We selected representative experimental studies that show the well-established effects of savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
-Using our model, we showed how each of these effects can be explained by the specifics of context inference, which integrates all the available information (e.g.
+Using the COIN model introduced by \cite{Heald_Contextual_2021}, we showed how each of these effects can be explained by the dynamics of context inference, which integrates all the available information (e.g.
 sensory cues, workspace location, reward and endpoint feedback), in some cases throughout many trials.
@@
 
-The same rationale explains the results by \cite{Herzfeld_memory_2014}: motor learning, which in our model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
+The same rationale explains the results by \cite{Herzfeld_memory_2014}: motor learning, which in the COIN model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
 This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
@@
 
+\subsubsection{Interference effects}
+As discussed in the Resuls section, the effect observed by \cite{Davidson_Scaling_2004} is explained by the model as an effect of slow context inference, instead of being a direct interference at the level of learning.
+As shown in \fref{fig:davidson-2004}B-C, the context inference account predicts that this effect would disappear if all contexts were significantly different from baseline, such that the baseline context never explains the observations.
+Removing the baseline context from a participant's context inference might be experimentally unfeasible, but other possibilities include making all adaptations bigger (e.g.
+bigger angles, stronger forces), and including contextual cues that rule out the baseline context.
+In the opposite direction, the model predicts that if all adaptations are smaller (i.e.
+closer to baseline), the differences between the two groups would increase.
+
 \subsubsection{Reducing the effect of volatility on learning}
-Experiments \citep{Herzfeld_memory_2014} have shown an effect of the volatility of the environment (i.e.
+Experiments \citep{Herzfeld_memory_2014,Heald_Contextual_2021} have shown an effect of the volatility of the environment (i.e.
 unpredictable switching between contexts) on measured learning rate.
-Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
+The model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
 \cite{Heald_Contextual_2021} showed that the COIN model explains this change in terms of the Kalman gain and the posterior probability over contexts (see their equation 14) \todo{Check this in more detail}.
 
-If this prediction is confirmed, our model would additionally suggest that human participants do not revisit the learned adaptation of the previous trial when a new observation comes in.
+If this prediction is confirmed, the model would additionally suggest that human participants do not revisit the learned adaptation of the previous trial when a new observation comes in.
 To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}: at trial $t$, the true context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
@@
 Does the participant update the internal model of A or of B?
-According to our model, participants incorrectly update A and, upon learning of their mistake at trial $t+1$, do not revert this update.
+According to the model, participants incorrectly update A and, upon learning of their mistake at trial $t+1$, do not revert this update.
 If this were not the case, the volatility of the environment would have no effect even without contextual cues, as the context at trial $t$ can almost always be identified at trial $t+1$.
@@
 \begin{equation}
-u_t  = \displaystyle\sum_{j}p(\zeta_{j,t} | ...) x_{j,t}
+u_t  = \displaystyle\sum_{j}p(\zeta_{j,t} | ...) x_{j,t} \label{eqn:dist-comm}
 \end{equation}

### 2022-05-17 adaptation/article/article.tex
Update on Overleaf.

@@
 
-\title{The effects of uncertain context inference on motor adaptation}
+\title{The effects of probabilistic context inference on motor adaptation}
 \shorttitle{Context-inference-dependent motor adaptation}
@@
 If the change is reverted, the adaptation is also quickly reverted.
-Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly.
+Humans are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly.
 Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
@@
 This model was used to show the effects of context inference on learning rates across different experiments.
-We expanded on that work by using the COIN model to show that the effects of context inference on motor adaptation and control go further than previously shown.
+We expanded on that work by using the COIN model to show that the effects of context inference on motor adaptation and control go even further than previously shown.
 Here, we used the COIN model to simulate classical motor adaptation experiments from previous works and showed that context inference, and how it is affected by the presence and reliability of feedback, effect a host of behavioral phenomena that had so far required multiple hypothesized mechanisms, lacking a unified explanation.
-Concretely, we show that the reliability of direct contextual information, as well as noisy sensory feedback, typical of many experiments, effect measurable changes in switching-task behavior, as well as in action selection, that stem directly from imprecise context inference.
+Concretely, we show that the reliability of direct contextual information, as well as noisy sensory feedback, typical of many experiments, effect measurable changes in switching-task behavior, as well as in action selection, that stem directly from probabilistic context inference.
 
@@
 This is known as motor adaptation.
-Moreover, human participants have been shown to adapt to different, even opposing, such changes during the course of a single experiment \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}.
+Moreover, human participants have been shown to adapt to different, even opposing, changes during the course of a single experiment \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}.
 Additionally, humans have been shown to dynamically switch between different learned adaptations \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.
@@
 
-Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced \cite{Heald_Contextual_2021}, called COIN (for context inference).
+Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced by \cite{Heald_Contextual_2021}, called COIN (for context inference).
 \cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, but is informed by it, establishing a loop whereby context inference also informs motor learning.
-With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning un previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.
+With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning in previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.
 
 In this work, we use the COIN model to show that the process of context inference underlies more behavioral phenomena than previously shown.
-we focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
+We focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
 More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured.
@@
 
-We show that the dynamics of context inference provide a principled mechanistic explanation for the experimental phenomena outlined above, providing a single account for motor adaptation under changing contexts, while relying on known mechanisms for adaptation.
+With these combined simulations and their qualitative comparison to the experimental phenomena outlined, we provide further evidence that context inference may be a single coherent and mechanistic account that underlies experimentally well-established motor adaption and history effects under changing contexts.
 
 \section{Results}
-Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how this model explains these experimental findings using the dynamics of context inference.
+Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how this model explains different experimental findings using the dynamics of context inference.
 We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.
 
-Before presenting these results, we make a brief description of the COIN model, leaving a more thorough explanation for the methods section.
+Before presenting these results, we briefly describe the COIN model, leaving a more thorough explanation for the methods section.
 Additionally, we present simulations using the model that show the effects of contextual cues and perceptual noise on context inference, which pave the way for the simulated experiments that we show later on.
@@
 inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities.
-In this work, we focus on switching behavior between known contexts, as well as in the perceptual aspect of context inference.
+In this work, we focus on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference.
 Additionally, we do not fit the COIN model to participant data, instead focusing on showing the way the COIN model explains known experimental phenomena.
@@
 \includegraphics[width=0.7\textwidth]{./figures/figure_1.png}
-\caption{Schematic representation of the model and simulations.
+\caption{Schematic representation of the model and illustrative simulations.
 (A) Generative model.
@@
 cursor position).
-The real context $\zeta_t$ is not directly observable, but influences $y_t$.
+The true context $\zeta_t$ is not directly observable, but influences $y_t$.
 The dark rectangles represent the prior distributions for the inferred adaptation level $x_t$ (Normal-Gamma distribution) and the current context $\zeta_t$ (discrete distribution with known $\pi$; see Methods).
 At every trial, the context is inferred, then motor adaptation is carried out and finally a motor command is issued; the flow of this process is indicated by the gray arrow in the background, while black arrows show the direction of information flow.
-(B) Simulations obtained with the model in (A), using an experimental setup similar to that by \cite{Davidson_Scaling_2004}.
-The plots are organized as follows: each pair of plots represent context inference (bottom; p(ctx)) and response/adaptation (top).
-In the context inference plot, the y-axis represents the posterior probability of each context $p(\zeta_t = j$; the black line represents the baseline context (i.e.
+(B)\todo{Explain vertical dashes} Simulations obtained with the model in (A), using a simulated experimental setup similar to that by \cite{Davidson_Scaling_2004}.
+A total of 3x3 experiments were simulated, with low, medium and high levels of both cue uncertainty and observation noise.
+Each pair of plots represent response/adaptation (top) and context inference $p(ctx)$ (bottom), for one specific level of cue uncertainty and observation noise.
+In the context inference plot, the y-axis represents the posterior probability of each context $p(\zeta_t = j)$; the black line represents the baseline context (i.e.
 no adaptation), the green line represents the only adaptation to be learned during the simulated experiment.
@@
 The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.
-Note, however, that the top plots are shown for a single run, for clarity (see main text).
-A total of 3x3 experiments were simulated, with low, medium and high levels of both cue uncertainty and observation noise.
+The top plots are shown for a single run, for clarity (see main text).
 All plots share the same x-axis, which represents the trial number.}
@@
 \end{figure}
-\todo{Probably remove the top plots, as they are hard to explain and add very little to the paper}
+\todo{Remove adaptation plots}
 These priors are not mathematically equivalent to those used by \cite{Heald_Contextual_2021}, but the prior samples they produce can be shown to be similar.
@@
 The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment.
-To show these effects, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to a visuomotor rotation of the cursor on a screen in relation to the movement of a pointing device.
+To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to a visuomotor rotation of the cursor on a screen in relation to the movement of a pointing device.
 
-To show the effects that will be the base for the results presented below, we manipulated the presence and reliability of contextual cues, as well as the perceptual noise added to the cursor.
+The key to an intuitive understanding of the results presented below, is to observe what happens when the presence and reliability of contextual cues is varied, as well as the perceptual noise added to the cursor.
 In \fref{fig:model}B, a 3x3 grid of results can be seen, where each simulation in this grid is a combination of low, medium or high contextual cue reliability (where low reliability is equivalent to presenting no cues), and low, medium or high perceptual noise.
@@
 In this section, we discuss savings and de-adaptation in terms of switching between contexts.
-We show that through context inference and how its affected by contextual cues and obsevation noise, savings are not immediate, but a relatively fast process that reflects context inference.
+We show that through context inference and how it is affected by contextual cues and observation noise, savings are not immediate, but a relatively fast process that reflects context inference.
 
@@
 For clarity, both experimental and simulated results are shown only up to trial 300, of the original 600.
-(B) An experiment by \cite{Oh_Minimizing_2019} similar to A, but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced.
+(B) An experiment by \cite{Oh_Minimizing_2019} similar to (A), but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced.
 O-A and A-O transitions (i.e.
@@
 This undershooting of responses (i.e.
-not doing the 40 degree rotation) happens despite participants being able to immediately and with high certainty identify the real context, as shown in the first-column panel.
+not doing the 40 degree rotation) happens despite participants being able to immediately and with high certainty identify the true context, as shown in the first-column panel.
 These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches.
@@
 \subsubsection{Uncertainty over contexts affects action selection}
-As with learning, action selection  affected by context inference.
+As with learning, action selection is affected by context inference.
 If the identity of the current context is known, the forward model for this context is used to select the current action.
@@
 
-In the context-inference account, the asymmetry is caused by the existance of the baseline context, which has a non-zero probability $p(\zeta_O | s_t ...)$.
+Under the COIN model, the asymmetry is caused by the existence of the baseline context, which has a non-zero probability $p(\zeta_O | s_t ...)$.
 When a new block of trials starts (e.g.
@@
 
-To confirm our hypothesis, we simulated variants of the experiments in which the context-inference account predicts that the difference between groups disappears.
+To confirm our hypothesis, we simulated variants of the experiments in which the COIN model predicts that the difference between groups disappears.
 First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments.
@@
 However, given the larger distance from baseline, the baseline context has a much lower attribution when the switch occurs compared to the original experiment.
-This change makes both simulated groups infer the correct context almost equally quickly, making the difference between their errors be much smaller compared to the original experiment.
+This change makes both simulated groups infer the correct context almost equally quickly, making the difference between their errors much smaller compared to the original experiment.
 Furthermore, we simulated an experiment with an identical structure to that of \cite{Davidson_Scaling_2004}, but eliminated the baseline context from the agent.
@@
 
-\cite{Vaswani_Decay_2013} studied in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
+This can be seen for example for \cite{Vaswani_Decay_2013}, who studied in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
 The authors found that during an E block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous block.
@@
 The most direct secondary contextual information comes in the form of reward and endpoint feedback.
-Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
-When a participants observe an unexpectedly large error, they can infer that the inferred context might be incorrect.
+For example, participants may be told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement.
+When participants observe an unexpectedly large error, they can infer that the inferred context might be incorrect.
 This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.
@@
 Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where participants were shown random errors during E trials, with a variance matching that of previously observed motor commands.
-The authors showed that by matching the errors expected by participants, they eliminated the slow tapering off observed in most E blocks.
+The authors showed that by matching the errors expected by participants, they eliminated the slow tapering-off observed in most E blocks.
 
-\subsection{Model predictions and future work}
-The basic principle behind the results we presented is that context inference is a process that develops over time and that carries with it uncertainty.
+\subsection{Model predictions}
+The basic principle behind the results we presented is that the COIN model describes a process that develops over time and that carries with it uncertainty.
 This uncertainty affects learning and behavior during motor adaptation, effecting phenomena that are directly observable during behavioral experiments.
+In the following, we discuss several testable predictions that are direct consequences of the model.
 
@@
 However, this component works exclusively by creating new contexts in which the forward models take the same form but have different parameter values.
-It is unclear whether the model can be adapted to create contexts in an online fashion that operate in an essentially different manner, as is the case of error-clamp trials, in which participants' responses do not affect the outcome and motor commands are issued based on criteria not directly related to the goal of the task (e.g.
+It is currently unclear whether the model can be adapted to create contexts in an online fashion that operate in an essentially different manner, as is the case of error-clamp trials, in which participants' responses do not affect the outcome and motor commands are issued based on criteria not directly related to the goal of the task (e.g.
 energy minimization or comfort maximization).
@@
 \subsection{Conclusions}
-The results we presented in this work show that many behavioral phenomena observed across different experiments can be explained by the uncertainty in context inference and its effects on learning and action selection.
+The results we presented in this work indicate that several well-established behavioral phenomena observed across different motor adaptation experiments can be explained by the uncertainty in context inference and its effects on learning and action selection.
 Together with the results by \cite{Heald_Contextual_2021}, these results suggest new venues of investigation for future works in motor adaptation and context-dependent behavior.
@@
 \subsubsection{Using the model}
-The free parameters of this model can be fit to participants' data, as was done by \cite{Heald_Contextual_2021}.
+The free parameters of this model can be fitted to participants' data, as was done by \cite{Heald_Contextual_2021}.
 In this work, we instead chose values for these parameters to show that the model is capable of explaining the experimental phenomena in the Results section.

### 2022-05-25 adaptation/article/article.tex
Final draft of the post-Heald version done.

@@
 inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities.
-In this work, we focus on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference.
-Additionally, we do not fit the COIN model to participant data, instead focusing on showing the way the COIN model explains known experimental phenomena.
-For these reasons, we chose not to use these components of the COIN model, instead fixing both the transition probabilities between contexts and the number of contexts based on the real values during the experiments, which were made available to participants \todo{Check this}.
+In this work, we focused on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference.
+More specifically, we simulated previous experiments and focused on the way the COIN model qualitatively explains experimental phenomena observed in these experiments.
+To do this, we fixed both the transition probabilities between contexts and the number of contexts based on the real values during the experiments, which were made available to participants.
 Fixing these values, in turn, allowed us to simplify the generative model, internal to the agent, using conjugate priors, with which the agent can do exact Bayesian inference.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{./figures/figure_1.png}
+\includegraphics[]{./figures/figure_1.png}
 \caption{Schematic representation of the model and illustrative simulations.
@@
 At every trial, the context is inferred, then motor adaptation is carried out and finally a motor command is issued; the flow of this process is indicated by the gray arrow in the background, while black arrows show the direction of information flow.
-(B)\todo{Explain vertical dashes} Simulations obtained with the model in (A), using a simulated experimental setup similar to that by \cite{Davidson_Scaling_2004}.
+(B) Simulations obtained with the model in (A), using a simulated experimental setup similar to that by \cite{Davidson_Scaling_2004}, in which the context changes at trial 20 (vertical, dashed lines).
 A total of 3x3 experiments were simulated, with low, medium and high levels of both cue uncertainty and observation noise.
-Each pair of plots represent response/adaptation (top) and context inference $p(ctx)$ (bottom), for one specific level of cue uncertainty and observation noise.
-In the context inference plot, the y-axis represents the posterior probability of each context $p(\zeta_t = j)$; the black line represents the baseline context (i.e.
+Each plot represents context inference $p(ctx)$ for one specific level of cue uncertainty and observation noise.
+The y-axis represents the posterior probability of each context $p(\zeta_t = j)$; the black line represents the baseline context (i.e.
 no adaptation), the green line represents the only adaptation to be learned during the simulated experiment.
-In the top plot (adaptation), the y-axis is the adaptation to be learned (e.g.
-the angle in a visuomotor rotation experiment), using the same color scheme (black for baseline).
-The purple line is the agent's response for the trial.
-The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.
-The top plots are shown for a single run, for clarity (see main text).
-All plots share the same x-axis, which represents the trial number.}
+The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.}
 \label{fig:model}
 \end{figure}
-\todo{Remove adaptation plots}
-These priors are not mathematically equivalent to those used by \cite{Heald_Contextual_2021}, but the prior samples they produce can be shown to be similar.
-For more details, see the supplementary materials.
-\todo{Do this}.
 
@@
 We discuss possible implementations of such an experiment in the Discussion section, alongside with other predictions made by the model.
-\todo{Promises, promises}
 
@@
 Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more quickly than groups 1.1 and 1.4.
-\todo{Danger!
-Danger!
-A keen observer might point out that the model might want to learn a new context, as COIN does.
-What to do?
-What to do?}
 
@@
 \section{Discussion}
-We expand on previous works \cite{Heald_Contextual_2021,Oh_Minimizing_2019} that introduced the idea that context inference is a process that informs and is informed by motor adaptation by showing that it explains behavioral phenomena that had previously required different specific, ad-hoc mechanisms outside of contextual motor adaptation.
+We showed that context inference as an active, continuous process, can explain many behavioral phenomena observed experimentally.
+In particular, we showed that the effects of the presence and reliability of contextual cues, as well as observation noise, can cause behavior that can be observed during context switching, as well as during times in which context inference is hindered, as is the case during error-clamp trials in many experiments.
 
-We selected representative experimental studies that show the well-established effects of savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
+To do this, we selected representative experimental studies that show the well-established effects of savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
 Using the COIN model introduced by \cite{Heald_Contextual_2021}, we showed how each of these effects can be explained by the dynamics of context inference, which integrates all the available information (e.g.
@@
 
-We used the recently-introduced COIN model \cite{Heald_Contextual_2021}, with a different choice of prior distributions that greatly simplified our simulations without sacrificing relevant parts of the original specification.
-While we focused on the Bayesian motor adaptation of the COIN model, we expect the results we presented here to hold using alternative models for motor adaptation, such as other Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}.
-An exception is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, as these models do not incorporate uncertainty in the estimate of the parameters of the forward models.
+With this, we expanded on previous works that introduced the idea that context inference is a process that informs and is informed by motor adaptation by showing that it explains behavioral phenomena that had previously required different specific, ad-hoc mechanisms outside of contextual motor adaptation.
 
@@
 The model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
-\cite{Heald_Contextual_2021} showed that the COIN model explains this change in terms of the Kalman gain and the posterior probability over contexts (see their equation 14) \todo{Check this in more detail}.
+\cite{Heald_Contextual_2021} showed that the COIN model explains this change in terms of the Kalman gain and the posterior probability over contexts (see their equation 14).
 
@@
 where $\pi_{\zeta_{t-1}}$ is the transition probability vector from context $\zeta_{t-1}$ to all other contexts.
-MARKOV CHAIN.
 The contextual cues (when present in an experiment) are assumed to be drawn depending on the context following:
@@
 \end{equation}
-where $Q_{\zeta_t}$ is the probability vector with which the contextual cue $q_t$ is shown to the agent in context $\zeta_t$.
+where $\Phi_{\zeta_t}$ is the probability vector with which the contextual cue $q_t$ is shown to the agent in context $\zeta_t$.
 As pointed out by \cite{Heald_Contextual_2021}, both $\Phi$ and $\pi$ are in principle infinite, but a task-relevant finite set can be used instead.
@@
 \begin{equation}
-p(x_{\zeta,t}) = \mathcal{N}(\mu_{\zeta,x}, \sigma_{\zeta,x})
+p(x_{\zeta,t}) = \mathcal{N}(\mu_{\zeta,x}, \sigma_{\zeta, x})
 \end{equation}
-where $\sigma$ is assumed to be shared across all contexts.
-Note that $\mu$ and $\sigma$ are parametrized by the parameters of the AR(1) process, namely $\mu_{\zeta,x} = d_\zeta / (1 - a_\zeta)$ and $\sigma_{\zeta, x} = \sigma_q / 1 - a_\zeta^2$, where $\sigma_q$ is a free parameter of the model.
+Note that $\mu_{\zeta, x}$ and $\sigma_{\zeta,x}$ are parametrized by the parameters of the AR(1) process, namely $\mu_{\zeta,x} = d_\zeta / (1 - a_\zeta)$ and $\sigma_{\zeta, x} = \sigma_q / 1 - a_\zeta^2$, where $\sigma_q$ is a free parameter of the model which is not context dependent.
 
@@
 Additionally, by fixing these parameters the agent is able to perform exact Bayesian inference at each trial using conjugate priors, replacing the MCMC approach used by \cite{Heald_Contextual_2021} due to the mathematical intractability of the full formulation.
-This, however, does not change the model and was done purely for computational efficiency.
+This, however, does not significantly change the model and was done purely for computational efficiency.
 In this section, we describe how we fixed parameters and the procedure for Bayesian inference.
@@
 For a fixed value of these parameters, the transition matrices also become fixed.
-In particular, setting the parameters to CALCULATE THIS, the transition matrices are such that contexts are assumed to be sticky, i.e.
-a transition from a context to itself is more likely than to any other context.
-In our simulations, we set the probability of self-transitioning (denoted $p_\zeta$) depending on the experiments (see below), to numbers that match the information given to the participants of each study.
+In our simulations, we set the probability of self-transitioning (denoted $p_\zeta$) depending on the experiments (see below), to numbers that match the experimental setup of each study.
 
@@
 Because the experiments we simulate do not include probabilistic or deceiving cues, contextual cues, when present, unequivocally reflect the current context, i.e.
-$p(q_t = i | c_t = j) = d_{ij}$, where $\delta_{ij}$ is the Kronecker delta, equaling zero unless $i = j$.
+$p(q_t = i | c_t = j) = d_{ij}$, where $\delta_{ij}$ is the Kronecker delta, equaling one when $i = j$, zero otherwise.
 
-Finally, we chose an equivalent parametrization for the priors over the hidden variables $x_{j,t}$, using the mean $\mu_x$ and standard deviation $\sigma_x$ directly instead of the AR(1) $a$ and $d$ parameters used by \cite{Heald_Contextual_2021}.
-Note that we dropped the $j$ dependency for clarity.
-This parametrization, in conjunction with the fixed parameters outlined above,  allows us to set priors over $\mu_x$ and $\sigma_x$ that enable exact inference over the latent variables.
-The priors are given by a Normal-Gamma distribution, typically used for priors of the mean and standard deviation of a Gaussian distribution [CITATION NEEDED]:
+Finally, we chose a different parametrization for the priors over the hidden variables $x_{j,t}$, using the mean $\mu_x$ and standard deviation $\sigma_x$ directly instead of the AR(1) $a$ and $d$ parameters used by \cite{Heald_Contextual_2021}.
+In what follows, we dropped the $j$ dependency for clarity.
+This parametrization, in conjunction with the fixed parameters outlined above, allows us to set priors over $\mu_{\zeta,x}$ and $\sigma_{\zeta,x}$ that enable exact inference over the latent variables:
 \begin{equation}
@@
 This formulation comes with four free parameters (i.e.
-the hyper-priors $\mu_0, \nu_0, \alpha_0, \beta_0$), in accordance with the original formulation (note that \cite{Heald_Contextual_2021} fixed the mean of the priors for $b$ to zero).
+the hyper-priors $\mu_{0,i}, \nu_{0,i}, \alpha_{0,i}, \beta_{0,i}$), in accordance with the original formulation (note that \cite{Heald_Contextual_2021} fixed the mean of the priors for $b$ to zero).
 While the two formulations are not mathematically identical, the effects of the hyper-priors for both are the same; we discuss these effects in the next section.
-Additionally, we found numerically that these two formulations could provide almost identical samples for given parametrizations (see SUPPLEMENTARY MATERIALS).
-\todo{Do this}
 
@@
   \begin{split}
-  \mu_\phi^{(t)} &= \frac{\nu_\phi^{(t-1)} \mu_\phi^{(t-1)} + q(\zeta_i)s_t}{\nu_\phi^{(t-1)} + q(\zeta_i)} \\
-  \nu_\phi^{(t)} &= \nu_\phi^{(t-1)} + q(\zeta_i) \\
-  \alpha_\phi^{(t)} &= \alpha_\phi^{(t-1)} + q(\zeta_i) / 2 \\
-  \beta_\phi^{(t)} &= \beta_\phi^{(t-1)} + \frac{q(\zeta_i)\nu_\phi^{(t-1)}}{\nu_\phi^{(t-1)} +
-    q(\zeta_i)}\frac{\left(s_t - \mu_\phi^{(t-1)}\right)^2}{2}  \label{eqn:update-full}
+  \mu_{\phi, i}^{(t)} &= \frac{\nu_{\phi,i}^{(t-1)} \mu_{\phi,i}^{(t-1)} + q(\zeta_i)s_t}{\nu_{\phi,t}^{(t-1)} + q(\zeta_i)} \\
+  \nu_{\phi,t}^{(t)} &= \nu_{\phi,t}^{(t-1)} + q(\zeta_i) \\
+  \alpha_{\phi,t}^{(t)} &= \alpha_{\phi,i}^{(t-1)} + q(\zeta_i) / 2 \\
+  \beta_{\phi,i}^{(t)} &= \beta_{\phi,i}^{(t-1)} + \frac{q(\zeta_i)\nu_{\phi,i}^{(t-1)}}{\nu_{\phi,i}^{(t-1)} +
+    q(\zeta_i)}\frac{\left(s_t - \mu_{\phi,t}^{(t-1)}\right)^2}{2}  \label{eqn:update-full}
   \end{split}
 \end{align}
+where $s_t$ represents the observations, in the form of the error between the observed and expected outcomes of the motor command.
 Note that the effect of the evidence (i.e.
@@
 
-We estimated the task parameters from the information provided in their respective publications; when direct information was not provided, we estimated it from the reported results; these estimations are not exact, but function as a proof of context.
+We estimated the task parameters from the information provided in their respective publications; when direct information was not provided, we estimated it from the reported results; these estimations are not exact, but function as a proof of concept.
 Agent parameter values are held constant for the different conditions or groups for each experiment, except those parameters that are expected to vary across conditions.
@@
 
-The agent has direct knowledge on the adaptation and observation noises of the task; we do not expect this to be the case in an experiment (instead participants would estimate it during their early trials).
-\todo{check simulations to see if I can make them identical; if not, change this}
-
 The star notation (e.g. $x_j^*$) denotes the real value used in the simulation of the task, which may be different from that assumed by the agent.
 
-\subsubsection{Interpreting the hyperparameters}
+For the simulations in \fref{fig:model}, the parameters were set as in the experiments by \cite{Davidson_Scaling_2004}, with two exceptions: (1) the cue noise, which is set to the values of 0, 0.003 and 0.33, for the low, medium and high values, respectively; and (2) the agent's observation noise $\sigma_r$, with values of 0.5, 1 and 2.
+
+\subsubsection{Interpreting the hyper-parameters}
 \label{sec:interpreting-hyperparameters}
 $\mu$ determines the initial estimate of the adaptation, in the same units as
-the necessary adaptation. $\nu$ encodes how stable this hyperprior is: higher
-values (e.g. 10,000) all but guarantee that the hypermean will not change its
-value after observations; In principle, enough evidence should still modify it,
-but that would not happen during an experiment. Smaller values (i.e. $\sim 1$)
-make the hypermean follow evidence more freely. Note that as more observations
-are accumulated, $\nu$ becomes bigger and bigger, stabilizing the value of the
-hypermean.
+the necessary adaptation. $\nu$ encodes how stable this hyper-prior is: higher
+values (e.g. 10,000) all but guarantee that the hyper-prior $\mu$ will not
+change its value after observations; In principle, enough evidence should still
+modify it, but that would not happen during an experiment. Smaller values
+(i.e. $\sim 1$) make $\mu$ follow evidence more freely. Note that as more
+observations are accumulated, $\nu$ becomes bigger and bigger, stabilizing the
+value of $\mu$.
 
-The hyperparameters $\alpha$ and $\beta$ have a more complex effect. Note that
+The hyper-parameters $\alpha$ and $\beta$ have a more complex effect. Note that
 the mean of a gamma distribution is $1 / (\alpha \nu)$; this mean is being used
-as the standard deviation of a Gaussian by the rest of the agent, which makes it
-an important measure of uncertainty. While setting the default hyperparameters,
-the values used are $\alpha = 0.5 / \sigma_0$ and $\beta = 0.5$, where
-$\sigma_0$ is the \textit{a priori} estimate of the standard deviation of the
-force exerted by the environment, which controls the initial learning rate. This
-makes the initial standard deviation equal $\sigma_0$, which makes it consistent
-with the fixed-force model. The 0.5 values ensure that uncertainty is large at
-the beginning and is greatly reduced during the experiment, but never to a point
-where it is so small that it makes trial-to-trial variation in the environment
-surprising. Changing this 0.5 would make the standard deviation change more
-quickly, making the model more or less precise in its predictions, independently
-of the volatility of the mean of the adaptation (via the hypermean).
+as the standard deviation of a Gaussian by the rest of the agent, which makes
+it an important measure of uncertainty. While setting the default
+hyper-parameters, the values used are $\alpha = 0.5 / \sigma_0$ and
+$\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the
+standard deviation of the force exerted by the environment, which controls the
+initial learning rate. This makes the initial standard deviation equal
+$\sigma_0$, which makes it consistent with the fixed-force model. The 0.5
+values ensure that uncertainty is large at the beginning and is greatly reduced
+during the experiment, but never to a point where it is so small that it makes
+trial-to-trial variation in the environment surprising. Changing this 0.5 would
+make the standard deviation change more quickly, making the model more or less
+precise in its predictions, independently of the volatility of the mean of the
+adaptation (via $\mu$).
 
@@
 deviation is 0.001 (compared that to the size of the adaptations in mechanical
-arm experiments, around 0.0125), and the hyperparameters of the standard
+arm experiments, around 0.0125), and the hyper-parameters of the standard
 deviation are stable during the experiment.

### 2022-05-25 adaptation/code/figures.py

@@ -7,6 +7,8 @@ is in the paper. Those that do not have one are not paper figures, but
 auxiliary functions."""
 
 import ipdb
+from itertools import product
+
 import numpy as np
 from matplotlib import pyplot as plt
 import pandas as pd
@@ -18,6 +20,7 @@ import simulations as sims
 import model
 import task_hold_hand as thh
 import pars
+import heald
 
 FIGURE_FOLDER = '../article/figures/'
 
@@ -27,7 +30,7 @@ mpl.rcParams['lines.linewidth'] = 1
 
 
 def model_showoff(fignum=1, show=True, do_a=True):
-    """Plots a bunch of arbitrary simulationsthat show how the parameters of
+    """Plots a bunch of arbitrary simulations that show how the parameters of
     the model work and affect inference. Also leaves an empty space on top
     to insert the diagram of the generative model by hand.
 
@@ -40,31 +43,29 @@ def model_showoff(fignum=1, show=True, do_a=True):
 
     """
     runs = 50
-    colors = ['black', 'tab:green', 'tab:blue']
+    highlight_color = 'lightgray'
+    colors = ['black', 'tab:olive', 'tab:blue']
     cycler = plt.cycler('color', colors)
     tasks, agents_pars = sims.model_showoff(plot=False)
     numbers = tasks.shape
-    figsize = (6, 8)
+    figsize = (5, 6)
     fig = plt.figure(fignum, clear=True, figsize=figsize)
-    height_ratios = np.ones(numbers[0] + 2)
+    height_ratios = 0.5 * np.ones(numbers[0] + 2)
     height_ratios[0] = 3
     height_ratios[1] = 0.3
     gsbig = gs.GridSpec(numbers[0] + 2, numbers[1], figure=fig,
                         height_ratios=height_ratios, hspace=0.1, wspace=0.1)
     axis_diagram = fig.add_subplot(gsbig[0, :])
-    ctx_switch = np.diff(tasks.reshape(-1)[0]['context_seq']).nonzero()[0][0] + 1
+    ctx_switch = np.diff(
+        tasks.reshape(-1)[0]['context_seq']).nonzero()[0][0] + 1
     num_trials = len(tasks.reshape(-1)[0]['context_seq'])
-    axes = np.empty((*numbers, 2), dtype=object)
-    small_hr = [2, 1]  # height ratios for adaptation/p(ctx)
+    axes = np.empty(numbers, dtype=object)
[…]

### 2022-05-25 adaptation/code/heald.py

@@ -1,0 +1,291 @@
+# -*- coding: utf-8 -*-
+# ../code/heald.py
+
+"""Some functions to test things for the Heald model and related stuff."""
+from itertools import product
+
+import pymc3 as pm
+import numpy as np
+from matplotlib import pyplot as plt
+import arviz as az
+
+plt.ion()
+
+
+def i_want_to_print_yes():
+    """yes"""
+    print('yes')
+
+
+def priors_over_x(obs=None, coin_args=None, mine_args=None):
+    """Check the priors over x_t and compare (a,b) to (mu, sd)."""
+    model_coin = pm.Model(name='COIN')
+    model_mine = pm.Model(name='Mine')
+
+    if coin_args is not None:
+        sigma_a, sigma_d, mu_a, sigma_q = coin_args
+        mu_b = 0
+    else:
+        sigma_a = 2
+        sigma_d = 2
+        mu_a = 0
+        mu_b = 0  # Fixed
+        sigma_q = 1
+    with model_coin:
+        ab = pm.MvNormal('ab', mu=np.array([mu_a, mu_b]),
+                         cov=np.diag([sigma_a, sigma_d]), shape=2)
+        bound_a = pm.Deterministic('a', pm.math.sigmoid(ab[0]))
+        mu_x = pm.Deterministic('mu_x', ab[1] / (1 - bound_a))
+        sd_x = pm.Deterministic('sd_x', sigma_q / (1 - bound_a ** 2))
+        if obs is None:
+            _ = pm.Normal('x', mu=mu_x, sd=sd_x)
+        else:
+            _ = pm.Normal('x', mu=mu_x, sd=sd_x, observed=obs)
+
+    if mine_args is not None:
+        mu_x, nu_x, alpha, beta = mine_args
+    else:
+        mu_x = 0
+        nu_x = 1
+        alpha = 36 / 2
+        beta = 10 / 2
+    with model_mine:
+        sd = pm.Gamma('sd_x', alpha=alpha, beta=beta)
+        mu = pm.Normal('mu_x', mu=mu_x, sd=sd / nu_x)
+        if obs is None:
+            _ = pm.Normal('x', mu=mu, sd=sd)
+        else:
+            _ = pm.Normal('x', mu=mu, sd=sd, observed=obs)
+
+    return model_coin, model_mine
+
+
+def sample_models(models):
+    """Samples from the given pymc3 models. Returns the samples"""
+    num_models = len(models)
[…]

### 2022-06-27 adaptation/article/article.tex
Update on Overleaf.
Final draft for re-submission

@@
 \usepackage{tabularx}
+\usepackage{lineno}
+\linenumbers
 
@@
 Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-Recently, the COIN computational model for motor adaptation and context inference was introduced, which contains explicit components of context inference and Bayesian motor adaptation.
+Recently, the COIN computational model for motor adaptation and context inference was introduced, which contains components for context inference and Bayesian motor adaptation.
 This model was used to show the effects of context inference on learning rates across different experiments.
@@
 \subsubsection{Action selection in error-clamp blocks}
-During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Spontaneous recovery, during which participants' behavior is consistent with a previously-encountered context (e.g.
-in an O-A-E experiment, consistent with A); this phase, when present, is seen during the early trials of the E block.
+During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Participants' behavior is consistent with a previously-encountered context (called spontaneous recover in O-A-B-E experiments, where behavior is consistent with A); this phase, when present, is seen during the early trials of the E block.
 (2) A slow return to baseline, which can last as long as hundreds of trials \citep{Brennan_Decay_2015}.
-However, the direction of the spontaneous recovery, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
+However, the direction of adaptation during the first phase, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the response vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.
 
@@
 In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
-To show this, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context for each simulated run (one line per run, 10 runs per group, all groups).
+Furthermore, variability in the lag between and within groups is explained by perceptual noise, which directly affects context inference.
+To show these effects, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context for each simulated run (one line per run, 10 runs per group, all groups).
 The start and speed of the drop depend on the run (i.e.
@@
 
-The same rationale explains the results by \cite{Herzfeld_memory_2014}: motor learning, which in the COIN model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
+The same rationale explains the results by \cite{Herzfeld_memory_2014}, as was shown by \cite{Heald_Contextual_2021}: motor learning, which in the COIN model is modulated by context inference, is minimal for errors close to 2 and -2 (see their figure 2E).
 This is because an error of 2 or -2 signals that the participant incorrectly identified the context (as adaptation has a magnitude of 1).
@@
 
+For the model predictions discussed below, it is important to keep in mind that different contextual cues are not equally effective at separating motor responses during learning and switching \cite{Howard_Gone_2012,Howard_ContextDependent_2010,Imamizu_Explicit_2007}.
+Because of this, the model predictions hinge on selecting the adequate type of contextual information that maximally helps the participants select the appropriate motor response.
+
 \subsubsection{Error-clamp as a known context}
@@
 If the E block is learned by participants during training, it might still be difficult for them to infer that an E block has started, which would create delays similar to those in \fref{fig:vaswani-2013}.
-However, the model predicts that if a visual cue is introduced that identifies the E block, participants would immediately switch to their baseline behavior, no longer displaying spontaneous recovery, lag, nor the slow return to baseline.
+However, the model predicts that if a visual cue is introduced that identifies the E block, participants would immediately switch to their baseline behavior, no longer displaying an adapted response, lag, nor the slow return to baseline.
 This immediate switch in the presence of contextual cues would persist even if endpoint feedback is manipulated as \cite{Vaswani_Decay_2013} did.
@@
 However, this component works exclusively by creating new contexts in which the forward models take the same form but have different parameter values.
-It is currently unclear whether the model can be adapted to create contexts in an online fashion that operate in an essentially different manner, as is the case of error-clamp trials, in which participants' responses do not affect the outcome and motor commands are issued based on criteria not directly related to the goal of the task (e.g.
+New mechanisms would be needed to allow the COIN model to create contexts in an online fashion that operate in an essentially different manner, as is the case of error-clamp trials, in which participants' responses do not affect the outcome and motor commands are issued based on criteria not directly related to the goal of the task (e.g.
 energy minimization or comfort maximization).
 
-\subsubsection{Interference effects}
+\subsubsection{Interference effects during context switching}
 As discussed in the Resuls section, the effect observed by \cite{Davidson_Scaling_2004} is explained by the model as an effect of slow context inference, instead of being a direct interference at the level of learning.

### 2022-09-04 adaptation/article/article.tex
Plos One submission ready

@@
 \abstract{Abstract}
-Humans have been shown to adapt their movements when sudden or gradual changes to the dynamics of the environment are introduced, a phenomenon called motor adaptation.
+Humans have been shown to adapt their movements when a sudden or gradual change to the dynamics of the environment are introduced, a phenomenon called motor adaptation.
 If the change is reverted, the adaptation is also quickly reverted.
 Humans are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly.
-Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations.
-Recently, the COIN computational model for motor adaptation and context inference was introduced, which contains components for context inference and Bayesian motor adaptation.
-This model was used to show the effects of context inference on learning rates across different experiments.
-We expanded on that work by using the COIN model to show that the effects of context inference on motor adaptation and control go even further than previously shown.
-Here, we used the COIN model to simulate classical motor adaptation experiments from previous works and showed that context inference, and how it is affected by the presence and reliability of feedback, effect a host of behavioral phenomena that had so far required multiple hypothesized mechanisms, lacking a unified explanation.
+Such switching relies on contextual information which is often noisy or misleading, affecting the switch between known adaptations.
+Recently, computational models for motor adaptation and context inference have been introduced, which contain components for context inference and Bayesian motor adaptation.
+These models were used to show the effects of context inference on learning rates across different experiments.
+We expanded on these works by using a simplificed version of the recently-introduced COIN model to show that the effects of context inference on motor adaptation and control go even further than previously shown.
+Here, we used this model to simulate classical motor adaptation experiments from previous works and showed that context inference, and how it is affected by the presence and reliability of feedback, effect a host of behavioral phenomena that had so far required multiple hypothesized mechanisms, lacking a unified explanation.
 Concretely, we show that the reliability of direct contextual information, as well as noisy sensory feedback, typical of many experiments, effect measurable changes in switching-task behavior, as well as in action selection, that stem directly from probabilistic context inference.
@@
 
-In this work, we use the COIN model to show that the process of context inference underlies more behavioral phenomena than previously shown.
+In this work show that the process of context inference underlies more behavioral phenomena than previously shown.
+To do this, we used a minimal model for motor adaptation that includes context inference, which we derived by simplifying the COIN model (henceforth called sCOIN).
 We focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
 More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured.
-We show that through context inference, switching behavior can display three main effects that have been previously attributed to hypothesized ad-hoc mechanisms: (1) The size of an adaptation dictates how quick and reliable switching between tasks is \cite{Oh_Minimizing_2019,Kim_Neural_2015}, which we explain in terms of the effects of perceptual noise on context inference.
+We show that through context inference, switching behavior can display three main effects that have been previously attributed to hypothesized ad-hoc mechanisms: (1) The size of an adaptation dictates how quick and reliable switching between tasks is \citep{Oh_Minimizing_2019,Kim_Neural_2015}, which we explain in terms of the effects of perceptual noise on context inference.
 (2) Previously-learned adaptations can interfere with switching behavior \citep{Davidson_Scaling_2004}, which we explain in terms of uncertain context inference.
@@
 which adaptations have been learned and for how long) affects switching during error-clamp trials \citep{Vaswani_Decay_2013}, which we also attribute to uncertain context inference.
-To do this, we used the COIN model to simulate the experimental setups and the decision-making agents (i.e.
+To do this, we used the sCOIN model to simulate the experimental setups and the decision-making agents (i.e.
 participants) during those experiments.
 
-With these combined simulations and their qualitative comparison to the experimental phenomena outlined, we provide further evidence that context inference may be a single coherent and mechanistic account that underlies experimentally well-established motor adaption and history effects under changing contexts.
+Importantly, the goal of this work is not to introduce a new model for contextual motor learning, but to use the existing ideas of the COIN model to show that context inference can explain more experimental phenomena than those explored by \cite{Heald_Contextual_2021}.
+
+With these combined simulations and the qualitative comparison to the experimental phenomena outlined above, we provide further evidence that context inference is a single, coherent and mechanistic account that underlies experimentally well-established motor adaption and history effects under changing contexts.
 
 \section{Results}
-Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how this model explains different experimental findings using the dynamics of context inference.
+Using the sCOIN model, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how this model explains different experimental findings using the dynamics of context inference.
 We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.
 
-Before presenting these results, we briefly describe the COIN model, leaving a more thorough explanation for the methods section.
-Additionally, we present simulations using the model that show the effects of contextual cues and perceptual noise on context inference, which pave the way for the simulated experiments that we show later on.
+Before presenting these results, we briefly describe the COIN model and the simplifications that led to the sCOIN version used in simulations.
+We leave a more thorough explanation of the models for the methods section.
+Additionally, we present simulations using the sCOIN model that show the effects of contextual cues and perceptual noise on context inference, which pave the way for the simulated experiments that we show later on.
 
@@
 inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities.
-In this work, we focused on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference.
-More specifically, we simulated previous experiments and focused on the way the COIN model qualitatively explains experimental phenomena observed in these experiments.
-To do this, we fixed both the transition probabilities between contexts and the number of contexts based on the real values during the experiments, which were made available to participants.
-Fixing these values, in turn, allowed us to simplify the generative model, internal to the agent, using conjugate priors, with which the agent can do exact Bayesian inference.
-The generative model we used here can be seen in \fref{fig:model}A, including the priors for both context inference and motor adaptation.
+Because we seeked to focus on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference, we chose to fix the participant-inferred transition probabilities between contexts, as well as the total number of contexts; in our simulations, we assume that participants have already inferred the real values of these quantities.
+Because we mainly focus on switching behavior, as well as error-clamp trials (both of which involve alread-learned adaptations), these simplifications to the model have minimal effects on our results.
+See Methods for more details on the sCOIN model, as well as the differences between COIN and sCOIN.
+
+By fixing the aforementioned values, the sCOIN model has a simpler generative model which allows the agent to perform exact Bayesian inference for motor adaptation.
+The inference process can be seen in \fref{fig:model}A, including the priors for both context inference and motor adaptation.
 For more details on these choices, see the Methods section.
@@
 \caption{Schematic representation of the model and illustrative simulations.
-(A) Generative model.
+(A) Inference done by the model at every trial.
 Clear circles represent the observables, i.e.
@@
 (B) Simulations obtained with the model in (A), using a simulated experimental setup similar to that by \cite{Davidson_Scaling_2004}, in which the context changes at trial 20 (vertical, dashed lines).
-A total of 3x3 experiments were simulated, with low, medium and high levels of both cue uncertainty and observation noise.
-Each plot represents context inference $p(ctx)$ for one specific level of cue uncertainty and observation noise.
+A total of 2x2 experiments were simulated, with low and high levels of both cue uncertainty and observation noise.
+In the left panel, the states inferred by the model for each of the 2x2 simulated experiments, where each color represents one experiment, with the same colors as the panels on the right.
+On the right, each plot represents context inference $p(ctx)$ for one specific level of cue uncertainty and observation noise.
 The y-axis represents the posterior probability of each context $p(\zeta_t = j)$; the black line represents the baseline context (i.e.
-no adaptation), the green line represents the only adaptation to be learned during the simulated experiment.
+no adaptation), the colored line (with the same colors as the panel on the left) represents the only adaptation to be learned during the simulated experiment.
 The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.}
@@
 The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment.
-To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to a visuomotor rotation of the cursor on a screen in relation to the movement of a pointing device.
+To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to the curl forces exerted by a mechanical arm on the participants' hands as they perform pointing movements.
 
-The key to an intuitive understanding of the results presented below, is to observe what happens when the presence and reliability of contextual cues is varied, as well as the perceptual noise added to the cursor.
-In \fref{fig:model}B, a 3x3 grid of results can be seen, where each simulation in this grid is a combination of low, medium or high contextual cue reliability (where low reliability is equivalent to presenting no cues), and low, medium or high perceptual noise.
+The key to an intuitive understanding of the results presented below is to observe what happens when the presence and reliability of contextual cues is varied, as well as the perceptual noise on the position of the hand.
+In \fref{fig:model}B, a 2x2 grid of results is shown: each simulation in this grid is a combination of low or high contextual cue uncertainty (where high uncertainty is equivalent to presenting no cues), and low or high perceptual noise (representing how well participants can detect deviations from the straight-line movements).
 
-The first column of \fref{fig:model} shows that, in the presence of reliable contextual cues, context inference is accurate, certain and fast to switch.
-However, as contextual cues become less reliable, switching between known contexts becomes slower, as seen in the posterior probabilities over contexts at and after trial 20 in the second and third columns.
+\fref{fig:model}B shows that, in the presence of reliable contextual cues, context inference is accurate, certain and fast to switch.
+However, as contextual cues become less reliable, switching between known contexts becomes slower, as seen in the posterior probabilities over contexts at and after trial 20 in the last column.
 Furthermore, as perceptual noise increases, switching becomes not only slower, but also more uncertain, with individual agents incorrectly missing the switch entirely.
-
-Additionally, the motor responses (shown in purple) can be seen to depend not only on the current estimate of the rotation angle (shown as the green line), but also the uncertainty over the current context.
-This is most evident in the high cue uncertainty and observation noise panel, in which the response varies substantially from the estimate of the angle (i.e.
-the purple line deviates from the green line), even when the real context has the highest estimated probability.
+While cue uncertainty and observation noise have an effect on the motor adaptation process, as seen on the left-most panel in \fref{fig:model}B, in all simulations the hidden state (i.e.
+the force exerted by the mechanical arm, $x_{j,t}$) is quickly inferred.
 
@@
 Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all.
-Let us define the baseline context O as that in which the mechanical arm exerts no force.
-Contexts A and B can be defined as those with clockwise or counter-clockwise, respectively.
+Let us define the baseline context $O$ as that in which the mechanical arm exerts no force.
+Contexts $A$ and $B$ can be defined as those with clockwise and counter-clockwise forces, respectively.
 Abusing notation, a usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions.
-Similarly, one can define context A/2, with the same direction of adaptation as A, but half the magnitude.
-Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter E.
+Similarly, one can define context $A/2$, with the same direction of adaptation as $A$, but half the magnitude.
+Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter $E$.
 
-With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of O-A-B-E, or O-A-(-A)-E, which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).
-With repeated contexts \citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1$-$A_1$-$O_2$-\ldots .
+With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of $O-A-B-E$, or $O-A-(-A)-E$, which means that the participant goes through a block of trials with no external force applied ($O$), a number of trials with a clockwise curl force ($A$), a block with counter-clockwise forces ($B$), and finally a block with error-clamp trials ($E$).
+With repeated contexts \citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1-A_1-O_2-\ldots$.
 
-\subsubsection{Savings and slow/fast switching}
+\subsubsection{Cue- and sensory feedback uncertainty affects switching behavior}
 The term 'savings' refer to the ability to remember a previously-learned adaptation and apply it without having to re-learn it.
 Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
-In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate than that observed during the first A block.
-The related concept of quick de-adaptation occurs in A-O transitions, where participants switch back to baseline without having to re-learn it.
+In an $O-A-O-A$ experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate than that observed during the first $A$ block.
+The related concept of quick de-adaptation occurs in $A-O$ transitions, where participants switch back to baseline without having to re-learn it.
 
-In this section, we discuss savings and de-adaptation in terms of switching between contexts.
+In this section, we discuss savings in terms of switching between contexts.
 We show that through context inference and how it is affected by contextual cues and observation noise, savings are not immediate, but a relatively fast process that reflects context inference.
+In particular, we show that the manifestations of savings on behavior are mediated by context inference, which could mask the presence of savings in cases where observations do not unequivocally identify a context.
 
@@
 The correspondence between simulations and experimental results can be seen in the switches between contexts in the center panel, i.e.
-when the solid gray line (representing the true context) switches between 40, 0 and -40; in both panels (second and third), the thick black line, representing the agent's adaptation, quickly follows these switches.
+when the solid gray line (representing the true context) switches between 40, 0 and -40; in the two right-most panels, the thick black line, representing the agent's adaptation, quickly follows these switches.
 
@@
 \includegraphics{./figures/figure_2.png}
-\caption{Savings and de-adaptation.
-Data from our simulations (first and second columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column).
-Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown.
-In the first column, the simulated context over inference is represented by the posterior probabilities over all available contexts.
+\caption{Switching between learned adaptations gated by context inference.
+Data from our simulations (first three columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column).
+Experimental and simulated data was, in all three experiments, averaged across all participants.
+In the first column, the simulated context inference is represented by the posterior probabilities over all available contexts.
 Each color (black, blue, green) represents a different context, with black always representing the baseline (i.e.
@@
 Vertical, dashed lines represent switches in the real context.
-In the second and third columns, the black line represents the adaptation (i.e.
+The inferred state (angle of the visuomotor rotation) is shown in the second column, with the same colors as in the first column.
+In the third and fourth columns, the black line represents the adaptation (i.e.
 response) displayed by the agent as a function of trial number.
@@
 the size of the true visuomotor rotation during the task.
-In all panels, the x-axis represents trial number, from 0 to 300.
-(A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O.
-Participants must adapt to the rotations during shooting movements, and the participant's response can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation.
-For clarity, both experimental and simulated results are shown only up to trial 300, of the original 600.
-(B) An experiment by \cite{Oh_Minimizing_2019} similar to (A), but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced.
-O-A and A-O transitions (i.e.
-adaptation and de-adaptation) can be observed.
+(A) Experiment by \cite{Kim_Neural_2015}.
+Both experimental and simulated results are shown only up to trial 300, of the original 600.
+(B) An experiment by \cite{Oh_Minimizing_2019}.
 Blue and black lines are as in A.
@@
 
-Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by their responses not being on par with the true rotation angle.
+In the second panel of \fref{fig:oh-2019}A, we show the inferred state (i.e.
+the inferred angle of rotation).
+Critically, it can be seen for the first four context switches (until about trial 200) that participants had not yet completely adapted to the rotation, as also evidenced by their responses not being on par with the true rotation angle.
 This undershooting of responses (i.e.
@@
 To expand on these results, we now turn to feedback and its effects on switching behavior.
-\cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visual rotation of 20 and 10 degrees, respectively.
-The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with the COIN model.
+\cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visuomotor rotation of 20 and 10 degrees, respectively.
+The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with the sCOIN model.
 Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A$.
-In subsequent $A - O$ transitions, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in their responses (black line in the last two columns of \fref{fig:oh-2019}B) closely following the switches in the black line (true rotation).
-In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up.
+In subsequent context transitions, participants showed immediate switching (with a one-trial lag) between $A$ and $O$ (both ways), which can be seen in their responses (black line in the last two columns of \fref{fig:oh-2019}B) closely following the switches in the true rotation.
+In the second experiment (\fref{fig:oh-2019}C), context switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up.
 As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C).
@@
 
-Note that the results from our simulations from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism.
+In contrast, \cite{Oh_Minimizing_2019} explained the results of their second experiment by positing that when adaptations were small, participants did not identify this as a new context and opted instead for a modification of their baseline model (i.e.
+how they move normally).
+Under this single-context explanation, however, savings do not exist, and adaptations need to be learned anew ever time a context changes.
+On the other hand, savings are present under the sCOIN model, their manifestations being diminished by the slower context inference.
+\cite{Oh_Minimizing_2019} analyzed savings during their second experiment and found that savings do exist, although greatly diminished compared to the first experiment.
+This is in favor of the dual-context model, as we present it here.
+\cite{Oh_Minimizing_2019} showed a slow decay during error-clamp trials in their experiments, which they considered evidence for the single-context account of the second experiment.
+However, as we show below, this can also be explained in a dual-context model as an effect of slow context inference.
 
 \subsubsection{Uncertainty over contexts affects action selection}
-As with learning, action selection is affected by context inference.
+As with learning, we show in this section that action selection is affected by context inference.
 If the identity of the current context is known, the forward model for this context is used to select the current action.
@@
 Experimental evidence supporting this view can be found in experiments with context switching.
-For example, \cite{Davidson_Scaling_2004} reported a curl-force experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A).
-After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A.
+For example, \cite{Davidson_Scaling_2004} reported a curl-force experiment in which participants had to switch between $A$ and $3A$ in one group, with a block sequence $A-3A-A-3A$, and from $A$ to $-A$ in another group, with a block sequence $A-(-A)-A-(-A)$.
+After $A$ and $3A$ (or $-A$ in the other group) had been learned in the first two blocks, the authors found that the switch from $3A$ to $A$ was faster than that from $-A$ to $A$.
 The authors interpreted this as evidence that switching between adaptations happens more quickly if it is in the same direction as the current adaptation (e.g.
@@
 
-Under the COIN model, the asymmetry is caused by the existence of the baseline context, which has a non-zero probability $p(\zeta_O | s_t ...)$.
+Under the sCOIN model, the asymmetry is caused by the existence of the baseline context, which has a non-zero probability $p(\zeta_O | s_t ...)$, as can be seen in \fref{fig:davidson-2004}A.
 When a new block of trials starts (e.g.
@@
 Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force).
-In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$.
-
-We simulated data with the model fitting the experimental structure in \cite{Davidson_Scaling_2004}.
-We show the results in \fref{fig:davidson-2004}A, alongside the experimental results from \cite{Davidson_Scaling_2004}.
-It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}.
+In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$ and the behavior lingers around $0$ until $p(\zeta_0|\ldots)$ drops back to zero.
 
@@
 \caption{Motor error when switching back to a previously-learned adaptation.
-(A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the COIN model are shown.
+(A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the SCOIN model are shown.
 In the first column, each panel represents context inference for one group of participants (top: group A; bottom: group 3A), with each line representing the posterior probability of a context (black for the baseline O).
@@
 All panels share the x-axis, representing the number of trials elapsed since the switch to the new context.
-All simulations were executed eight times per group, following the number of participants in the experiments; shaded areas represent the standard deviation across these eight simulations.
+All simulations were executed 32 times per group, to obtain a reliable mean; shaded areas represent the standard deviation across all simulations.
 (B) Simulated results for an experiment similar to \cite{Davidson_Scaling_2004}, but changing the contexts seen by the two groups from -A to -2A for the first group, and from 3A to 4A for the second group.
@@
 
-To confirm our hypothesis, we simulated variants of the experiments in which the COIN model predicts that the difference between groups disappears.
+To confirm this explanation, we simulated variants of the experiment in which the sCOIN model predicts that the difference between groups diminishes or disappears.
 First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments.
@@
 As in the original experiment, the second contexts ($-2A$ for one group, $4A$ for the other) are equally spaced from the first context.
-However, given the larger distance from baseline, the baseline context has a much lower attribution when the switch occurs compared to the original experiment.
+However, given the larger distance from baseline, the baseline context has the same probability for both groups after the switch back to $A$.
 This change makes both simulated groups infer the correct context almost equally quickly, making the difference between their errors much smaller compared to the original experiment.
@@
 The results can be seen in \fref{fig:davidson-2004}C, where the switches between contexts are made identically by the two groups.
-We discuss possible implementations of such an experiment in the Discussion section, alongside with other predictions made by the model.
 
 \subsubsection{Action selection in error-clamp blocks}
-During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Participants' behavior is consistent with a previously-encountered context (called spontaneous recover in O-A-B-E experiments, where behavior is consistent with A); this phase, when present, is seen during the early trials of the E block.
+During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Participants' behavior is consistent with a previously-encountered context (called spontaneous recover in $O-A-B-E$ experiments, where behavior is consistent with $A$); this phase, when present, is seen during the early trials of the E block.
 (2) A slow return to baseline, which can last as long as hundreds of trials \citep{Brennan_Decay_2015}.
@@
 
-This can be seen for example for \cite{Vaswani_Decay_2013}, who studied in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
-The authors found that during an E block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous block.
+This can be seen for example in \citep{Vaswani_Decay_2013}, where the authors studied in detail human behavior during an error-clamp block in a shooting movement paradigm with a mechanical arm.
+The authors found that during an $E$ block at the end of each experiment, there was a lag of a few trials (depending on participant) before their motor behavior changed from that of the previous block.
 After that, the exerted force slowly dropped towards zero throughout tens of trials, but never reaching values around zero.
-Participants were divided into four groups, each of which going through a different block sequence: (1) A-E, (2) O-A-E, (3) (-A/2)-A-E, and (4) (-A)-E.
+Participants were divided into four groups, each of which going through a different block sequence: (1.1) $A-E$, (1.2) $O-A-E$, (1.3) $(-A/2)-A-E$, and (1.4) $(-A)-E$.
 No pauses were made during the experiment nor were there any contextual cues, so transitions between blocks were not signaled to participants.
@@
 
-In \fref{fig:vaswani-2013}A, we show data simulated with the COIN model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
-The displayed adaptation is shown during the E trials for the three experimental groups in the experiment.
+In \fref{fig:vaswani-2013}A, we show data simulated with the sCOIN model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}.
+As in the experiments by \cite{Vaswani_Decay_2013}, we simulated error-clamp trials by forcing the observed error to zero, regardless of the action taken by the model.
+The displayed adaptation is shown during the $E$ trials for the three experimental groups in the experiment.
 It can be seen that group 1.3 (i.e.
-the participants who had learned in the -A/2 context in addition to A) more quickly recognized a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.
+the participants who had learned in the $-A/2$ context in addition to $A$) more quickly recognized a change in context and lowered the force applied on the mechanical handle, as can be seen in the experimental data.
 
@@
 (A) Simulated adaptation during the error-clamp trials for the three groups of participants in \cite{Vaswani_Decay_2013}, using the same colors.
-The groups differ in the sequence of adaptations: 1.1 performed an A-E sequence; 1.2, O-A-E; 1.3, (-A/2)-A-E; and 1.4, (-A)-E.
 Following \cite{Vaswani_Decay_2013}, group 1.4 is not shown in A and B, as their behavior is identical to group 1.1.
@@
 As in (A), solid lines represent the average across all runs and shaded areas represent the standard deviation.
-(D) Simulations: Visualization of the lag before a change in context is detected by the agent during the E trials.
+(D) Simulations: Visualization of the lag before a change in context is detected by the agent during the $E$ trials.
 Each line represents one run (10 runs per group).}
@@
 This causes the agent to infer the combination of some of the known contexts that best fits the observations.
-Depending on the contexts previously learned by the agent: groups 1.1 and 1.4 display the same behavior, where the previous context (A and -A, respectively) slowly dwindles.
+Depending on the contexts previously learned by the agent: groups 1.1 and 1.4 display the same behavior, where the previous context ($A$ and $-A$, respectively) slowly dwindles.
 These agents will slowly lower the force applied.
-In contrast, group 1.3 has learned the additional -A/2 context, which has a non-zero posterior probability during E trials, pushing the agent's adaptation force more quickly towards zero.
-Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more quickly than groups 1.1 and 1.4.
+In contrast, group 1.3 has learned the additional $-A/2$ context, which has a non-zero posterior probability during $E$ trials, pushing the agent's adaptation force more quickly towards zero.
+Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during $E$ trials, making the agent reduce its force during $E$ trials slightly more quickly than groups 1.1 and 1.4.
 
-Context inference also explains the variability in the lag before adaptation starts to drop to zero after the E block starts.
-In the model, this lag is due to a period in which context inference has not ``caught on'' to the change of context; during this period, behavior remains consistent with that of the previously-observed block, as can be seen in the experimental data as well.
-Furthermore, variability in the lag between and within groups is explained by perceptual noise, which directly affects context inference.
-To show these effects, we show in \fref{fig:vaswani-2013}D the drop in the inferred probability of the previous context for each simulated run (one line per run, 10 runs per group, all groups).
-The start and speed of the drop depend on the run (i.e.
-the participant), reproducing the different lags observed in experimental data.
+\cite{Vaswani_Decay_2013} also found participant-specific delays before the decay to baseline started after the error-clamp phase begun.
+As can be seen in \fref{fig:vaswani-2013}D, each simulated participant follows a different path of decay, with variations caused directly by perceptual noise.
+In our simulations, however, it is clear that no such systematic lag can be directly observed, which is most noticeable when looking at context inference (\fref{fig:vaswani-2013}C), which begins the switch as soon as the error-clamp trials begin.
+This can be further observed in \fref{fig:vaswani-2013}D, where we plot each simulated participant (one run of the simulations, color coded as before); because observation noise was chosen randomly for each run, some runs appear to contain a large delay before the decay begins.
+This falls in line with the experiments by \cite{Brennan_Decay_2015}, who found that the lag observed by \cite{Vaswani_Decay_2013} disappeared when controling for correlations in perceptual noise, as well as by using a balanced experimental design and unbiased analysis.
 
@@
 
-To do this, we selected representative experimental studies that show the well-established effects of savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues.
-Using the COIN model introduced by \cite{Heald_Contextual_2021}, we showed how each of these effects can be explained by the dynamics of context inference, which integrates all the available information (e.g.
+To do this, we selected representative experimental studies that show the well-established effects of savings, spontaneous recovery and the effects of sensory cues.
+Using a simplified version of the COIN model introduced by \cite{Heald_Contextual_2021}, we showed how each of these effects can be explained by the dynamics of context inference, which integrates all the available information (e.g.
 sensory cues, workspace location, reward and endpoint feedback), in some cases throughout many trials.
@@
 It is in these cases where the effects of context inference are most noticeable.
-While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases.
+While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicitly stochastic cases.
 Indeed, as we noted in the Results section, many experiments inadvertently include partial contextual information used by participants.
@@
 
-A subtler source of information can be found in long pauses between blocks of adaptation trials, after which an unprompted partial return to baseline has been observed \cite{Ethier_Spontaneous_2008}.
-This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occurred, prompting participants to rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context A, but now includes a component of the baseline O, as it is the most common one in everyday life.
+A subtler source of information can be found in long pauses between blocks of adaptation trials, after which an unprompted partial return to baseline has been observed \citep{Ethier_Spontaneous_2008}.
+This can be explained by context inference, as a long pause could prompt participants to infer that a switch had occurred, prompting participants to rely on their belief of the underlying probability of observing any of the known contexts, which is dominated by the previously observed context $A$, but now includes a component of the baseline $O$, as it is the most common one in everyday life.
 
-Error-clamp (E) trials present another insight.
+Error-clamp ($E$) trials present another insight.
 If error is kept at zero, one could assume that participants would continue doing what they were doing before, as there is no reason (no observed error) to infer a change in context.
@@
 This prompts participants to re-evaluate their inferred context, which can partially activate a previously-observed context, as we showed in \fref{fig:vaswani-2013}.
-\cite{Pekny_Protection_2011} found similar results, demonstrating that the duration of the previously-observed adaptation block also affects behavior in the E block.
+\cite{Pekny_Protection_2011} found similar results, demonstrating that the duration of the previously-observed adaptation block also affects behavior in the $E$ block.
 Additionally, \cite{Criscimagna-Hemminger_Consolidation_2008} showed that introducing long periods before the E block begins lowers the initial force that participants exerted on the mechanical arm during the E block; longer periods of time make context inference revert to the prior expectation that a new baseline block begins, because participants are free to move their arm about during the pause.
@@
 In our account, if all information indicating a change in context is removed from the experiment, participants would continue to behave as they were in the previous block.
-Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where participants were shown random errors during E trials, with a variance matching that of previously observed motor commands.
-The authors showed that by matching the errors expected by participants, they eliminated the slow tapering-off observed in most E blocks.
+Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where participants were shown random errors during $E$ trials, with a variance matching that of previously observed motor commands.
+The authors showed that by matching the errors expected by participants, they eliminated the slow tapering-off observed in most $E$ blocks.
 
 \subsection{Model predictions}
-The basic principle behind the results we presented is that the COIN model describes a process that develops over time and that carries with it uncertainty.
+The basic principle behind the results we presented is that the sCOIN model describes a process that develops over time and that carries with it uncertainty.
 This uncertainty affects learning and behavior during motor adaptation, effecting phenomena that are directly observable during behavioral experiments.
@@
 
-For the model predictions discussed below, it is important to keep in mind that different contextual cues are not equally effective at separating motor responses during learning and switching \cite{Howard_Gone_2012,Howard_ContextDependent_2010,Imamizu_Explicit_2007}.
+For the model predictions discussed below, it is important to keep in mind that different contextual cues are not equally effective at separating motor responses during learning and switching \citep{Howard_Gone_2012,Howard_ContextDependent_2010,Imamizu_Explicit_2007}.
 Because of this, the model predictions hinge on selecting the adequate type of contextual information that maximally helps the participants select the appropriate motor response.
@@
 We expect that the same effect would be observed in error-clamp trials.
-If the E block is learned by participants during training, it might still be difficult for them to infer that an E block has started, which would create delays similar to those in \fref{fig:vaswani-2013}.
-However, the model predicts that if a visual cue is introduced that identifies the E block, participants would immediately switch to their baseline behavior, no longer displaying an adapted response, lag, nor the slow return to baseline.
+If the $E$ block is learned by participants during training, it might still be difficult for them to infer that an $E$ block has started, which would create delays similar to those in \fref{fig:vaswani-2013}.
+However, the model predicts that if a visual cue is introduced that identifies the $E$ block, participants would immediately switch to their baseline behavior, no longer displaying an adapted response, lag, nor the slow return to baseline.
 This immediate switch in the presence of contextual cues would persist even if endpoint feedback is manipulated as \cite{Vaswani_Decay_2013} did.
@@
 In the opposite direction, the model predicts that if all adaptations are smaller (i.e.
-closer to baseline), the differences between the two groups would increase.
-
-\subsubsection{Reducing the effect of volatility on learning}
-Experiments \citep{Herzfeld_memory_2014,Heald_Contextual_2021} have shown an effect of the volatility of the environment (i.e.
-unpredictable switching between contexts) on measured learning rate.
-The model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.
-\cite{Heald_Contextual_2021} showed that the COIN model explains this change in terms of the Kalman gain and the posterior probability over contexts (see their equation 14).
-
-If this prediction is confirmed, the model would additionally suggest that human participants do not revisit the learned adaptation of the previous trial when a new observation comes in.
-To see this, consider the following scenario from the experiments by \cite{Herzfeld_memory_2014}: at trial $t$, the true context was B but the participant inferred context A, issued a motor command consistent with context A and then observed the outcome at $t + 1$.
-When the outcome is observed, it becomes clear to the participant that the context was B.
-Does the participant update the internal model of A or of B?
-According to the model, participants incorrectly update A and, upon learning of their mistake at trial $t+1$, do not revert this update.
-If this were not the case, the volatility of the environment would have no effect even without contextual cues, as the context at trial $t$ can almost always be identified at trial $t+1$.
+closer to baseline), the differences between the two groups would increase, although such differences might become impossible to detect due to different sources of noise in the data.
 
@@
 \section{Methods}
-\subsection{The COIN model}
-In this work, we used the recently-introduced COIN model \citep{Heald_Contextual_2021}, adapted to the experiments that we covered in our simulations.
+\subsection{The COIN and sCOIN models}
+In this work, we used a simplified version of the recently-introduced COIN model \citep{Heald_Contextual_2021}, adapted to the experiments that we covered in our simulations.
 In this section, we give a brief introduction to the COIN model and, in the subsequent subsection, describe how we adapted the model to the experimental tasks.
@@
 the parameters of the force field in mechanical-arm experiments).
-The context is represented by a latent, categorical variable $\zeta_t$:
+The context is represented by a latent, categorical variable $\zeta_t$, which is assumed to evolve over time according to:
 \begin{equation}
@@
 
-The context-dependent adaptation is represented by the latent variable $x_{\zeta,t}$ and assumed to arise from an autoregressive process AR(1), with a stationary Gaussian distribution of unknown mean and variance:
+The context-dependent adaptation is represented by the latent variable $x_{\zeta,t}$ and assumed to arise from an autoregressive process AR(1):
+\begin{equation}
+x_{\zeta,t} = a_\zeta x_{t-1} + b_\zeta + \omega_\zeta
+\end{equation}
+where $a_\zeta$ and $b_\zeta$ are unknown, context-dependent parameters and $\omega$ is a Gaussian noise term of zero mean and unknown standard deviation $\sigma_{\zeta, x}$.
+This AR(1) process is assumed to have existed before the experiment begins and to have a stationary Gaussian distribution of unknown mean and variance:
 \begin{equation}
@@
 \end{equation}
-Note that $\mu_{\zeta, x}$ and $\sigma_{\zeta,x}$ are parametrized by the parameters of the AR(1) process, namely $\mu_{\zeta,x} = d_\zeta / (1 - a_\zeta)$ and $\sigma_{\zeta, x} = \sigma_q / 1 - a_\zeta^2$, where $\sigma_q$ is a free parameter of the model which is not context dependent.
+Note that $\mu_{\zeta, x}$ and $\sigma_{\zeta,x}$ are parametrized by the parameters of the AR(1) process, namely $\mu_{\zeta,x} = d_\zeta / (1 - a_\zeta)$ and $\sigma_{\zeta, x} = \sigma_q / (1 - a_\zeta^2)$, where $\sigma_q$ is a free parameter of the model which is not context dependent.
 
@@
 \begin{equation}
-u_t  = \displaystyle\sum_{j}p(\zeta_{j,t} | ...) x_{j,t} \label{eqn:dist-comm}
+u_t  = \displaystyle\sum_{j}p(\zeta_{j,t} | q_t...) x_{j,t} \label{eqn:dist-comm}
 \end{equation}
+where $p(\zeta_{j,t} | q_t ...)$ is the predictive probability.
+
 To include motor noise (independent from estimation uncertainty), as well as carry over the uncertainty over $x_{j,t}$, we instead sample motor commands from a Gaussian centered on this mean, with a standard deviation $\sigma_u$, which is a free parameter of the model.
 
-\subsubsection{Using the model}
+\subsubsection{Simplified COIN model}
 The free parameters of this model can be fitted to participants' data, as was done by \cite{Heald_Contextual_2021}.
@@
 For a fixed value of these parameters, the transition matrices also become fixed.
-In our simulations, we set the probability of self-transitioning (denoted $p_\zeta$) depending on the experiments (see below), to numbers that match the experimental setup of each study.
+In our simulations, we set the probability of self-transitioning (denoted $p_\zeta$) depending on the experiments (see below), to numbers that approximate the experimental setup of each study.
 
@@
 This is done through a set of cue probability vectors that are generated via a parametric distribution, whose parameters are fitted to participants' data.
-Because the experiments we simulate do not include probabilistic or deceiving cues, contextual cues, when present, unequivocally reflect the current context, i.e.
+For experiments that do not include probabilistic or deceiving cues, contextual cues, when present, unequivocally reflect the current context, i.e.
 $p(q_t = i | c_t = j) = d_{ij}$, where $\delta_{ij}$ is the Kronecker delta, equaling one when $i = j$, zero otherwise.
+For the simulations of \fref{fig:model}, where contextual cues are probabilistic, cue uncertainty is implemented as $p(c_t = i | q_t = i) = 1 - \eta$, where $\eta$ is the cue uncertainty, and $p(c_t = i | q_t = j) = \eta / (N_c-1) \forall i \ne j$, where $N_c$ is the total number of contexts in the experiment.
 
-Finally, we chose a different parametrization for the priors over the hidden variables $x_{j,t}$, using the mean $\mu_x$ and standard deviation $\sigma_x$ directly instead of the AR(1) $a$ and $d$ parameters used by \cite{Heald_Contextual_2021}.
-In what follows, we dropped the $j$ dependency for clarity.
-This parametrization, in conjunction with the fixed parameters outlined above, allows us to set priors over $\mu_{\zeta,x}$ and $\sigma_{\zeta,x}$ that enable exact inference over the latent variables:
+Using the above, the probability of a context for the state feedback for a trial after the cue has been observed is given by:
+\begin{equation}
+p(c_t | q_t, y_{1:t-1}) \propto p(c_t | q_t) p(c_t | c_{t-1}) p(c_{t-1} | y_{1:t-1})
+\end{equation}
+where $p(c_t | c_{t-1})$ is given by the context self-transition ($p_\zeta$ in \tref{tab:parameters} below) such that:
+\begin{equation}
+p(c_t = i | c_{t-1} = j) = \begin{cases}
+                             p_\zeta &\quad \text{if } i = j \\
+                             \frac{1 - p_\zeta}{N_c - 1} &\quad \text{otherwise}
+                           \end{cases}
+\end{equation}
+
+Finally, for the hidden variables $x_{j,t}$ we chose a stationary Gaussian distribution with unknown mean $\mu_x$ and standard deviation $\sigma_x$, instead of the AR(1) $a$ and $d$ parameters used by \cite{Heald_Contextual_2021}.
+As a consequence, the sCOIN model does not have intrinsic memory decay, instead relying on the dynamics of context inference to explain the slow decay of memories during error-clamp trials \citep[e.g.][]{Scheidt_Persistence_2000,Vaswani_Decay_2013,Brennan_Decay_2015}.
+
+Using Bayesian inference, the model infers the values of $\mu_x$ and $\sigma_x$ using a Gaussian likelihood and NormalGamma priors, which allowed us to use exact inference.
+The likelihood of the data is given by the prediction error of the observations (which drives learning):
+\begin{equation}
+p(y_t | x_t) = \mathcal{N}(y - \hat y, \hat \sigma)
+\end{equation}
+where $\hat y$ is the predicted observation given the previous observation and the previous action, and $\hat \sigma$ is the expected standard deviation of the predicted observation, given by the updated parameters of the model (discussed below).
+
+We set priors over $\mu_{\zeta,x}$ and $\sigma_{\zeta,x}$ that enable exact inference over the latent variables $x$ (in what follows, we dropped the $j$ dependency for clarity):
 \begin{equation}
@@
   \begin{split}
-  \mu_{\phi, i}^{(t)} &= \frac{\nu_{\phi,i}^{(t-1)} \mu_{\phi,i}^{(t-1)} + q(\zeta_i)s_t}{\nu_{\phi,t}^{(t-1)} + q(\zeta_i)} \\
-  \nu_{\phi,t}^{(t)} &= \nu_{\phi,t}^{(t-1)} + q(\zeta_i) \\
-  \alpha_{\phi,t}^{(t)} &= \alpha_{\phi,i}^{(t-1)} + q(\zeta_i) / 2 \\
-  \beta_{\phi,i}^{(t)} &= \beta_{\phi,i}^{(t-1)} + \frac{q(\zeta_i)\nu_{\phi,i}^{(t-1)}}{\nu_{\phi,i}^{(t-1)} +
-    q(\zeta_i)}\frac{\left(s_t - \mu_{\phi,t}^{(t-1)}\right)^2}{2}  \label{eqn:update-full}
+  \mu_{\phi, i}^{(t)} &= \frac{\nu_{\phi,i}^{(t-1)} \mu_{\phi,i}^{(t-1)} + p(\zeta_i | q_t, ...)s_t}{\nu_{\phi,t}^{(t-1)} + p(\zeta_i | q_t, ...)} \\
+  \nu_{\phi,t}^{(t)} &= \nu_{\phi,t}^{(t-1)} + p(\zeta_i | q_t, ...) \\
+  \alpha_{\phi,t}^{(t)} &= \alpha_{\phi,i}^{(t-1)} + p(\zeta_i | q_t, ...) / 2 \\
+  \beta_{\phi,i}^{(t)} &= \beta_{\phi,i}^{(t-1)} + \frac{p(\zeta_i | q_t, ...)\nu_{\phi,i}^{(t-1)}}{\nu_{\phi,i}^{(t-1)} +
+    p(\zeta_i | q_t, ...)}\frac{\left(s_t - \mu_{\phi,t}^{(t-1)}\right)^2}{2}  \label{eqn:update-full}
   \end{split}
@@
 Note that the effect of the evidence (i.e.
-observations) on the inference over the context-dependent hidden states is gated by the probability of each context $p(\zeta_i)$, as in \cite[][supplementary materials]{Heald_Contextual_2021}.
+observations) on the inference over the context-dependent hidden states is gated by the probability of each context $p(\zeta_i | q_t, ...)$, as in \cite[][supplementary materials]{Heald_Contextual_2021}.
 
@@
   \cline{1-11}
-  & & & Kim (2015) & \multicolumn{2}{c|}{Oh (2009)} & \multicolumn{2}{c|}{Davidson (2004)} & \multicolumn{3}{c|}{Vaswani (2013)} \\
+  & & & Kim (2015) & \multicolumn{2}{c|}{Oh (2019)} & \multicolumn{2}{c|}{Davidson (2004)} & \multicolumn{3}{c|}{Vaswani (2013)} \\
   \cline{2-11}
@@
   \cline{2-11}
-  & & Adaptation noise & 0.01 & \multicolumn{2}{c|}{1} & \multicolumn{2}{c|}{0.5} & \multicolumn{3}{c|}{0.01} \\
+  & $\sigma_a$ & Adaptation noise & 0.01 & \multicolumn{2}{c|}{1} & \multicolumn{2}{c|}{0.5} & \multicolumn{3}{c|}{0.1} \\
   \cline{2-11}
-  & $\sigma_r^*$ & Obs. noise & 3 & \multicolumn{2}{c|}{1.5} & \multicolumn{2}{c|}{0.1} & \multicolumn{3}{c|}{0.01} \\
+  & $\sigma_r^*$ & Obs. noise & 3 & \multicolumn{2}{c|}{2.5} & \multicolumn{2}{c|}{0.1} & \multicolumn{3}{c|}{0.1} \\
   \cline{1-11}
@@
   \cline{2-2} \cline{4-11}
-  & $\nu_0$ & & 1e4, 10, 10 & \multicolumn{2}{c|}{0, 0} & \multicolumn{2}{c|}{1e4, 0.5, 0.5} & \multicolumn{2}{c|}{1e4, 1} & 1e4, 1, 1 \\
+  & $\nu_0$ & & 1e4, 1, 1 & \multicolumn{2}{c|}{1e4, 1} & \multicolumn{2}{c|}{1e4, 1, 1} & \multicolumn{2}{c|}{1e4, 1} & 1e4, 1, 1 \\
   \cline{2-2} \cline{4-11}
-  & $\alpha_0$ & & 25e4, 0.25, 0.25 & \multicolumn{2}{c|}{22e4, 2.2} & \multicolumn{2}{c|}{33e4, 33, 33} & \multicolumn{2}{c|}{5e5, 5} & 15e5, 5, 5 \\
+  & $\alpha_0$ & & 25e3, 0.25, 0.25 & \multicolumn{2}{c|}{22e3, 2.2} & \multicolumn{2}{c|}{33e3, 4e2, 4e2} & \multicolumn{2}{c|}{5e4, 5} & 15e4, 5, 5 \\
   \cline{2-2} \cline{4-11}
-  & $\beta_0$ & & 1e5, 2, 2 & \multicolumn{2}{c|}{1e5, 20} & \multicolumn{2}{c|}{1e5, 200, 200} & \multicolumn{2}{c|}{1e5, 2} & 1e5, 2, 2 \\
+  & $\beta_0$ & & 1e5, 2, 2 & \multicolumn{2}{c|}{1e5, 20} & \multicolumn{2}{c|}{1e5, 23e2, 23e2} & \multicolumn{2}{c|}{1e5, 2} & 1e5, 2, 2 \\
   \cline{2-11}
-  & $\sigma_u$ & Motor noise & 0.001 & \multicolumn{4}{c|}{2} & \multicolumn{3}{c|}{0.17} \\
+  & $\sigma_u$ & Motor noise & 1 & \multicolumn{4}{c|}{2} & \multicolumn{3}{c|}{0.17} \\
   \cline{2-11}
-  & $\sigma_r$& Obs. noise & 3 & \multicolumn{2}{c|}{3.5} & \multicolumn{2}{c|}{2.5} & \multicolumn{3}{c|}{0.1} \\
+  & $\sigma_r$& Obs. noise & 3 & \multicolumn{2}{c|}{2.5} & \multicolumn{2}{c|}{2.5} & \multicolumn{3}{c|}{0.1} \\
   \cline{1-11}
   \end{tabular}}
-\caption{Model and simulation parameters.}
+\caption{Model and simulation parameters. The star notation (e.g. $x_j^*$) denotes the real value used in the simulation of the task, which may be different from that assumed by the agent.}
 \label{tab:parameters}
@@
 
-The context-dependent hyper-priors $\{\mu_0, \nu_0, \alpha_0, \beta_0\}$ effect a learning rate and, as such, were set such that the learning rate for the baseline context is nearly zero, following the assumption that participants will not change the way they perform these movements outside of the lab.
-The values for contexts outside of baseline were set identically within each experiment.
+Because the sCOIN model does not have a mechanism for the online creation of new contexts, relying instead of a fixed number of contexts, the number of existing contexts was set according to each experiment.
+For the experiments by \cite{Kim_Neural_2015}, to aid in learning of the two adaptations, the $\mu_0$ hyperparameters were set to -1 and 1 (plus the baseline of zero), which lead to the model learning the -40 and 40 visuomotor rotation angles, respectively.
+As the experiments by \cite{Oh_Minimizing_2019} have only one adaptation, this was not necessary and the new context was initiated with $\mu_0 = 0$.
+For the rest of the simulated experiments the focus was not on learning, but on the switching between known contexts, therefore we started simulations with models that had already learned the adaptations, setting the learned values to the real values used in each experiment.
 
-The star notation (e.g. $x_j^*$) denotes the real value used in the simulation of the task, which may be different from that assumed by the agent.
+The hyperparameters $\alpha_0$ and $\beta_0$ were set first for the baseline context such that the expected standard deviation of observations $\beta / \alpha$ roughly matched the observation noise in the task, i.e.
+$\beta_0 / \alpha_0 \sim \sigma_r + \sigma_u$, while keeping the values for $\beta_0$ and $\alpha_0$ very high, which, together with the high $\nu_0$ values, ensure that learning in this context is very slow.
+For the other contexts, the ratio $\beta_0 /  \alpha_0$ was set to be higher than the baseline, while keeping the individual values $\alpha_0$ and $\beta_0$ much lower, to speed up learning.
 
-For the simulations in \fref{fig:model}, the parameters were set as in the experiments by \cite{Davidson_Scaling_2004}, with two exceptions: (1) the cue noise, which is set to the values of 0, 0.003 and 0.33, for the low, medium and high values, respectively; and (2) the agent's observation noise $\sigma_r$, with values of 0.5, 1 and 2.
+The exact values for $\alpha_0$ and $\beta_0$ were set for each experiment such that $\beta_0 / \alpha_0 = 2 * (\sigma_r \sigma_a)$, where $\sigma_a$ is the standard deviation of the adaptation.
+The rationale behind this choice is that $\sigma_r$ and $\sigma_a$ determine the noise in the observations made by the model at each trial, and their sum is the value of $\beta / \alpha$ to which the leraning process converges with enough trials.
+We multiplied it by 2 in order to help in learning, specifically to make the \textit{a priori} standard deviation higher for the untrained contexts than for the baseline context.
+
+Of important note is the difference between the true observation noise and the expected observation noise in the simulations for the \cite{Davidson_Scaling_2004} experiments.
+The expected observation noise $\sigma_r$ was set to a higher value to reflect the fact that feedback in curl-force mechanical arm experiments, while devoid of any added noise, is more difficult for people to use to inform adaptation than in other types of experiments due to the nonlinear nature of the force.
+This fact is reflected in the high number of trials necessary for full adaptation in these experiments as compared to, for example, visuomotor rotation experiments.
+
+For the simulations in \fref{fig:model}, the parameters were set as in the experiments by \cite{Davidson_Scaling_2004}, with two exceptions: (1) the cue uncertainty, which is set to the values of 0 and 0.33, for the low and high values, respectively; and (2) the agent's observation noise $\sigma_r$, with values of 0.5 and 2.
 
@@
 \label{sec:interpreting-hyperparameters}
-$\mu$ determines the initial estimate of the adaptation, in the same units as
-the necessary adaptation. $\nu$ encodes how stable this hyper-prior is: higher
-values (e.g. 10,000) all but guarantee that the hyper-prior $\mu$ will not
-change its value after observations; In principle, enough evidence should still
-modify it, but that would not happen during an experiment. Smaller values
-(i.e. $\sim 1$) make $\mu$ follow evidence more freely. Note that as more
-observations are accumulated, $\nu$ becomes bigger and bigger, stabilizing the
-value of $\mu$.
+$\mu$ determines the initial estimate of the adaptation, in the same units as the necessary adaptation.
+$\nu$ encodes how stable this hyper-prior is: higher values (e.g.
+10,000) all but guarantee that the hyper-prior $\mu$ will not change its value after observations; In principle, enough evidence should still modify it, but that would not happen during an experiment.
+Smaller values (i.e.
+$\sim 1$) make $\mu$ follow evidence more freely.
+Note that as more observations are accumulated, $\nu$ becomes bigger and bigger, stabilizing the value of $\mu$.
 
-The hyper-parameters $\alpha$ and $\beta$ have a more complex effect. Note that
-the mean of a gamma distribution is $1 / (\alpha \nu)$; this mean is being used
-as the standard deviation of a Gaussian by the rest of the agent, which makes
-it an important measure of uncertainty. While setting the default
-hyper-parameters, the values used are $\alpha = 0.5 / \sigma_0$ and
-$\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the
-standard deviation of the force exerted by the environment, which controls the
-initial learning rate. This makes the initial standard deviation equal
-$\sigma_0$, which makes it consistent with the fixed-force model. The 0.5
-values ensure that uncertainty is large at the beginning and is greatly reduced
-during the experiment, but never to a point where it is so small that it makes
-trial-to-trial variation in the environment surprising. Changing this 0.5 would
-make the standard deviation change more quickly, making the model more or less
-precise in its predictions, independently of the volatility of the mean of the
-adaptation (via $\mu$).
+The hyper-parameters $\alpha$ and $\beta$ have a more complex effect.
+Note that the mean of a gamma distribution is $\beta / \alpha$; this mean is being used as the standard deviation of a Gaussian by the rest of the agent, which makes it an important measure of uncertainty.
+While setting the default hyper-parameters, the values used are, e.g., $\alpha = 0.5 / \sigma_0$ and $\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the standard deviation of the force exerted by the environment, which controls the initial learning rate.
+This makes the initial standard deviation equal $\sigma_0$, which makes it consistent with the fixed-force model.
+The 0.5 values ensure that uncertainty is large at the beginning and is greatly reduced during the experiment, but never to a point where it is so small that it makes trial-to-trial variation in the environment surprising.
+Changing this 0.5 would make the standard deviation change more quickly, making the model more or less precise in its predictions, independently of the volatility of the mean of the adaptation (via $\mu$).
 
-The baseline model defaults to different values that make it a lot more
-stable. The hyper-standard deviation of the mean is set to 10,000, which makes
-the mean entirely stable during the duration of the experiment. The values of
-$\alpha$ and $\beta$ are fixed regardless of $\sigma_0$ such that the standard
-deviation is 0.001 (compared that to the size of the adaptations in mechanical
-arm experiments, around 0.0125), and the hyper-parameters of the standard
-deviation are stable during the experiment.
+The baseline model defaults to different values that make it a lot more stable.
+The hyper-standard deviation of the mean is set to 10,000, which makes the mean entirely stable during the duration of the experiment.
+The values of $\alpha$ and $\beta$ are fixed regardless of $\sigma_0$ such that the standard deviation is 0.001 (compared that to the size of the adaptations in mechanical arm experiments, around 0.0125), and the hyper-parameters of the standard deviation are stable during the experiment.
 

### 2022-09-04 adaptation/code/figures.py

@@ -7,7 +7,6 @@ is in the paper. Those that do not have one are not paper figures, but
 auxiliary functions."""
 
 import ipdb
-from itertools import product
 
 import numpy as np
 from matplotlib import pyplot as plt
@@ -20,7 +19,6 @@ import simulations as sims
 import model
 import task_hold_hand as thh
 import pars
-import heald
 
 FIGURE_FOLDER = '../article/figures/'
 
@@ -44,8 +42,7 @@ def model_showoff(fignum=1, show=True, do_a=True):
     """
     runs = 50
     highlight_color = 'lightgray'
-    colors = ['black', 'tab:olive', 'tab:blue']
-    cycler = plt.cycler('color', colors)
+    colors = ['black', 'tab:olive', 'tab:blue', 'tab:orange', 'tab:green']
     tasks, agents_pars = sims.model_showoff(plot=False)
     numbers = tasks.shape
     figsize = (5, 6)
@@ -53,16 +50,23 @@ def model_showoff(fignum=1, show=True, do_a=True):
     height_ratios = 0.5 * np.ones(numbers[0] + 2)
     height_ratios[0] = 3
     height_ratios[1] = 0.3
-    gsbig = gs.GridSpec(numbers[0] + 2, numbers[1], figure=fig,
-                        height_ratios=height_ratios, hspace=0.1, wspace=0.1)
+    width_ratios = np.ones(numbers[1] + 2)
+    width_ratios[0] = 2
+    width_ratios[1] = 0.35
+    gsbig = gs.GridSpec(numbers[0] + 2, numbers[1] + 2, figure=fig,
+                        height_ratios=height_ratios, width_ratios=width_ratios,
+                        hspace=0.1, wspace=0.1)
     axis_diagram = fig.add_subplot(gsbig[0, :])
     ctx_switch = np.diff(
         tasks.reshape(-1)[0]['context_seq']).nonzero()[0][0] + 1
     num_trials = len(tasks.reshape(-1)[0]['context_seq'])
     axes = np.empty(numbers, dtype=object)
+    axes_state = fig.add_subplot(gsbig[2:, 0])
     for ix_row in np.arange(2, numbers[0] + 2):
-        for ix_col in np.arange(numbers[1]):
-            axes[ix_row - 2, ix_col] = fig.add_subplot(gsbig[ix_row, ix_col])
+        for ix_col in np.arange(2, numbers[1] + 2):
+            axes[ix_row - 2, ix_col -
[…]

### 2022-09-04 adaptation/code/model.py

@@ -1,8 +1,6 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/model.py
 
-import ipdb
-
 import numpy as np
 from scipy import stats
 import pandas as pd
@@ -48,7 +46,7 @@ class LeftRightAgent(object):
     # Free parameters with default values. Can be overwritten by __init__
     action_sd = pars['action_sd']  # SD of a Gaussian for action uncertainty
     cue_noise = 0.1  # If ix_cue is observed, the posterior over the
-                     # corresponding context is 1 - 2 * cue_noise.
+    # corresponding context is 1 - 2 * cue_noise.
     obs_sd = pars['obs_noise']
     angles = np.array([0, 0.01, -0.01])
     force_sds = pars['force_sd']  # np.array([0.01, 0.2, 0.2])
@@ -105,7 +103,9 @@ class LeftRightAgent(object):
         if sample_action is not None:
             self.sample_action = sample_action
         if prior_over_contexts is None:
-            self.prior_over_contexts = np.ones(self.num_contexts) / self.num_contexts
+            self.prior_over_contexts = 0.2 / self.num_contexts * np.ones(
+                self.num_contexts) / self.num_contexts
+            self.prior_over_contexts[0] = 0.8
         else:
             self.prior_over_contexts = prior_over_contexts
 
@@ -115,9 +115,7 @@ class LeftRightAgent(object):
         self.hand_position_history = [0]
         self.hand_position = 0
         self.action_history = [0]
-        exp_log_context = 0.2 / self.num_contexts * np.ones_like(self.angles)
-        exp_log_context[0] = 0.8
-        self.log_context = np.log(exp_log_context)
+        self.log_context = np.log(self.prior_over_contexts)
         self.log_context_history = [self.log_context]
         self.cue_history = [-1]
 
@@ -138,7 +136,8 @@ class LeftRightAgent(object):
         """
         self.is_reset = True
         if priors is None:
-            self.log_context = - np.log(self.num_contexts) * np.ones(self.num_contexts)
+            self.log_context = - \
+                np.log(self.num_contexts) * np.ones(self.num_contexts)
         else:
             if not np.isclose(np.sum(priors), 1):
                 raise ValueError('Provided priors do not add to 1')
@@ -188,7 +187,8 @@ class LeftRightAgent(object):
         # sampled_context = np.random.choice(np.arange(len(probs)), p=probs)
         # sampled_decision = bests[sampled_context] + \
[…]

### 2022-09-04 adaptation/code/simulations.py

@@ -191,7 +191,8 @@ def context_inference(pandata):
     """
     pandata = pandata.copy()
     # pandata.dropna(axis=0, how='any', inplace=True)
-    pandata['context_error'] = pandata['ix_context'].astype(float) != pandata['con_t']
+    pandata['context_error'] = pandata['ix_context'].astype(
+        float) != pandata['con_t']
     labels, _ = _define_grid()
     return pandata.groupby(labels + ['agent']).sum()['context_error']
 
@@ -243,7 +244,8 @@ def interactive_plot(pandata, axes=None, fignum=3):
         if flag_noindex:
             axes[0].set_title(indices[ix_ix])
         else:
-            title = ''.join([label + ': {}\n' for label in labels]) + 'Agent: {}\n'
+            title = ''.join(
+                [label + ': {}\n' for label in labels]) + 'Agent: {}\n'
             axes[0].set_title(title.format(*indices[ix_ix]))
         plt.draw()
         input = yield None
@@ -363,7 +365,7 @@ def plot_adaptation(pandata, axis=None, fignum=5):
               if column.startswith('mag_sd')]
     for color_x, error_x, magmu_x in zip(colors, errors, magmu):
         axis.plot(magmu_x, color=color_x, label='{} model'.format(color_x))
-        axis.fill_between(trial, magmu_x - 2 * error_x, magmu_x + 2 * error_x,
+        axis.fill_between(trial, magmu_x - error_x, magmu_x + error_x,
                           color=color_x, alpha=0.1)
     magmu = np.array(magmu)
     yrange = np.array([magmu.min(), magmu.max()]) * 1.1
@@ -568,7 +570,7 @@ def kim_2015(plot=True, axis=None, fignum=11):
                   'context_noise': 0.1,
                   'force_sds': np.ones(3),
                   'max_force': 60,
-                  'hyper_sd': 1,
+                  'hyper_sd': 10,
                   'obs_sd': 3,
                   'all_learn': True,
                   'learn_rate': 10,
@@ -604,7 +606,7 @@ def oh_2019(plot=True, axes=None, fignum=10):
 
     """
     task_20 = task_pars.copy()
-    task_20['obs_noise'] = 1.5
+    task_20['obs_noise'] = 2.8
     task_20['force_noise'] = 1 * np.ones(2)
     task_20['forces'] = [[0, 0], [1, 20]]
[…]

### 2022-09-04 adaptation/code/task_hold_hand.py

@@ -1,6 +1,5 @@
 # -*- coding: utf-8 -*-
 # ./adaptation/task_hold_hand.py
-import ipdb
 from datetime import datetime
 
 import numpy as np
@@ -80,12 +79,25 @@ def trial(c_trial, hand_position, agent, pars):
     c_hand_position = hand_position
     c_obs = sample_observation(hand_position, pars)
     action = agent.one_trial(c_obs, cue=cue)
+    # agent.cue_history.append(cue)
+    # agent.hand_position = c_hand_position
+    # agent.hand_position_history.append(c_obs)
+    # p_con = agent.log_context
+    # agent.infer_context(None, cue)
+    # # agent.update_magnitudes()
+    # action = agent.make_decision()
+    # agent.is_reset = False
     if context == pars['clamp_index']:
         n_hand_position = 0
         force = -action
     else:
         force = sample_force(context, pars)
         n_hand_position = c_hand_position + force + action
+    # agent.hand_position = n_hand_position
+    # agent.log_context = p_con
+    # # del agent.log_context_history[-1]
+    # agent.infer_context(n_hand_position, cue=cue, rewrite=True)
+    # agent.update_magnitudes()
     outs = [action, force, c_hand_position, n_hand_position,
             context]
     return outs

### 2023-01-17 adaptation/article/article.tex
Some reviewrs' comments addressed.

@@
 \usepackage{lineno}
+\usepackage{xcolor}  % For highlighting new content
+\usepackage{soul}    % For highlighting new content
 \linenumbers
@@
 % For revision
-\DeclareRobustCommand{\newcontent}[1]{#1}
+\DeclareRobustCommand{\newcontent}[1]{\hl{#1}}
 
@@
 a mechanical arm exerts a force on the participant's hand), experimenters are able to observe motor adaptation through the lens of motor error.
-Across many different motor adaptation experiments \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Davidson_Scaling_2004}, well-established phenomena have been observed: (i) the ability to recall previously-learned skills, called savings; (ii) the ability to return to unmodified dynamics, termed de-adaptation; (iii) the interference in motor learning between opposing manipulations in dynamics, called anterograde interference; (iv) spontaneous display of behavior consistent with a previously-learned adaptation, during trials where errors are forced to be zero, called spontaneous recovery.
+Across many different motor adaptation experiments \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Davidson_Scaling_2004}, well-established phenomena have been observed: (i) the ability to recall previously-learned skills \newcontent{upon re-exposure to a previous altered dynamic}, called savings; (ii) the ability to return to unmodified dynamics, termed de-adaptation; (iii) the interference in motor learning between opposing manipulations in dynamics, called anterograde interference; (iv) spontaneous display of behavior consistent with a previously-learned adaptation, during trials where errors are forced to be zero, called spontaneous recovery.
 
-To explain these phenomena, a number of computational models have been introduced, which adapt their motor commands after observing motor errors.
-The most studied are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, but Bayesian accounts have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between forward models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
+To explain these phenomena, a number of computational models have been introduced, \newcontent{in which motor commands are adapted based on observed motor errors}.
+The most well-studied models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, \newcontent{while Bayesian accounts have also been shown to provide} an alternative explanation for savings and quick de-adaptation in the form of switching between forward models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.
 
@@
 Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced by \cite{Heald_Contextual_2021}, called COIN (for context inference).
-\cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, but is informed by it, establishing a loop whereby context inference also informs motor learning.
+\cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, \newcontent{but both informs and is informed by it}.
 With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning in previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.
 
-In this work show that the process of context inference underlies more behavioral phenomena than previously shown.
-To do this, we used a minimal model for motor adaptation that includes context inference, which we derived by simplifying the COIN model (henceforth called sCOIN).
-We focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.
+In this work we show that the process of context inference underlies more behavioral phenomena than previously shown.
+\newcontent{We believe that these phenomena stem directly from the process of active inference and the uncertainty that comes with it in a Bayesian setting, as done by} \cite{Heald_Contextual_2021}, \newcontent{and not from the specific of the generative model used.
+To show how this Bayesian approach accounts for these phenomena}, we used a minimal model for motor adaptation that includes context inference, which we derived by simplifying the COIN model (henceforth called sCOIN).
+We focused on the effects of uncertain contextual information on switching behavior, especially during error-clamp trials, in which errors are forced to zero by experimenters.
 More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured.
@@
 
+\subsection{Experimental designs}
+In the following sections, we discuss the experiments done by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004} and \cite{Vaswani_Decay_2013}.
+In all experiments, participants were asked to do reaching or shooting movements from a starting position to a known target, and the experimental manipulations can be clasified into two categories: (1) curl forces with mechanical arms, and (2) visuomotor rotations.
+In this section, we briefly describe these two manipulations and leave the details of the experimental design to the sections below where each study is discussed.
+
+In the experiments with mechanical arms by \cite{Davidson_Scaling_2004,Vaswani_Decay_2013}, participants performed the movements while holding the handle of a mechanical arm; the arm exerted a force on the handle on each trial which depended on the speed of the participant's hand, creating a force that was perpendicular to the direction of movement, called a curl force (see \fref{fig:experiments}A).
+In these experiments, the participants' view of their hand and the mechanical arm was blocked, and they had to rely on a representation of their hand on a screen in front of them, which included the starting position and the target.
+The experiments by \cite{Davidson_Scaling_2004,Vaswani_Decay_2013} provided an accurate position of the participant's hand, with no noise added to it, and the cursor representing the hand is visible throughout the experiment; note that \cite{Vaswani_Decay_2013} added noise to the cursor on the screen during their third experiment, but only their first experiment is simulated here.
+
+\begin{figure}
+\centering
+\includegraphics[]{./figures/experiments.png}
+\caption{Mechanical arm and visuomotor rotation experiments.
+Green and blue circles represent the starting point and the target, respectively.
+Targets are placed on the circumference on both experimental paradigms.
+(A) Mechanical arm experiment.
+During adaptation trials, the force exerted by the arm on the handle pushes the hand away from the straight line.
+In error-clamp trials, the mechanical arm creates resistance against movements perpendicular to the straight line to the target.
+(B) Visuomotor rotation experiment.
+The correspondence between hand movements and the cursor is rotated (visuomotor rotation).
+The error is measured as the arc between the target and the place at which the trajectory (dotted line) crossed the circumference.}
+\label{fig:experiments}
+\end{figure}
+
+When the mechanical arm was set to exert no force, movements made by human participants were close to straight lines between the starting position and the target.
+When the curl force was applied, movements deviated from this straight-line path, creating curved movements that nevertheless reached the target, as participants adjusted their trajectory before the end of the movement.
+As participants repeated the movements over many trials, they learned to adapt to these curl forces and went back to performing near straight-line movements.
+During error-clamp trials, the mechanical arm engaged a very stiff spring that caused participants' movements to be a straight line, regardless of the participants' applied forces (see \fref{fig:experiments}A).
+This allowed experimenters to present a zero-error feedback to the participant, while at the same time measuring the forces participants applied onto the spring.
+
+Because participants could see the cursor, they adapted their movements in-flight, always reaching the target.
+Errors were measured as deviations from the straight line connecting starting and target positions.
+
+In the experiments with visuomotor experiments by \cite{Kim_Neural_2015,Oh_Minimizing_2019}, participants moved a cursor on a screen by using a joystick \cite{Kim_Neural_2015}, or on the surface of a mirror using the pen of a digitizer pad \cite{Oh_Minimizing_2019}.
+Movements were done from a starting position in the middle of the screen to targets appearing on a circumference of 10cm in radius.
+The experimental manipulation took the form of a rotation between the hand movements and the cursor of the screen (see \fref{fig:experiments}B), centered on the starting position.
+A clockwise rotation of 20 degrees means that a hand movement forward translated into a movement towards the right on the screen or mirror.
+\cite{Kim_Neural_2015} used the convention that clockwise rotations are negative, while \cite{Oh_Minimizing_2019} used the opposite.
+
+\cite{Oh_Minimizing_2019} added Gaussian noise to the visuomotor rotation at each trial.
+With this, each trial had a different visuomotor rotation, centered at 10 or 20 degrees, and with a standard deviation of 0.5 or 4 degrees, depending on the condition.
+In this work, we focused in the comparison between conditions 1a and 3, with means of 20 and 10 degrees, respectively, and standard deviation of 0.5 degrees.
+
+In contrast to experiments with a mechanical arm, movements in visuomotor rotation experiments are always close to a straight line.
+Errors are measured in degrees, as the deviation between the target and the point in the 10cm circumference where the cursor crossed it (see \fref{fig:experiments}B).
+
 \subsection{Modeling context-dependent adaptation}
 We focused on three main components of the COIN model: (1) context inference, (2) motor adaptation and (3) action selection.
-The processes defined by these component occur in this order, and each component informs the ones that follow.
+The processes defined by these components occur in this order, and each component informs the ones that follow.
 
@@
 Because we seeked to focus on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference, we chose to fix the participant-inferred transition probabilities between contexts, as well as the total number of contexts; in our simulations, we assume that participants have already inferred the real values of these quantities.
-Because we mainly focus on switching behavior, as well as error-clamp trials (both of which involve alread-learned adaptations), these simplifications to the model have minimal effects on our results.
+Because we mainly focus on switching behavior, as well as error-clamp trials (both of which involve already-learned adaptations), these simplifications to the model have minimal effects on our results.
 See Methods for more details on the sCOIN model, as well as the differences between COIN and sCOIN.
@@
 \includegraphics[]{./figures/figure_1.png}
-\caption{Schematic representation of the model and illustrative simulations.
+\caption{Schematic representation of the sCOIN model and illustrative simulations.
 (A) Inference done by the model at every trial.
 Clear circles represent the observables, i.e.
-motor commands $u_t$ and direct observations $y_t$ (e.g.
+\newcontent{the information available to the model (and the brain), and used for adaptation}: motor commands $u_t$ and direct observations $y_t$ (e.g.
 cursor position).
@@
 The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment.
-To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to the curl forces exerted by a mechanical arm on the participants' hands as they perform pointing movements.
+To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}: \newcontent{participants performed reaching movements from a central position to fixed targets on a circumference, then back to the central position.
+All targets were positioned on the horizontal plane, at shoulder height.
+Movements were performed holding a robotic arm which exerted velocity-dependent rotary forces, which, when ignored, would move the participant's hand away from the straight-line path to the target.
+See the following section for more details on the experiments by} \cite{Davidson_Scaling_2004}.
+The results of these simulations can be seen in \fref{fig:model}B.
 
 The key to an intuitive understanding of the results presented below is to observe what happens when the presence and reliability of contextual cues is varied, as well as the perceptual noise on the position of the hand.
+\newcontent{To do this in our simulations, we added noise to the categorical variable that encodes contextual cues, and to the observations (see Methods for details), at two different levels (low and high noise).
+We simulated 50 participants, with 30 trials per participant}.
 In \fref{fig:model}B, a 2x2 grid of results is shown: each simulation in this grid is a combination of low or high contextual cue uncertainty (where high uncertainty is equivalent to presenting no cues), and low or high perceptual noise (representing how well participants can detect deviations from the straight-line movements).
@@
 
-As we show below, these effects are at the heart of the behavioral phenomena observed in the experiments by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004}, and \cite{Vaswani_Decay_2013}, which we directly simulate in this work, alongside others that we discuss in the Discussion section.
+As we show below, these effects are at the heart of the behavioral phenomena observed in the experiments by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004}, and \cite{Vaswani_Decay_2013}, \newcontent{and our simulated data qualitatively matches their results, as well as} others that we discuss in the Discussion section.
 
@@
 Similarly, one can define context $A/2$, with the same direction of adaptation as $A$, but half the magnitude.
-Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter $E$.
+Finally, many experiments include a block of error-clamp trials at the end of the experiment, \newcontent{in which the mechanical arm forces the participant's hand into a straight-line trajectory towards the target by counteracting any lateral forces}, eliminating any error for the trial; we represent these with the letter $E$.
 
 With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of $O-A-B-E$, or $O-A-(-A)-E$, which means that the participant goes through a block of trials with no external force applied ($O$), a number of trials with a clockwise curl force ($A$), a block with counter-clockwise forces ($B$), and finally a block with error-clamp trials ($E$).
-With repeated contexts \citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1-A_1-O_2-\ldots$.
+\newcontent{When participants are exposed to the same context multiple times }\citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1-A_1-O_2-\ldots$, \newcontent{where $A_n$ is context $A$, presented to the participant for the n-th time.}
 
 \subsubsection{Cue- and sensory feedback uncertainty affects switching behavior}
-The term 'savings' refer to the ability to remember a previously-learned adaptation and apply it without having to re-learn it.
+The term 'savings' refers to the ability to remember a previously-learned adaptation, \newcontent{observed as an instant recall of the previously-learned adaptation} \cite[e.g.][]{Kim_Neural_2015,Oh_Minimizing_2019}, or as accelerated re-learning \cite[e.g.][]{Kojima_Memory_2004}; in the case of instant recall, the phenomenon is often called 'switching'.
 Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}.
@@
 In \fref{fig:oh-2019}A we show the results of simulations with the model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}.
-The correspondence between simulations and experimental results can be seen in the switches between contexts in the center panel, i.e.
-when the solid gray line (representing the true context) switches between 40, 0 and -40; in the two right-most panels, the thick black line, representing the agent's adaptation, quickly follows these switches.
+The correspondence between simulations and experimental results can be seen in the switches between contexts in \newcontent{the third and fourth columns}, where the solid gray line (representing the true context) switches between 40, 0 and -40, and the thick black line, representing the agent's adaptation, quickly follows these switches.
 
@@
 Data from our simulations (first three columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column).
-Experimental and simulated data was, in all three experiments, averaged across all participants.
+Experimental and simulated data was, in all three experiments, averaged across all participants (20, 11 and 11 participants, respectively).
+We simulated 1, 11 and 11 participants, respectively, as the simulations for \cite{Kim_Neural_2015} had almost no variability.
 In the first column, the simulated context inference is represented by the posterior probabilities over all available contexts.
-Each color (black, blue, green) represents a different context, with black always representing the baseline (i.e.
+Each color (black, orange, green) represents a different context, with black always representing the baseline (i.e.
 no adaptation).
@@
 how they move normally).
-Under this single-context explanation, however, savings do not exist, and adaptations need to be learned anew ever time a context changes.
+Under this single-context explanation, however, savings do not exist, and adaptations need to be learned anew every time a context changes.
 On the other hand, savings are present under the sCOIN model, their manifestations being diminished by the slower context inference.
@@
 \caption{Motor error when switching back to a previously-learned adaptation.
-(A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the SCOIN model are shown.
-In the first column, each panel represents context inference for one group of participants (top: group A; bottom: group 3A), with each line representing the posterior probability of a context (black for the baseline O).
+(A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the sCOIN model are shown.
+In the first column, each panel represents context inference for one group of participants (top: group $-A$; bottom: group $3A$), with each line representing the posterior probability of a context (black for the baseline O).
 The second column represents the error made by our simulated agent after returning to the previously-learned context, with blue and yellow representing groups -A and 3A, respectively.
-The last panel represents the same data, from the \cite{Davidson_Scaling_2004} experiment.
+The last panel represents the same data from eight participants per group (mean + SEM), from the \cite{Davidson_Scaling_2004} experiment.
 All panels share the x-axis, representing the number of trials elapsed since the switch to the new context.
-All simulations were executed 32 times per group, to obtain a reliable mean; shaded areas represent the standard deviation across all simulations.
+All simulations were executed 32 times per group, to obtain a reliable mean; shaded areas represent the SEM across all simulations.
 (B) Simulated results for an experiment similar to \cite{Davidson_Scaling_2004}, but changing the contexts seen by the two groups from -A to -2A for the first group, and from 3A to 4A for the second group.
@@
 First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments.
-To do this, one group adapts in a $O-A-(-2A)$ paradigm, while the other group adapts in a $0-A-4A$ paradigm.
+To do this, one group adapts in a $O-A-(-2A)$ paradigm, while the other group adapts in a $O-A-4A$ paradigm.
 As in the original experiment, the second contexts ($-2A$ for one group, $4A$ for the other) are equally spaced from the first context.
@@
 
-\cite{Vaswani_Decay_2013} also found participant-specific delays before the decay to baseline started after the error-clamp phase begun.
+\cite{Vaswani_Decay_2013} also found participant-specific delays before the decay to baseline started after the error-clamp phase began.
 As can be seen in \fref{fig:vaswani-2013}D, each simulated participant follows a different path of decay, with variations caused directly by perceptual noise.

### 2023-01-17 adaptation/code/figures.py

@@ -327,9 +327,9 @@ def _one_oh(pandota, oh_reps, colors, context_color, adapt, axes):
     sns.lineplot(data=pandota, x='trial', y='adapt', ax=axes[2], color='black')
     for c_con, c_state, c_color in zip(ohcon, magmustr, colors):
         sns.lineplot(data=pandota, x='trial', y=c_con, color=c_color,
-                     ax=axes[0])
+                     ax=axes[0], errorbar='se')
         sns.lineplot(data=pandota, x='trial', y=c_state, color=c_color,
-                     ax=axes[1])
+                     ax=axes[1], errorbar='se')
     for axis in axes:
         axis.set_xlabel('Trial')
         axis.set_ylabel('')
@@ -468,25 +468,25 @@ def _davidson_trio(repeats, color_list, agent_pars, tasks, names, axes, labels,
     ymax = data.loc[data['trial'] > ran[0], 'pos(t)'].max()
     ymin = data.loc[data['trial'] > ran[0], 'pos(t)'].min()
     sns.lineplot(data=data, x='trial', y='pos(t)', ax=axes[0],
-                 hue='Group', palette=colors, ci='sd')
+                 hue='Group', palette=colors, errorbar='se')
     datum_a = data.query('Group == @names[0]')
     datum_b = data.query('Group == @names[1]')
     sns.lineplot(data=datum_a, x='trial', y='con0', ax=axes[1],
-                 color='black', label=labels[0][0])
+                 color='black', label=labels[0][0], errorbar='se')
     sns.lineplot(data=datum_a,
                  x='trial', y='con1', ax=axes[1], label=labels[0][1],
-                 color='tab:green')
+                 color='tab:green', errorbar='se')
     sns.lineplot(data=datum_a,
                  x='trial', y='con2', ax=axes[1], label=labels[0][2],
-                 color='tab:blue')
+                 color='tab:blue', errorbar='se')
     sns.lineplot(data=datum_b, x='trial', y='con0', ax=axes[2],
-                 color='black', label=labels[1][0])
+                 color='black', label=labels[1][0], errorbar='se')
     sns.lineplot(data=datum_b,
                  x='trial', y='con1', ax=axes[2], label=labels[1][1],
-                 color='tab:green')
+                 color='tab:green', errorbar='se')
     sns.lineplot(data=datum_b,
[…]
